### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@aa51b8efb1b7e6328ebc9aba1416f8aa295c4ca8 🚀

%!TEX TS-program = xelatex
\documentclass[11pt]{article}

\usepackage[english]{babel}

\usepackage{amsmath,amssymb,amsfonts}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{stix2}
\usepackage[scaled]{helvet}
\usepackage[scaled]{inconsolata}

\usepackage{lastpage}

\usepackage{setspace}

\usepackage{ccicons}

\usepackage[hang,flushmargin]{footmisc}

\usepackage{geometry}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\makeatletter
\newcounter{tableno}
\newenvironment{tablenos:no-prefix-table-caption}{
  \caption@ifcompatibility{}{
    \let\oldthetable\thetable
    \let\oldtheHtable\theHtable
    \renewcommand{\thetable}{tableno:\thetableno}
    \renewcommand{\theHtable}{tableno:\thetableno}
    \stepcounter{tableno}
    \captionsetup{labelformat=empty}
  }
}{
  \caption@ifcompatibility{}{
    \captionsetup{labelformat=default}
    \let\thetable\oldthetable
    \let\theHtable\oldtheHtable
    \addtocounter{table}{-1}
  }
}
\makeatother

\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\let\PBS=\PreserveBackslash

\usepackage[breaklinks=true]{hyperref}
\hypersetup{colorlinks,%
citecolor=blue,%
filecolor=blue,%
linkcolor=blue,%
urlcolor=blue}
\usepackage{url}

\usepackage{caption}
\setcounter{secnumdepth}{0}
\usepackage{cleveref}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}

\usepackage{longtable}
\usepackage{booktabs}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

\usepackage{lineno}
\usepackage[nolists,noheads]{endfloat}

\pagestyle{plain}

\tolerance=1
\emergencystretch=\maxdimen
\hyphenpenalty=10000
\hbadness=10000

\doublespacing

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily The coevolutionary mosaic of
bat-betacoronaviruses spillover risk}
\vskip 5em

%
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}\quad %
Peregrin\,Took%
%
\,\textsuperscript{3,4}\quad %
Merriadoc\,Brandybuck%
%
\,\textsuperscript{5,4,‡}

\textsuperscript{1}\,Université de
Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
Sciences\quad \textsuperscript{3}\,Inn of the Prancing
Pony\quad \textsuperscript{4}\,Fellowship of the
Ring\quad \textsuperscript{5}\,Green Dragon Inn

\textsuperscript{‡}\,These authors contributed equally to the work\\

\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\

\vfill
This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
Last revision: \emph{\today}

\clearpage
\thispagestyle{empty}

\vfill

        {\bfseries Purpose:}\,This template provides a series of scripts
to render a markdown document into an interactive website and a series
of PDFs.\\%
        {\bfseries Motivation:}\,It makes collaborating on text with
GitHub easier, and means that we never need to think about the
output.\\%
        {\bfseries Internals:}\,GitHub actions and a series of python
scritpts. The markdown is handled with \texttt{pandoc}.\\%

\vfill

\clearpage
\linenumbers
\pagestyle{normal}

This template uses \texttt{pandoc} (and a few additional python glue
scripts) to facilitate the production of scientific articles using a
standard markdown file. The objective is to ensure that standard
markdown (with the important exception of the \texttt{pandoc-crossref}
citation markup) will be rendered into an interactive website (which
allows collaborative annotations with the \texttt{hypothes.is}
platform), a ``draft'' style PDF (double-spaced, numbered lines, figures
at the end), and a ``preprint'' style PDF (with slightly more
reader-friendly pagination).

The core bit of configuration is the \texttt{metadata.json} file, which
handles information about authorship, affiliations, the abstract,
keywords, etc. All documents will be deployed to \texttt{gh-pages}
\emph{only} on push events from the \texttt{main} branch. All of the
artifacts will be built when doing pull requests, so you can check that
merging a branch is \emph{not} going to cause the compilation of the
documents to fail; indeed, you can download the artifacts produced
during the run, to check the PDF and html files. The website is only
updated from the \texttt{main} branch.

The workflow is \emph{very} GitHub based, and so the manuscript file
\emph{is} the \texttt{README.md} - this is not going to be a huge issue
as 90\% of the markdown is standard, with the exception of the citations
and mathematics, so this will render (mostly) like a normal README file.

\hypertarget{deploying-the-template}{%
\section{Deploying the template}\label{deploying-the-template}}

The process of deploying this template has been \emph{greatly}
streamlined from previous versions:

\begin{itemize}
\tightlist
\item
  Click on the ``Use this template'' button
\item
  Edit \texttt{README.md} with your own text, commit, and push
\item
  This push will trigger the first build - the builds are only active on
  the \texttt{main} branch (\emph{not} \texttt{master}!), and on pull
  requests
\item
  Go to \texttt{http://you.github.io/repo-name/} to view the html
  version, and get access to the PDFs
\item
  Add your references to the \texttt{references.bib} file
\item
  Edit the \texttt{metadata.json} file to add the title, abstract,
  authors
\end{itemize}

In particular, note that \emph{you do not need} to create a personnal
access token to deploy to \texttt{gh-pages} (from where the website is
served).

\hypertarget{the-metadata-file}{%
\section{The metadata file}\label{the-metadata-file}}

\hypertarget{general-information}{%
\subsection{General information}\label{general-information}}

The title is a field in the \texttt{metadata.json}:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{\{}
    \DataTypeTok{"title"}\FunctionTok{:} \StringTok{"Preprint template"}
\FunctionTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{authorship}{%
\subsection{Authorship}\label{authorship}}

Authors are listed as objects in the \texttt{authors} block. Each author
is specified as follows:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{\{}
      \DataTypeTok{"familyname"}\FunctionTok{:} \StringTok{"Bob"}\FunctionTok{,}
      \DataTypeTok{"givennames"}\FunctionTok{:} \StringTok{"Alice"}\FunctionTok{,}
      \DataTypeTok{"email"}\FunctionTok{:} \StringTok{"alice.bob@u.edu"}\FunctionTok{,}
      \DataTypeTok{"orcid"}\FunctionTok{:} \StringTok{"0000{-}0000{-}0000{-}0001"}\FunctionTok{,}
      \DataTypeTok{"affiliations"}\FunctionTok{:} \OtherTok{[}
        \StringTok{"Affiliation 1"}\OtherTok{,}
        \StringTok{"Affiliation 2"}
      \OtherTok{]}\FunctionTok{,}
      \DataTypeTok{"status"}\FunctionTok{:} \OtherTok{[}\StringTok{"corresponding"}\OtherTok{,} \StringTok{"equal"}\OtherTok{]}
    \FunctionTok{\}}
\end{Highlighting}
\end{Shaded}

The \texttt{email} field is recommended for all authors. The
\texttt{status} field is only useful for the corresponding author, and
to denote equal contributions. These informations are rendered on the
initial page. If an \texttt{orcid} is given, it will be linked on the
HTML and PDF versions.

Note that there is \emph{no need} to number the affiliations - a small
python script will take care of this automatically.

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

This template supports three types of abstracts, indicated in the
metadata file as \texttt{abstract}:

A regular \texttt{abstract} is defined as

\begin{Shaded}
\begin{Highlighting}[]
\ErrorTok{"abstract":} \ErrorTok{"A} \ErrorTok{very} \ErrorTok{long} \ErrorTok{string"}
\end{Highlighting}
\end{Shaded}

An itemized abstract is an array of strings, each representing a bullet
point:

\begin{Shaded}
\begin{Highlighting}[]
\ErrorTok{"abstract":} \OtherTok{[}
    \StringTok{"Point 1"}\OtherTok{,}
    \StringTok{"Point 2"}
\OtherTok{]}
\end{Highlighting}
\end{Shaded}

A structured abstract is an object with key-value pairs :

\begin{Shaded}
\begin{Highlighting}[]
\ErrorTok{"abstract":} \FunctionTok{\{}
    \DataTypeTok{"Location"}\FunctionTok{:} \StringTok{"Worldwide"}\FunctionTok{,}
    \DataTypeTok{"Organisms"}\FunctionTok{:} \StringTok{"Mammals"}
\FunctionTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{citation-style}{%
\subsection{Citation style}\label{citation-style}}

The \texttt{citationstyle} key corresponds to the name, with
\texttt{.csl} ommited, of a CSL stylesheet stored in the
\href{https://github.com/citation-style-language/}{citation style
language} repository. Note that there is no difference between main and
dependent styles, the build engine will take the correct steps to get
the correct style. The default is
\texttt{"citationstyle":\ "ecology-letters"}. There is a longer section
about references management later on.

\hypertarget{references-management}{%
\section{References management}\label{references-management}}

The references are managed by \texttt{pandoc}. Note that we \emph{do
not} use \texttt{pandoc-citeproc}, which was an external module for
older \texttt{pandoc} versions. References \emph{must} be stored in a
\texttt{references.bib} file, and that it would make sense to order it
alphabetically by key.

We use \href{https://www.zotero.org/}{Zotero} for references management,
and for the lab's manuscripts, we work from folders in a shared library
(with a folder for every manuscript).

It is recommedned to use the
\href{https://retorque.re/zotero-better-bibtex/}{Better BibTeX} plugin
for citation key generations, and auto-export of the shared library to
the \texttt{references.bib} file. We use a citation key format meant to
convey information on the author (first author full name), date (complet
year), and title (first three letters of the first two non-stop words).
It must be set in the Better BibTeX preferences as (you might need to
remove the line changes):

\begin{verbatim}
[auth:fold]
[year]
[title:fold:nopunctordash:skipwords:lower:select=1,1:substring=1,3:capitalize]
[title:fold:nopunctordash:skipwords:lower:select=2,2:substring=1,3:capitalize]
\end{verbatim}

It is a good idea to configure Better BibTeX to auto-export on change,
and to remove a lot of fields that are not strictly speaking required
for references. The list of fields we usually ignore is:

\begin{verbatim}
abstract,copyright,annotation,file,pmid,month,shorttitle,keywords
\end{verbatim}

The citations are done using the normal markdown syntax, where
\texttt{@Elton1927AniEco} produces Elton (1927), and
\texttt{{[}@Camerano1880EquViv{]}} produces (Camerano 1880).

\hypertarget{figures-tables-and-other-floats}{%
\section{Figures, Tables, and other
floats}\label{figures-tables-and-other-floats}}

Note that you can wrap the text of legends for both figures and tables.
This avoids the issue of having very long lines.

\hypertarget{mathematics}{%
\subsection{Mathematics}\label{mathematics}}

The following equation

\begin{equation}\protect\hypertarget{eq:eq1}{}{J'(p) = \frac{1}{\text{log}(S)}\times\left(-\sum p \times \text{log}(p)\right)}\label{eq:eq1}\end{equation}

is produced using

\begin{Shaded}
\begin{Highlighting}[]
\SpecialStringTok{$$J\textquotesingle{}(p) = }\SpecialCharTok{\textbackslash{}frac}\SpecialStringTok{\{1\}\{}\SpecialCharTok{\textbackslash{}text}\NormalTok{\{log\}}\SpecialStringTok{(S)\}}\SpecialCharTok{\textbackslash{}times}\SpecialStringTok{ ... $$}\NormalTok{ \{\#eq:eq1\}}
\end{Highlighting}
\end{Shaded}

and can be referenced using \texttt{@eq:eq1}, which will result in
eq.~\ref{eq:eq1}. Note that because we use \texttt{pandoc-crossref}, the
label ``eq.'' will be generated automatically.

\hypertarget{tables}{%
\subsection{Tables}\label{tables}}

Table legends go on the line after the table itself. To generate a
reference to the table, use \texttt{\{\#tbl:id\}} -- then, in the text,
you can use \texttt{\{@tbl:id\}} to refer to the table. For example, the
table below is tbl.~\ref{tbl:id}. You can remove the \emph{table} in
front by using \texttt{!@tbl:id}, or force it to be capitalized with
\texttt{\textbackslash{}*tbl:id}.

\hypertarget{tbl:id}{}
\begin{longtable}[]{@{}rrrrl@{}}
\caption{\label{tbl:id}This is a table, and its identifier is
\texttt{id} -- we can refer to it using \texttt{\{@tbl:id\}}. Note that
even if the table legend is written below the table itself, it will
appear on top in the PDF document.}\tabularnewline
\toprule
Sepal.Length & Sepal.Width & Petal.Length & Petal.Width &
Species\tabularnewline
\midrule
\endfirsthead
\toprule
Sepal.Length & Sepal.Width & Petal.Length & Petal.Width &
Species\tabularnewline
\midrule
\endhead
5.1 & 3.5 & 1.4 & 0.2 & setosa\tabularnewline
5.0 & 3.6 & 1.4 & 0.2 & setosa\tabularnewline
5.4 & 3.9 & 1.7 & 0.4 & setosa\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{figures}{%
\section{Figures}\label{figures}}

Figures can have a legend -- all figures \emph{must} be in the
\texttt{figures/} folder of the project, as it is also used for the
website. We recommend to use good resolution images, rather than PDFs,
or at least to have multiple versions available.

\begin{verbatim}
![This is the legend of the figure...](figures/figure.png){#fig:figure}
\end{verbatim}

\begin{figure}
\hypertarget{fig:figure}{%
\centering
\includegraphics{figures/figure.png}
\caption{This is the legend of the figure, which will be shown in the
margin in preprint mode, and underneath the figure in draft mode. The
legend can contain references, etc. It is advised to use a resolution of
at least 600dpi for the figures.}\label{fig:figure}
}
\end{figure}

We can now use \texttt{@fig:figure} to refer to fig.~\ref{fig:figure}.

\hypertarget{example-text}{%
\section{Example text}\label{example-text}}

Connectance, defined as the ratio of realized interactions on the total
number of potential interactions, is one of the most common descriptor
of network structure. In a bipartite network with \(T\) species at the
top, and \(B\) at the bottom, having a total of \(L\) interactions, it
is defined as \(Co = L/(T\times B)\). Connectance has a lower bound, as
the network cannot have fewer interactions that the number of species in
its more speciose level -- the minimal connectance is therefore
\(c_m = \text{max}(T,B)\). This makes the connectance of networks of
different sizes difficult to compare, especially since bipartite
networks tends to have a low connectance. For this reason, we used a
corrected version of connectance, defined as

\begin{equation}\protect\hypertarget{eq:cstar}{}{Co^\star=\frac{L-c_m}{T\times B-c_m} \,.}\label{eq:cstar}\end{equation}

\hypertarget{this-is-a-subsection}{%
\subsection{This is a subsection}\label{this-is-a-subsection}}

This takes values between 0 (the network has the minimal number of
interactions) and 1 (all species are connected), but is robust to
variations in species richness.

\hypertarget{this-is-another-subsection}{%
\subsection{This is another
subsection}\label{this-is-another-subsection}}

This takes values between 0 (the network has the minimal number of
interactions) and 1 (all species are connected), but is robust to
variations in species richness.

\hypertarget{some-non-standard-maths}{%
\subsection{Some non-standard maths}\label{some-non-standard-maths}}

The phylogenetic reconstruction of \(\hat{\mathscr{L}}\) and
\(\hat{\mathscr{R}}\) has an associated uncertainty, represented by the
breadth of the uniform distribution associated to each of their entries.
Therefore, we can use this information to assemble a
\emph{probabilistic} metaweb in the sense of
(\textbf{Poisot2016StrPro?}), \emph{i.e.} in which every interaction is
represented as a single, independent, Bernoulli event of probability
\(p\).

Specifically, we have adopted the following approach. For every entry in
\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\), we draw a value from
its distribution. This results in one instance of the possible left
(\(\hat{\mathscr{l}}\)) and right (\(\hat{\mathscr{r}}\)) subspaces for
the Canadian metaweb. These can be multiplied, to produce one matrix of
real values. Because the entries in \(\hat{\mathscr{l}}\) and
\(\hat{\mathscr{r}}\) are in the same space where \(\mathscr{L}\) and
\(\mathscr{R}\) were originally predicted, it follows that the threshold
\(\rho\) estimated for the European metaweb also applies. We use this
information to produce one random Canadian metaweb,
\(N = \hat{\mathscr{L}}\)\(\hat{\mathscr{R}}' \ge \rho\).

Because the intervals around some trait values can be broad (in fact,
probably broader than what they would actually be, see \emph{e.g.}
\textbf{Garland1999IntPhy?}), we repeat the above process
\(2\times 10^5\) times, which results in a probabilistic metaweb \(P\),
where the probability of an interaction (here conveying our degree of
trust that it exists given the inferred trait distributions) is given by
the number of times where it appears across all random draws \(N\),
divided by the number of samples. An interaction with \(P_{i,j} = 1\)
means that these two species were predicted to interact in all
\(2\times 10^5\) random draws, etc..

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Camerano1880EquViv}{}%
Camerano, L. (1880). Dell'equilibrio dei viventi merce la reciproca
distruzione. \emph{Atti Della R. Accad. Delle Sci. Torino}, 15,
393--414.

\leavevmode\hypertarget{ref-Elton1927AniEco}{}%
Elton, C.S. (1927). \emph{Animal ecology}. University of Chicago Press.

\end{CSLReferences}

\end{document}

### 2022-04-06 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@aa51b8efb1b7e6328ebc9aba1416f8aa295c4ca8 🚀

--- 2022-04-06 draft.tex
@@
 %!TEX TS-program = xelatex
-\documentclass[11pt]{article}
+\documentclass[10pt,oneside]{article}
+\usepackage[fontsize=9pt]{scrextend}
 
@@
 \newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\usepackage[table,dvipsnames]{xcolor}
 
-\usepackage{lineno}
-\usepackage[nolists,noheads]{endfloat}
+\geometry{includemp,
+            letterpaper,
+            top=2.4cm,
+            bottom=2.4cm,
+            left=1.0cm,
+            right=1.0cm,
+            marginparwidth=5cm,
+            marginparsep=1.0cm}
+
+\usepackage[singlelinecheck=off]{caption}
+
+\captionsetup{
+  font={small},
+  labelfont={bf},
+  format=plain,
+  labelsep=quad
+}
+
+\usepackage{floatrow}
+
+\floatsetup[figure]{margins=hangright,
+              facing=no,
+              capposition=beside,
+              capbesideposition={center,outside},
+              floatwidth=\textwidth}
+
+% \floatsetup[table]{margins=hangright,
+%              facing=no,
+%              capposition=beside,
+%              capbesideposition={center,outside},
+%              floatwidth=\textwidth}
 
@@
 
-\tolerance=1
-\emergencystretch=\maxdimen
-\hyphenpenalty=10000
-\hbadness=10000
+\setcounter{secnumdepth}{5}
 
-\doublespacing
+\usepackage{titlesec}
+
+\titleformat{\section}[block]
+{\normalfont\large\sffamily}
+{\thesection}{.5em}{\titlerule\\[.8ex]\bfseries}
+
+\titleformat{\subsection}[runin]
+{\normalfont\fontseries{b}\selectfont\filright\sffamily}
+{\thesubsection.}{.5em}{}
+
+\titleformat{\subsubsection}[runin]
+{\normalfont\itshape\rmfamily\bfseries}{\thesubsubsection}{1em}{}
+
+\fancypagestyle{firstpage}
+{
+   \fancyhf{}
+   \renewcommand{\headrulewidth}{0pt}
+   \fancyfoot[R]{\footnotesize\ccby}
+   \fancyfoot[L]{\footnotesize\sffamily\today}
+}
 
@@
 }
+
+\usepackage{tikz}
 \begin{document}
-\raggedright
-\thispagestyle{empty}
-{\Large\bfseries\sffamily The coevolutionary mosaic of
-bat-betacoronaviruses spillover risk}
-\vskip 5em
+\tikz [remember picture, overlay] %
+\node [shift={(-0.6in,1.1cm)},scale=0.2,opacity=0.4] at (current page.south east)[anchor=south east]{\includegraphics{logo}};%
+\pagestyle{normal}
+\thispagestyle{firstpage}
 
+\newcommand{\colorRule}[3][black]{\textcolor[HTML]{#1}{\rule{#2}{#3}}}
+
+\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
+bat-betacoronaviruses spillover risk}}}
+
+\medskip
+\begin{flushleft}
+{\small
 %
@@
 %
-\,\textsuperscript{1,2,‡}\quad %
+\,\textsuperscript{1,2,‡}, %
 Peregrin\,Took%
 %
-\,\textsuperscript{3,4}\quad %
+\,\textsuperscript{3,4}, %
 Merriadoc\,Brandybuck%
@@
 \,\textsuperscript{5,4,‡}
-
-\textsuperscript{1}\,Université de
-Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
-Sciences\quad \textsuperscript{3}\,Inn of the Prancing
-Pony\quad \textsuperscript{4}\,Fellowship of the
-Ring\quad \textsuperscript{5}\,Green Dragon Inn
-
+\vskip 1em
+\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
+Centre for Biodiversity Sciences; \textsuperscript{3}\,Inn of the
+Prancing Pony; \textsuperscript{4}\,Fellowship of the
+Ring; \textsuperscript{5}\,Green Dragon Inn\\
 \textsuperscript{‡}\,These authors contributed equally to the work\\
-
+\vskip 1em
 \textbf{Correspondance to:}\\
 Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
+}
+\end{flushleft}
 
-\vfill
-This work is released by its authors under a CC-BY 4.0 license\hfill\ccby\\
-Last revision: \emph{\today}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
+\noindent
 
-\clearpage
-\thispagestyle{empty}
-
-\vfill
+\marginpar{\vskip 1em\flushright
+{\small{\bfseries Keywords}:\par
+pandoc\\pandoc-crossref\\github actions\\}
+}
 
@@
 
-\vfill
-
-\clearpage
-\linenumbers
-\pagestyle{normal}
+\vskip 2em
+\makebox[0pt][l]{\colorRule[CCCCCC]{2.0\textwidth}{0.5pt}}
+\vskip 2em
 

### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ce714f31455d6ef4bdd57d7ff199cef40f1e81a6 🚀

@@
 
-This template uses \texttt{pandoc} (and a few additional python glue
-scripts) to facilitate the production of scientific articles using a
-standard markdown file. The objective is to ensure that standard
-markdown (with the important exception of the \texttt{pandoc-crossref}
-citation markup) will be rendered into an interactive website (which
-allows collaborative annotations with the \texttt{hypothes.is}
-platform), a ``draft'' style PDF (double-spaced, numbered lines, figures
-at the end), and a ``preprint'' style PDF (with slightly more
-reader-friendly pagination).
+Spillover risk is not unidimensional. From the standpoint of an animal
+community, i.e.~a pool of possible hosts, there are a multiplicity of
+ecological factors that come into play (Plowright et al.~2017). The
+global richness of hosts is one such component commonly
+mentioned/analysed (Anthony et al.~2017), but there is an argument to be
+made that species who are not competent hosts of a specific virus genus
+may not factor into this (Plowright et al.~2015), or that species who
+are assumed to share viruses at different rates should be weighted
+accordingly (Albery et al.~2020). In mammals, key functional traits (for
+which phylogeny is a reasonable proxy) are determinants of the spillover
+potential (Olival et al.~2017); these include, notably, TK. Finally,
+especially when the pool of potential hosts spans the entire globe,
+there may be local host pools that are highly unique; not having been
+observed in other locations, these can act on the overall risk either by
+providing novel contact opportunities, reflecting unique
+host-environment combinations (Engering et al.~2013), or facilitating
+rapid changes in specialism (Agosta et al.~2010). In the specific case
+of generalist pathogens (as betacoronavirus clearly are), there is
+conceptual and empirical support to the idea that these community- level
+mechanisms are even more important in driving the overall risk (Power
+and Mitchell 2004).
 
-The core bit of configuration is the \texttt{metadata.json} file, which
-handles information about authorship, affiliations, the abstract,
-keywords, etc. All documents will be deployed to \texttt{gh-pages}
-\emph{only} on push events from the \texttt{main} branch. All of the
-artifacts will be built when doing pull requests, so you can check that
-merging a branch is \emph{not} going to cause the compilation of the
-documents to fail; indeed, you can download the artifacts produced
-during the run, to check the PDF and html files. The website is only
-updated from the \texttt{main} branch.
+Bats are important reservoir hosts for different classes of
+microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
+threaten human health. Especially concerning is the fact that bats are
+reservoirs for a variety of emerging viruses (Calisher 2006), making
+balancing the needs for bat conservation and disease prevention a
+potentially difficult act, especially in more densely populated areas
+(REF). Chiropterans emerged around 64 million years ago and are one of
+the most diverse mammalian orders, with an estimated richness of more
+than 12000 species, (Peixoto F et al, 2018) and 14325 known species
+Simmons \& Cirranello. They exhibit a broad variety of habitat use,
+behaviour, and feeding strategies, resulting in their playing an
+essential role in the delivery of several ecosystem services (Kasso
+2013), including economic benefits. Over two-thirds of bats are either
+obligate or facultative insectivorous mammals, therefore playing an
+important role in the regulation of insect pests that can affect crops
+(Williams-Guillen 2011), and vectors of diseases that put a risk on
+human health (Gonsalves 2013 ). Because bats are globally distributed
+and have a long evolutionary history, phylogeographic and biogeographic
+approaches are required to shed light on the extant distribution of
+coevolutionary processes between bats and the pathogens they carry. As a
+consequence, not all areas are facing a risk of human spillover, and
+those who do might not be facing risks of the same nature and magnitude.
 
-The workflow is \emph{very} GitHub based, and so the manuscript file
-\emph{is} the \texttt{README.md} - this is not going to be a huge issue
-as 90\% of the markdown is standard, with the exception of the citations
-and mathematics, so this will render (mostly) like a normal README file.
+Yet a comprehensive assessment of the risk of spillover of
+betacoronaviruses from bat hosts to humans is limited by the fact that
+we do not know the full diversity of viruses associated with every bat
+species. Predictive models can help fill in some of these gaps, by
+recommending hosts based on known host-virus associations BECKER REF. In
+this paper, we examine the biogeographic structure of
+bats-betacoronaviruses associations, based on a curated dataset of known
+and predicted hosts. We turn these associations into a spatially
+explicit additive mapping of zoonotic risk components, which reveals
+extreme heterogeneity of risk at the global scale; furthermore, we
+identify the Amazon and South-Eastern Asia as hotspots of phylogenetic
+distinctiveness of betacoronaviruses; surprisingly, current data suggest
+that viral sharing between hosts in high in the Amazon and low in
+South-Eastern Asia, which has the potential to result in different
+evolutionary dynamics between these two regions.
 
-\hypertarget{deploying-the-template}{%
-\section{Deploying the template}\label{deploying-the-template}}
+TK summary of the results
 
-The process of deploying this template has been \emph{greatly}
-streamlined from previous versions:
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
-\begin{itemize}
-\tightlist
-\item
-  Click on the ``Use this template'' button
-\item
-  Edit \texttt{README.md} with your own text, commit, and push
-\item
-  This push will trigger the first build - the builds are only active on
-  the \texttt{main} branch (\emph{not} \texttt{master}!), and on pull
-  requests
-\item
-  Go to \texttt{http://you.github.io/repo-name/} to view the html
-  version, and get access to the PDFs
-\item
-  Add your references to the \texttt{references.bib} file
-\item
-  Edit the \texttt{metadata.json} file to add the title, abstract,
-  authors
-\end{itemize}
+\hypertarget{known-betacoronavirus-hosts}{%
+\subsection{Known betacoronavirus
+hosts}\label{known-betacoronavirus-hosts}}
 
-In particular, note that \emph{you do not need} to create a personnal
-access token to deploy to \texttt{gh-pages} (from where the website is
-served).
+We downloaded the βCoV reservoir database from
+https://www.viralemergence.org/betacov on Aug.~2021. This database was
+assembled by a combination of data mining, literature surveys, and
+application of an ensemble recommender system classifying hosts as
+either ``Suspected'' or ``Unlikely'' (REF BECKER). The hosts considered
+for this study were all hosts with a known record of a betacoronavirus,
+and all those with a ``Suspected'' status in the ensemble model. This
+resulted in a list of TK TP unique host species.
 
-\hypertarget{the-metadata-file}{%
-\section{The metadata file}\label{the-metadata-file}}
+\hypertarget{bats-occurrences}{%
+\subsection{Bats occurrences}\label{bats-occurrences}}
 
-\hypertarget{general-information}{%
-\subsection{General information}\label{general-information}}
+We downloaded the rangemap of every extant bat species that was either
+classified as an empirically documented or a suspected host of
+beta-coronaviruses (Becker et al.~2020), according to recent IUCN data
+(IUCN 2021). The range maps were subsequently rasterized at a resolution
+of approximately TK TP. For every pixel in the resulting raster where at
+least one bat host of betacoronavirus was present, we extract the
+species pool, which was used to calculate the following risk assessment
+components: phylogenetic diversity, bat compositional uniqueness, and
+predicted viral sharing risk.
 
-The title is a field in the \texttt{metadata.json}:
+\hypertarget{bats-phylogeography}{%
+\subsection{Bats phylogeography}\label{bats-phylogeography}}
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{\{}
-    \DataTypeTok{"title"}\FunctionTok{:} \StringTok{"Preprint template"}
-\FunctionTok{\}}
-\end{Highlighting}
-\end{Shaded}
+For every pixel, we measured Faith's Phylogenetic Diversity (Faith 1992)
+based on a recent synthetic tree with robust time calibration, covering
+about 6000 mammalian species (Upham et al.~2019). Faith's PD measures
+the sum of unique branches from an arbitrary root to a set of tips, and
+comparatively larger values indicate a more phylogenetic diverse species
+pool. We measured phylogenetic diversity starting from the root of the
+entire tree (and not from Chiroptera); this bears no consequences on the
+resulting values, since all branches leading up to Chiroptera are only
+counted one per species pool, and (as we explain when describing the
+assembly of the composite risk map), all individual risk components are
+ranged in {[}0,1{]}. This measure incorporates a richness component,
+which we chose not to correct for; the interpretation of the
+phylogenetic diversity is therefore a weighted species richness, that
+accounts for phylogenetic over/under-dispersal in some places.
 
-\hypertarget{authorship}{%
-\subsection{Authorship}\label{authorship}}
+\hypertarget{bats-compositional-uniqueness}{%
+\subsection{Bats compositional
+uniqueness}\label{bats-compositional-uniqueness}}
 
-Authors are listed as objects in the \texttt{authors} block. Each author
-is specified as follows:
+For every species pool, we measured its Local Contribution to
+Beta-Diversity (Legendre and De Cáceres 2013); LCBD works from a
+species-data matrix (traditionally noted as Y), where species are rows
+and sites are columns, and a value of 1 indicates occurrence. We
+extracted the Y matrix assuming that every pixel represents a unique
+location, and following best practices (Legendre and Condit 2019)
+transformed it using Hellinger's distance to account for unequal bat
+richness at different pixels. The correction of raw community data is
+particularly important for two reasons: first, it prevents the artifact
+of richer sites having higher importance; second, it removes the effect
+of overall species richness, which is already incorporated in the
+phylogenetic diversity component. High values of LCBD indicate that the
+pixel has a community that is on average more dissimilar in species
+composition than what is expected knowing the entire matrix, i.e.~a more
+unique community.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{\{}
-      \DataTypeTok{"familyname"}\FunctionTok{:} \StringTok{"Bob"}\FunctionTok{,}
-      \DataTypeTok{"givennames"}\FunctionTok{:} \StringTok{"Alice"}\FunctionTok{,}
-      \DataTypeTok{"email"}\FunctionTok{:} \StringTok{"alice.bob@u.edu"}\FunctionTok{,}
-      \DataTypeTok{"orcid"}\FunctionTok{:} \StringTok{"0000{-}0000{-}0000{-}0001"}\FunctionTok{,}
-      \DataTypeTok{"affiliations"}\FunctionTok{:} \OtherTok{[}
-        \StringTok{"Affiliation 1"}\OtherTok{,}
-        \StringTok{"Affiliation 2"}
-      \OtherTok{]}\FunctionTok{,}
-      \DataTypeTok{"status"}\FunctionTok{:} \OtherTok{[}\StringTok{"corresponding"}\OtherTok{,} \StringTok{"equal"}\OtherTok{]}
-    \FunctionTok{\}}
-\end{Highlighting}
-\end{Shaded}
+\hypertarget{viral-sharing-between-hosts}{%
+\subsection{Viral sharing between
+hosts}\label{viral-sharing-between-hosts}}
 
-The \texttt{email} field is recommended for all authors. The
-\texttt{status} field is only useful for the corresponding author, and
-to denote equal contributions. These informations are rendered on the
-initial page. If an \texttt{orcid} is given, it will be linked on the
-HTML and PDF versions.
+For all bat hosts of betacoronaviruses, we extracted their predicted
+viral sharing network (Albery et al.~2020). This network stores pairwise
+values of viral community similarity. To project viral sharing values
+into a single value for every pixel, we averaged the pairwise scores.
+High values of the average sharing propensity means that this specific
+extant bat assemblage is likely to be proficient at exchanging viruses.
 
-Note that there is \emph{no need} to number the affiliations - a small
-python script will take care of this automatically.
+\hypertarget{composite-risk-map}{%
+\subsection{Composite risk map}\label{composite-risk-map}}
 
-\hypertarget{abstract}{%
-\subsection{Abstract}\label{abstract}}
+To visualize the aggregated risk at the global scale, we combine the
+three individual risk components (phylogenetic diversity, compositional
+uniqueness, and viral sharing) using an additive color model (Seekell et
+al.~2018). In this approach, every risk component gets assigned a
+component in the RGB color model (phylogenetic diversity is green,
+compositional uniqueness is red, and viral sharing is blue). In order to
+achieve a valid RGB measure, all components are re-scaled to the
+{[}0,1{]} interval. This additive model conveys both the intensity of
+the overall risk, but also the nature of the risk as colors diverge
+towards combinations of values for three risk components.
 
-This template supports three types of abstracts, indicated in the
-metadata file as \texttt{abstract}:
+\hypertarget{viral-phylogeography}{%
+\subsection{Viral phylogeography}\label{viral-phylogeography}}
 
-A regular \texttt{abstract} is defined as
+We used the following query to pull all betacoronavirus sequence data
+from the GenBank Nucleotide database except SARS-CoV-2;
+(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
+NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
+OR sars-cov-2{[}All Fields{]}). We added a single representative
+sequence for SARS-CoV-2 and manually curated to remove sequences without
+the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
+indicating recombinant or laboratory strains including ``patent,''
+``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
+taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
+syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
+betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
+(Katoh and Standley 2013, Supplemental X) and a maximum likelihood tree
+reconstructed in IQ-TREE v 1.6.12 (Nguyen et al.~2015) with ModelFinder
+(Kalyaanamoorthy et al.~2017) ultrafast bootstrap approximation (Hoang
+et al. 2018) and the following parameters (STEPH WILL ADD, Supplemental
+X).
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\ErrorTok{"abstract":} \ErrorTok{"A} \ErrorTok{very} \ErrorTok{long} \ErrorTok{string"}
-\end{Highlighting}
-\end{Shaded}
+\hypertarget{viral-evolutionary-diversification}{%
+\subsection{Viral evolutionary
+diversification}\label{viral-evolutionary-diversification}}
 
-An itemized abstract is an array of strings, each representing a bullet
-point:
+We first tested the hypothesis that hotspots of viral diversification
+would track hotspots of bat diversification. To do so, we plotted the
+number of known bat hosts (specifically only those included in the
+phylogeny, so there was a 1:1 correspondence between data sources)
+against the ``mean evolutionary distinctiveness'' of the associated
+viruses. To calculate this, we derived the fair proportions evolutionary
+distinctiveness (Isaac et al., 2007) for each of the viruses in the
+tree, then averaged these at the bat species level, projected these
+values onto their geographic distributions, and averaged across every
+bat found in a given pixel. As such, this can be thought of as a map of
+the mean evolutionary distinctiveness of the known viral community
+believed to be associated with a particular subset of bats present.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\ErrorTok{"abstract":} \OtherTok{[}
-    \StringTok{"Point 1"}\OtherTok{,}
-    \StringTok{"Point 2"}
-\OtherTok{]}
-\end{Highlighting}
-\end{Shaded}
+\hypertarget{co-distribution-of-hosts-and-viral-hotspots}{%
+\subsection{Co-distribution of hosts and viral
+hotspots}\label{co-distribution-of-hosts-and-viral-hotspots}}
 
-A structured abstract is an object with key-value pairs :
+Subsequently, we tested the hypothesis that the biogeography of bat
+betacoronaviruses should track the biogeography of their hosts. To test
+this idea, we loosely adapted a method from Kreft \& Jetz (2010), who
+proposed a phylogenetic method for the delineation of animal
+biogeographic regions. In their original method, a distance matrix -
+where each row or column represents a geographic raster's grid cell, and
+the dissimilarity values are the ``beta diversity similarity'' of their
+community assemble - undergoes non-metric multidimensional scaling
+(NMDS); the first two axes of the NMDS are projected geographically
+using a four-color bivariate map.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\ErrorTok{"abstract":} \FunctionTok{\{}
-    \DataTypeTok{"Location"}\FunctionTok{:} \StringTok{"Worldwide"}\FunctionTok{,}
-    \DataTypeTok{"Organisms"}\FunctionTok{:} \StringTok{"Mammals"}
-\FunctionTok{\}}
-\end{Highlighting}
-\end{Shaded}
+Here, we build on this idea with an entirely novel methodology. First,
+we measure the phylogenetic distance between the different viruses in
+the betacoronavirus tree by using the cophenetic function in `ape';
+subsequently, we take a principal components analysis of that distance
+matrix (readily interchangeable for NMDS in this case) to project the
+viral tree into an n-dimensional space. We then take the first two
+principal components and, as with the evolutionary distinctiveness
+analysis, aggregated these to a mean host value and projected them using
+a four-color bivariate map.
 
-\hypertarget{citation-style}{%
-\subsection{Citation style}\label{citation-style}}
+\hypertarget{outbreaks-data-geo-referencing}{%
+\subsection{Outbreaks data
+geo-referencing}\label{outbreaks-data-geo-referencing}}
 
-The \texttt{citationstyle} key corresponds to the name, with
-\texttt{.csl} ommited, of a CSL stylesheet stored in the
-\href{https://github.com/citation-style-language/}{citation style
-language} repository. Note that there is no difference between main and
-dependent styles, the build engine will take the correct steps to get
-the correct style. The default is
-\texttt{"citationstyle":\ "ecology-letters"}. There is a longer section
-about references management later on.
+Finally, we provide a summary visualization of what available
+information describes the spillover of zoonotic betacoronaviruses of bat
+origin where data was available before and up through the COVID-19
+pandemic. The SARS-CoV-2 outbreak was georeferenced to the initial case
+cluster in Wuhan, China; SARS-CoV was georeferenced based on the cave
+with the closest known viruses circulating in nature (Hu et al.~2017
+PLoS Pathogens), and a nearby location where serological (antibody)
+evidence has indicated human exposure to SARS-like viruses (Wang et
+al.~2018 Virologica Sinica). For MERS-CoV, we presented the index cases
+available from a recently-published compendium of MERS-CoV cases
+(Ramshaw et al.~2019); these are largely if not all presumed to be
+camel-to-human transmission, and the precise origin point of MERS-CoV in
+bats is uncertain. Not shown is a recent case of a recombinant canine
+coronavirus that showed the ability to infect humans, both because this
+study was published after the beginning of the COVID-19 pandemic and
+because bats' involvement in this cycle of transmission has been
+marginal to non-existent.
 
-\hypertarget{references-management}{%
-\section{References management}\label{references-management}}
-
-The references are managed by \texttt{pandoc}. Note that we \emph{do
-not} use \texttt{pandoc-citeproc}, which was an external module for
-older \texttt{pandoc} versions. References \emph{must} be stored in a
-\texttt{references.bib} file, and that it would make sense to order it
-alphabetically by key.
-
-We use \href{https://www.zotero.org/}{Zotero} for references management,
-and for the lab's manuscripts, we work from folders in a shared library
-(with a folder for every manuscript).
-
-It is recommedned to use the
-\href{https://retorque.re/zotero-better-bibtex/}{Better BibTeX} plugin
-for citation key generations, and auto-export of the shared library to
-the \texttt{references.bib} file. We use a citation key format meant to
-convey information on the author (first author full name), date (complet
-year), and title (first three letters of the first two non-stop words).
-It must be set in the Better BibTeX preferences as (you might need to
-remove the line changes):
-
-\begin{verbatim}
-[auth:fold]
-[year]
-[title:fold:nopunctordash:skipwords:lower:select=1,1:substring=1,3:capitalize]
-[title:fold:nopunctordash:skipwords:lower:select=2,2:substring=1,3:capitalize]
-\end{verbatim}
-
-It is a good idea to configure Better BibTeX to auto-export on change,
-and to remove a lot of fields that are not strictly speaking required
-for references. The list of fields we usually ignore is:
-
-\begin{verbatim}
-abstract,copyright,annotation,file,pmid,month,shorttitle,keywords
-\end{verbatim}
-
-The citations are done using the normal markdown syntax, where
-\texttt{@Elton1927AniEco} produces Elton (1927), and
-\texttt{{[}@Camerano1880EquViv{]}} produces (Camerano 1880).
-
-\hypertarget{figures-tables-and-other-floats}{%
-\section{Figures, Tables, and other
-floats}\label{figures-tables-and-other-floats}}
-
-Note that you can wrap the text of legends for both figures and tables.
-This avoids the issue of having very long lines.
-
-\hypertarget{mathematics}{%
-\subsection{Mathematics}\label{mathematics}}
-
-The following equation
-
-\begin{equation}\protect\hypertarget{eq:eq1}{}{J'(p) = \frac{1}{\text{log}(S)}\times\left(-\sum p \times \text{log}(p)\right)}\label{eq:eq1}\end{equation}
-
-is produced using
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\SpecialStringTok{$$J\textquotesingle{}(p) = }\SpecialCharTok{\textbackslash{}frac}\SpecialStringTok{\{1\}\{}\SpecialCharTok{\textbackslash{}text}\NormalTok{\{log\}}\SpecialStringTok{(S)\}}\SpecialCharTok{\textbackslash{}times}\SpecialStringTok{ ...
-$$}\NormalTok{ \{\#eq:eq1\}}
-\end{Highlighting}
-\end{Shaded}
-
-and can be referenced using \texttt{@eq:eq1}, which will result in
-eq.~\ref{eq:eq1}. Note that because we use \texttt{pandoc-crossref}, the
-label ``eq.'' will be generated automatically.
-
-\hypertarget{tables}{%
-\subsection{Tables}\label{tables}}
-
-Table legends go on the line after the table itself. To generate a
-reference to the table, use \texttt{\{\#tbl:id\}} -- then, in the text,
-you can use \texttt{\{@tbl:id\}} to refer to the table. For example, the
-table below is tbl.~\ref{tbl:id}. You can remove the \emph{table} in
-front by using \texttt{!@tbl:id}, or force it to be capitalized with
-\texttt{\textbackslash{}*tbl:id}.
-
-\hypertarget{tbl:id}{}
-\begin{longtable}[]{@{}rrrrl@{}}
-\caption{\label{tbl:id}This is a table, and its identifier is
-\texttt{id} -- we can refer to it using \texttt{\{@tbl:id\}}. Note that
-even if the table legend is written below the table itself, it will
-appear on top in the PDF document.}\tabularnewline
-\toprule
-Sepal.Length & Sepal.Width & Petal.Length & Petal.Width &
-Species\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-Sepal.Length & Sepal.Width & Petal.Length & Petal.Width &
-Species\tabularnewline
-\midrule
-\endhead
-5.1 & 3.5 & 1.4 & 0.2 & setosa\tabularnewline
-5.0 & 3.6 & 1.4 & 0.2 & setosa\tabularnewline
-5.4 & 3.9 & 1.7 & 0.4 & setosa\tabularnewline
-\bottomrule
-\end{longtable}
-
-\hypertarget{figures}{%
-\section{Figures}\label{figures}}
-
-Figures can have a legend -- all figures \emph{must} be in the
-\texttt{figures/} folder of the project, as it is also used for the
-website. We recommend to use good resolution images, rather than PDFs,
-or at least to have multiple versions available.
-
-\begin{verbatim}
-![This is the legend of the figure...](figures/figure.png){#fig:figure}
-\end{verbatim}
-
-\begin{figure}
-\hypertarget{fig:figure}{%
-\centering
-\includegraphics{figures/figure.png}
-\caption{This is the legend of the figure, which will be shown in the
-margin in preprint mode, and underneath the figure in draft mode. The
-legend can contain references, etc. It is advised to use a resolution of
-at least 600dpi for the figures.}\label{fig:figure}
-}
-\end{figure}
-
-We can now use \texttt{@fig:figure} to refer to fig.~\ref{fig:figure}.
-
-\hypertarget{example-text}{%
-\section{Example text}\label{example-text}}
-
-Connectance, defined as the ratio of realized interactions on the total
-number of potential interactions, is one of the most common descriptor
-of network structure. In a bipartite network with \(T\) species at the
-top, and \(B\) at the bottom, having a total of \(L\) interactions, it
-is defined as \(Co = L/(T\times B)\). Connectance has a lower bound, as
-the network cannot have fewer interactions that the number of species in
-its more speciose level -- the minimal connectance is therefore
-\(c_m = \text{max}(T,B)\). This makes the connectance of networks of
-different sizes difficult to compare, especially since bipartite
-networks tends to have a low connectance. For this reason, we used a
-corrected version of connectance, defined as
-
-\begin{equation}\protect\hypertarget{eq:cstar}{}{Co^\star=\frac{L-c_m}{T\times B-c_m} \,.}\label{eq:cstar}\end{equation}
-
-\hypertarget{this-is-a-subsection}{%
-\subsection{This is a subsection}\label{this-is-a-subsection}}
-
-This takes values between 0 (the network has the minimal number of
-interactions) and 1 (all species are connected), but is robust to
-variations in species richness.
-
-\hypertarget{this-is-another-subsection}{%
-\subsection{This is another
-subsection}\label{this-is-another-subsection}}
-
-This takes values between 0 (the network has the minimal number of
-interactions) and 1 (all species are connected), but is robust to
-variations in species richness.
-
-\hypertarget{some-non-standard-maths}{%
-\subsection{Some non-standard maths}\label{some-non-standard-maths}}
-
-The phylogenetic reconstruction of \(\hat{\mathscr{L}}\) and
-\(\hat{\mathscr{R}}\) has an associated uncertainty, represented by the
-breadth of the uniform distribution associated to each of their entries.
-Therefore, we can use this information to assemble a
-\emph{probabilistic} metaweb in the sense of
-(\textbf{Poisot2016StrPro?}), \emph{i.e.} in which every interaction is
-represented as a single, independent, Bernoulli event of probability
-\(p\).
-
-Specifically, we have adopted the following approach. For every entry in
-\(\hat{\mathscr{L}}\) and \(\hat{\mathscr{R}}\), we draw a value from
-its distribution. This results in one instance of the possible left
-(\(\hat{\mathscr{l}}\)) and right (\(\hat{\mathscr{r}}\)) subspaces for
-the Canadian metaweb. These can be multiplied, to produce one matrix of
-real values. Because the entries in \(\hat{\mathscr{l}}\) and
-\(\hat{\mathscr{r}}\) are in the same space where \(\mathscr{L}\) and
-\(\mathscr{R}\) were originally predicted, it follows that the threshold
-\(\rho\) estimated for the European metaweb also applies. We use this
-information to produce one random Canadian metaweb,
-\(N = \hat{\mathscr{L}}\)\(\hat{\mathscr{R}}' \ge \rho\).
-
-Because the intervals around some trait values can be broad (in fact,
-probably broader than what they would actually be, see \emph{e.g.}
-\textbf{Garland1999IntPhy?}), we repeat the above process
-\(2\times 10^5\) times, which results in a probabilistic metaweb \(P\),
-where the probability of an interaction (here conveying our degree of
-trust that it exists given the inferred trait distributions) is given by
-the number of times where it appears across all random draws \(N\),
-divided by the number of samples. An interaction with \(P_{i,j} = 1\)
-means that these two species were predicted to interact in all
-\(2\times 10^5\) random draws, etc..
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\hypertarget{ref-Camerano1880EquViv}{}%
-Camerano, L. (1880). Dell'equilibrio dei viventi merce la reciproca
-distruzione. \emph{Atti Della R. Accad. Delle Sci. Torino}, 15,
-393--414.
-
-\leavevmode\hypertarget{ref-Elton1927AniEco}{}%
-Elton, C.S. (1927). \emph{Animal ecology}. University of Chicago Press.
-
-\end{CSLReferences}
+\hypertarget{results}{%
+\section{Results}\label{results}}
 

### 2022-04-06 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ce714f31455d6ef4bdd57d7ff199cef40f1e81a6 🚀

@@
 {\small{\bfseries Keywords}:\par
-pandoc\\pandoc-crossref\\github actions\\}
+bats\\betacoronavirus\\disease ecology\\geographic mosaic theory of
+coevolution\\phylogenetic diversity\\viral sharing\\SARS-CoV-2\\}
 }

### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@23e2605f8bc7387ccc41cb2ad8ab052ac79e7759 🚀

@@
 
+\hypertarget{host-distribution}{%
+\subsection{Host distribution}\label{host-distribution}}
+
+Chiroptera are an hyperdiverse group, distributed in a large part if the
+world, and are an important reservoir for different strains of
+betacoronaviruses (Drexler et al., 2014); this has attracted attention
+to areas where high diversity of bats can be an important issue for
+human health (Calisher et al., 2006). Accordingly, we collected the IUCN
+rangemaps for known hosts of betacoronaviruses, to illustrate where
+hotspots of host diversity are. These results are presented in Fig xx.a.
+As per our current knowledge of which bats are hosts of
+betacoronaviruses, these hotspots are primarily South-East Asia, parts
+of Europe, and to a lesser extent sub-saharan Africa. Even the subset of
+chiroptera that are hosts of betacoronaviruses fits the evolutionary
+timeline of the group. Chiropterans can be classified as Microchiroptera
+and macrochiroptera, where macrochiroptera have an older history from an
+evolutionary perspective compared to macrochiroptera (Springer, 2013;
+Teeling et al., 2005). South-East Asia has a high diversity of bats
+(Kingston, 2010), and our results show that part of that diversity
+includes betacoronavirus hosts. High density of hosts sharing the same
+virus (albeit possibly different strains) calls into question the
+evolution of the bat antiviral immune system and its co-evolution with
+viruses, which may result in distinct immunological responses in
+different area, as evidenced in other bat species (Banerjee et al.,
+2020)
+
+\hypertarget{viral-evolutionary-distinctiveness}{%
+\subsection{Viral evolutionary
+distinctiveness}\label{viral-evolutionary-distinctiveness}}
+
+Higher host diversity may not result in a higher viral diversity; for
+this reason, we quantified and mapped the evolutionary distinctiveness
+of betacoronaviruses, based on \ldots. Viral evolutionary
+distinctiveness largely tracks host diversity, particularly in southern
+China but oddly not throughout the rest of southeast Asia, perhaps
+indicating that many distinctive viruses remain to be discovered in this
+region (an idea that is unsurprising given the growing realization,
+around the emergence of SARS-CoV-2, that a unique lineage of similar
+viruses are widespread in bats but still mostly undescribed). The most
+distinct betacoronaviruses are found in South America, a region with a
+comparatively lower number of hosts; this suggests that the South
+American bat-betacoronvirus complex has been more isolated, and is
+probably undergoing a different co-evolutionary dynamic. Alternatively,
+this distinctiveness hostpot may be a product of under-sampling:
+South-America is one of the places where the fewest betacoronaviruses
+have been discovered (Anthony et al., 2017), and adding more viruses
+would bring the distinctiveness of known sequences down. Previous work
+has suggested the Americas may be a hotspot of both undiscovered bat
+viruses in general (Olival) and coronavirus specifically (Anthony),
+though not necessarily betacoronaviruses, and particularly not those in
+clades with notable zoonotic potential (c.f. Anthony).
+
+\hypertarget{geographic-mosaic-of-bat-betacoronavirus-risk}{%
+\subsection{Geographic Mosaic of bat-betacoronavirus
+risk}\label{geographic-mosaic-of-bat-betacoronavirus-risk}}
+
+In order to turn the hypotheses based on the Geographic Mosaic Theory of
+Coevolution into a measure of risk, we overlapped three components:
+viral sharing, i.e.~the chance that two bats will share viruses overall;
+Local Contribution to Beta Diversity, i.e.~the fact that a bat community
+is compositionally unique compared to the average compositional
+similarity across the entire system; finally, the phylogenetic
+diversity, i.e.~how dispersed the bats in a location are within the tree
+of life. These results are presented using an additive color mapping in
+Figure xx, and lead to the definition of broad biogeographic regions of
+risk, where the same color represents the same type of risk. Pairwise
+maps of the three components are present in supplementary materials.
+
+From the perspective of spillover risk, the most important combination
+of factors is a high phylogenetic diversity of hosts with low viral
+sharing; this, essentially, means that very different betacoronavirus
+could co-exist within the same place. This is particularly the case
+given that betacoronaviruses often evolve and even achieve host shifts
+through recombination, which requires the co-occurrence of sufficiently
+distinct viruses to be a major driver of emergence. In Fig. xx, this
+corresponds to yellow to pale green areas, which are essentially limited
+to South-Eastern Asia, and to some part of Sub-Saharan Africa. Adopting
+a geographic mosaic theory perspective on risk, other regions of the
+world are of lesser concern.
+
+Available data on bat betacoronavirus spillover into humans (TP overlay
+on the figure) is limited and circumstantial at best for these purposes,
+but our risk maps suggest that the areas predicted by prior expectations
+about host biogeography correspond loosely to those where previous
+emergence events have been recorded. Areas with high bat diversity and
+high turnover may facilitate the evolutionary radiation of viruses,
+matching previous findings that the diversification of bat coronaviruses
+is driven largely by host shifts (inter-genus or higher levels of
+cross-species transmission) and, to a lesser degree, cospeciation and
+sharing (intra-genus cross-species transmission; Anthony et al.~2017).
+This diversification - while not an actual risk factor for spillover
+itself - likely increases the random chance of a virus with the raw
+genomic components required for the potential to infect humans.
+
+\hypertarget{global-distribution-of-spillover-risk}{%
+\subsection{Global distribution of spillover
+risk}\label{global-distribution-of-spillover-risk}}
+
+Based on the previous result, we extracted the yellow component of the
+risk map (TP add methods), to provide a single measure of risk varying
+between 0 and 1. This measure is presented in Fig. xxA. However, this
+maps the potential risk, which must be weighed by the potential for
+contacts with humans. As a proxy for this measure, we used the
+proportion of build/urban land from the EarthEnv dataset: this is a
+reasonable proxy for the density of humans per unit area, which
+increases the probability of pathogen spread more widely (Hazarie et
+al., 2021). Since human activity is required to amplify the frequency of
+virus encounters and thus create areas of viral amplification, mapping
+the potential risk against measures of land use is required to generate
+a more actionable assessment of risk. This map is presented in Fig. xxB.
+Most of South America and Europe are at low risk, as although densely
+populated, settlements tend to be in areas with lower potential risk.
+However, this mapping reveals that South-East Asia, the Indian
+subcontinent, and parts of sub-Saharan Africa, are at high risk due to
+the overlap between built areas and bat communities representing more
+opportunities for cross-species transmission of betacoronaviruses.
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Driven by the need to understand the ecological factors involved in the
+emergence of viral pathogens, we spatially mapped bat-betacoronavirus
+interactions worldwide, using (i) a database of known betacov
+hosts(Becker et al., 2020), and (ii) range maps for the hosts according
+to IUCN (IUCN 2021). To reflect the fact that the risk posed by viruses
+has many ecological origins, we quantified the phylogenetic diversity of
+hosts, their compositional uniqueness, and the expected viral sharing.
+Because these components of risk matter when contrasted to human
+density, we compared them to a proxy, namely the proportion of each
+pixel that is covered by urban or built land. This provides a synthetic
+risk map, allowing to identifying of hotspots where the
+bat-betacoronavirus system may originate viruses in humans. SE Asia is
+one of the regions with the highest risk since, according to our
+results, several of its conditions could increase the risk of
+transmission of the virus.
+
+Bats are found worldwide and are one of the most diverse groups among
+mammals (Moratelli \& Calisher, 2015). Previous research (Anthony et
+al., 2017; Mollentze \& Streicker, 2020) states that locally diverse bat
+communities could maintain more viruses and hence, a higher probability
+of having a pathogen that could represent a risk for human health. This
+probability involves multiple factors, among which the relatedness of
+hosts (which can make the jumps easier (Longdon et al., 2011; Mollentze
+et al., 2020; Wolfe et al., 2007), and the overall tendency of hosts
+within a locality to share viruses, which may limit viral diversity
+because of within-host competition (Leeks et al., 2018; Sallinen et al.,
+2020). Species richness, therefore, is not a sufficient measure of viral
+risk. This is exemplified in our results, where both South America and
+South-Eastern Asia have a high species richness of betacov hosts, but
+only the latter region has a high risk. Specifically, because previous
+studies propose that Asia is important when it comes to understanding
+the evolutionary origin of various mammalian taxa (Beard C K, 1988).
+Including bats (Yu et al., 2014), which could support the relationship
+between evolutionary time and the development of an immune system with
+characteristics that allow them to be better adapted to infection by
+emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
+related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
+et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
+
+Our study focuses largely on the biogeography of hosts. Yet, we know
+that viruses with high host plasticity, that is, the ability of a given
+virus to adapt to various taxonomic orders and ecological groups
+(Kreuder Johnson et al., 2015); are more likely to amplify viral
+spillover, followed by secondary human-to-human transmission, and
+geographical spread (Hazarie et al., 2021). High viral host plasticity
+is an especially important trait for RNA viruses such as betacov
+(Kreuder Johnson et al., 2015; Haddad et al., 2021). Indeed, our
+analysis of viral sequences reveals that Latin America is a hotspot of
+viral distinctiveness, suggesting that this part of the bats-betacov
+system may be undergoing independent evolutionary dynamics (related
+species sharing viruses that are different from the rest of the global
+pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
+richness is high but sharing is low; this suggests a different type of
+evolutionary dynamics (unrelated viruses coevolving with evolutionarily
+distinct hosts, generating high diversity locally).
+
+This diversity of hosts and how the exchange of viruses occurs between
+species, is largely affected by the different environmental changes, as
+the case of sarbecovirus bats reservoirs (Muylaert et al., 2021) where
+they are affected by the area of the cave or the alteration of the
+forest, which could result in modifications of host distribution.
+Additionally, our results highlight the importance of Asia as a betacov
+hotspot, which is consistent with recent studies (Muylaert et al.,
+2021), where projections on this area suggest that new future events of
+sarbecovirus viral exchange might be easily spread among species or
+humans.
+
+There are several factors that drive changes in the diversity of bats
+(Alves et al., 2018), but human activities' effects on the ecosystem
+(like modifications of land use) could significantly decrease it.
+Therefore, it can be suggested that changes in the diversity of betacovs
+in bats are linked to their biogeographic variation, and human
+population density and other anthropogenic factors are decisive
+moderators for its implications in public health. With the increase of
+contact between humans and potential hosts, we also increase the risk of
+emergence of novel diseases (Johnson et al., 2020), as previous studies
+on RNA viruses suggest the importance of host phylogeography at the time
+of virus dispersal (Gryseels et al., 2017).
+
+One of these scenarios where interaction between bats and humans can
+occur can be seed dispersal in tropical agroecosystems. It opens the
+discussion of whether the fruits thrown by bats not only disperse seeds
+but could also be a source of indirect interaction between viruses of
+bat origin and humans (Deshpande et al., 2022) . This represents a
+challenge for conservation strategies and disease ecology since we have
+areas with potential zoonotic viruses and bat-human interaction.
+However, it must still be taken into account the quantification of real
+exposure from several scenarios, where there can be directly or
+indirectly bat - human interaction.
+
+Comparing scenarios of high viral exchange vs low viral exchange, open
+the discussion to consider if the best scenario is where viruses easily
+adapted to multiple hosts but with low virulence or easily ignored by
+the immune system of the host, or where we have viruses specialized to a
+specific host, but highly virulent when invade a new host. Accordingly,
+the understanding of viral-host interactions from a taxonomic and
+phylogenetic contributes to improving zoonoses surveillance programs.
+
+\textbf{Acknowledgements}: We acknowledge that this study was conducted
+on land within the traditional unceded territory of the Saint Lawrence
+Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
+nations. This work was supported by funding to the Viral Emergence
+Research Initiative (VERENA) consortium including NSF BII 2021909 and a
+grant from Institut de Valorisation des Données (IVADO). This research
+was enabled in part by support provided by Calcul Québec
+(www.calculquebec.ca) and Compute Canada (www.computecanada.ca). NF is
+funded by the NSERC BIOS² CREATE program. TP and NF are funded by the
+Courtois Foundation. RLM was supported by Bryce Carmine and Anne Carmine
+(née Percival), through the Massey University Foundation.
+
 \end{document}

### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c838ae68a102dec3d41800888666f71828b1239a 🚀

@@
 community, i.e.~a pool of possible hosts, there are a multiplicity of
-ecological factors that come into play (Plowright et al.~2017). The
+ecological factors that come into play (Plowright et al. 2017). The
 global richness of hosts is one such component commonly
@@
 made that species who are not competent hosts of a specific virus genus
-may not factor into this (Plowright et al.~2015), or that species who
+may not factor into this (Plowright et al. 2015), or that species who
 are assumed to share viruses at different rates should be weighted
@@
 providing novel contact opportunities, reflecting unique
-host-environment combinations (Engering et al.~2013), or facilitating
-rapid changes in specialism (Agosta et al.~2010). In the specific case
+host-environment combinations (Engering et al. 2013), or facilitating
+rapid changes in specialism (Agosta et al. 2010). In the specific case
 of generalist pathogens (as betacoronavirus clearly are), there is
@@
 
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
+Agosta, S. J. et al. 2010. How specialists can be generalists: Resolving
+the {``parasite paradox''} and implications for emerging infectious
+disease. - Zoologia (Curitiba) 27: 151--162.
+
+\leavevmode\hypertarget{ref-Engering2013PatHos}{}%
+Engering, A. et al. 2013. Pathogen--host--environment interplay and
+disease emergence. - Emerging Microbes \& Infections 2: e5.
+
+\leavevmode\hypertarget{ref-Plowright2015EcoDyn}{}%
+Plowright, R. K. et al. 2015. Ecological dynamics of emerging bat virus
+spillover. - Proceedings of the Royal Society B: Biological Sciences
+282: 20142124.
+
+\leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%
+Plowright, R. K. et al. 2017. Pathways to zoonotic spillover. - Nature
+Reviews Microbiology 15: 502--510.
+
+\end{CSLReferences}
+
 \end{document}

### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c76c179bbea5c8313459ae2046bec2ff4f52c611 🚀

@@
 Spillover risk is not unidimensional. From the standpoint of an animal
-community, i.e.~a pool of possible hosts, there are a multiplicity of
-ecological factors that come into play (Plowright et al. 2017). The
-global richness of hosts is one such component commonly
-mentioned/analysed (Anthony et al.~2017), but there is an argument to be
+community, \emph{i.e.} a pool of possible hosts, it is driven by a
+multiplicity of factors (Plowright et al. 2017). The global richness of
+hosts is one such component commonly mentioned/analysed (see \emph{e.g.}
+Anthony et al. 2017 for coronaviruses), but there is an argument to be
 made that species who are not competent hosts of a specific virus genus
@@
 are assumed to share viruses at different rates should be weighted
-accordingly (Albery et al.~2020). In mammals, key functional traits (for
+accordingly (Albery et al. 2020). In mammals, key functional traits (for
 which phylogeny is a reasonable proxy) are determinants of the spillover
-potential (Olival et al.~2017); these include, notably, TK. Finally,
+potential (Olival et al. 2017); these include, notably, body mass and
+affinity for urban environments (Albery et al. 2022). Finally,
 especially when the pool of potential hosts spans the entire globe,
@@
 
+\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
+Albery, G. F. et al. 2020. Predicting the global mammalian viral sharing
+network using phylogeography. - Nature Communications 11: 2260.
+
+\leavevmode\hypertarget{ref-Albery2022UrbMam}{}%
+Albery, G. F. et al. 2022. Urban-adapted mammal species have more known
+pathogens. in press.
+
+\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
+Anthony, S. J. et al. 2017. Global patterns in coronavirus diversity. -
+Virus Evolution in press.
+
 \leavevmode\hypertarget{ref-Engering2013PatHos}{}%
@@
 disease emergence. - Emerging Microbes \& Infections 2: e5.
+
+\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
+Olival, K. J. et al. 2017. Host and viral traits predict zoonotic
+spillover from mammals. - Nature 546: 646--650.
 
@@
 
+\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
+Power, A. G. and Mitchell, C. E. 2004. Pathogen Spillover in Disease
+Epidemics. - The American Naturalist 164: S79--S89.
+
 \end{CSLReferences}

### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@dad08218838f5926c9b134a7afbd8009ad71a986 🚀

@@
 Spillover risk is not unidimensional. From the standpoint of an animal
-community, \emph{i.e.} a pool of possible hosts, it is driven by a
+community, \emph{i.e.} a pool of suitable hosts, it is driven by a
 multiplicity of factors (Plowright et al. 2017). The global richness of
@@
 Anthony et al. 2017 for coronaviruses), but there is an argument to be
-made that species who are not competent hosts of a specific virus genus
-may not factor into this (Plowright et al. 2015), or that species who
-are assumed to share viruses at different rates should be weighted
-accordingly (Albery et al. 2020). In mammals, key functional traits (for
-which phylogeny is a reasonable proxy) are determinants of the spillover
-potential (Olival et al. 2017); these include, notably, body mass and
-affinity for urban environments (Albery et al. 2022). Finally,
-especially when the pool of potential hosts spans the entire globe,
-there may be local host pools that are highly unique; not having been
-observed in other locations, these can act on the overall risk either by
-providing novel contact opportunities, reflecting unique
+made that species who are not competent (or know) hosts of a specific
+virus genus may not factor into this (Plowright et al. 2015), or that
+species who are assumed to share viruses at different rates should be
+weighted accordingly (Albery et al. 2020). In mammals, key functional
+traits (for which phylogeny is a reasonable proxy) are determinants of
+the spillover potential (Olival et al. 2017); these include, notably,
+body mass and affinity for urban environments (Albery et al. 2022).
+Finally, especially when the pool of potential hosts spans the entire
+globe, there may be local host pools that are highly unique; not having
+been observed in other locations, these can act on the overall risk
+either by providing novel contact opportunities, reflecting unique
 host-environment combinations (Engering et al. 2013), or facilitating
-rapid changes in specialism (Agosta et al. 2010). In the specific case
-of generalist pathogens (as betacoronavirus clearly are), there is
+rapid evolutionary changes in specialism of their pathogens (Agosta et
+al. 2010). In the specific case of generalist pathogens, there is
 conceptual and empirical support to the idea that these community- level
@@
 microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
-threaten human health. Especially concerning is the fact that bats are
-reservoirs for a variety of emerging viruses (Calisher 2006), making
-balancing the needs for bat conservation and disease prevention a
-potentially difficult act, especially in more densely populated areas
-(REF). Chiropterans emerged around 64 million years ago and are one of
-the most diverse mammalian orders, with an estimated richness of more
-than 12000 species, (Peixoto F et al, 2018) and 14325 known species
-Simmons \& Cirranello. They exhibit a broad variety of habitat use,
-behaviour, and feeding strategies, resulting in their playing an
-essential role in the delivery of several ecosystem services (Kasso
-2013), including economic benefits. Over two-thirds of bats are either
-obligate or facultative insectivorous mammals, therefore playing an
-important role in the regulation of insect pests that can affect crops
-(Williams-Guillen 2011), and vectors of diseases that put a risk on
-human health (Gonsalves 2013 ). Because bats are globally distributed
-and have a long evolutionary history, phylogeographic and biogeographic
-approaches are required to shed light on the extant distribution of
-coevolutionary processes between bats and the pathogens they carry. As a
-consequence, not all areas are facing a risk of human spillover, and
-those who do might not be facing risks of the same nature and magnitude.
+threaten human health. Chiropterans emerged around 64 million years ago
+and are one of the most diverse mammalian orders, with an estimated
+richness of more than 12000 species, (Peixoto F et al, 2018) and 14325
+known species (Simmons \& Cirranello). They exhibit a broad variety of
+habitat use, behaviour, and feeding strategies, resulting in their
+playing an essential role in the delivery of several ecosystem services
+tied to important ecosystem-derived benefits (Kasso 2013). For example,
+over two-thirds of bats are know to be either obligate or facultative
+insectivorous mammals, therefore playing an important role in the
+regulation of insect pests that can affect crops (Williams-Guillen
+2011), and vectors of diseases that put a risk on human health
+(Gonsalves 2013). Because bats are globally distributed and have a long
+evolutionary history, phylogeographic and biogeographic approaches are
+required to shed light on the extant distribution of coevolutionary
+processes between bats and the pathogens they carry. Not all areas in
+which bats, viruses, and human are co-occuring are facing a risk of
+spillover towards human populations, and the areas in which this risk
+exist may not be facing risks of the same nature and magnitude.
 
-Yet a comprehensive assessment of the risk of spillover of
-betacoronaviruses from bat hosts to humans is limited by the fact that
-we do not know the full diversity of viruses associated with every bat
-species. Predictive models can help fill in some of these gaps, by
-recommending hosts based on known host-virus associations BECKER REF. In
-this paper, we examine the biogeographic structure of
+In this paper, we examine the biogeographic structure of
 bats-betacoronaviruses associations, based on a curated dataset of known
-and predicted hosts. We turn these associations into a spatially
-explicit additive mapping of zoonotic risk components, which reveals
-extreme heterogeneity of risk at the global scale; furthermore, we
-identify the Amazon and South-Eastern Asia as hotspots of phylogenetic
+and recently discovered hosts. This work is important both as a
+description of the bats-betacoronavirus complex, but aslo because more
+broadly, bats are known reservoirs for a variety of emerging viruses
+(Calisher 2006), making balancing the needs for bat conservation and
+disease prevention a potentially difficult act and a source of
+human-wildlife conflicst, especially in more densely populated areas
+(REF). By drawing on concepts from the Geographic Mosaic Theory of
+Coevolution (REF), we turn these associations into a spatially explicit
+additive mapping of zoonotic risk components, which reveals extreme
+heterogeneity of risk at the global scale; furthermore, we identify the
+Amazon and South-Eastern Asia as hotspots of phylogenetic
 distinctiveness of betacoronaviruses; surprisingly, current data suggest
@@
 evolutionary dynamics between these two regions.
-
-TK summary of the results
 

### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b3e97c387fcd2d3bb3787550a5e90c67f911bbe2 🚀

@@
 
-We downloaded the βCoV reservoir database from
-https://www.viralemergence.org/betacov on Aug.~2021. This database was
-assembled by a combination of data mining, literature surveys, and
-application of an ensemble recommender system classifying hosts as
-either ``Suspected'' or ``Unlikely'' (REF BECKER). The hosts considered
-for this study were all hosts with a known record of a betacoronavirus,
-and all those with a ``Suspected'' status in the ensemble model. This
-resulted in a list of TK TP unique host species.
+We downloaded the data on bats hosts of betacoronaviruses assembled by
+Becker et al. (2022) from
+\texttt{https://www.viralemergence.org/betacov} on Aug.~2021, and
+filtered it to ``known'' hosts (established before the emergence of
+SARS-CoV-2) and ``novel'' hosts (confirmed through sampling since the
+emergence of SARS-CoV-2). This database was assembled by a combination
+of data mining and literature surveys.
 
@@
 
+\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
+Becker, D. J. et al. 2022. Optimising predictive models to prioritise
+viral discovery in zoonotic reservoirs. - The Lancet Microbe in press.
+
 \leavevmode\hypertarget{ref-Engering2013PatHos}{}%

### 2022-04-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@0380a385b3465bcaad0a778806c03312f02b8a8e 🚀

@@
 %
+Norma\,Forero Rocio Munoz%
+%
+\,\textsuperscript{1,2}\quad %
+Renata L.\,Muylaert%
+%
+\,\textsuperscript{3}\quad %
+Stephanie N.\,Seifert%
+%
+\,\textsuperscript{4}\quad %
+Gregory F.\,Albery%
+%
+\,\textsuperscript{5}\quad %
+Colin J.\,Carlson%
+%
+\,\textsuperscript{5}\quad %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
 %
-\,\textsuperscript{1,2,‡}\quad %
-Peregrin\,Took%
-%
-\,\textsuperscript{3,4}\quad %
-Merriadoc\,Brandybuck%
-%
-\,\textsuperscript{5,4,‡}
+\,\textsuperscript{1,2,‡}
 
@@
 Montréal\quad \textsuperscript{2}\,Québec Centre for Biodiversity
-Sciences\quad \textsuperscript{3}\,Inn of the Prancing
-Pony\quad \textsuperscript{4}\,Fellowship of the
-Ring\quad \textsuperscript{5}\,Green Dragon Inn
+Sciences\quad \textsuperscript{3}\,Molecular Epidemiology and Public
+Health Laboratory, Hopkirk Research Institute, Massey University,
+Palmerston North, New Zealand\quad \textsuperscript{4}\,Paul G. Allen
+School for Global Health, Washington State University, Pullman, WA,
+United States\quad \textsuperscript{5}\,???
 

### 2022-04-07 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@0380a385b3465bcaad0a778806c03312f02b8a8e 🚀

@@
 %
+Norma\,Forero Rocio Munoz%
+%
+\,\textsuperscript{1,2}, %
+Renata L.\,Muylaert%
+%
+\,\textsuperscript{3}, %
+Stephanie N.\,Seifert%
+%
+\,\textsuperscript{4}, %
+Gregory F.\,Albery%
+%
+\,\textsuperscript{5}, %
+Colin J.\,Carlson%
+%
+\,\textsuperscript{5}, %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
 %
-\,\textsuperscript{1,2,‡}, %
-Peregrin\,Took%
-%
-\,\textsuperscript{3,4}, %
-Merriadoc\,Brandybuck%
-%
-\,\textsuperscript{5,4,‡}
+\,\textsuperscript{1,2,‡}
 \vskip 1em
 \textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
-Centre for Biodiversity Sciences; \textsuperscript{3}\,Inn of the
-Prancing Pony; \textsuperscript{4}\,Fellowship of the
-Ring; \textsuperscript{5}\,Green Dragon Inn\\
+Centre for Biodiversity Sciences; \textsuperscript{3}\,Molecular
+Epidemiology and Public Health Laboratory, Hopkirk Research Institute,
+Massey University, Palmerston North, New
+Zealand; \textsuperscript{4}\,Paul G. Allen School for Global Health,
+Washington State University, Pullman, WA, United
+States; \textsuperscript{5}\,???\\
 \textsuperscript{‡}\,These authors contributed equally to the work\\

### 2022-04-07 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a2dd8bb55c0522d53f67019ac72781b93953140b 🚀

@@
 \vfill
-
-        {\bfseries Purpose:}\,This template provides a series of scripts
-to render a markdown document into an interactive website and a series
-of PDFs.\\%
-        {\bfseries Motivation:}\,It makes collaborating on text with
-GitHub easier, and means that we never need to think about the
-output.\\%
-        {\bfseries Internals:}\,GitHub actions and a series of python
-scritpts. The markdown is handled with \texttt{pandoc}.\\%
+Coming soon
 

### 2022-04-07 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a2dd8bb55c0522d53f67019ac72781b93953140b 🚀

@@
 
-        {\bfseries Purpose:}\,This template provides a series of scripts
-to render a markdown document into an interactive website and a series
-of PDFs.\\%
-        {\bfseries Motivation:}\,It makes collaborating on text with
-GitHub easier, and means that we never need to think about the
-output.\\%
-        {\bfseries Internals:}\,GitHub actions and a series of python
-scritpts. The markdown is handled with \texttt{pandoc}.\\%
+Coming soon
 

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@642638bfc147d135a1075a90473a98c2a061f374 🚀

@@
 and are one of the most diverse mammalian orders, with an estimated
-richness of more than 12000 species, (Peixoto F et al, 2018) and 14325
-known species (Simmons \& Cirranello). They exhibit a broad variety of
-habitat use, behaviour, and feeding strategies, resulting in their
-playing an essential role in the delivery of several ecosystem services
-tied to important ecosystem-derived benefits (Kasso 2013). For example,
-over two-thirds of bats are know to be either obligate or facultative
-insectivorous mammals, therefore playing an important role in the
-regulation of insect pests that can affect crops (Williams-Guillen
-2011), and vectors of diseases that put a risk on human health
-(Gonsalves 2013). Because bats are globally distributed and have a long
+richness of more than 12000 species, (Peixoto et al. 2018) and 14325
+known species (Simmons and Cirranello 2020). They exhibit a broad
+variety of habitat use, behaviour, and feeding strategies, resulting in
+their playing an essential role in the delivery of several ecosystem
+services tied to important ecosystem-derived benefits (Kasso and
+Balakrishnan 2013). For example, over two-thirds of bats are know to be
+either obligate or facultative insectivorous mammals, therefore playing
+an important role in the regulation of insect pests that can affect
+crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016), and
+vectors of diseases that put a risk on human health (Gonsalves et al.
+2013a, b). Because bats are globally distributed and have a long
 evolutionary history, phylogeographic and biogeographic approaches are
@@
 
+\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
+Gonsalves, L. et al. 2013a. Do mosquitoes influence bat activity in
+coastal habitats? - Wildlife Research 40: 10--24.
+
+\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
+Gonsalves, L. et al. 2013b. Mosquito Consumption by Insectivorous Bats:
+Does Size Matter? - PLOS ONE 8: e77183.
+
+\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
+Kasso, M. and Balakrishnan, M. 2013. Ecological and Economic Importance
+of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
@@
 spillover from mammals. - Nature 546: 646--650.
+
+\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
+Peixoto, F. P. et al. 2018. A synthesis of ecological and evolutionary
+determinants of bat diversity across spatial scales. - BMC ecology 18:
+18.
 
@@
 
+\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
+Simmons, N. B. and Cirranello, A. L. 2020. Bat Species of the World: A
+taxonomic and geographic database.
+
+\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
+2016. Bats in the Anthropocene: Conservation of Bats in a Changing World
+(CC Voigt and T Kingston, Eds.). - Springer International Publishing.
+
+\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
+Williams-Guillén, K. et al. 2008. Bats Limit Insects in a Neotropical
+Agroforestry System. - Science 320: 70--70.
+
 \end{CSLReferences}

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@5e66493683d008c6662dbbd79b7ddd910f641b51 🚀

@@
 description of the bats-betacoronavirus complex, but aslo because more
-broadly, bats are known reservoirs for a variety of emerging viruses
-(Calisher 2006), making balancing the needs for bat conservation and
-disease prevention a potentially difficult act and a source of
-human-wildlife conflicst, especially in more densely populated areas
-(REF). By drawing on concepts from the Geographic Mosaic Theory of
-Coevolution (REF), we turn these associations into a spatially explicit
+broadly, bats are known reservoirs for a variety of emerging viruses and
+pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
+the needs for bat conservation and disease prevention a potentially
+difficult act and a source of human-wildlife conflicts, especially in
+more densely populated areas (Stone et al. 2015, Rego et al. 2015). By
+drawing on concepts from the Geographic Mosaic Theory of Coevolution
+(Thompson 2005), we turn these associations into a spatially explicit
 additive mapping of zoonotic risk components, which reveals extreme
@@
 
+\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
+Calisher, C. H. et al. 2006. Bats: Important Reservoir Hosts of Emerging
+Viruses. - Clinical Microbiology Reviews 19: 531--545.
+
 \leavevmode\hypertarget{ref-Engering2013PatHos}{}%
@@
 of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.
+
+\leavevmode\hypertarget{ref-Melaun2014BatPot}{}%
+Melaun, C. et al. 2014. Bats as Potential Reservoir Hosts for
+Vector-Borne Diseases. - In: Klimpel, S. and Mehlhorn, H. (eds), Bats
+(Chiroptera) as Vectors of Diseases and Parasites: Facts and Myths.
+Parasitology Research Monographs. Springer, pp. 25--61.
 
@@
 
+\leavevmode\hypertarget{ref-Rego2015AssHum}{}%
+Rego, K. M. da C. et al. 2015. Assessing human-bat interactions around a
+protected area in northeastern Brazil. - Journal of Ethnobiology and
+Ethnomedicine 11: 80.
+
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
@@
 taxonomic and geographic database.
+
+\leavevmode\hypertarget{ref-Stone2015ManCon}{}%
+Stone, E. et al. 2015. Managing Conflict between Bats and Humans: The
+Response of Soprano Pipistrelles (Pipistrellus pygmaeus) to Exclusion
+from Roosts in Houses. - PLoS ONE 10: e0131825.
+
+\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
+Thompson, J. N. 2005. The Geographic Mosaic of Coevolution. - University
+Of Chicago Press.
 

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@db274119a671a5a4aedb594967efcbc6a1d2cc1c 🚀

@@
 Amazon and South-Eastern Asia as hotspots of phylogenetic
-distinctiveness of betacoronaviruses; surprisingly, current data suggest
-that viral sharing between hosts in high in the Amazon and low in
-South-Eastern Asia, which has the potential to result in different
-evolutionary dynamics between these two regions.
+distinctiveness of betacoronaviruses (Anthony et al. 2017);
+surprisingly, current data suggest that viral sharing between hosts in
+high in the Amazon and low in South-Eastern Asia, which has the
+potential to result in different evolutionary dynamics between these two
+regions.
 

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@0164ce425d534ca57568c97d08a18db32ffcf41c 🚀

@@
 Becker et al. (2022) from
-\texttt{https://www.viralemergence.org/betacov} on Aug.~2021, and
+\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
 filtered it to ``known'' hosts (established before the emergence of
 SARS-CoV-2) and ``novel'' hosts (confirmed through sampling since the
-emergence of SARS-CoV-2). This database was assembled by a combination
-of data mining and literature surveys.
+emergence of SARS-CoV-2). The original database was assembled by a
+combination of data mining and literature surveys, including automated
+alerts on the ``bats'' and ``coronavirus'' keywords to identify novel
+empirical evidence of bats-betacoronaviruses associations.
 
@@
 We downloaded the rangemap of every extant bat species that was either
-classified as an empirically documented or a suspected host of
-beta-coronaviruses (Becker et al.~2020), according to recent IUCN data
-(IUCN 2021). The range maps were subsequently rasterized at a resolution
-of approximately TK TP. For every pixel in the resulting raster where at
-least one bat host of betacoronavirus was present, we extract the
-species pool, which was used to calculate the following risk assessment
-components: phylogenetic diversity, bat compositional uniqueness, and
-predicted viral sharing risk.
+classified as an empirically documented host of beta-coronaviruses from
+the previous step, according to recent IUCN data (IUCN 2021). The range
+maps were subsequently rasterized using the \texttt{rasterize} function
+from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
+approximately \textbf{TK TP}. For every pixel in the resulting raster
+where at least one bat host of betacoronavirus was present, we extract
+the species pool (list of all bat species), which was used to calculate
+the following risk assessment components: phylogenetic diversity, bat
+compositional uniqueness, and predicted viral sharing risk.
 
@@
 based on a recent synthetic tree with robust time calibration, covering
-about 6000 mammalian species (Upham et al.~2019). Faith's PD measures
+about 6000 mammalian species (Upham et al. 2019). Faith's PD measures
 the sum of unique branches from an arbitrary root to a set of tips, and
@@
 
+\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
+Faith, D. P. 1992. Conservation evaluation and phylogenetic diversity. -
+Biological Conservation 61: 1--10.
+
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
@@
 Does Size Matter? - PLOS ONE 8: e77183.
+
+\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
+IUCN 2021. The IUCN Red List of Threatened Species.
 
@@
 
+\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
+Rouault, E. et al. 2022. GDAL/OGR Geospatial Data Abstraction software
+Library. - Zenodo.
+
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
@@
 
+\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
+Upham, N. S. et al. 2019. Inferring the mammal tree: Species-level sets
+of phylogenies for questions in ecology, evolution, and conservation. -
+PLOS Biology 17: e3000494.
+
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@28c85b6732c2e44e4483ff56e8f0bb79c1f65c3f 🚀

@@
 Beta-Diversity (Legendre and De Cáceres 2013); LCBD works from a
-species-data matrix (traditionally noted as Y), where species are rows
-and sites are columns, and a value of 1 indicates occurrence. We
-extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices (Legendre and Condit 2019)
-transformed it using Hellinger's distance to account for unequal bat
-richness at different pixels. The correction of raw community data is
-particularly important for two reasons: first, it prevents the artifact
-of richer sites having higher importance; second, it removes the effect
-of overall species richness, which is already incorporated in the
-phylogenetic diversity component. High values of LCBD indicate that the
-pixel has a community that is on average more dissimilar in species
-composition than what is expected knowing the entire matrix, i.e.~a more
-unique community.
+species-data matrix (traditionally noted as \(\mathbf{Y}\)), where
+species are rows and sites are columns, and a value of 1 indicates
+occurrence. We extracted the Y matrix assuming that every pixel
+represents a unique location, and following best practices (Legendre and
+Condit 2019) transformed it using Hellinger's distance to account for
+unequal bat richness at different pixels. The correction of raw
+community data is particularly important for two reasons: first, it
+prevents the artifact of richer sites having higher importance; second,
+it removes the effect of overall species richness, which is already
+incorporated in the phylogenetic diversity component. High values of
+LCBD indicate that the pixel has a community that is on average more
+dissimilar in species composition than what is expected knowing the
+entire matrix, i.e.~a more unique community. Recent results by Dansereau
+et al. (2022) shows that LCBD measures are robust with regards to
+spatial scale, and are therefore applicable at the global scale.
 
@@
 
+\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
+Dansereau, G. et al. 2022. Evaluating ecological uniqueness over broad
+spatial extents using species distribution modelling. - Oikos n/a:
+e09063.
+
 \leavevmode\hypertarget{ref-Engering2013PatHos}{}%
@@
 of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.
+
+\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
+Legendre, P. and De Cáceres, M. 2013. Beta diversity as the variance of
+community data: Dissimilarity coefficients and partitioning (H Morlon,
+Ed.). - Ecology Letters 16: 951--963.
+
+\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
+Legendre, P. and Condit, R. 2019. Spatial and temporal analysis of beta
+diversity in the Barro Colorado Island forest dynamics plot, Panama. -
+Forest Ecosystems 6: 7.
 

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f3162065c1002a2b64bc983127ad9961ce2932c0 🚀

@@
 For all bat hosts of betacoronaviruses, we extracted their predicted
-viral sharing network (Albery et al.~2020). This network stores pairwise
+viral sharing network (Albery et al. 2020). This network stores pairwise
 values of viral community similarity. To project viral sharing values
@@
 uniqueness, and viral sharing) using an additive color model (Seekell et
-al.~2018). In this approach, every risk component gets assigned a
+al. 2018). In this approach, every risk component gets assigned a
 component in the RGB color model (phylogenetic diversity is green,
@@
 achieve a valid RGB measure, all components are re-scaled to the
-{[}0,1{]} interval. This additive model conveys both the intensity of
-the overall risk, but also the nature of the risk as colors diverge
-towards combinations of values for three risk components.
+{[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
+diversity, and no compositional uniqueness is black, and a pixel with
+maximal values for each is white. This additive model conveys both the
+intensity of the overall risk, but also the nature of the risk as colors
+diverge towards combinations of values for three risk components. Out of
+the possible combinations, the most risky in terms or rapid
+diversification and spillover potential is high phylogenetic diversity
+and low viral sharing (Gomulkiewicz et al. 2000, Cavender-Bares et al.
+2009), in that this allows multiple independent host-virus
+coevolutionary dynamics to take place in the same location. In the
+colorimetric space, this correspond to yellow -- because the HSV space
+is more amenable to calculations for feature extraction (see \emph{e.g.}
+Keke et al. 2010), we measured the risk level by calculating the angular
+distance of the hue of each pixel to a reference value of 60, and
+weighted this risk level by the value component. Specifically, given a
+pixel with colorimetric coordinates \((h,s,v)\), its ranged weighted
+risk value is
+
+\[
+v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right)\right|}{2\pi}\right]\,,
+\]
+
+where X is \(atan(cos(deg2rad(60.0)), sin(deg2rad(60.0))\), a constant
+approximately equal to \(0.5235\).
 
@@
 
+\leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
+Cavender-Bares, J. et al. 2009. The merging of community ecology and
+phylogenetic biology. - Ecol. Lett. 12: 693--715.
+
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
@@
 Biological Conservation 61: 1--10.
+
+\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
+Gomulkiewicz, R. et al. 2000. Hot Spots, Cold Spots, and the Geographic
+Mosaic Theory of Coevolution. - The American Naturalist 156: 156--174.
 
@@
 of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.
+
+\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
+Keke, S. et al. 2010. Study on skin color image segmentation used by
+Fuzzy-c-means arithmetic. - 2010 Seventh International Conference on
+Fuzzy Systems and Knowledge Discovery 2: 612--615.
 
@@
 
+\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
+Seekell, D. A. et al. 2018. A geography of lake carbon cycling. -
+Limnology and Oceanography Letters 3: 49--56.
+
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e10565101f449c7cd01760eebd7624592f2c71d9 🚀

@@
 \[
-v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right)\right|}{2\pi}\right]\,,
+v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right|}{2\pi}\right]\,,
 \]
 
-where X is \(atan(cos(deg2rad(60.0)), sin(deg2rad(60.0))\), a constant
+where X is \(atan(cos(deg2rad(60.0)), sin(deg2rad(60.0)))\), a constant
 approximately equal to \(0.5235\).
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness (Isaac et al., 2007) for each of the viruses in the
-tree, then averaged these at the bat species level, projected these
-values onto their geographic distributions, and averaged across every
-bat found in a given pixel. As such, this can be thought of as a map of
-the mean evolutionary distinctiveness of the known viral community
-believed to be associated with a particular subset of bats present.
+distinctiveness (Isaac et al. 2007) for each of the viruses in the tree,
+then averaged these at the bat species level, projected these values
+onto their geographic distributions, and averaged across every bat found
+in a given pixel. As such, this can be thought of as a map of the mean
+evolutionary distinctiveness of the known viral community believed to be
+associated with a particular subset of bats present.
 
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from Kreft \& Jetz (2010), who
-proposed a phylogenetic method for the delineation of animal
+this idea, we loosely adapted a method from (Kreft and Jetz 2007, 2010),
+who proposed a phylogenetic method for the delineation of animal
 biogeographic regions. In their original method, a distance matrix -
@@
 (NMDS); the first two axes of the NMDS are projected geographically
-using a four-color bivariate map.
-
-Here, we build on this idea with an entirely novel methodology. First,
-we measure the phylogenetic distance between the different viruses in
-the betacoronavirus tree by using the cophenetic function in `ape';
+using a four-color bivariate map. Here, we build on this idea with an
+entirely novel methodology. First, we measure the phylogenetic distance
+between the different viruses in the betacoronavirus tree by using the
+cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
 subsequently, we take a principal components analysis of that distance
@@
 
+\leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
+Isaac, N. J. B. et al. 2007. Mammals on the EDGE: Conservation
+Priorities Based on Threat and Phylogeny. - PLOS ONE 2: e296.
+
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
@@
 Fuzzy Systems and Knowledge Discovery 2: 612--615.
+
+\leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
+Kreft, H. and Jetz, W. 2007. Global patterns and determinants of
+vascular plant diversity. - Proceedings of the National Academy of
+Sciences 104: 5925--5930.
+
+\leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
+Kreft, H. and Jetz, W. 2010. A framework for delineating biogeographical
+regions based on species distributions. - Journal of Biogeography 37:
+2029--2053.
 
@@
 spillover from mammals. - Nature 546: 646--650.
+
+\leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
+Paradis, E. and Schliep, K. 2019. Ape 5.0: An environment for modern
+phylogenetics and evolutionary analyses in R. - Bioinformatics 35:
+526--528.
 

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6a2b0037727a9b03bf181a2cd910b10aeee05a00 🚀

@@
 world, and are an important reservoir for different strains of
-betacoronaviruses (Drexler et al., 2014); this has attracted attention
-to areas where high diversity of bats can be an important issue for
-human health (Calisher et al., 2006). Accordingly, we collected the IUCN
-rangemaps for known hosts of betacoronaviruses, to illustrate where
-hotspots of host diversity are. These results are presented in Fig xx.a.
-As per our current knowledge of which bats are hosts of
-betacoronaviruses, these hotspots are primarily South-East Asia, parts
-of Europe, and to a lesser extent sub-saharan Africa. Even the subset of
-chiroptera that are hosts of betacoronaviruses fits the evolutionary
-timeline of the group. Chiropterans can be classified as Microchiroptera
-and macrochiroptera, where macrochiroptera have an older history from an
-evolutionary perspective compared to macrochiroptera (Springer, 2013;
-Teeling et al., 2005). South-East Asia has a high diversity of bats
-(Kingston, 2010), and our results show that part of that diversity
-includes betacoronavirus hosts. High density of hosts sharing the same
-virus (albeit possibly different strains) calls into question the
-evolution of the bat antiviral immune system and its co-evolution with
-viruses, which may result in distinct immunological responses in
-different area, as evidenced in other bat species (Banerjee et al.,
-2020)
+betacoronaviruses (Drexler et al. 2014); this has attracted attention to
+areas where high diversity of bats can be an important issue for human
+health (Calisher et al. 2006, Moratelli and Calisher 2015). Accordingly,
+we collected the IUCN rangemaps for known hosts of betacoronaviruses, to
+illustrate where hotspots of host diversity are. These results are
+presented in Fig xx.a. As per our current knowledge of which bats are
+hosts of betacoronaviruses, these hotspots are primarily South-East
+Asia, parts of Europe, and to a lesser extent sub-saharan Africa. Even
+the subset of chiroptera that are hosts of betacoronaviruses fits the
+evolutionary timeline of the group. Chiropterans can be classified as
+Microchiroptera and macrochiroptera, where macrochiroptera have an older
+history from an evolutionary perspective compared to macrochiroptera
+(Teeling et al. 2005, Springer 2013). South-East Asia has a high
+diversity of bats (Kingston, 2010), and our results show that part of
+that diversity includes betacoronavirus hosts. High density of hosts
+sharing the same virus (albeit possibly different strains) calls into
+question the evolution of the bat antiviral immune system and its
+co-evolution with viruses, which may result in distinct immunological
+responses in different area, as evidenced in other bat species (Banerjee
+et al. 2020).
 
@@
 
+\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
+Banerjee, A. et al. 2020. Novel Insights Into Immune Systems of Bats. -
+Frontiers in Immunology 11: 26.
+
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
@@
 e09063.
+
+\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
+Drexler, J. F. et al. 2014. Ecology, evolution and classification of bat
+coronaviruses in the aftermath of SARS. - Antiviral Research 101:
+45--56.
 
@@
 
+\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
+Moratelli, R. and Calisher, C. H. 2015. Bats and zoonotic viruses: Can
+we confidently link bats with emerging deadly viruses? - Memórias do
+Instituto Oswaldo Cruz 110: 1--22.
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
@@
 
+\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
+Springer, M. S. 2013. Phylogenetics: Bats United, Microbats Divided. -
+Current Biology 23: R999--R1001.
+
 \leavevmode\hypertarget{ref-Stone2015ManCon}{}%
@@
 from Roosts in Houses. - PLoS ONE 10: e0131825.
+
+\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
+Teeling, E. C. et al. 2005. A Molecular Phylogeny for Bats Illuminates
+Biogeography and the Fossil Record. - Science (New York, N.Y.) 307:
+580--584.
 

### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f6794af7cdf35fdcb3aa85893d1c3ce5eb4daa84 🚀

@@
 
-where X is \(atan(cos(deg2rad(60.0)), sin(deg2rad(60.0)))\), a constant
-approximately equal to \(0.5235\).
+where X is
+\(\text{atan}\left(\text{cos}(\text{rad}(60)), \text{sin}(\text{rad}(60))\right)\),
+a constant approximately equal to \(0.5235\).
 
@@
 South-America is one of the places where the fewest betacoronaviruses
-have been discovered (Anthony et al., 2017), and adding more viruses
+have been discovered (Anthony et al. 2017), and adding more viruses
 would bring the distinctiveness of known sequences down. Previous work
 has suggested the Americas may be a hotspot of both undiscovered bat
-viruses in general (Olival) and coronavirus specifically (Anthony),
-though not necessarily betacoronaviruses, and particularly not those in
-clades with notable zoonotic potential (c.f. Anthony).
+viruses in general (Olival et al. 2017, Allen et al. 2017) and
+coronavirus specifically (Anthony et al. 2017), though not necessarily
+betacoronaviruses, and particularly not those in clades with notable
+zoonotic potential.
 
@@
 
+\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
+Allen, T. et al. 2017. Global hotspots and correlates of emerging
+zoonotic diseases. - Nature Communications in press.
+
 \leavevmode\hypertarget{ref-Anthony2017GloPat}{}%

### 2022-04-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@07bf7fcfd0bbd09f9ae2dcefa0c47eeaf7e27313 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@29ac99b98a2aab44e530d1a13ee42e5a12e4ae6c 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@8dd7d67ed0c1ccefb16231d48b0addedccf9764e 🚀

@@
 and are one of the most diverse mammalian orders, with an estimated
-richness of more than 12000 species, (Peixoto et al. 2018) and 14325
-known species (Simmons and Cirranello 2020). They exhibit a broad
-variety of habitat use, behaviour, and feeding strategies, resulting in
-their playing an essential role in the delivery of several ecosystem
-services tied to important ecosystem-derived benefits (Kasso and
-Balakrishnan 2013). For example, over two-thirds of bats are know to be
-either obligate or facultative insectivorous mammals, therefore playing
-an important role in the regulation of insect pests that can affect
-crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016), and
-vectors of diseases that put a risk on human health (Gonsalves et al.
-2013a, b). Because bats are globally distributed and have a long
-evolutionary history, phylogeographic and biogeographic approaches are
-required to shed light on the extant distribution of coevolutionary
-processes between bats and the pathogens they carry. Not all areas in
-which bats, viruses, and human are co-occuring are facing a risk of
-spillover towards human populations, and the areas in which this risk
-exist may not be facing risks of the same nature and magnitude.
+richness of more than 14000 species (Peixoto et al. 2018, Simmons and
+Cirranello 2020). They exhibit a broad variety of habitat use,
+behaviour, and feeding strategies, resulting in their playing an
+essential role in the delivery of several ecosystem services tied to
+important ecosystem-derived benefits (Kasso and Balakrishnan 2013). For
+example, over two-thirds of bats are know to be either obligate or
+facultative insectivorous mammals, therefore playing an important role
+in the regulation of insect pests that can affect crops
+(Williams-Guillén et al. 2008, Voigt and Kingston 2016), and vectors of
+diseases that put a risk on human health (Gonsalves et al. 2013a, b).
+Because bats are globally distributed and have a long evolutionary
+history, phylogeographic and biogeographic approaches are required to
+shed light on the extant distribution of coevolutionary processes
+between bats and the pathogens they carry. Not all areas in which bats,
+viruses, and human are co-occuring are facing a risk of spillover
+towards human populations, and the areas in which this risk exist may
+not be facing risks of the same nature and magnitude.
 
@@
 
-\hypertarget{viral-phylogeography}{%
-\subsection{Viral phylogeography}\label{viral-phylogeography}}
+\hypertarget{viral-phylogeography-and-evolutionary-diversification}{%
+\subsection{Viral phylogeography and evolutionary
+diversification}\label{viral-phylogeography-and-evolutionary-diversification}}
 
@@
 betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
-(Katoh and Standley 2013, Supplemental X) and a maximum likelihood tree
-reconstructed in IQ-TREE v 1.6.12 (Nguyen et al.~2015) with ModelFinder
-(Kalyaanamoorthy et al.~2017) ultrafast bootstrap approximation (Hoang
-et al. 2018) and the following parameters (STEPH WILL ADD, Supplemental
-X).
-
-\hypertarget{viral-evolutionary-diversification}{%
-\subsection{Viral evolutionary
-diversification}\label{viral-evolutionary-diversification}}
+(\textbf{Katoh and Standley 2013}, parameters in text?) and a maximum
+likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al.~2015)
+with ModelFinder (Kalyaanamoorthy et al.~2017) ultrafast bootstrap
+approximation (Hoang et al. 2018) and the following parameters
+(\textbf{STEPH WILL ADD}, parameters in text?).
 
@@
 
-\hypertarget{outbreaks-data-geo-referencing}{%
-\subsection{Outbreaks data
-geo-referencing}\label{outbreaks-data-geo-referencing}}
-
-Finally, we provide a summary visualization of what available
-information describes the spillover of zoonotic betacoronaviruses of bat
-origin where data was available before and up through the COVID-19
-pandemic. The SARS-CoV-2 outbreak was georeferenced to the initial case
-cluster in Wuhan, China; SARS-CoV was georeferenced based on the cave
-with the closest known viruses circulating in nature (Hu et al.~2017
-PLoS Pathogens), and a nearby location where serological (antibody)
-evidence has indicated human exposure to SARS-like viruses (Wang et
-al.~2018 Virologica Sinica). For MERS-CoV, we presented the index cases
-available from a recently-published compendium of MERS-CoV cases
-(Ramshaw et al.~2019); these are largely if not all presumed to be
-camel-to-human transmission, and the precise origin point of MERS-CoV in
-bats is uncertain. Not shown is a recent case of a recombinant canine
-coronavirus that showed the ability to infect humans, both because this
-study was published after the beginning of the COVID-19 pandemic and
-because bats' involvement in this cycle of transmission has been
-marginal to non-existent.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{results-and-discussion}{%
+\section{Results and discussion}\label{results-and-discussion}}
 
@@
 
-Chiroptera are an hyperdiverse group, distributed in a large part if the
-world, and are an important reservoir for different strains of
-betacoronaviruses (Drexler et al. 2014); this has attracted attention to
-areas where high diversity of bats can be an important issue for human
-health (Calisher et al. 2006, Moratelli and Calisher 2015). Accordingly,
-we collected the IUCN rangemaps for known hosts of betacoronaviruses, to
-illustrate where hotspots of host diversity are. These results are
-presented in Fig xx.a. As per our current knowledge of which bats are
-hosts of betacoronaviruses, these hotspots are primarily South-East
-Asia, parts of Europe, and to a lesser extent sub-saharan Africa. Even
-the subset of chiroptera that are hosts of betacoronaviruses fits the
-evolutionary timeline of the group. Chiropterans can be classified as
-Microchiroptera and macrochiroptera, where macrochiroptera have an older
-history from an evolutionary perspective compared to macrochiroptera
-(Teeling et al. 2005, Springer 2013). South-East Asia has a high
-diversity of bats (Kingston, 2010), and our results show that part of
-that diversity includes betacoronavirus hosts. High density of hosts
-sharing the same virus (albeit possibly different strains) calls into
-question the evolution of the bat antiviral immune system and its
-co-evolution with viruses, which may result in distinct immunological
-responses in different area, as evidenced in other bat species (Banerjee
-et al. 2020).
+Bats are found worldwide and are one of the most diverse groups among
+mammals (Moratelli \& Calisher, 2015), and one of the main animal
+reservoir for different strains of betacoronaviruses (Drexler et al.
+2014). This has attracted attention to areas where high diversity of
+bats, and therefore presumably high diversity of betacoronaviruses, can
+be an important issue for human health (Calisher et al. 2006, Moratelli
+and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
+known hosts of betacoronaviruses, to illustrate where hotspots of host
+diversity are. These results are presented in Fig xx.a. As per our
+current knowledge of which bats are hosts of betacoronaviruses, these
+hotspots are primarily South-East Asia, parts of Europe, and to a lesser
+extent sub-saharan Africa. Even the subset of chiroptera that are hosts
+of betacoronaviruses fits the evolutionary timeline of the group.
+Chiropterans can be classified as Microchiroptera and macrochiroptera,
+where macrochiroptera have an older history from an evolutionary
+perspective compared to macrochiroptera (Teeling et al. 2005, Springer
+2013).
+
+South-East Asia has a high diversity of bats (\textbf{Kingston, 2010}),
+and our results show that part of that diversity includes
+betacoronavirus hosts. High density of hosts sharing the same virus
+(albeit possibly different strains) calls into question the evolution of
+the bat antiviral immune system and its co-evolution with viruses, which
+may result in distinct immunological responses in different areas, as
+evidenced in other bat species (Banerjee et al. 2020). Immune
+characteristics that allow bats to be better adapted to infection by
+emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
+related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
+et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
+Considering whether viruses easily adapted to multiple hosts have lower
+virulence on these hosts, or lower ability to jump to hosts with
+different immune characteristics, should yield valuable additional
+predictors for the total risk of spillover. Previous research
+(\textbf{Anthony et al., 2017; Mollentze \& Streicker, 2020}) states
+that locally diverse bat communities could maintain more viruses and
+hence, a higher probability of having a pathogen that could represent a
+risk for human health; locally diverse, virus-rich bats communities
+could represent an increased risk of spillover under climate change
+(\textbf{Ice ice berg berg}). This probability involves multiple
+factors, among which the relatedness of hosts (which can make the jumps
+easier (\textbf{Longdon et al., 2011; Mollentze et al., 2020; Wolfe et
+al., 2007}), and the overall tendency of hosts within a locality to
+share viruses, which may limit viral diversity because of within-host
+competition (Leeks et al., 2018; Sallinen et al., 2020). All things
+considered, the richness of known betacoronaviruses hosts is not a
+sufficient predictor of spillover risk.
 
@@
 
-Higher host diversity may not result in a higher viral diversity; for
-this reason, we quantified and mapped the evolutionary distinctiveness
-of betacoronaviruses, based on \ldots. Viral evolutionary
+Higher host diversity may not result in a higher viral diversity, for
+example if all hosts share the same viruses, or share closely
+evolutionarily related strains. For this reason, we quantified and
+mapped the evolutionary distinctiveness of betacoronaviruses, based on
+their position in a molecular phylogeny. Viral evolutionary
 distinctiveness largely tracks host diversity, particularly in southern
-China but oddly not throughout the rest of southeast Asia, perhaps
-indicating that many distinctive viruses remain to be discovered in this
-region (an idea that is unsurprising given the growing realization,
-around the emergence of SARS-CoV-2, that a unique lineage of similar
-viruses are widespread in bats but still mostly undescribed). The most
-distinct betacoronaviruses are found in South America, a region with a
-comparatively lower number of hosts; this suggests that the South
-American bat-betacoronvirus complex has been more isolated, and is
-probably undergoing a different co-evolutionary dynamic. Alternatively,
-this distinctiveness hostpot may be a product of under-sampling:
-South-America is one of the places where the fewest betacoronaviruses
-have been discovered (Anthony et al. 2017), and adding more viruses
-would bring the distinctiveness of known sequences down. Previous work
-has suggested the Americas may be a hotspot of both undiscovered bat
-viruses in general (Olival et al. 2017, Allen et al. 2017) and
-coronavirus specifically (Anthony et al. 2017), though not necessarily
-betacoronaviruses, and particularly not those in clades with notable
-zoonotic potential.
+China but, oddly, not throughout the rest of southeast Asia. This
+indicates, perhaps, that many distinctive viruses remain to be
+discovered in this region (an idea that is unsurprising given the
+growing realization, around the emergence of SARS-CoV-2, that a unique
+lineage of similar viruses are widespread in bats but still mostly
+undescribed). The most distinct betacoronaviruses are found in South
+America, a region with a comparatively lower number of hosts; this
+suggests that the South American bat-betacoronvirus complex has been
+more isolated, and is probably undergoing a different co-evolutionary
+dynamic. Alternatively, this distinctiveness hostpot may be a product of
+under-sampling: South-America is one of the places where the fewest
+betacoronaviruses have been discovered (Anthony et al. 2017), and adding
+more viruses would bring the distinctiveness of known sequences down.
+Previous work has suggested the Americas may be a hotspot of both
+undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
+2017) and coronavirus specifically (Anthony et al. 2017), though not
+necessarily betacoronaviruses, and particularly not those in clades with
+notable zoonotic potential.
 
@@
 
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Driven by the need to understand the ecological factors involved in the
-emergence of viral pathogens, we spatially mapped bat-betacoronavirus
-interactions worldwide, using (i) a database of known betacov
-hosts(Becker et al., 2020), and (ii) range maps for the hosts according
-to IUCN (IUCN 2021). To reflect the fact that the risk posed by viruses
-has many ecological origins, we quantified the phylogenetic diversity of
-hosts, their compositional uniqueness, and the expected viral sharing.
-Because these components of risk matter when contrasted to human
-density, we compared them to a proxy, namely the proportion of each
-pixel that is covered by urban or built land. This provides a synthetic
-risk map, allowing to identifying of hotspots where the
-bat-betacoronavirus system may originate viruses in humans. SE Asia is
-one of the regions with the highest risk since, according to our
-results, several of its conditions could increase the risk of
-transmission of the virus.
-
-Bats are found worldwide and are one of the most diverse groups among
-mammals (Moratelli \& Calisher, 2015). Previous research (Anthony et
-al., 2017; Mollentze \& Streicker, 2020) states that locally diverse bat
-communities could maintain more viruses and hence, a higher probability
-of having a pathogen that could represent a risk for human health. This
-probability involves multiple factors, among which the relatedness of
-hosts (which can make the jumps easier (Longdon et al., 2011; Mollentze
-et al., 2020; Wolfe et al., 2007), and the overall tendency of hosts
-within a locality to share viruses, which may limit viral diversity
-because of within-host competition (Leeks et al., 2018; Sallinen et al.,
-2020). Species richness, therefore, is not a sufficient measure of viral
-risk. This is exemplified in our results, where both South America and
-South-Eastern Asia have a high species richness of betacov hosts, but
-only the latter region has a high risk. Specifically, because previous
-studies propose that Asia is important when it comes to understanding
-the evolutionary origin of various mammalian taxa (Beard C K, 1988).
-Including bats (Yu et al., 2014), which could support the relationship
-between evolutionary time and the development of an immune system with
-characteristics that allow them to be better adapted to infection by
-emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
-related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
-et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
+Finally, we provide a summary visualization of what available
+information describes the spillover of zoonotic betacoronaviruses of bat
+origin where data was available before and up through the COVID-19
+pandemic. The SARS-CoV-2 outbreak was georeferenced to the initial case
+cluster in Wuhan, China; SARS-CoV was georeferenced based on the cave
+with the closest known viruses circulating in nature (Hu et al.~2017
+PLoS Pathogens), and a nearby location where serological (antibody)
+evidence has indicated human exposure to SARS-like viruses (Wang et
+al.~2018 Virologica Sinica). For MERS-CoV, we presented the index cases
+available from a recently-published compendium of MERS-CoV cases
+(Ramshaw et al.~2019); these are largely if not all presumed to be
+camel-to-human transmission, and the precise origin point of MERS-CoV in
+bats is uncertain. Not shown is a recent case of a recombinant canine
+coronavirus that showed the ability to infect humans, both because this
+study was published after the beginning of the COVID-19 pandemic and
+because bats' involvement in this cycle of transmission has been
+marginal to non-existent.
 
@@
 
-This diversity of hosts and how the exchange of viruses occurs between
-species, is largely affected by the different environmental changes, as
-the case of sarbecovirus bats reservoirs (Muylaert et al., 2021) where
-they are affected by the area of the cave or the alteration of the
-forest, which could result in modifications of host distribution.
-Additionally, our results highlight the importance of Asia as a betacov
-hotspot, which is consistent with recent studies (Muylaert et al.,
-2021), where projections on this area suggest that new future events of
-sarbecovirus viral exchange might be easily spread among species or
-humans.
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+Driven by the need to understand the ecological factors involved in the
+emergence of viral pathogens, we spatially mapped bat-betacoronavirus
+interactions worldwide, using (i) a database of known betacov
+hosts(Becker et al., 2020), and (ii) range maps for the hosts according
+to IUCN (IUCN 2021). To reflect the fact that the risk posed by viruses
+has many ecological origins, we quantified the phylogenetic diversity of
+hosts, their compositional uniqueness, and the expected viral sharing.
+Because these components of risk matter when contrasted to human
+density, we compared them to a proxy, namely the proportion of each
+pixel that is covered by urban or built land. This provides a synthetic
+risk map, allowing to identifying of hotspots where the
+bat-betacoronavirus system may originate viruses in humans. SE Asia is
+one of the regions with the highest risk since, according to our
+results, several of its conditions could increase the risk of
+transmission of the virus.
+
+Species richness, therefore, is not a sufficient measure of viral risk.
+This is exemplified in our results, where both South America and
+South-Eastern Asia have a high species richness of betacov hosts, but
+only the latter region has a high risk. Specifically, because previous
+studies propose that Asia is important when it comes to understanding
+the evolutionary origin of various mammalian taxa (Beard C K, 1988).
 
@@
 
+This diversity of hosts and how the exchange of viruses occurs between
+species, is largely affected by the different environmental changes, as
+the case of sarbecovirus bats reservoirs (Muylaert et al., 2021) where
+they are affected by the area of the cave or the alteration of the
+forest, which could result in modifications of host distribution.
+Additionally, our results highlight the importance of Asia as a betacov
+hotspot, which is consistent with recent studies (Muylaert et al.,
+2021), where projections on this area suggest that new future events of
+sarbecovirus viral exchange might be easily spread among species or
+humans.
+
 One of these scenarios where interaction between bats and humans can
@@
 indirectly bat - human interaction.
-
-Comparing scenarios of high viral exchange vs low viral exchange, open
-the discussion to consider if the best scenario is where viruses easily
-adapted to multiple hosts but with low virulence or easily ignored by
-the immune system of the host, or where we have viruses specialized to a
-specific host, but highly virulent when invade a new host. Accordingly,
-the understanding of viral-host interactions from a taxonomic and
-phylogenetic contributes to improving zoonoses surveillance programs.
 

### 2022-04-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@934aa5f485bf2bc6afb6fe14cb605aefcb6df284 🚀

@@
 In order to turn the hypotheses based on the Geographic Mosaic Theory of
-Coevolution into a measure of risk, we overlapped three components:
-viral sharing, i.e.~the chance that two bats will share viruses overall;
-Local Contribution to Beta Diversity, i.e.~the fact that a bat community
-is compositionally unique compared to the average compositional
-similarity across the entire system; finally, the phylogenetic
-diversity, i.e.~how dispersed the bats in a location are within the tree
-of life. These results are presented using an additive color mapping in
-Figure xx, and lead to the definition of broad biogeographic regions of
-risk, where the same color represents the same type of risk. Pairwise
-maps of the three components are present in supplementary materials.
+Coevolution into a measure of risk, we overlapped three components of
+spillover risk: viral sharing, \emph{i.e.} the chance that two bats will
+share viruses overall; Local Contribution to Beta Diversity, \emph{i.e.}
+the fact that a bat community is compositionally unique compared to the
+average compositional similarity across the entire system; finally, the
+phylogenetic diversity, \emph{i.e.} how dispersed the bats in a location
+are within the tree of life. These results are presented using an
+additive color mapping in Figure xx, and lead to the definition of broad
+biogeographic regions of risk, where the same color represents the same
+type of risk. Pairwise maps of the three components are present in
+supplementary materials.
+
+\begin{figure}
+\hypertarget{fig:trivariate}{%
+\centering
+\includegraphics{figures/risk_trivariate.png}
+\caption{This is the legend of the figure\ldots{}}\label{fig:trivariate}
+}
+\end{figure}
 
@@
 world are of lesser concern.
+
+\begin{figure}
+\hypertarget{fig:risk}{%
+\centering
+\includegraphics{figures/risk_map.png}
+\caption{This is the legend of the figure\ldots{}}\label{fig:risk}
+}
+\end{figure}
 
@@
 opportunities for cross-species transmission of betacoronaviruses.
+
+\begin{figure}
+\hypertarget{fig:compound}{%
+\centering
+\includegraphics{figures/risk_compounded.png}
+\caption{This is the legend of the figure\ldots{}}\label{fig:compound}
+}
+\end{figure}
 

### 2022-04-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@88087d31d74f610c94b69763691a8aafa08f8ecf 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@434b7dd7e4792516ac36f190400a2dbd4c2c5a1e 🚀

@@
 
+\begin{figure}
+\hypertarget{fig:richness}{%
+\centering
+\includegraphics{figures/combined_richness.png}
+\caption{Top panel: relative diversoty of known bat hosts of
+betacoronaviruses. This map shows that the region with the largest
+number of possible hosts is South-Eastern Asia. Bottom panel: congruence
+between the evolutionary distinctiveness of the hosts (grey to blue) and
+the viruses (grey to red). By contrast to the richness map, this reveals
+that South America has the most evolutionary distinct hosts \emph{and}
+viruses, whereas South-Eastern Asia has mostly distinct viruses. This is
+congruent with know results about New World bats being evolutionary
+distinct, and suggests that they similarly have distinct
+viruses.}\label{fig:richness}
+}
+\end{figure}
+
 South-East Asia has a high diversity of bats (\textbf{Kingston, 2010}),
@@
 notable zoonotic potential.
+
+\begin{figure}
+\hypertarget{fig:biogeo}{%
+\centering
+\includegraphics{figures/combined_biogeo.png}
+\caption{This is the legend of the figure\ldots{}}\label{fig:biogeo}
+}
+\end{figure}
 

### 2022-04-29 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@88087d31d74f610c94b69763691a8aafa08f8ecf 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@434b7dd7e4792516ac36f190400a2dbd4c2c5a1e 🚀

@@
 
+\begin{figure}
+\hypertarget{fig:biogeo}{%
+\centering
+\includegraphics{figures/combined_biogeo.png}
+\caption{This is the legend of the figure\ldots{}}\label{fig:biogeo}
+}
+\end{figure}
+
 \hypertarget{geographic-mosaic-of-bat-betacoronavirus-risk}{%

### 2022-04-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b82956c6e0ed0ab5ef5a4d80a336662ab4fb5fb4 🚀

@@
 
-\hypertarget{host-distribution}{%
-\subsection{Host distribution}\label{host-distribution}}
+\hypertarget{host-distribution-does-not-match-virus-distinctiveness}{%
+\subsection{Host distribution does not match virus
+distinctiveness}\label{host-distribution-does-not-match-virus-distinctiveness}}
 
@@
 known hosts of betacoronaviruses, to illustrate where hotspots of host
-diversity are. These results are presented in Fig xx.a. As per our
-current knowledge of which bats are hosts of betacoronaviruses, these
-hotspots are primarily South-East Asia, parts of Europe, and to a lesser
-extent sub-saharan Africa. Even the subset of chiroptera that are hosts
-of betacoronaviruses fits the evolutionary timeline of the group.
-Chiropterans can be classified as Microchiroptera and macrochiroptera,
-where macrochiroptera have an older history from an evolutionary
-perspective compared to macrochiroptera (Teeling et al. 2005, Springer
-2013).
+diversity are. These results are presented in fig.~\ref{fig:richness}
+(top). As per our current knowledge of which bats are hosts of
+betacoronaviruses, these hotspots are primarily South-East Asia, parts
+of Europe, and to a lesser extent parts of Africa in the -25-0 range of
+latitudes. Previous research (\textbf{Anthony et al., 2017; Mollentze \&
+Streicker, 2020}) states that locally diverse bat communities could
+maintain more viruses and hence, a higher probability of having a
+pathogen that could represent a risk for human health; locally diverse,
+virus-rich bats communities could represent an increased risk of
+spillover under climate change (\textbf{Ice ice berg berg}). This
+probability involves multiple factors, among which the relatedness of
+hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
+Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
+of hosts within a locality to share viruses, which may limit viral
+diversity because of within-host competition (Leeks et al., 2018;
+Sallinen et al., 2020). All things considered, the richness of known
+betacoronaviruses hosts is not a sufficient predictor of spillover risk,
+and we should move towards a more coevolutionary understanding of this
+system.
 
@@
 
-South-East Asia has a high diversity of bats (\textbf{Kingston, 2010}),
-and our results show that part of that diversity includes
-betacoronavirus hosts. High density of hosts sharing the same virus
-(albeit possibly different strains) calls into question the evolution of
-the bat antiviral immune system and its co-evolution with viruses, which
-may result in distinct immunological responses in different areas, as
-evidenced in other bat species (Banerjee et al. 2020). Immune
-characteristics that allow bats to be better adapted to infection by
-emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
-related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
-et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
-Considering whether viruses easily adapted to multiple hosts have lower
-virulence on these hosts, or lower ability to jump to hosts with
-different immune characteristics, should yield valuable additional
-predictors for the total risk of spillover. Previous research
-(\textbf{Anthony et al., 2017; Mollentze \& Streicker, 2020}) states
-that locally diverse bat communities could maintain more viruses and
-hence, a higher probability of having a pathogen that could represent a
-risk for human health; locally diverse, virus-rich bats communities
-could represent an increased risk of spillover under climate change
-(\textbf{Ice ice berg berg}). This probability involves multiple
-factors, among which the relatedness of hosts (which can make the jumps
-easier (\textbf{Longdon et al., 2011; Mollentze et al., 2020; Wolfe et
-al., 2007}), and the overall tendency of hosts within a locality to
-share viruses, which may limit viral diversity because of within-host
-competition (Leeks et al., 2018; Sallinen et al., 2020). All things
-considered, the richness of known betacoronaviruses hosts is not a
-sufficient predictor of spillover risk.
+High density of hosts sharing the same virus (albeit possibly different
+strains) calls into question the evolution of the bat antiviral immune
+system and its co-evolution with viruses, which may result in distinct
+immunological responses in different areas, as evidenced in other bat
+species (Banerjee et al. 2020). Immune characteristics that allow bats
+to be better adapted to infection by emerging viruses (Gorbunova et al.,
+2020; Irving et al., 2021) may be related to a wide variety of diets
+(Jones et al., 2022; Moreno Santillán et al., 2021; Banerjee et al.,
+2020; Schneeberger et al., 2013). Considering whether viruses easily
+adapted to multiple hosts have lower virulence on these hosts, or lower
+ability to jump to hosts with different immune characteristics, should
+yield valuable additional predictors for the total risk of spillover.
+The subset of chiroptera that are hosts of betacoronaviruses fits the
+evolutionary timeline of the group. Chiropterans can be classified as
+Microchiroptera and macrochiroptera, where macrochiroptera have an older
+history from an evolutionary perspective compared to macrochiroptera
+(Teeling et al. 2005, Springer 2013). Specifically, we would expect that
+the New World group of bats, being more evolutionary distinct, would
+also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
+(bottom) reveals it to be the case. By contrast, South-Eastern Asia has
+a lot of non-evolutionary distinct bats, but evolutionary-distinct
+viruses.
 
-\hypertarget{viral-evolutionary-distinctiveness}{%
-\subsection{Viral evolutionary
-distinctiveness}\label{viral-evolutionary-distinctiveness}}
-
-Higher host diversity may not result in a higher viral diversity, for
-example if all hosts share the same viruses, or share closely
-evolutionarily related strains. For this reason, we quantified and
-mapped the evolutionary distinctiveness of betacoronaviruses, based on
-their position in a molecular phylogeny. Viral evolutionary
-distinctiveness largely tracks host diversity, particularly in southern
-China but, oddly, not throughout the rest of southeast Asia. This
-indicates, perhaps, that many distinctive viruses remain to be
-discovered in this region (an idea that is unsurprising given the
-growing realization, around the emergence of SARS-CoV-2, that a unique
-lineage of similar viruses are widespread in bats but still mostly
-undescribed). The most distinct betacoronaviruses are found in South
-America, a region with a comparatively lower number of hosts; this
+Viral evolutionary distinctiveness largely tracks host diversity,
+particularly in southern China but, oddly, not throughout the rest of
+southeast Asia. This indicates, perhaps, that many distinctive viruses
+remain to be discovered in this region (an idea that is unsurprising
+given the growing realization, around the emergence of SARS-CoV-2, that
+a unique lineage of similar viruses are widespread in bats but still
+mostly undescribed). The most distinct betacoronaviruses are found in
+South America, a region with a comparatively lower number of hosts; this
 suggests that the South American bat-betacoronvirus complex has been
@@
 
+\hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
+\subsection{The phylogeographic regions of hosts and their viruses
+overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}
+
 \begin{figure}
@@
 
-\hypertarget{geographic-mosaic-of-bat-betacoronavirus-risk}{%
-\subsection{Geographic Mosaic of bat-betacoronavirus
-risk}\label{geographic-mosaic-of-bat-betacoronavirus-risk}}
+\hypertarget{coevolution-informed-spillover-risk-is-different-in-space}{%
+\subsection{Coevolution-informed spillover risk is different in
+space}\label{coevolution-informed-spillover-risk-is-different-in-space}}
 
@@
 
-\hypertarget{global-distribution-of-spillover-risk}{%
-\subsection{Global distribution of spillover
-risk}\label{global-distribution-of-spillover-risk}}
+\hypertarget{human-occupancy-drives-different-levels-of-effective-risk-globally}{%
+\subsection{Human occupancy drives different levels of effective risk
+globally}\label{human-occupancy-drives-different-levels-of-effective-risk-globally}}
 

### 2022-04-29 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b82956c6e0ed0ab5ef5a4d80a336662ab4fb5fb4 🚀

@@
 
-South-East Asia has a high diversity of bats (\textbf{Kingston, 2010}),
-and our results show that part of that diversity includes
-betacoronavirus hosts. High density of hosts sharing the same virus
-(albeit possibly different strains) calls into question the evolution of
-the bat antiviral immune system and its co-evolution with viruses, which
-may result in distinct immunological responses in different areas, as
-evidenced in other bat species (Banerjee et al. 2020). Immune
-characteristics that allow bats to be better adapted to infection by
-emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
-related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
-et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).
-Considering whether viruses easily adapted to multiple hosts have lower
-virulence on these hosts, or lower ability to jump to hosts with
-different immune characteristics, should yield valuable additional
-predictors for the total risk of spillover. Previous research
-(\textbf{Anthony et al., 2017; Mollentze \& Streicker, 2020}) states
-that locally diverse bat communities could maintain more viruses and
-hence, a higher probability of having a pathogen that could represent a
-risk for human health; locally diverse, virus-rich bats communities
-could represent an increased risk of spillover under climate change
-(\textbf{Ice ice berg berg}). This probability involves multiple
-factors, among which the relatedness of hosts (which can make the jumps
-easier (\textbf{Longdon et al., 2011; Mollentze et al., 2020; Wolfe et
-al., 2007}), and the overall tendency of hosts within a locality to
-share viruses, which may limit viral diversity because of within-host
-competition (Leeks et al., 2018; Sallinen et al., 2020). All things
-considered, the richness of known betacoronaviruses hosts is not a
-sufficient predictor of spillover risk.
-
-\hypertarget{viral-evolutionary-distinctiveness}{%
-\subsection{Viral evolutionary
-distinctiveness}\label{viral-evolutionary-distinctiveness}}
-
-Higher host diversity may not result in a higher viral diversity, for
-example if all hosts share the same viruses, or share closely
-evolutionarily related strains. For this reason, we quantified and
-mapped the evolutionary distinctiveness of betacoronaviruses, based on
-their position in a molecular phylogeny. Viral evolutionary
-distinctiveness largely tracks host diversity, particularly in southern
-China but, oddly, not throughout the rest of southeast Asia. This
-indicates, perhaps, that many distinctive viruses remain to be
-discovered in this region (an idea that is unsurprising given the
-growing realization, around the emergence of SARS-CoV-2, that a unique
-lineage of similar viruses are widespread in bats but still mostly
-undescribed). The most distinct betacoronaviruses are found in South
-America, a region with a comparatively lower number of hosts; this
+High density of hosts sharing the same virus (albeit possibly different
+strains) calls into question the evolution of the bat antiviral immune
+system and its co-evolution with viruses, which may result in distinct
+immunological responses in different areas, as evidenced in other bat
+species (Banerjee et al. 2020). Immune characteristics that allow bats
+to be better adapted to infection by emerging viruses (Gorbunova et al.,
+2020; Irving et al., 2021) may be related to a wide variety of diets
+(Jones et al., 2022; Moreno Santillán et al., 2021; Banerjee et al.,
+2020; Schneeberger et al., 2013). Considering whether viruses easily
+adapted to multiple hosts have lower virulence on these hosts, or lower
+ability to jump to hosts with different immune characteristics, should
+yield valuable additional predictors for the total risk of spillover.
+The subset of chiroptera that are hosts of betacoronaviruses fits the
+evolutionary timeline of the group. Chiropterans can be classified as
+Microchiroptera and macrochiroptera, where macrochiroptera have an older
+history from an evolutionary perspective compared to macrochiroptera
+(Teeling et al. 2005, Springer 2013). Specifically, we would expect that
+the New World group of bats, being more evolutionary distinct, would
+also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
+(bottom) reveals it to be the case. By contrast, South-Eastern Asia has
+a lot of non-evolutionary distinct bats, but evolutionary-distinct
+viruses.
+
+Viral evolutionary distinctiveness largely tracks host diversity,
+particularly in southern China but, oddly, not throughout the rest of
+southeast Asia. This indicates, perhaps, that many distinctive viruses
+remain to be discovered in this region (an idea that is unsurprising
+given the growing realization, around the emergence of SARS-CoV-2, that
+a unique lineage of similar viruses are widespread in bats but still
+mostly undescribed). The most distinct betacoronaviruses are found in
+South America, a region with a comparatively lower number of hosts; this
 suggests that the South American bat-betacoronvirus complex has been

### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e8f7f88e898980f4260ab03e9f584f170625e1fa 🚀

@@
 
-\hypertarget{host-distribution-does-not-match-virus-distinctiveness}{%
-\subsection{Host distribution does not match virus
-distinctiveness}\label{host-distribution-does-not-match-virus-distinctiveness}}
+\hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
+\subsection{Host richness does not predict virus
+distinctiveness}\label{host-richness-does-not-predict-virus-distinctiveness}}
 
-Bats are found worldwide and are one of the most diverse groups among
-mammals (Moratelli \& Calisher, 2015), and one of the main animal
-reservoir for different strains of betacoronaviruses (Drexler et al.
-2014). This has attracted attention to areas where high diversity of
-bats, and therefore presumably high diversity of betacoronaviruses, can
-be an important issue for human health (Calisher et al. 2006, Moratelli
-and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
-known hosts of betacoronaviruses, to illustrate where hotspots of host
-diversity are. These results are presented in fig.~\ref{fig:richness}
-(top). As per our current knowledge of which bats are hosts of
-betacoronaviruses, these hotspots are primarily South-East Asia, parts
-of Europe, and to a lesser extent parts of Africa in the -25-0 range of
-latitudes. Previous research (\textbf{Anthony et al., 2017; Mollentze \&
-Streicker, 2020}) states that locally diverse bat communities could
-maintain more viruses and hence, a higher probability of having a
-pathogen that could represent a risk for human health; locally diverse,
-virus-rich bats communities could represent an increased risk of
-spillover under climate change (\textbf{Ice ice berg berg}). This
-probability involves multiple factors, among which the relatedness of
-hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
-Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
-of hosts within a locality to share viruses, which may limit viral
-diversity because of within-host competition (Leeks et al., 2018;
-Sallinen et al., 2020). All things considered, the richness of known
-betacoronaviruses hosts is not a sufficient predictor of spillover risk,
-and we should move towards a more coevolutionary understanding of this
-system.
+Bats are found worldwide and are both one of the most diverse groups
+among mammals (\textbf{Moratelli \& Calisher, 2015}), and one of the
+main animal reservoir for different strains of betacoronaviruses
+(Drexler et al. 2014). This has attracted attention to areas where high
+diversity of bats, and therefore presumably high diversity of
+betacoronaviruses, can be an important issue for human health (Calisher
+et al. 2006, Moratelli and Calisher 2015). By overlaying the IUCN
+rangempas for confirmed bat hosts of betacoronaviruses
+{[}fig.~\ref{fig:richness}; top{]}, we see that the the main hotspots of
+host richness are primarily South-Eastern Asia, parts of Southern
+Europe, and to a lesser extent parts of Africa in the -25-0 range of
+latitudes. The description of host richness is an important first step
+towards understanding risk, as previous research (\textbf{Anthony et
+al., 2017; Mollentze \& Streicker, 2020}) states that locally diverse
+bat communities could maintain more viruses and hence, a higher
+probability of having a pathogen that could represent a risk for human
+health.
 
@@
 \includegraphics{figures/combined_richness.png}
-\caption{Top panel: relative diversoty of known bat hosts of
+\caption{Top panel: relative diversity of known bat hosts of
 betacoronaviruses. This map shows that the region with the largest
@@
 
-High density of hosts sharing the same virus (albeit possibly different
-strains) calls into question the evolution of the bat antiviral immune
-system and its co-evolution with viruses, which may result in distinct
-immunological responses in different areas, as evidenced in other bat
-species (Banerjee et al. 2020). Immune characteristics that allow bats
-to be better adapted to infection by emerging viruses (Gorbunova et al.,
-2020; Irving et al., 2021) may be related to a wide variety of diets
-(Jones et al., 2022; Moreno Santillán et al., 2021; Banerjee et al.,
-2020; Schneeberger et al., 2013). Considering whether viruses easily
-adapted to multiple hosts have lower virulence on these hosts, or lower
-ability to jump to hosts with different immune characteristics, should
-yield valuable additional predictors for the total risk of spillover.
-The subset of chiroptera that are hosts of betacoronaviruses fits the
-evolutionary timeline of the group. Chiropterans can be classified as
-Microchiroptera and macrochiroptera, where macrochiroptera have an older
-history from an evolutionary perspective compared to macrochiroptera
-(Teeling et al. 2005, Springer 2013). Specifically, we would expect that
-the New World group of bats, being more evolutionary distinct, would
+Nevertheless, locally diverse and virus-rich bat communities could
+represent an increased risk of spillover under climate change through
+the creation of novel interactions (\textbf{Ice ice berg berg}), and
+therefore the diversity of betacoronavirus strains should similarly be
+ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
+evolutionary distinctiveness of bats and viruses -- this reveals a
+slightly different portrait than bat richness alone. Chiropterans can be
+classified, from a macro-evolutionary standpoint, as microchiroptera and
+macrochiroptera, where macrochiroptera have an older history from an
+evolutionary perspective compared to macrochiroptera (Teeling et al.
+2005, Springer 2013). Specifically, we would expect that the so-called
+``New World'' group of bats, being more evolutionary distinct, would
 also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
-(bottom) reveals it to be the case. By contrast, South-Eastern Asia has
-a lot of non-evolutionary distinct bats, but evolutionary-distinct
-viruses.
+(bottom) reveals it to be the case, and this region harbors a distinct
+bat-betacoronavirus complex. By contrast, South-Eastern Asia has a lot
+of non-evolutionary distinct bats, but evolutionary-distinct viruses.
 
-Viral evolutionary distinctiveness largely tracks host diversity,
-particularly in southern China but, oddly, not throughout the rest of
-southeast Asia. This indicates, perhaps, that many distinctive viruses
-remain to be discovered in this region (an idea that is unsurprising
-given the growing realization, around the emergence of SARS-CoV-2, that
-a unique lineage of similar viruses are widespread in bats but still
-mostly undescribed). The most distinct betacoronaviruses are found in
-South America, a region with a comparatively lower number of hosts; this
-suggests that the South American bat-betacoronvirus complex has been
-more isolated, and is probably undergoing a different co-evolutionary
-dynamic. Alternatively, this distinctiveness hostpot may be a product of
-under-sampling: South-America is one of the places where the fewest
-betacoronaviruses have been discovered (Anthony et al. 2017), and adding
-more viruses would bring the distinctiveness of known sequences down.
-Previous work has suggested the Americas may be a hotspot of both
-undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
-2017) and coronavirus specifically (Anthony et al. 2017), though not
-necessarily betacoronaviruses, and particularly not those in clades with
-notable zoonotic potential.
+It is noteworthy that outside of South America, viral evolutionary
+distinctiveness does not accurately tracks host diversity, with some
+areas having over-distinct viruses (southern China but, oddly, not the
+rest of southeast Asia). There are a number of likely explanations.
+First, given the richness of bats in southeast Asia, many
+betacoronaviruses likely remain to be discovered in this region. Indeed,
+global predictions by (\textbf{Becker?}) highlight that southeast Asia
+is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
+would likely result in additiona viral discoveries. This idea is
+unsurprising given the growing realization, around the emergence of
+SARS-CoV-2, that a unique lineage of similar viruses are widespread in
+bats but still mostly undescribed. The most distinct
+bats/betacoronavirus complex is found in South America, a region with a
+comparatively lower number of hosts; this matches with the isolation
+through viariance of the host group, and may highlight a different
+co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
+be a product of under-sampling: South-America is one of the places where
+the fewest betacoronaviruses have been discovered (Anthony et al. 2017,
+Olival et al. 2017, Allen et al. 2017), resulting in sparser
+phylogenetic tree, thereby artificially inflating distinctiveness.
+Adding more viruses would bring the distinctiveness of known sequences
+down.
 
@@
 overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}
+
+Despite the difference in evolutionary distinctiveness globally, there
+are reasons to expect that the phylogeography of bats and
+betacoronaviruses should show some degree of congruence. High density of
+hosts sharing the same virus (albeit possibly different strains) can
+drive or result from evolution of the bat antiviral immune system,
+resulting in spatially distinct immunological responses, as evidenced in
+several bat species (Banerjee et al. 2020). Immune characteristics that
+allow bats to be better adapted to infection by emerging viruses
+(Gorbunova et al., 2020; Irving et al., 2021) may be related to a wide
+variety of diets (Jones et al., 2022; Moreno Santillán et al., 2021;
+Banerjee et al., 2020; Schneeberger et al., 2013), themselves likely to
+be driven by spatial effect, especially at the local scale -- bats,
+indeed, occupy a variety of environments.
 
@@
 
+In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
+signal of bats (top) and viruses (bottom) in space; the distinct
+groupings (represented by different colors symbolizing positions in the
+subspace formed by the first two axes of the PCoA) are essentially
+equivalent between the two groups, and can be coarsely delineated as
+southeast Asia, Eurasia above a northing of 25, and Africa and south
+America. These results suggest that, although the evolutionary
+distinctiveness of the bat/betacoronavirus complex varies spatially, the
+system shows an important degree of spatial consistency, with a reduced
+number of bioregions. Available information describing the spillover of
+zoonotic betacoronaviruses of bat origin where data was available before
+and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
+in Wuhan, China; SARS-CoV in \textbf{XXX} based on the presence of
+closest known viruses circulating in nature, and a nearby location where
+serological (antibody) evidence has indicated human exposure to
+SARS-like viruses (Wang et al.~2018 Virologica Sinica); MERS-CoV in
+\textbf{XXX} based on index cases available from a recently-published
+compendium of cases (Ramshaw et al.~2019). For the latest event, most if
+not all index cases are presumed to be camel-to-human transmission, and
+the precise origin point (if it exists) of MERS-CoV in bats is
+uncertain. Recent recombinant canine coronavirus spillover events in
+Haiti (\textbf{ref}) and Europe (\textbf{ref}) are not relevant here, as
+bats' involvement in these cycles of transmission have been supposed to
+be non-existent.
+
 \hypertarget{coevolution-informed-spillover-risk-is-different-in-space}{%
@@
 
-In order to turn the hypotheses based on the Geographic Mosaic Theory of
-Coevolution into a measure of risk, we overlapped three components of
-spillover risk: viral sharing, \emph{i.e.} the chance that two bats will
-share viruses overall; Local Contribution to Beta Diversity, \emph{i.e.}
-the fact that a bat community is compositionally unique compared to the
-average compositional similarity across the entire system; finally, the
-phylogenetic diversity, \emph{i.e.} how dispersed the bats in a location
-are within the tree of life. These results are presented using an
-additive color mapping in Figure xx, and lead to the definition of broad
-biogeographic regions of risk, where the same color represents the same
-type of risk. Pairwise maps of the three components are present in
-supplementary materials.
+As host richness, joint disctinveness, or phylogeographic structure
+suggest that the bat/betacoronavirus complex is globally fragmented
+enough to give rise to both different levels of risk (as evidenced by
+the spatial location of spillover events) and different types of
+co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
+Coevolution \textbf{REF} to provide a measure of risk accounting for
+multiple processes. In fig.~\ref{fig:trivariate}, we overlapped three
+components of spillover risk: viral sharing, \emph{i.e.} the chance that
+two bats will share viruses overall; Local Contribution to Beta
+Diversity, \emph{i.e.} the fact that a bat community is compositionally
+unique compared to the average compositional similarity across the
+entire system; finally, host phylogenetic diversity, \emph{i.e.} how
+dispersed the bats in a location are within the tree of life. This
+approach leads to the definition of broad biogeographic regions of risk,
+where the same color represents the same type of risk. By way of
+constrat to figures fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo},
+these regions do not necessarilly overlap with previous spatial
+partitions of the bat/betacoronavirus complex.
 
@@
 
-Finally, we provide a summary visualization of what available
-information describes the spillover of zoonotic betacoronaviruses of bat
-origin where data was available before and up through the COVID-19
-pandemic. The SARS-CoV-2 outbreak was georeferenced to the initial case
-cluster in Wuhan, China; SARS-CoV was georeferenced based on the cave
-with the closest known viruses circulating in nature (Hu et al.~2017
-PLoS Pathogens), and a nearby location where serological (antibody)
-evidence has indicated human exposure to SARS-like viruses (Wang et
-al.~2018 Virologica Sinica). For MERS-CoV, we presented the index cases
-available from a recently-published compendium of MERS-CoV cases
-(Ramshaw et al.~2019); these are largely if not all presumed to be
-camel-to-human transmission, and the precise origin point of MERS-CoV in
-bats is uncertain. Not shown is a recent case of a recombinant canine
-coronavirus that showed the ability to infect humans, both because this
-study was published after the beginning of the COVID-19 pandemic and
-because bats' involvement in this cycle of transmission has been
-marginal to non-existent.
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
 
@@
 distinct hosts, generating high diversity locally).
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
 
@@
 indirectly bat - human interaction.
+
+This probability involves multiple factors, among which the relatedness
+of hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
+Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
+of hosts within a locality to share viruses, which may limit viral
+diversity because of within-host competition (Leeks et al., 2018;
+Sallinen et al., 2020). All things considered, the richness of known
+betacoronaviruses hosts is not a sufficient predictor of spillover risk,
+and we should move towards a more coevolutionary understanding of this
+system.
+
+Considering whether viruses easily adapted to multiple hosts have lower
+virulence on these hosts, or lower ability to jump to hosts with
+different immune characteristics, should yield valuable additional
+predictors for the total risk of spillover.
 

### 2022-04-30 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e8f7f88e898980f4260ab03e9f584f170625e1fa 🚀

@@
 
-\hypertarget{host-distribution-does-not-match-virus-distinctiveness}{%
-\subsection{Host distribution does not match virus
-distinctiveness}\label{host-distribution-does-not-match-virus-distinctiveness}}
-
-Bats are found worldwide and are one of the most diverse groups among
-mammals (Moratelli \& Calisher, 2015), and one of the main animal
-reservoir for different strains of betacoronaviruses (Drexler et al.
-2014). This has attracted attention to areas where high diversity of
-bats, and therefore presumably high diversity of betacoronaviruses, can
-be an important issue for human health (Calisher et al. 2006, Moratelli
-and Calisher 2015). Accordingly, we collected the IUCN rangemaps for
-known hosts of betacoronaviruses, to illustrate where hotspots of host
-diversity are. These results are presented in fig.~\ref{fig:richness}
-(top). As per our current knowledge of which bats are hosts of
-betacoronaviruses, these hotspots are primarily South-East Asia, parts
-of Europe, and to a lesser extent parts of Africa in the -25-0 range of
-latitudes. Previous research (\textbf{Anthony et al., 2017; Mollentze \&
-Streicker, 2020}) states that locally diverse bat communities could
-maintain more viruses and hence, a higher probability of having a
-pathogen that could represent a risk for human health; locally diverse,
-virus-rich bats communities could represent an increased risk of
-spillover under climate change (\textbf{Ice ice berg berg}). This
-probability involves multiple factors, among which the relatedness of
-hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
-Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
-of hosts within a locality to share viruses, which may limit viral
-diversity because of within-host competition (Leeks et al., 2018;
-Sallinen et al., 2020). All things considered, the richness of known
-betacoronaviruses hosts is not a sufficient predictor of spillover risk,
-and we should move towards a more coevolutionary understanding of this
-system.
+\hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
+\subsection{Host richness does not predict virus
+distinctiveness}\label{host-richness-does-not-predict-virus-distinctiveness}}
+
+Bats are found worldwide and are both one of the most diverse groups
+among mammals (\textbf{Moratelli \& Calisher, 2015}), and one of the
+main animal reservoir for different strains of betacoronaviruses
+(Drexler et al. 2014). This has attracted attention to areas where high
+diversity of bats, and therefore presumably high diversity of
+betacoronaviruses, can be an important issue for human health (Calisher
+et al. 2006, Moratelli and Calisher 2015). By overlaying the IUCN
+rangempas for confirmed bat hosts of betacoronaviruses
+{[}fig.~\ref{fig:richness}; top{]}, we see that the the main hotspots of
+host richness are primarily South-Eastern Asia, parts of Southern
+Europe, and to a lesser extent parts of Africa in the -25-0 range of
+latitudes. The description of host richness is an important first step
+towards understanding risk, as previous research (\textbf{Anthony et
+al., 2017; Mollentze \& Streicker, 2020}) states that locally diverse
+bat communities could maintain more viruses and hence, a higher
+probability of having a pathogen that could represent a risk for human
+health.
 
@@
 
-High density of hosts sharing the same virus (albeit possibly different
-strains) calls into question the evolution of the bat antiviral immune
-system and its co-evolution with viruses, which may result in distinct
-immunological responses in different areas, as evidenced in other bat
-species (Banerjee et al. 2020). Immune characteristics that allow bats
-to be better adapted to infection by emerging viruses (Gorbunova et al.,
-2020; Irving et al., 2021) may be related to a wide variety of diets
-(Jones et al., 2022; Moreno Santillán et al., 2021; Banerjee et al.,
-2020; Schneeberger et al., 2013). Considering whether viruses easily
-adapted to multiple hosts have lower virulence on these hosts, or lower
-ability to jump to hosts with different immune characteristics, should
-yield valuable additional predictors for the total risk of spillover.
-The subset of chiroptera that are hosts of betacoronaviruses fits the
-evolutionary timeline of the group. Chiropterans can be classified as
-Microchiroptera and macrochiroptera, where macrochiroptera have an older
-history from an evolutionary perspective compared to macrochiroptera
-(Teeling et al. 2005, Springer 2013). Specifically, we would expect that
-the New World group of bats, being more evolutionary distinct, would
+Nevertheless, locally diverse and virus-rich bat communities could
+represent an increased risk of spillover under climate change through
+the creation of novel interactions (\textbf{Ice ice berg berg}), and
+therefore the diversity of betacoronavirus strains should similarly be
+ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
+evolutionary distinctiveness of bats and viruses -- this reveals a
+slightly different portrait than bat richness alone. Chiropterans can be
+classified, from a macro-evolutionary standpoint, as microchiroptera and
+macrochiroptera, where macrochiroptera have an older history from an
+evolutionary perspective compared to macrochiroptera (Teeling et al.
+2005, Springer 2013). Specifically, we would expect that the so-called
+``New World'' group of bats, being more evolutionary distinct, would
 also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
-(bottom) reveals it to be the case. By contrast, South-Eastern Asia has
-a lot of non-evolutionary distinct bats, but evolutionary-distinct
-viruses.
-
-Viral evolutionary distinctiveness largely tracks host diversity,
-particularly in southern China but, oddly, not throughout the rest of
-southeast Asia. This indicates, perhaps, that many distinctive viruses
-remain to be discovered in this region (an idea that is unsurprising
-given the growing realization, around the emergence of SARS-CoV-2, that
-a unique lineage of similar viruses are widespread in bats but still
-mostly undescribed). The most distinct betacoronaviruses are found in
-South America, a region with a comparatively lower number of hosts; this
-suggests that the South American bat-betacoronvirus complex has been
-more isolated, and is probably undergoing a different co-evolutionary
-dynamic. Alternatively, this distinctiveness hostpot may be a product of
-under-sampling: South-America is one of the places where the fewest
-betacoronaviruses have been discovered (Anthony et al. 2017), and adding
-more viruses would bring the distinctiveness of known sequences down.
-Previous work has suggested the Americas may be a hotspot of both
-undiscovered bat viruses in general (Olival et al. 2017, Allen et al.
-2017) and coronavirus specifically (Anthony et al. 2017), though not
-necessarily betacoronaviruses, and particularly not those in clades with
-notable zoonotic potential.
+(bottom) reveals it to be the case, and this region harbors a distinct
+bat-betacoronavirus complex. By contrast, South-Eastern Asia has a lot
+of non-evolutionary distinct bats, but evolutionary-distinct viruses.
+
+It is noteworthy that outside of South America, viral evolutionary
+distinctiveness does not accurately tracks host diversity, with some
+areas having over-distinct viruses (southern China but, oddly, not the
+rest of southeast Asia). There are a number of likely explanations.
+First, given the richness of bats in southeast Asia, many
+betacoronaviruses likely remain to be discovered in this region. Indeed,
+global predictions by (\textbf{Becker?}) highlight that southeast Asia
+is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
+would likely result in additiona viral discoveries. This idea is
+unsurprising given the growing realization, around the emergence of
+SARS-CoV-2, that a unique lineage of similar viruses are widespread in
+bats but still mostly undescribed. The most distinct
+bats/betacoronavirus complex is found in South America, a region with a
+comparatively lower number of hosts; this matches with the isolation
+through viariance of the host group, and may highlight a different
+co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
+be a product of under-sampling: South-America is one of the places where
+the fewest betacoronaviruses have been discovered (Anthony et al. 2017,
+Olival et al. 2017, Allen et al. 2017), resulting in sparser
+phylogenetic tree, thereby artificially inflating distinctiveness.
+Adding more viruses would bring the distinctiveness of known sequences
+down.
 
@@
 
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
 Driven by the need to understand the ecological factors involved in the
@@
 
+This probability involves multiple factors, among which the relatedness
+of hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
+Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
+of hosts within a locality to share viruses, which may limit viral
+diversity because of within-host competition (Leeks et al., 2018;
+Sallinen et al., 2020). All things considered, the richness of known
+betacoronaviruses hosts is not a sufficient predictor of spillover risk,
+and we should move towards a more coevolutionary understanding of this
+system.
+
+Considering whether viruses easily adapted to multiple hosts have lower
+virulence on these hosts, or lower ability to jump to hosts with
+different immune characteristics, should yield valuable additional
+predictors for the total risk of spillover.
+
 \textbf{Acknowledgements}: We acknowledge that this study was conducted

### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b6ae14646be7624385df869d6e5eb43e8f23ce1f 🚀

@@
 through recombination, which requires the co-occurrence of sufficiently
-distinct viruses to be a major driver of emergence. In Fig. xx, this
-corresponds to yellow to pale green areas, which are essentially limited
-to South-Eastern Asia, and to some part of Sub-Saharan Africa. Adopting
-a geographic mosaic theory perspective on risk, other regions of the
-world are of lesser concern.
+distinct viruses to be a major driver of emergence. In
+fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
+areas, which are essentially limited to South-Eastern Asia, and to some
+part of Sub-Saharan Africa. Adopting a geographic mosaic theory
+perspective on risk, other regions of the world are of lesser concern
+fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
+diversity or distinctiveness. The simple rationale behind it is that the
+acquisition of viral data is rarely disconnected from the acquisition of
+host data; furthermore, there are more sources of information on hosts
+than on viruses, allowing to develop a host-centric perspective on risk.
+Areas with high bat diversity and high turnover \emph{may} facilitate
+the evolutionary radiation of viruses, matching previous findings that
+the diversification of bat coronaviruses is driven largely by host
+shifts (inter-genus or higher levels of cross-species transmission) and,
+to a lesser degree, cospeciation and sharing (intra-genus cross-species
+transmission; \textbf{Anthony et al.~2017}). This diversification is not
+an actual risk factor for spillover itself, but acts downstream of a
+spillover event by increasing the random chance of the emergence of a
+virus with the raw genomic components required for the potential to
+infect humans.
 
@@
 \end{figure}
-
-Available data on bat betacoronavirus spillover into humans (TP overlay
-on the figure) is limited and circumstantial at best for these purposes,
-but our risk maps suggest that the areas predicted by prior expectations
-about host biogeography correspond loosely to those where previous
-emergence events have been recorded. Areas with high bat diversity and
-high turnover may facilitate the evolutionary radiation of viruses,
-matching previous findings that the diversification of bat coronaviruses
-is driven largely by host shifts (inter-genus or higher levels of
-cross-species transmission) and, to a lesser degree, cospeciation and
-sharing (intra-genus cross-species transmission; Anthony et al.~2017).
-This diversification - while not an actual risk factor for spillover
-itself - likely increases the random chance of a virus with the raw
-genomic components required for the potential to infect humans.
 

### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@038b5ced3967567216335ab65613b6abe255f450 🚀

@@
 \vfill
-Coming soon
+Driven by the need to understand the ecological factors involved in the
+emergence of betacoronavirus through bat hosts, we develop an approach
+to the assesment of spillover risk based on the Geographic Mosaic Theory
+of Coevolution. In doing so, we provide a global mapping of the risk
+posed by viruses, reflecting the fact that this risk has many ecological
+and evolutionary origins. Our framework reveals that reservoir richness
+alone is not a sufficient predictor of risk, and offers alternative
+insights based on viral sharing, host compositional uniqueness, and host
+phylogenetic diversity. By comparing our aggregated measure of risk to a
+proxy for human density, namely the proportion of each pixel that is
+covered by urban or built land, we provide a synthetic risk map,
+allowing the identification of hotspots where the bat-betacoronavirus
+system may originate novel viruses in humans.
 
@@
 variety of diets (Jones et al., 2022; Moreno Santillán et al., 2021;
-Banerjee et al., 2020; Schneeberger et al., 2013), themselves likely to
-be driven by spatial effect, especially at the local scale -- bats,
-indeed, occupy a variety of environments.
+Banerjee et al., 2020; Schneeberger et al., 2013; Muylaert et al.,
+2021), themselves likely to be driven by spatial effect, especially at
+the local scale -- bats, indeed, occupy a variety of environments.
 
@@
 \includegraphics{figures/risk_map.png}
-\caption{This is the legend of the figure\ldots{}}\label{fig:risk}
+\caption{Extraction of a measure of risk based on the colorimetric space
+from fig.~\ref{fig:trivariate}. The risk is a composite measure of the
+color value and angular distance to the yellow hue, as defined in the
+methods, ranged in the unit space. Based on this analyses, regions at
+high risk of spillover are southeast Asia and
+Madagascar.}\label{fig:risk}
 }
@@
 risk map (TP add methods), to provide a single measure of risk varying
-between 0 and 1. This measure is presented in Fig. xxA. However, this
-maps the potential risk, which must be weighed by the potential for
-contacts with humans. As a proxy for this measure, we used the
-proportion of build/urban land from the EarthEnv dataset: this is a
+between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
+However, this maps the potential risk, which must be weighed by the
+potential for contacts with humans. As a proxy for this measure, we used
+the proportion of build/urban land from the EarthEnv dataset: this is a
 reasonable proxy for the density of humans per unit area, which
@@
 the potential risk against measures of land use is required to generate
-a more actionable assessment of risk. This map is presented in Fig. xxB.
-Most of South America and Europe are at low risk, as although densely
-populated, settlements tend to be in areas with lower potential risk.
-However, this mapping reveals that South-East Asia, the Indian
-subcontinent, and parts of sub-Saharan Africa, are at high risk due to
-the overlap between built areas and bat communities representing more
-opportunities for cross-species transmission of betacoronaviruses.
+a more actionable assessment of risk. This map is presented in
+fig.~\ref{fig:compound}. Most of South America and Europe are at low
+risk, as although densely populated, settlements tend to be in areas
+with lower potential risk. However, this mapping reveals that South-East
+Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
+high risk due to the overlap between built areas and bat communities
+representing more opportunities for cross-species transmission of
+betacoronaviruses.
 
@@
 \includegraphics{figures/risk_compounded.png}
-\caption{This is the legend of the figure\ldots{}}\label{fig:compound}
+\caption{Overlap of the percent of each pixel occupied by urbanized
+structures, representing the degree of settlement, on the risk map.
+Darker pixels correspond to more risk, in that the GMTC-derived risk of
+fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
+human populations. This approach increases the relative risk of several
+regions in Africa, and concenreates highlights the risk in India,
+southeast China, and the arabic peninsula.}\label{fig:compound}
 }
@@
 evolutionary dynamics (unrelated viruses coevolving with evolutionarily
-distinct hosts, generating high diversity locally).
-
-Driven by the need to understand the ecological factors involved in the
-emergence of viral pathogens, we spatially mapped bat-betacoronavirus
-interactions worldwide, using (i) a database of known betacov
-hosts(Becker et al., 2020), and (ii) range maps for the hosts according
-to IUCN (IUCN 2021). To reflect the fact that the risk posed by viruses
-has many ecological origins, we quantified the phylogenetic diversity of
-hosts, their compositional uniqueness, and the expected viral sharing.
-Because these components of risk matter when contrasted to human
-density, we compared them to a proxy, namely the proportion of each
-pixel that is covered by urban or built land. This provides a synthetic
-risk map, allowing to identifying of hotspots where the
-bat-betacoronavirus system may originate viruses in humans. SE Asia is
-one of the regions with the highest risk since, according to our
-results, several of its conditions could increase the risk of
-transmission of the virus.
-
-Species richness, therefore, is not a sufficient measure of viral risk.
-This is exemplified in our results, where both South America and
-South-Eastern Asia have a high species richness of betacov hosts, but
-only the latter region has a high risk. Specifically, because previous
-studies propose that Asia is important when it comes to understanding
-the evolutionary origin of various mammalian taxa (Beard C K, 1988).
+distinct hosts, generating high diversity locally). Both of these areas
+should be priority areas for sampling, especially since
+(\textbf{Becker?}) advance that they harbor undiscovered hosts of
+beta-coronaviruses. This diversity of hosts, and the mechanisms by which
+the exchange of viruses occurs between species, is largely affected by
+the local environmental conditions and environmental change.
 
@@
 on RNA viruses suggest the importance of host phylogeography at the time
-of virus dispersal (Gryseels et al., 2017).
-
-This diversity of hosts and how the exchange of viruses occurs between
-species, is largely affected by the different environmental changes, as
-the case of sarbecovirus bats reservoirs (Muylaert et al., 2021) where
-they are affected by the area of the cave or the alteration of the
-forest, which could result in modifications of host distribution.
-Additionally, our results highlight the importance of Asia as a betacov
-hotspot, which is consistent with recent studies (Muylaert et al.,
-2021), where projections on this area suggest that new future events of
-sarbecovirus viral exchange might be easily spread among species or
-humans.
-
-One of these scenarios where interaction between bats and humans can
-occur can be seed dispersal in tropical agroecosystems. It opens the
-discussion of whether the fruits thrown by bats not only disperse seeds
-but could also be a source of indirect interaction between viruses of
-bat origin and humans (Deshpande et al., 2022) . This represents a
-challenge for conservation strategies and disease ecology since we have
-areas with potential zoonotic viruses and bat-human interaction.
-However, it must still be taken into account the quantification of real
-exposure from several scenarios, where there can be directly or
-indirectly bat - human interaction.
-
-This probability involves multiple factors, among which the relatedness
-of hosts (which can make the jumps easier (\textbf{Longdon et al., 2011;
-Mollentze et al., 2020; Wolfe et al., 2007}), and the overall tendency
-of hosts within a locality to share viruses, which may limit viral
-diversity because of within-host competition (Leeks et al., 2018;
-Sallinen et al., 2020). All things considered, the richness of known
-betacoronaviruses hosts is not a sufficient predictor of spillover risk,
-and we should move towards a more coevolutionary understanding of this
-system.
-
-Considering whether viruses easily adapted to multiple hosts have lower
-virulence on these hosts, or lower ability to jump to hosts with
-different immune characteristics, should yield valuable additional
-predictors for the total risk of spillover.
+of virus dispersal (Gryseels et al., 2017). One of these scenarios where
+interaction between bats and humans can occur can be seed dispersal in
+tropical agroecosystems. It opens the discussion of whether the fruits
+thrown by bats not only disperse seeds but could also be a source of
+indirect interaction between viruses of bat origin and humans (Deshpande
+et al., XX). This represents a challenge for conservation strategies and
+disease ecology since we have areas with both potential for the
+acquisition of zoonotic viruses and bat-human interaction; in
+particular, the challenge lies in the fact that actual exposure must be
+quantified from several scenarios, including both direct and indirect
+bat - human interaction.
 

### 2022-04-30 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@038b5ced3967567216335ab65613b6abe255f450 🚀

@@
 
-Coming soon
+Driven by the need to understand the ecological factors involved in the
+emergence of betacoronavirus through bat hosts, we develop an approach
+to the assesment of spillover risk based on the Geographic Mosaic Theory
+of Coevolution. In doing so, we provide a global mapping of the risk
+posed by viruses, reflecting the fact that this risk has many ecological
+and evolutionary origins. Our framework reveals that reservoir richness
+alone is not a sufficient predictor of risk, and offers alternative
+insights based on viral sharing, host compositional uniqueness, and host
+phylogenetic diversity. By comparing our aggregated measure of risk to a
+proxy for human density, namely the proportion of each pixel that is
+covered by urban or built land, we provide a synthetic risk map,
+allowing the identification of hotspots where the bat-betacoronavirus
+system may originate novel viruses in humans.
 

### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@513ca395296bee28e4275a7dae38173958673138 🚀

@@
 allow bats to be better adapted to infection by emerging viruses
-(Gorbunova et al., 2020; Irving et al., 2021) may be related to a wide
-variety of diets (Jones et al., 2022; Moreno Santillán et al., 2021;
-Banerjee et al., 2020; Schneeberger et al., 2013; Muylaert et al.,
-2021), themselves likely to be driven by spatial effect, especially at
-the local scale -- bats, indeed, occupy a variety of environments.
+(Gorbunova et al., 2020; Irving et al., 2021), in addition to being
+hardcoded in their genome (\textbf{Jebb et al.~six\ldots{}}), may be
+related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
+et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013; Muylaert
+et al., 2021), themselves likely to be driven by spatial effects,
+especially at the local scale -- bats, indeed, occupy a variety of
+environments, and therefore display a variety of adaptations to these
+environments.
 
@@
 \includegraphics{figures/combined_biogeo.png}
-\caption{This is the legend of the figure\ldots{}}\label{fig:biogeo}
+\caption{Phylogeographic regions of bats (top) and viruses (bottom)
+based on the joint analysis of their occurrence and phylogenetic
+relatedness. The different colors show tendencies to separate alongside
+the first two components of a PCoA. Note that the PCoA for the bats and
+viruses are independent, and so cannot be compared directly -- that
+being said, the regions can be compared across maps.}\label{fig:biogeo}
 }
@@
 \includegraphics{figures/risk_trivariate.png}
-\caption{This is the legend of the figure\ldots{}}\label{fig:trivariate}
+\caption{Trivariate additive mapping of the components of risk in the
+red/green/blue, where high virus sharing is encoded in the blue channel,
+host phylogenetic diversity in the green channel, and compositional
+uniqueness in the red channel. A pixel that would maximize all measures
+would be a pure white (specifically RGB(1.0, 1.0. 1.0)), and a pixel
+with the lowest possible values would be pure black (specifically
+RGB(0.0, 0.0, 0.0)). Therefore, lighter values (the sum of the three
+channels gets closer to 3) indicate higher risk, and the color indicates
+the proportional distribution of the factors making up the total
+risk.}\label{fig:trivariate}
 }

### 2022-04-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a58f35f21758c1dd7d0ce50ff5ef6ef8a368bfd2 🚀

@@
 human populations. This approach increases the relative risk of several
-regions in Africa, and concenreates highlights the risk in India,
-southeast China, and the arabic peninsula.}\label{fig:compound}
+regions in Africa, and highlights the risk in India, southeast China,
+and the arabic peninsula where areas of high to moderate risk overlap
+with areas of denser population.}\label{fig:compound}
 }

### 2022-05-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a246124feaeeda2ef7c9aff0b272e1e5264cc10b 🚀

@@
 Driven by the need to understand the ecological factors involved in the
-emergence of betacoronavirus through bat hosts, we develop an approach
-to the assesment of spillover risk based on the Geographic Mosaic Theory
-of Coevolution. In doing so, we provide a global mapping of the risk
-posed by viruses, reflecting the fact that this risk has many ecological
-and evolutionary origins. Our framework reveals that reservoir richness
-alone is not a sufficient predictor of risk, and offers alternative
+emergence of betacoronavirus (the genus causing the SARS and MERS
+disease in human) through bat hosts, we develop an approach to the
+assesment of spillover risk based on the Geographic Mosaic Theory of
+Coevolution. In doing so, we provide a global mapping of the spillover
+risk posed by betacoronaviruses, reflecting the fact that this risk has
+many ecological and evolutionary origins. Our framework reveals that
+reservoir richness alone, although a component of viral hazard, is not a
+sufficiently integrative predictor of risk. We offer alternative
 insights based on viral sharing, host compositional uniqueness, and host

### 2022-05-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@21da99f6e7268f26f8e019dd4c4ae5cb5f973aec 🚀

@@
 
-Spillover risk is not unidimensional. From the standpoint of an animal
-community, \emph{i.e.} a pool of suitable hosts, it is driven by a
-multiplicity of factors (Plowright et al. 2017). The global richness of
-hosts is one such component commonly mentioned/analysed (see \emph{e.g.}
-Anthony et al. 2017 for coronaviruses), but there is an argument to be
-made that species who are not competent (or know) hosts of a specific
-virus genus may not factor into this (Plowright et al. 2015), or that
+Spillover risk is multidimensional and complicated. From the standpoint
+of an animal community, \emph{i.e.} a pool of suitable hosts, it is
+driven by a multiplicity of factors (Plowright et al. 2017). Proxies for
+the global richness of hosts is one such component commonly analysed
+(see \emph{e.g.} Anthony et al. 2017 for coronaviruses), but there is an
+argument to be made that species who are not competent (or known) hosts
+of a specific virus genus may not factor into this (Plowright et al.
+2015); this is especially true as competence data increases predictive
+power when the taxonomic scope of hosts of a viral family increases
+(\textbf{Becker2020InfInt?}, \textbf{Mull2022VirIso?}). Similarly, host
 species who are assumed to share viruses at different rates should be
@@
 rapid evolutionary changes in specialism of their pathogens (Agosta et
-al. 2010). In the specific case of generalist pathogens, there is
-conceptual and empirical support to the idea that these community- level
-mechanisms are even more important in driving the overall risk (Power
-and Mitchell 2004).
+al. 2010). In the specific case of generalist pathogens (which is the
+case many viruses in the betacoronavirus genus, see \emph{e.g.}
+\textbf{MacLean2021NatSel?}), there is conceptual and empirical support
+to the idea that these community-level mechanisms are even more
+important in driving the overall risk (Power and Mitchell 2004).
 
 Bats are important reservoir hosts for different classes of
-microorganisms (Chu 2008, Donaldson 2010, Li 2010), some of which can
-threaten human health. Chiropterans emerged around 64 million years ago
-and are one of the most diverse mammalian orders, with an estimated
-richness of more than 14000 species (Peixoto et al. 2018, Simmons and
-Cirranello 2020). They exhibit a broad variety of habitat use,
-behaviour, and feeding strategies, resulting in their playing an
-essential role in the delivery of several ecosystem services tied to
-important ecosystem-derived benefits (Kasso and Balakrishnan 2013). For
-example, over two-thirds of bats are know to be either obligate or
-facultative insectivorous mammals, therefore playing an important role
-in the regulation of insect pests that can affect crops
-(Williams-Guillén et al. 2008, Voigt and Kingston 2016), and vectors of
-diseases that put a risk on human health (Gonsalves et al. 2013a, b).
-Because bats are globally distributed and have a long evolutionary
-history, phylogeographic and biogeographic approaches are required to
-shed light on the extant distribution of coevolutionary processes
-between bats and the pathogens they carry. Not all areas in which bats,
-viruses, and human are co-occuring are facing a risk of spillover
-towards human populations, and the areas in which this risk exist may
-not be facing risks of the same nature and magnitude.
+microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
+threaten human health (\textbf{Letko2020BatVir?},
+\textbf{VanBrussel2022ZooDis?}). Chiropterans emerged around 64 million
+years ago and are one of the most diverse mammalian orders, with an
+estimated richness of more than 1400 species (Peixoto et al. 2018,
+Simmons and Cirranello 2020). They exhibit a broad variety of habitat
+use, behaviour, and feeding strategies, putting them at key positions in
+the delivery and provisioning of several ecosystem services, tied to
+important ecosystem-derived benefits to human (Kasso and Balakrishnan
+2013). For example, bats are an essential component of many
+seed-dispersal networks (\textbf{Mello2011MisPar?}). Over two-thirds of
+bats are know to be either obligate or facultative insectivores,
+therefore playing an important role in the regulation of insect pests
+that can affect crops (Williams-Guillén et al. 2008, Voigt and Kingston
+2016), and vectors of pathogens that put a risk on human health
+(Gonsalves et al. 2013a, b). Because bats are globally distributed and
+have a long evolutionary history, phylogeographic and biogeographic
+approaches are required to shed light on the contemporary distribution
+of coevolutionary processes between bats and the pathogens they host.
+Not all areas in which bats, viruses, and human are co-occuring are
+facing a risk of spillover towards human populations, and the areas in
+which this risk exist may not be facing risks of the same nature and
+magnitude.
 
 In this paper, we examine the biogeographic structure of
-bats-betacoronaviruses associations, based on a curated dataset of known
-and recently discovered hosts. This work is important both as a
-description of the bats-betacoronavirus complex, but aslo because more
+bats-betacoronaviruses associations, based on a curated dataset of all
+confirmed bat hosts of betacoronaviruses. This work is important both as
+a description of the bats-betacoronavirus complex, but aslo because more
 broadly, bats are known reservoirs for a variety of emerging viruses and
-pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
-the needs for bat conservation and disease prevention a potentially
-difficult act and a source of human-wildlife conflicts, especially in
-more densely populated areas (Stone et al. 2015, Rego et al. 2015). By
-drawing on concepts from the Geographic Mosaic Theory of Coevolution
-(Thompson 2005), we turn these associations into a spatially explicit
-additive mapping of zoonotic risk components, which reveals extreme
-heterogeneity of risk at the global scale; furthermore, we identify the
-Amazon and South-Eastern Asia as hotspots of phylogenetic
-distinctiveness of betacoronaviruses (Anthony et al. 2017);
-surprisingly, current data suggest that viral sharing between hosts in
-high in the Amazon and low in South-Eastern Asia, which has the
+other pathogens (Calisher et al. 2006, Melaun et al. 2014), making
+balancing the needs for bat conservation and disease prevention most
+likely very difficult and a source of human-wildlife conflicts,
+especially in more densely populated areas (Stone et al. 2015, Rego et
+al. 2015). By drawing on concepts from the Geographic Mosaic Theory of
+Coevolution (Thompson 2005), we turn these associations into a spatially
+explicit additive mapping of zoonotic risk components, revealing the
+extreme heterogeneity of risk at the global scale. We identify the
+Amazon and South-Eastern Asia as hotspots where the phylogenetic
+distinctiveness of betacoronaviruses is the highest (Anthony et al.
+2017); surprisingly, current data suggest that viral sharing between
+hosts is high in the Amazon and low in South-Eastern Asia, which has the
 potential to result in different evolutionary dynamics between these two

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@3bee2b6e9b1d6cc01c9e2dc9e73ad7bf73394406 🚀

@@
 power when the taxonomic scope of hosts of a viral family increases
-(\textbf{Becker2020InfInt?}, \textbf{Mull2022VirIso?}). Similarly, host
-species who are assumed to share viruses at different rates should be
-weighted accordingly (Albery et al. 2020). In mammals, key functional
-traits (for which phylogeny is a reasonable proxy) are determinants of
-the spillover potential (Olival et al. 2017); these include, notably,
-body mass and affinity for urban environments (Albery et al. 2022).
-Finally, especially when the pool of potential hosts spans the entire
-globe, there may be local host pools that are highly unique; not having
-been observed in other locations, these can act on the overall risk
-either by providing novel contact opportunities, reflecting unique
+(Becker et al. 2020, Mull et al. 2022). Similarly, host species who are
+assumed to share viruses at different rates should be weighted
+accordingly (Albery et al. 2020). In mammals, key functional traits (for
+which phylogeny is a reasonable proxy) are determinants of the spillover
+potential (Olival et al. 2017); these include, notably, body mass and
+affinity for urban environments (Albery et al. 2022). Finally,
+especially when the pool of potential hosts spans the entire globe,
+there may be local host pools that are highly unique; not having been
+observed in other locations, these can act on the overall risk either by
+providing novel contact opportunities, reflecting unique
 host-environment combinations (Engering et al. 2013), or facilitating
@@
 al. 2010). In the specific case of generalist pathogens (which is the
-case many viruses in the betacoronavirus genus, see \emph{e.g.}
-\textbf{MacLean2021NatSel?}), there is conceptual and empirical support
-to the idea that these community-level mechanisms are even more
-important in driving the overall risk (Power and Mitchell 2004).
+case many viruses in the betacoronavirus genus, see \emph{e.g.} MacLean
+et al. 2021), there is conceptual and empirical support to the idea that
+these community-level mechanisms are even more important in driving the
+overall risk (Power and Mitchell 2004).
 
@@
 microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
-threaten human health (\textbf{Letko2020BatVir?},
-\textbf{VanBrussel2022ZooDis?}). Chiropterans emerged around 64 million
-years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species (Peixoto et al. 2018,
-Simmons and Cirranello 2020). They exhibit a broad variety of habitat
-use, behaviour, and feeding strategies, putting them at key positions in
-the delivery and provisioning of several ecosystem services, tied to
-important ecosystem-derived benefits to human (Kasso and Balakrishnan
-2013). For example, bats are an essential component of many
-seed-dispersal networks (\textbf{Mello2011MisPar?}). Over two-thirds of
-bats are know to be either obligate or facultative insectivores,
-therefore playing an important role in the regulation of insect pests
-that can affect crops (Williams-Guillén et al. 2008, Voigt and Kingston
-2016), and vectors of pathogens that put a risk on human health
-(Gonsalves et al. 2013a, b). Because bats are globally distributed and
-have a long evolutionary history, phylogeographic and biogeographic
-approaches are required to shed light on the contemporary distribution
-of coevolutionary processes between bats and the pathogens they host.
-Not all areas in which bats, viruses, and human are co-occuring are
-facing a risk of spillover towards human populations, and the areas in
-which this risk exist may not be facing risks of the same nature and
-magnitude.
+threaten human health (Letko et al. 2020, Van Brussel and Holmes 2022).
+Chiropterans emerged around 64 million years ago and are one of the most
+diverse mammalian orders, with an estimated richness of more than 1400
+species (Peixoto et al. 2018, Simmons and Cirranello 2020). They exhibit
+a broad variety of habitat use, behaviour, and feeding strategies,
+putting them at key positions in the delivery and provisioning of
+several ecosystem services, tied to important ecosystem-derived benefits
+to human (Kasso and Balakrishnan 2013). For example, bats are an
+essential component of many seed-dispersal networks (Mello et al. 2011).
+Over two-thirds of bats are know to be either obligate or facultative
+insectivores, therefore playing an important role in the regulation of
+insect pests that can affect crops (Williams-Guillén et al. 2008, Voigt
+and Kingston 2016), and vectors of pathogens that put a risk on human
+health (Gonsalves et al. 2013a, b). Because bats are globally
+distributed and have a long evolutionary history, phylogeographic and
+biogeographic approaches are required to shed light on the contemporary
+distribution of coevolutionary processes between bats and the pathogens
+they host. Not all areas in which bats, viruses, and human are
+co-occuring are facing a risk of spillover towards human populations,
+and the areas in which this risk exist may not be facing risks of the
+same nature and magnitude.
 
@@
 latitudes. The description of host richness is an important first step
-towards understanding risk, as previous research (\textbf{Anthony et
-al., 2017; Mollentze \& Streicker, 2020}) states that locally diverse
-bat communities could maintain more viruses and hence, a higher
-probability of having a pathogen that could represent a risk for human
-health.
+towards understanding risk, as previous research (Anthony et al. 2017,
+Mollentze and Streicker 2020) states that locally diverse bat
+communities could maintain more viruses and hence, a higher probability
+of having a pathogen that could represent a risk for human health.
 
@@
 to a lesser degree, cospeciation and sharing (intra-genus cross-species
-transmission; \textbf{Anthony et al.~2017}). This diversification is not
-an actual risk factor for spillover itself, but acts downstream of a
+transmission; Anthony et al. 2017). This diversification is not an
+actual risk factor for spillover itself, but acts downstream of a
 spillover event by increasing the random chance of the emergence of a
@@
 \end{figure}
+
+From another perspective, areas of high host uniqueness and virus
+sharing (red-to-pink) could provide hotspots of betacoronavirus risk
+through mixing of unique viruses (via codivergence) and in turn
+recombination. Under our framework, such a hotspot was identified in
+Madagascar, where most bat species are endemic following evolutionary
+divergence from sister species in both African and Asian continents
+(\emph{e.g.} Shi et al. 2014). Recent surveillance (Kettenburg et al.
+2022) has identified a novel betacoronavirus (in the subgenus
+\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
+(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
+strong proof of principle in model predictions.
 
@@
 
+\leavevmode\hypertarget{ref-Becker2020InfInt}{}%
+Becker, D. J. et al. 2020. Beyond Infection: Integrating Competence into
+Reservoir Host Prediction. - Trends in Ecology \& Evolution 35:
+1062--1065.
+
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
@@
 
+\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
+Kettenburg, G. et al. 2022. Full Genome Nobecovirus Sequences From
+Malagasy Fruit Bats Define a Unique Evolutionary History for This
+Coronavirus Clade. - Frontiers in Public Health in press.
+
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
@@
 Forest Ecosystems 6: 7.
+
+\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
+Letko, M. et al. 2020. Bat-borne virus diversity, spillover and
+emergence. - Nature Reviews Microbiology 18: 461--471.
+
+\leavevmode\hypertarget{ref-MacLean2021NatSel}{}%
+MacLean, O. A. et al. 2021. Natural selection in the evolution of
+SARS-CoV-2 in bats created a generalist virus and highly capable human
+pathogen. - PLOS Biology 19: e3001115.
 
@@
 
+\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
+Mello, M. A. R. et al. 2011. The Missing Part of Seed Dispersal
+Networks: Structure and Robustness of Bat-Fruit Interactions. - PLOS ONE
+6: e17395.
+
+\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
+Mollentze, N. and Streicker, D. G. 2020. Viral zoonotic risk is
+homogenous among taxonomic orders of mammalian and avian reservoir
+hosts. - Proceedings of the National Academy of Sciences 117:
+9423--9430.
+
 \leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
@@
 
+\leavevmode\hypertarget{ref-Mull2022VirIso}{}%
+Mull, N. et al. 2022. Virus isolation data improve host predictions for
+New World rodent orthohantaviruses. - Journal of Animal Ecology in
+press.
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
@@
 
+\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
+Shi, J. J. et al. 2014. A Deep Divergence Time between Sister Species of
+Eidolon (Pteropodidae) with Evidence for Widespread Panmixia. - Acta
+Chiropterologica 16: 279--292.
+
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
@@
 
+\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
+Van Brussel, K. and Holmes, E. C. 2022. Zoonotic disease and virome
+diversity in bats. - Current Opinion in Virology 52: 192--202.
+
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ec18cd6940b2ac05776c605d198846cc9cd2f92c 🚀

@@
 reasonable proxy for the density of humans per unit area, which
-increases the probability of pathogen spread more widely (Hazarie et
-al., 2021). Since human activity is required to amplify the frequency of
+increases the probability of pathogen spread more widely (Hazarie et al.
+2021). Since human activity is required to amplify the frequency of
 virus encounters and thus create areas of viral amplification, mapping
@@
 virus to adapt to various taxonomic orders and ecological groups
-(Kreuder Johnson et al., 2015); are more likely to amplify viral
+(Kreuder Johnson et al. 2015), are more likely to amplify viral
 spillover, followed by secondary human-to-human transmission, and
-geographical spread (Hazarie et al., 2021). High viral host plasticity
-is an especially important trait for RNA viruses such as betacov
-(Kreuder Johnson et al., 2015; Haddad et al., 2021). Indeed, our
-analysis of viral sequences reveals that Latin America is a hotspot of
-viral distinctiveness, suggesting that this part of the bats-betacov
-system may be undergoing independent evolutionary dynamics (related
-species sharing viruses that are different from the rest of the global
-pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
-richness is high but sharing is low; this suggests a different type of
-evolutionary dynamics (unrelated viruses coevolving with evolutionarily
-distinct hosts, generating high diversity locally). Both of these areas
-should be priority areas for sampling, especially since
-(\textbf{Becker?}) advance that they harbor undiscovered hosts of
-beta-coronaviruses. This diversity of hosts, and the mechanisms by which
-the exchange of viruses occurs between species, is largely affected by
-the local environmental conditions and environmental change.
+geographical spread (Hazarie et al. 2021). High viral host plasticity is
+an especially important trait for RNA viruses like betacoronaviruses
+(Haddad et al. 2021). Indeed, our analysis of viral sequences reveals
+that Latin America is a hotspot of viral distinctiveness, suggesting
+that this part of the bats-betacoronavirus system may be undergoing
+independent evolutionary dynamics (related species sharing viruses that
+are different from the rest of the global pool). The other hotspot of
+viral distinctiveness is S.E. Asia, in which richness is high but
+sharing is low; this suggests a different type of evolutionary dynamics
+(unrelated viruses coevolving with evolutionarily distinct hosts,
+generating high diversity locally). Both of these areas should be
+priority areas for sampling, especially since Becker et al. (2022)
+advance that they harbor undiscovered hosts of beta-coronaviruses. This
+diversity of hosts, and the mechanisms by which the exchange of viruses
+occurs between species, is largely affected by the local environmental
+conditions and environmental change.
 
 There are several factors that drive changes in the diversity of bats
-(Alves et al., 2018), but human activities' effects on the ecosystem
+(Alves et al. 2018), but human activities' effects on the ecosystem
 (like modifications of land use) could significantly decrease it.
@@
 contact between humans and potential hosts, we also increase the risk of
-emergence of novel diseases (Johnson et al., 2020), as previous studies
+emergence of novel diseases (Johnson et al. 2020), as previous studies
 on RNA viruses suggest the importance of host phylogeography at the time
-of virus dispersal (Gryseels et al., 2017). One of these scenarios where
+of virus dispersal (Gryseels et al. 2017). One of these scenarios where
 interaction between bats and humans can occur can be seed dispersal in
@@
 indirect interaction between viruses of bat origin and humans (Deshpande
-et al., XX). This represents a challenge for conservation strategies and
-disease ecology since we have areas with both potential for the
-acquisition of zoonotic viruses and bat-human interaction; in
-particular, the challenge lies in the fact that actual exposure must be
-quantified from several scenarios, including both direct and indirect
-bat - human interaction.
+et al. 2022). This represents a challenge for conservation strategies
+and disease ecology since some areas can haveboth potential for the
+acquisition of zoonotic viruses and bat-human interactions; in
+particular, the challenge lies in the fact that actual exposure must
+then be quantified accounting for several transmission scenarios,
+including both direct and indirect bat - human interaction.
 
@@
 
+\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
+Alves, D. M. C. C. et al. 2018. Geographic variation in the relationship
+between large-scale environmental determinants and bat species richness.
+- Basic and Applied Ecology 27: 1--8.
+
 \leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
@@
 
+\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
+Deshpande, K. et al. 2022. Forbidden fruits? Ecosystem services from
+seed dispersal by fruit bats in the context of latent zoonotic risk. -
+Oikos (Copenhagen, Denmark): oik.08359.
+
 \leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
@@
 
+\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
+Gryseels, S. et al. 2017. When Viruses Don't Go Viral: The Importance of
+Host Phylogeographic Structure in the Spatial Spread of Arenaviruses (JH
+Kuhn, Ed.). - PLOS Pathogens 13: e1006073.
+
+\leavevmode\hypertarget{ref-Haddad2021SarPos}{}%
+Haddad, D. et al. 2021. SARS-CoV-2: Possible recombination and emergence
+of potentially more virulent strains (H Attoui, Ed.). - PLOS ONE 16:
+e0251368.
+
+\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
+Hazarie, S. et al. 2021. Interplay between population density and
+mobility in determining the spread of epidemics in cities. -
+Communications Physics 4: 191.
+
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
@@
 IUCN 2021. The IUCN Red List of Threatened Species.
+
+\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
+Johnson, C. K. et al. 2020. Global shifts in mammalian population trends
+reveal key predictors of virus spillover risk. - Proceedings of the
+Royal Society B: Biological Sciences 287: 20192736.
 
@@
 2029--2053.
+
+\leavevmode\hypertarget{ref-KreuderJohnson2015SpiPan}{}%
+Kreuder Johnson, C. et al. 2015. Spillover and pandemic properties of
+zoonotic viruses with high host plasticity. - Scientific Reports 5:
+14830.
 

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6f3e9e1ff2f027a1487065eab454118d787ca67e 🚀

@@
 and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
-in Wuhan, China; SARS-CoV in \textbf{XXX} based on the presence of
+in Wuhan, China; SARS-CoV in Guangdong, China based on the presence of
 closest known viruses circulating in nature, and a nearby location where
 serological (antibody) evidence has indicated human exposure to
-SARS-like viruses (Wang et al.~2018 Virologica Sinica); MERS-CoV in
-\textbf{XXX} based on index cases available from a recently-published
-compendium of cases (Ramshaw et al.~2019). For the latest event, most if
-not all index cases are presumed to be camel-to-human transmission, and
-the precise origin point (if it exists) of MERS-CoV in bats is
-uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti (\textbf{ref}) and Europe (\textbf{ref}) are not relevant here, as
+SARS-like viruses (Wang et al. 2018); MERS-CoV in Saudi Arabia based on
+index cases available from a recently-published compendium of cases
+(Ramshaw et al. 2019). For the latest event, most if not all index cases
+are presumed to be camel-to-human transmission, and the precise origin
+point (if it exists) of MERS-CoV in bats is uncertain. Recent
+recombinant canine coronavirus spillover events in Haiti (Lednicky et
+al. 2021) and Europe (Vlasova et al. 2022) are not relevant here, as
 bats' involvement in these cycles of transmission have been supposed to
-be non-existent.
+be non-existent. These index cases fall within different phylogeographic
+bioregions (fig.~\ref{fig:biogeo}), which further highlight the issue
+that different host-virus sub-systems may lead to widespread emergence.
 
@@
 representing more opportunities for cross-species transmission of
-betacoronaviruses.
+betacoronaviruses. In looking for the origins of SARS in China, Xu et
+al. (2004) present serological evidence that strongest human-animal
+contact results in higher risk of virus exposure, regardless of the
+animal species, but that different types of contact had different
+impacts. Ideally, finer-grained information about human activity (rather
+than human presence through anthropisation) could allow to partition
+this risk further.
 
@@
 
+\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
+Lednicky, J. A. et al. 2021. Isolation of a Novel Recombinant Canine
+Coronavirus from a Visitor to Haiti: Further Evidence of Transmission of
+Coronaviruses of Zoonotic Origin to Humans. - Clinical Infectious
+Diseases: An Official Publication of the Infectious Diseases Society of
+America: ciab924.
+
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
@@
 
+\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
+Ramshaw, R. E. et al. 2019. A database of geopositioned Middle East
+Respiratory Syndrome Coronavirus occurrences. - Scientific Data 6: 318.
+
 \leavevmode\hypertarget{ref-Rego2015AssHum}{}%
@@
 
+\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
+Vlasova, A. N. et al. 2022. Animal alphacoronaviruses found in human
+patients with acute respiratory illness in different countries. -
+Emerging Microbes \& Infections 11: 699--702.
+
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
@@
 
+\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
+Wang, N. et al. 2018. Serological Evidence of Bat SARS-Related
+Coronavirus Infection in Humans, China. - Virologica Sinica 33:
+104--107.
+
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
@@
 
+\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
+Xu, R.-H. et al. 2004. Epidemiologic Clues to SARS Origin in China. -
+Emerging Infectious Diseases 10: 1030--1037.
+
 \end{CSLReferences}

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9b5e8dc75883cba38d2bbae428327ca82b972ccd 🚀

@@
 
-Spillover risk is multidimensional and complicated. From the standpoint
-of an animal community, \emph{i.e.} a pool of suitable hosts, it is
-driven by a multiplicity of factors (Plowright et al. 2017). Proxies for
-the global richness of hosts is one such component commonly analysed
-(see \emph{e.g.} Anthony et al. 2017 for coronaviruses), but there is an
-argument to be made that species who are not competent (or known) hosts
-of a specific virus genus may not factor into this (Plowright et al.
-2015); this is especially true as competence data increases predictive
-power when the taxonomic scope of hosts of a viral family increases
-(Becker et al. 2020, Mull et al. 2022). Similarly, host species who are
-assumed to share viruses at different rates should be weighted
-accordingly (Albery et al. 2020). In mammals, key functional traits (for
-which phylogeny is a reasonable proxy) are determinants of the spillover
-potential (Olival et al. 2017); these include, notably, body mass and
-affinity for urban environments (Albery et al. 2022). Finally,
-especially when the pool of potential hosts spans the entire globe,
-there may be local host pools that are highly unique; not having been
-observed in other locations, these can act on the overall risk either by
-providing novel contact opportunities, reflecting unique
-host-environment combinations (Engering et al. 2013), or facilitating
-rapid evolutionary changes in specialism of their pathogens (Agosta et
-al. 2010). In the specific case of generalist pathogens (which is the
-case many viruses in the betacoronavirus genus, see \emph{e.g.} MacLean
-et al. 2021), there is conceptual and empirical support to the idea that
-these community-level mechanisms are even more important in driving the
-overall risk (Power and Mitchell 2004).
-
-Bats are important reservoir hosts for different classes of
-microorganisms (Chu 2008, Donaldson 2010, Li 2010), many of which can
-threaten human health (Letko et al. 2020, Van Brussel and Holmes 2022).
-Chiropterans emerged around 64 million years ago and are one of the most
-diverse mammalian orders, with an estimated richness of more than 1400
-species (Peixoto et al. 2018, Simmons and Cirranello 2020). They exhibit
-a broad variety of habitat use, behaviour, and feeding strategies,
-putting them at key positions in the delivery and provisioning of
-several ecosystem services, tied to important ecosystem-derived benefits
-to human (Kasso and Balakrishnan 2013). For example, bats are an
-essential component of many seed-dispersal networks (Mello et al. 2011).
-Over two-thirds of bats are know to be either obligate or facultative
-insectivores, therefore playing an important role in the regulation of
-insect pests that can affect crops (Williams-Guillén et al. 2008, Voigt
-and Kingston 2016), and vectors of pathogens that put a risk on human
-health (Gonsalves et al. 2013a, b). Because bats are globally
-distributed and have a long evolutionary history, phylogeographic and
-biogeographic approaches are required to shed light on the contemporary
-distribution of coevolutionary processes between bats and the pathogens
-they host. Not all areas in which bats, viruses, and human are
-co-occuring are facing a risk of spillover towards human populations,
-and the areas in which this risk exist may not be facing risks of the
-same nature and magnitude.
+Spillover risk is multidimensional and complicated. Within a pool of
+competent hosts, it is driven by a multiplicity of factors (Plowright et
+al. 2017). Although proxies for the local richness of hosts is commonly
+analysed (see \emph{e.g.} Anthony et al. 2017 for coronaviruses), there
+is an argument to be made that species who are not (known as) competent
+hosts of a specific virus genus may not factor into this (Plowright et
+al. 2015), calling for species-level information. This is especially
+true as competence data increases predictive power when the taxonomic
+scope of hosts of a viral family increases (Becker et al. 2020, Mull et
+al. 2022). Similarly, host species who share viruses at different rates
+should be weighted accordingly (Albery et al. 2020). In mammals, key
+functional traits (for which phylogeny is a reasonable proxy) are
+determinants of the spillover potential (Olival et al. 2017); these
+include, notably, body mass, and affinity for urban environments (Albery
+et al. 2022). Finally, especially when the pool of potential hosts spans
+the entire globe, there may be local host pools that are highly unique;
+not having been observed in other locations, these can act on the
+overall risk either by providing novel contact opportunities, reflecting
+unique host-environment combinations (Engering et al. 2013), or
+facilitating rapid evolutionary changes in specialism of their pathogens
+(Agosta et al. 2010). In the specific case of generalist pathogens
+(which is the case many viruses in the betacoronavirus genus, see
+\emph{e.g.} MacLean et al. 2021), there is conceptual and empirical
+support to the idea that these community-level mechanisms are even more
+important in driving the overall risk (Power and Mitchell 2004).
 
@@
 bats-betacoronaviruses associations, based on a curated dataset of all
-confirmed bat hosts of betacoronaviruses. This work is important both as
-a description of the bats-betacoronavirus complex, but aslo because more
-broadly, bats are known reservoirs for a variety of emerging viruses and
-other pathogens (Calisher et al. 2006, Melaun et al. 2014), making
-balancing the needs for bat conservation and disease prevention most
-likely very difficult and a source of human-wildlife conflicts,
-especially in more densely populated areas (Stone et al. 2015, Rego et
-al. 2015). By drawing on concepts from the Geographic Mosaic Theory of
-Coevolution (Thompson 2005), we turn these associations into a spatially
-explicit additive mapping of zoonotic risk components, revealing the
-extreme heterogeneity of risk at the global scale. We identify the
-Amazon and South-Eastern Asia as hotspots where the phylogenetic
-distinctiveness of betacoronaviruses is the highest (Anthony et al.
-2017); surprisingly, current data suggest that viral sharing between
-hosts is high in the Amazon and low in South-Eastern Asia, which has the
-potential to result in different evolutionary dynamics between these two
-regions.
+confirmed bat hosts of betacoronaviruses. By drawing on concepts from
+the Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005), we
+turn these associations into a spatially explicit additive mapping of
+zoonotic risk components, revealing the extreme heterogeneity of risk at
+the global scale. Explicitely framing the notion of spillover risk based
+on propositions from the GMTC (which is to say, based on a framework
+linking interactions between species to change within species) is a
+novel idea, that should be relatively general. Indeed, it only assumes
+the action of well described evolutionary mechanisms. The benefit of
+this approach is to provide the potential for a more dynamic and nuanced
+understanding of risk: not only on ecological timescales, but also by
+providing clues about which areas can change over micro-evolutionary
+timescales. This provides a way to look at spatial structure by
+accounting for more notions than species richness/similarity, but also a
+way to identify spatial areas of higher risk.
+
+We identify the Amazon and South-Eastern Asia as hotspots where the
+phylogenetic distinctiveness of betacoronaviruses is the highest
+(Anthony et al. 2017); surprisingly, current data suggest that viral
+sharing between hosts is high in the Amazon and low in South-Eastern
+Asia, which has the potential to result in different evolutionary
+dynamics between these two regions, hinting at different futures for
+their viral communities. This work is important both as a description of
+the bats-betacoronavirus complex, but also because more broadly, bats
+are known reservoirs for a variety of emerging viruses and other
+pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
+the needs for bat conservation and disease prevention most likely very
+difficult and a source of human-wildlife conflicts, especially in more
+densely populated areas (Stone et al. 2015, Rego et al. 2015).
 
@@
 
+Bats are important reservoir hosts for different classes of
+microorganisms, many of which a threat to human health (Letko et al.
+2020, Van Brussel and Holmes 2022). Chiropterans emerged around 64
+million years ago and are one of the most diverse mammalian orders, with
+an estimated richness of more than 1400 species (Peixoto et al. 2018,
+Simmons and Cirranello 2020). They exhibit a broad variety of habitat
+use, behaviour, and feeding strategies, putting them at key positions in
+the delivery and provisioning of several ecosystem services, tied to
+important ecosystem-derived benefits to human (Kasso and Balakrishnan
+2013). For example, bats are an essential component of many
+seed-dispersal networks (Mello et al. 2011). Over two-thirds of bats are
+know to be either obligate or facultative insectivores, therefore
+playing an important role in the regulation of insect pests that can
+affect crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016),
+and vectors of pathogens that put a risk on human health (Gonsalves et
+al. 2013a, b). Because bats are globally distributed and have a long
+evolutionary history, phylogeographic and biogeographic approaches are
+required to shed light on the contemporary distribution of
+coevolutionary processes between bats and the pathogens they host. Not
+all areas in which bats, viruses, and human are co-occuring are facing a
+risk of spillover towards human populations, and the areas in which this
+risk exist may not be facing risks of the same nature and magnitude.
+
 There are several factors that drive changes in the diversity of bats

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@3386f7c9b624ab21772c3543697456665d27c11e 🚀

@@
 
-Based on the previous result, we extracted the yellow component of the
-risk map (TP add methods), to provide a single measure of risk varying
+Based on the previous result, we extracted the risk component from the
+composite map (see Methods), to provide a single measure of risk varying
 between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e964b8d8ff0c2847e33b1bf3933454c77c226731 🚀

@@
 represent an increased risk of spillover under climate change through
-the creation of novel interactions (\textbf{Ice ice berg berg}), and
-therefore the diversity of betacoronavirus strains should similarly be
-ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
-evolutionary distinctiveness of bats and viruses -- this reveals a
-slightly different portrait than bat richness alone. Chiropterans can be
-classified, from a macro-evolutionary standpoint, as microchiroptera and
-macrochiroptera, where macrochiroptera have an older history from an
-evolutionary perspective compared to macrochiroptera (Teeling et al.
-2005, Springer 2013). Specifically, we would expect that the so-called
-``New World'' group of bats, being more evolutionary distinct, would
-also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
-(bottom) reveals it to be the case, and this region harbors a distinct
+the creation of novel interactions (Carlson et al. 2022), and therefore
+the diversity of betacoronavirus strains should similarly be ccounted
+for. In fig.~\ref{fig:richness} (bottom), we contrast the evolutionary
+distinctiveness of bats and viruses -- this reveals a slightly different
+portrait than bat richness alone. Chiropterans can be classified, from a
+macro-evolutionary standpoint, as microchiroptera and macrochiroptera,
+where macrochiroptera have an older history from an evolutionary
+perspective compared to macrochiroptera (Teeling et al. 2005, Springer
+2013). Specifically, we would expect that the so-called ``New World''
+group of bats, being more evolutionary distinct, would also have
+evolutionary distinct viruses. Indeed fig.~\ref{fig:richness} (bottom)
+reveals it to be the case, and this region harbors a distinct
 bat-betacoronavirus complex. By contrast, South-Eastern Asia has a lot
@@
 
+\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
+Carlson, C. J. et al. 2022. Climate change increases cross-species viral
+transmission risk. - Nature: 1--1.
+
 \leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9fbaf07f388a6fd032325943e161934ee12c9d73 🚀

@@
 betacoronaviruses likely remain to be discovered in this region. Indeed,
-global predictions by (\textbf{Becker?}) highlight that southeast Asia
+global predictions by Becker et al. (2022) highlight that southeast Asia
 is a likely hostpot of unconfirmed hosts of betacoronaviruses, which

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e57162f06f949bbb2947b87ea8c1580e70b187be 🚀

@@
 allow bats to be better adapted to infection by emerging viruses
-(Gorbunova et al., 2020; Irving et al., 2021), in addition to being
-hardcoded in their genome (\textbf{Jebb et al.~six\ldots{}}), may be
-related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
-et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013; Muylaert
-et al., 2021), themselves likely to be driven by spatial effects,
-especially at the local scale -- bats, indeed, occupy a variety of
-environments, and therefore display a variety of adaptations to these
+(Gorbunova et al. 2020, Irving et al. 2021), in addition to being
+hardcoded in their genome (Jebb et al. 2020), may be related to a wide
+variety of diets (Banerjee et al. 2020, Moreno Santillán et al. 2021,
+Muylaert et al. 2021), themselves likely to be driven by spatial
+effects, especially at the local scale -- bats, indeed, occupy a variety
+of environments, and therefore display a variety of adaptations to these
 environments.
@@
 
+\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
+Gorbunova, V. et al. 2020. The World Goes Bats: Living Longer and
+Tolerating Viruses. - Cell Metabolism 32: 31--43.
+
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
@@
 
+\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
+Irving, A. T. et al. 2021. Lessons from the host defences of bats, a
+unique viral reservoir. - Nature 589: 363--370.
+
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
@@
 IUCN 2021. The IUCN Red List of Threatened Species.
+
+\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
+Jebb, D. et al. 2020. Six reference-quality genomes reveal evolution of
+bat adaptations. - Nature 583: 578--584.
 
@@
 
+\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
+Moreno Santillán, D. D. et al. 2021. Large-scale genome sampling reveals
+unique immunity and metabolic adaptations in bats. - Molecular Ecology:
+mec.16027.
+
 \leavevmode\hypertarget{ref-Mull2022VirIso}{}%
@@
 
+\leavevmode\hypertarget{ref-Muylaert2021PreFut}{}%
+Muylaert, R. L. et al. 2021. Present and future distribution of bat
+hosts of sarbecoviruses: Implications for conservation and public
+health.
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a486162393bdd46cba57058d9855197a42bd9b0c 🚀

@@
 co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
-Coevolution \textbf{REF} to provide a measure of risk accounting for
+Coevolution (Thompson 2005) to provide a measure of risk accounting for
 multiple processes. In fig.~\ref{fig:trivariate}, we overlapped three

### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9589ee267f240de2621beeb994006c7e11f55c85 🚀

@@
 Bats are found worldwide and are both one of the most diverse groups
-among mammals (\textbf{Moratelli \& Calisher, 2015}), and one of the
-main animal reservoir for different strains of betacoronaviruses
-(Drexler et al. 2014). This has attracted attention to areas where high
-diversity of bats, and therefore presumably high diversity of
-betacoronaviruses, can be an important issue for human health (Calisher
-et al. 2006, Moratelli and Calisher 2015). By overlaying the IUCN
-rangempas for confirmed bat hosts of betacoronaviruses
-{[}fig.~\ref{fig:richness}; top{]}, we see that the the main hotspots of
-host richness are primarily South-Eastern Asia, parts of Southern
-Europe, and to a lesser extent parts of Africa in the -25-0 range of
-latitudes. The description of host richness is an important first step
-towards understanding risk, as previous research (Anthony et al. 2017,
-Mollentze and Streicker 2020) states that locally diverse bat
-communities could maintain more viruses and hence, a higher probability
-of having a pathogen that could represent a risk for human health.
+among mammals and one of the main animal reservoir for different strains
+of betacoronaviruses (Drexler et al. 2014). This has attracted attention
+to areas where high diversity of bats, and therefore presumably high
+diversity of betacoronaviruses, can be an important issue for human
+health (Calisher et al. 2006, Moratelli and Calisher 2015). By
+overlaying the IUCN rangempas for confirmed bat hosts of
+betacoronaviruses {[}fig.~\ref{fig:richness}; top{]}, we see that the
+the main hotspots of host richness are primarily South-Eastern Asia,
+parts of Southern Europe, and to a lesser extent parts of Africa in the
+-25-0 range of latitudes. The description of host richness is an
+important first step towards understanding risk, as previous research
+(Anthony et al. 2017, Mollentze and Streicker 2020) states that locally
+diverse bat communities could maintain more viruses and hence, a higher
+probability of having a pathogen that could represent a risk for human
+health.
 

### 2022-06-17 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@fa58cdffbc410db83637c21cff76d58ca3bd6ac3 🚀

@@
 {\Large\bfseries\sffamily The coevolutionary mosaic of
-bat-betacoronaviruses spillover risk}
+\emph{Betacoronavirus} spillover risk from bat hosts}
 \vskip 5em
@@
 (Agosta et al. 2010). In the specific case of generalist pathogens
-(which is the case many viruses in the betacoronavirus genus, see
+(which is the case many viruses in the \emph{Betacoronavirus} genus, see
 \emph{e.g.} MacLean et al. 2021), there is conceptual and empirical
@@
 We identify the Amazon and South-Eastern Asia as hotspots where the
-phylogenetic distinctiveness of betacoronaviruses is the highest
+phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
 (Anthony et al. 2017); surprisingly, current data suggest that viral
@@
 their viral communities. This work is important both as a description of
-the bats-betacoronavirus complex, but also because more broadly, bats
+the bats-betacoronaviruses complex, but also because more broadly, bats
 are known reservoirs for a variety of emerging viruses and other
@@
 \hypertarget{known-betacoronavirus-hosts}{%
-\subsection{Known betacoronavirus
-hosts}\label{known-betacoronavirus-hosts}}
-
-We downloaded the data on bats hosts of betacoronaviruses assembled by
-Becker et al. (2022) from
+\subsection{\texorpdfstring{Known \emph{Betacoronavirus}
+hosts}{Known Betacoronavirus hosts}}\label{known-betacoronavirus-hosts}}
+
+We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
+by Becker et al. (2022) from
 \texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
 filtered it to ``known'' hosts (established before the emergence of
-SARS-CoV-2) and ``novel'' hosts (confirmed through sampling since the
-emergence of SARS-CoV-2). The original database was assembled by a
-combination of data mining and literature surveys, including automated
-alerts on the ``bats'' and ``coronavirus'' keywords to identify novel
-empirical evidence of bats-betacoronaviruses associations.
+SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
+competence assays since the emergence of SARS-CoV-2). The original
+database was assembled by a combination of data mining and literature
+surveys, including automated alerts on the ``bats'' and ``coronavirus''
+keywords to identify novel empirical evidence of bats-betacoronaviruses
+associations; this yielded a total of 126 known hosts, 47 of which were
+novel hosts.
 
@@
 
-We downloaded the rangemap of every extant bat species that was either
-classified as an empirically documented host of beta-coronaviruses from
-the previous step, according to recent IUCN data (IUCN 2021). The range
-maps were subsequently rasterized using the \texttt{rasterize} function
-from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
-approximately \textbf{TK TP}. For every pixel in the resulting raster
-where at least one bat host of betacoronavirus was present, we extract
-the species pool (list of all bat species), which was used to calculate
-the following risk assessment components: phylogenetic diversity, bat
-compositional uniqueness, and predicted viral sharing risk.
-
-\hypertarget{bats-phylogeography}{%
-\subsection{Bats phylogeography}\label{bats-phylogeography}}
+We downloaded the rangemap of every current bat species that was
+classified as an empirically documented host of \emph{Betacoronavirus}
+from the previous step, according to recent IUCN data (IUCN 2021). The
+range maps were subsequently rasterized using the \texttt{rasterize}
+function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
+approximately 100kmx100km. For every pixel in the resulting raster where
+at least one bat host of \emph{Betacoronavirus} was present, we extract
+the species pool (list of all competent bat hosts), which was used to
+calculate the following risk assessment components: bat phylogenetic
+diversity, bat compositional uniqueness, and predicted viral sharing
+risk.
+
+\hypertarget{bats-phylogenetic-diversity}{%
+\subsection{Bats phylogenetic
+diversity}\label{bats-phylogenetic-diversity}}
 
@@
 
-For all bat hosts of betacoronaviruses, we extracted their predicted
-viral sharing network (Albery et al. 2020). This network stores pairwise
-values of viral community similarity. To project viral sharing values
-into a single value for every pixel, we averaged the pairwise scores.
-High values of the average sharing propensity means that this specific
-extant bat assemblage is likely to be proficient at exchanging viruses.
+For all bat hosts of \emph{Betacoronavirus}, we extracted their
+predicted viral sharing network, generated from a previously published
+generalized additive mixed model of virus sharing by a tensor function
+of phylogenetic distance and geographic range overlap (Albery et al.
+2020). This network stores pairwise values of viral community
+similarity. To project viral sharing values into a single value for
+every pixel, we averaged the pairwise scores. High values of the average
+sharing propensity means that this specific extant bat assemblage is
+likely to be proficient at exchanging viruses.
 
@@
 
-We used the following query to pull all betacoronavirus sequence data
-from the GenBank Nucleotide database except SARS-CoV-2;
+To next represent phylogeography of betacoronaviruses in bats, we
+aggregated and analyzed betacoronavirus sequence data. We used the
+following query to pull all \emph{Betacoronavirus} sequence data from
+the GenBank Nucleotide database except SARS-CoV-2;
 (``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
@@
 betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
-(\textbf{Katoh and Standley 2013}, parameters in text?) and a maximum
-likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al.~2015)
-with ModelFinder (Kalyaanamoorthy et al.~2017) ultrafast bootstrap
-approximation (Hoang et al. 2018) and the following parameters
-(\textbf{STEPH WILL ADD}, parameters in text?).
+{[}Katoh and Standley (2013); Algorithm FFT-NS-2, Scoring matrix 200PAM
+/ k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
+likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al. 2015)
+with ModelFinder (Kalyaanamoorthy et al. 2017) ultrafast bootstrap
+approximation (Hoang et al. 2018) with a general time reversible model
+with empirical base frequencies and the 5-discrete-rate-category
+FreeRaye model of nucleotide substitution (GTR+F+R5).
 
@@
 entirely novel methodology. First, we measure the phylogenetic distance
-between the different viruses in the betacoronavirus tree by using the
+between the different viruses in the betacoronaviruses tree by using the
 cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
@@
 
-Bats are found worldwide and are both one of the most diverse groups
-among mammals and one of the main animal reservoir for different strains
-of betacoronaviruses (Drexler et al. 2014). This has attracted attention
-to areas where high diversity of bats, and therefore presumably high
+Bats, the second most diverse groups of mammals, are found worldwide and
+serve as the main animal reservoir for different strains of
+betacoronaviruses (Drexler et al. 2014). This has attracted attention to
+areas where high diversity of bats, and therefore presumably high
 diversity of betacoronaviruses, can be an important issue for human
 health (Calisher et al. 2006, Moratelli and Calisher 2015). By
-overlaying the IUCN rangempas for confirmed bat hosts of
+overlaying the IUCN range maps for confirmed bat hosts of
 betacoronaviruses {[}fig.~\ref{fig:richness}; top{]}, we see that the
-the main hotspots of host richness are primarily South-Eastern Asia,
-parts of Southern Europe, and to a lesser extent parts of Africa in the
--25-0 range of latitudes. The description of host richness is an
-important first step towards understanding risk, as previous research
-(Anthony et al. 2017, Mollentze and Streicker 2020) states that locally
-diverse bat communities could maintain more viruses and hence, a higher
-probability of having a pathogen that could represent a risk for human
-health.
+the main hotspots (both in terms of size and higher values) of host
+richness are primarily South-Eastern Asia, parts of Southern Europe, and
+to a lesser extent parts of Africa in the -25-0 range of latitudes. The
+description of host richness is an important first step towards
+understanding risk, as previous research (Anthony et al. 2017, Mollentze
+and Streicker 2020) states that locally diverse bat communities could
+maintain more viruses and hence, a higher probability of having a
+pathogen that could represent a risk for human health.
 
@@
 that South America has the most evolutionary distinct hosts \emph{and}
-viruses, whereas South-Eastern Asia has mostly distinct viruses. This is
-congruent with know results about New World bats being evolutionary
-distinct, and suggests that they similarly have distinct
-viruses.}\label{fig:richness}
+viruses, whereas South-Eastern Asia and the Rift Valley region have
+mostly distinct viruses. This is congruent with know results about New
+World bats being evolutionary distinct, and suggests that they similarly
+have distinct viruses.}\label{fig:richness}
 }
@@
 the creation of novel interactions (Carlson et al. 2022), and therefore
-the diversity of betacoronavirus strains should similarly be ccounted
-for. In fig.~\ref{fig:richness} (bottom), we contrast the evolutionary
-distinctiveness of bats and viruses -- this reveals a slightly different
-portrait than bat richness alone. Chiropterans can be classified, from a
-macro-evolutionary standpoint, as microchiroptera and macrochiroptera,
-where macrochiroptera have an older history from an evolutionary
-perspective compared to macrochiroptera (Teeling et al. 2005, Springer
-2013). Specifically, we would expect that the so-called ``New World''
-group of bats, being more evolutionary distinct, would also have
-evolutionary distinct viruses. Indeed fig.~\ref{fig:richness} (bottom)
-reveals it to be the case, and this region harbors a distinct
-bat-betacoronavirus complex. By contrast, South-Eastern Asia has a lot
+the diversity of \emph{Betacoronavirus} strains should similarly be
+ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
+evolutionary distinctiveness of bats and viruses -- this reveals a
+slightly different portrait than bat richness alone. Chiropterans can be
+classified, from a macro-evolutionary standpoint, as Yangochiroptera in
+the ``New World,'' and Yinpterochiroptera elsewhere (Teeling et al.
+2005, Springer 2013). Specifically, we would expect that the so-called
+``New World'' group of bats, being more evolutionary distinct, would
+also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
+(bottom) reveals it to be the case, and this region harbors a distinct
+bat-betacoronaviruses complex. By contrast, South-Eastern Asia has a lot
 of non-evolutionary distinct bats, but evolutionary-distinct viruses.
@@
 It is noteworthy that outside of South America, viral evolutionary
-distinctiveness does not accurately tracks host diversity, with some
-areas having over-distinct viruses (southern China but, oddly, not the
+distinctiveness does not accurately track host diversity, with some
+areas having over-distinct viruses (eastern China but, oddly, not the
 rest of southeast Asia). There are a number of likely explanations.
@@
 is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
-would likely result in additiona viral discoveries. This idea is
-unsurprising given the growing realization, around the emergence of
-SARS-CoV-2, that a unique lineage of similar viruses are widespread in
-bats but still mostly undescribed. The most distinct
-bats/betacoronavirus complex is found in South America, a region with a
-comparatively lower number of hosts; this matches with the isolation
+would likely result in additional viral discoveries. This idea is
+unsurprising given the growing realization, especially since the
+emergence of SARS-CoV-2, that unique lineages of similar viruses are
+widespread in bats but still mostly undescribed. The most distinct
+bats-betacoronaviruses complex is found in South America, a region with
+a comparatively lower number of hosts; this matches with the isolation
 through viariance of the host group, and may highlight a different
@@
 be a product of under-sampling: South-America is one of the places where
-the fewest betacoronaviruses have been discovered (Anthony et al. 2017,
-Olival et al. 2017, Allen et al. 2017), resulting in sparser
-phylogenetic tree, thereby artificially inflating distinctiveness.
-Adding more viruses would bring the distinctiveness of known sequences
-down.
+the fewest \emph{Betacoronavirus} sequences have been discovered
+(Anthony et al. 2017, Olival et al. 2017, Allen et al. 2017), resulting
+in sparser phylogenetic tree, thereby artificially inflating
+distinctiveness. Adding more viruses would bring the distinctiveness of
+known sequences down. Finally, South-America is the range of
+Phyllostomidae, a group of bats that underwent explosive diversification
+events (Villalobos and Arita 2010), which may drive the emergence of
+multiple viral lineages.
 
@@
 are reasons to expect that the phylogeography of bats and
-betacoronaviruses should show some degree of congruence. High density of
-hosts sharing the same virus (albeit possibly different strains) can
-drive or result from evolution of the bat antiviral immune system,
-resulting in spatially distinct immunological responses, as evidenced in
-several bat species (Banerjee et al. 2020). Immune characteristics that
-allow bats to be better adapted to infection by emerging viruses
-(Gorbunova et al. 2020, Irving et al. 2021), in addition to being
-hardcoded in their genome (Jebb et al. 2020), may be related to a wide
-variety of diets (Banerjee et al. 2020, Moreno Santillán et al. 2021,
-Muylaert et al. 2021), themselves likely to be driven by spatial
-effects, especially at the local scale -- bats, indeed, occupy a variety
-of environments, and therefore display a variety of adaptations to these
-environments.
+betacoronaviruses should show some degree of congruence (Letko et al.
+2020, Van Brussel and Holmes 2022). In particular, this should be the
+case if viruses can circulate among hosts and co-evolve with local hosts
+communities, making their evolutionary process more than a byproduct of
+host evolution. High density of hosts sharing the same virus (albeit
+possibly different strains) can drive or result from evolution of the
+bat antiviral immune system, resulting in spatially distinct
+immunological responses, as evidenced in several bat species (Banerjee
+et al. 2020). Immune characteristics that allow bats to be better
+adapted to infection by emerging viruses (Gorbunova et al. 2020, Irving
+et al. 2021), in addition to being hardcoded in their genome (Jebb et
+al. 2020), may be related to a wide variety of diets (Banerjee et al.
+2020, Moreno Santillán et al. 2021), themselves likely to be driven by
+spatial effects, especially at the local scale -- bats, indeed, occupy a
+variety of environments, and therefore display a variety of adaptations
+to these environments (Muylaert et al. 2022).
 
@@
 America. These results suggest that, although the evolutionary
-distinctiveness of the bat/betacoronavirus complex varies spatially, the
-system shows an important degree of spatial consistency, with a reduced
-number of bioregions. Available information describing the spillover of
-zoonotic betacoronaviruses of bat origin where data was available before
-and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
-in Wuhan, China; SARS-CoV in Guangdong, China based on the presence of
-closest known viruses circulating in nature, and a nearby location where
-serological (antibody) evidence has indicated human exposure to
-SARS-like viruses (Wang et al. 2018); MERS-CoV in Saudi Arabia based on
-index cases available from a recently-published compendium of cases
-(Ramshaw et al. 2019). For the latest event, most if not all index cases
-are presumed to be camel-to-human transmission, and the precise origin
-point (if it exists) of MERS-CoV in bats is uncertain. Recent
-recombinant canine coronavirus spillover events in Haiti (Lednicky et
-al. 2021) and Europe (Vlasova et al. 2022) are not relevant here, as
-bats' involvement in these cycles of transmission have been supposed to
-be non-existent. These index cases fall within different phylogeographic
-bioregions (fig.~\ref{fig:biogeo}), which further highlight the issue
-that different host-virus sub-systems may lead to widespread emergence.
+distinctiveness of the bat-betacoronaviruses complex varies spatially,
+the system shows an important degree of spatial consistency, with a
+reduced number of bioregions. Available information describing the
+spillover of zoonotic betacoronaviruses of bat origin where data was
+available before and up through the COVID-19 pandemic puts spillover
+events of SARS-CoV-2 in Wuhan, China; SARS-CoV in Guangdong, China based
+on the presence of closest known viruses circulating in nature, and a
+nearby location where serological (antibody) evidence has indicated
+human exposure to SARS-like viruses (Wang et al. 2018); MERS-CoV in
+Saudi Arabia based on index cases available from a recently-published
+compendium of cases (Ramshaw et al. 2019). For the latest event, most if
+not all index cases are presumed to be camel-to-human transmission, and
+the precise origin point (if it exists) of MERS-CoV in bats is
+uncertain. Recent recombinant canine coronavirus spillover events in
+Haiti (Lednicky et al. 2021) and Europe (Vlasova et al. 2022) are not
+relevant here, as bats' involvement in these cycles of transmission have
+been supposed to be non-existent. These index cases fall within
+different phylogeographic bioregions (fig.~\ref{fig:biogeo}), which
+further highlight the issue that different host-virus sub-systems may
+lead to widespread emergence.
 
@@
 
-As host richness, joint disctinveness, or phylogeographic structure
-suggest that the bat/betacoronavirus complex is globally fragmented
+As host richness, joint distinctiveness, or phylogeographic structure
+suggest that the bat-betacoronaviruses complex is globally fragmented
 enough to give rise to both different levels of risk (as evidenced by
@@
 these regions do not necessarilly overlap with previous spatial
-partitions of the bat/betacoronavirus complex.
+partitions of the bat-betacoronaviruses complex.
 
@@
 uniqueness in the red channel. A pixel that would maximize all measures
-would be a pure white (specifically RGB(1.0, 1.0. 1.0)), and a pixel
-with the lowest possible values would be pure black (specifically
-RGB(0.0, 0.0, 0.0)). Therefore, lighter values (the sum of the three
-channels gets closer to 3) indicate higher risk, and the color indicates
-the proportional distribution of the factors making up the total
-risk.}\label{fig:trivariate}
+(highest possible risk) would be a pure white (specifically RGB(1.0,
+1.0. 1.0)), and a pixel with the lowest possible values would be pure
+black (specifically RGB(0.0, 0.0, 0.0)). Therefore, lighter values (the
+sum of the three channels gets closer to 3) indicate higher risk, and
+the color indicates the proportional distribution of the factors making
+up the total risk.}\label{fig:trivariate}
 }
@@
 of factors is a high phylogenetic diversity of hosts with low viral
-sharing; this, essentially, means that very different betacoronavirus
+sharing; this, essentially, means that very different betacoronaviruses
 could co-exist within the same place. This is particularly the case
@@
 acquisition of viral data is rarely disconnected from the acquisition of
-host data; furthermore, there are more sources of information on hosts
-than on viruses, allowing to develop a host-centric perspective on risk.
-Areas with high bat diversity and high turnover \emph{may} facilitate
-the evolutionary radiation of viruses, matching previous findings that
-the diversification of bat coronaviruses is driven largely by host
-shifts (inter-genus or higher levels of cross-species transmission) and,
-to a lesser degree, cospeciation and sharing (intra-genus cross-species
+host data. There are more sources of information on hosts than on
+viruses, allowing to develop a host-centric perspective on risk
+(although this estimate would more accurate with viral traits related to
+\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
+high bat diversity and high turnover \emph{may} facilitate the
+evolutionary radiation of viruses, matching previous findings that the
+diversification of bat coronaviruses is driven largely by host shifts
+(inter-genus or higher levels of cross-species transmission) and, to a
+lesser degree, cospeciation and sharing (intra-genus cross-species
 transmission; Anthony et al. 2017). This diversification is not an
@@
 \includegraphics{figures/risk_map.png}
-\caption{Extraction of a measure of risk based on the colorimetric space
-from fig.~\ref{fig:trivariate}. The risk is a composite measure of the
-color value and angular distance to the yellow hue, as defined in the
-methods, ranged in the unit space. Based on this analyses, regions at
-high risk of spillover are southeast Asia and
-Madagascar.}\label{fig:risk}
+\caption{Extraction of a measure of \emph{Betacoronavirus} spillover
+risk from bat hosts based on the colorimetric space from
+fig.~\ref{fig:trivariate}. The risk is a composite measure of the color
+value and angular distance to the yellow hue, as defined in the methods,
+ranged in the unit space. Based on this analyses, regions at high risk
+of spillover are southeast Asia and Madagascar.}\label{fig:risk}
 }
@@
 From another perspective, areas of high host uniqueness and virus
-sharing (red-to-pink) could provide hotspots of betacoronavirus risk
-through mixing of unique viruses (via codivergence) and in turn
+sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
+risk through mixing of unique viruses (via codivergence) and in turn
 recombination. Under our framework, such a hotspot was identified in
@@
 (\emph{e.g.} Shi et al. 2014). Recent surveillance (Kettenburg et al.
-2022) has identified a novel betacoronavirus (in the subgenus
+2022) has identified a novel \emph{Betacoronavirus} (in the subgenus
 \emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
@@
 composite map (see Methods), to provide a single measure of risk varying
-between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
-However, this maps the potential risk, which must be weighed by the
+between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
+this map represents the potential risk, it must be weighed by the
 potential for contacts with humans. As a proxy for this measure, we used
@@
 a more actionable assessment of risk. This map is presented in
-fig.~\ref{fig:compound}. Most of South America and Europe are at low
-risk, as although densely populated, settlements tend to be in areas
-with lower potential risk. However, this mapping reveals that South-East
-Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
-high risk due to the overlap between built areas and bat communities
-representing more opportunities for cross-species transmission of
-betacoronaviruses. In looking for the origins of SARS in China, Xu et
-al. (2004) present serological evidence that strongest human-animal
-contact results in higher risk of virus exposure, regardless of the
-animal species, but that different types of contact had different
-impacts. Ideally, finer-grained information about human activity (rather
-than human presence through anthropisation) could allow to partition
-this risk further.
+fig.~\ref{fig:compound}. Most of South America and Europe are at
+comparatively lower risk, as although densely populated, settlements
+tend to be in areas with lower potential risk. Regions like Malaysia and
+the North coast of Australia have a high risk component, but should
+represent a relatively lower effective risk due to low human density.
+However, this mapping reveals that South-East Asia, the Indian
+subcontinent, and parts of sub-Saharan Africa, are at high risk due to
+the overlap between built areas and bat communities representing more
+opportunities for cross-species transmission of betacoronaviruses. In
+looking for the origins of SARS in China, Xu et al. (2004) present
+serological evidence that strongest human-animal contact results in
+higher risk of virus exposure, regardless of the animal species, but
+that different types of contact had different impacts. Ideally,
+finer-grained information about human activity (rather than human
+presence through anthropisation) could allow to partition this risk
+further, alebit at the cost of more hypotheses required to estimate the
+amount of risk represented by each activity. Our map of risk overlays
+with recent results from Cohen et al. (2022) -- areas of purported high
+risk/diversitifcation potential (Madagascar, South-America) overlay with
+sampling gaps for \emph{Betacoronavirus}, stressing the need for
+spatially targeted monitoring and discovery.
 
@@
 \caption{Overlap of the percent of each pixel occupied by urbanized
-structures, representing the degree of settlement, on the risk map.
-Darker pixels correspond to more risk, in that the GMTC-derived risk of
+structures, representing the degree of settlement, on the spillover risk
+map (where the risk comes only from wildlife, and ignores multi-hosts
+chains of transmissions including non-bats hosts). Darker pixels
+correspond to more risk, in that the GMTC-derived risk of
 fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
@@
 regions in Africa, and highlights the risk in India, southeast China,
-and the arabic peninsula where areas of high to moderate risk overlap
+and the Arabian peninsula where areas of high to moderate risk overlap
 with areas of denser population.}\label{fig:compound}
@@
 that Latin America is a hotspot of viral distinctiveness, suggesting
-that this part of the bats-betacoronavirus system may be undergoing
+that this part of the bats-betacoronaviruses complex may be undergoing
 independent evolutionary dynamics (related species sharing viruses that
@@
 (unrelated viruses coevolving with evolutionarily distinct hosts,
-generating high diversity locally). Both of these areas should be
-priority areas for sampling, especially since Becker et al. (2022)
-advance that they harbor undiscovered hosts of beta-coronaviruses. This
-diversity of hosts, and the mechanisms by which the exchange of viruses
-occurs between species, is largely affected by the local environmental
-conditions and environmental change.
+generating high diversity locally, see \emph{e.g.} Latinne et al. 2020).
+Both of these areas should be priority areas for sampling, especially
+since Becker et al. (2022) advance that they harbor undiscovered hosts
+of beta-coronaviruses. This diversity of hosts, and the mechanisms by
+which the exchange of viruses occurs between species, is largely
+affected by the local environmental conditions and environmental change.
 
@@
 know to be either obligate or facultative insectivores, therefore
-playing an important role in the regulation of insect pests that can
-affect crops (Williams-Guillén et al. 2008, Voigt and Kingston 2016),
-and vectors of pathogens that put a risk on human health (Gonsalves et
-al. 2013a, b). Because bats are globally distributed and have a long
-evolutionary history, phylogeographic and biogeographic approaches are
-required to shed light on the contemporary distribution of
-coevolutionary processes between bats and the pathogens they host. Not
-all areas in which bats, viruses, and human are co-occuring are facing a
-risk of spillover towards human populations, and the areas in which this
-risk exist may not be facing risks of the same nature and magnitude.
+actively contributing for agricultural pest control (Williams-Guillén et
+al. 2008, Voigt and Kingston 2016), and vectors of pathogens that put a
+risk on human health (Gonsalves et al. 2013a, b). Because bats are
+globally distributed and have a long evolutionary history,
+phylogeographic and biogeographic approaches are required to shed light
+on the contemporary distribution of coevolutionary processes between
+bats and the pathogens they host. Not all areas in which bats, viruses,
+and human are co-occuring are facing a risk of spillover towards human
+populations, and the areas in which this risk exist may not be facing
+risks of the same nature and magnitude.
 
@@
 
+\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
+Cohen, L. E. et al. 2022. Sampling strategies and pre-pandemic
+surveillance gaps for bat coronaviruses.: 2022.06.15.496296.
+
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
@@
 
+\leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
+Hoang, D. T. et al. 2018. UFBoot2: Improving the Ultrafast Bootstrap
+Approximation. - Molecular Biology and Evolution 35: 518--522.
+
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
@@
 
+\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
+Kalyaanamoorthy, S. et al. 2017. ModelFinder: Fast model selection for
+accurate phylogenetic estimates. - Nature Methods 14: 587--589.
+
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
@@
 of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.
+
+\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
+Katoh, K. and Standley, D. M. 2013. MAFFT Multiple Sequence Alignment
+Software Version 7: Improvements in Performance and Usability. -
+Molecular Biology and Evolution 30: 772--780.
 
@@
 14830.
+
+\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
+Latinne, A. et al. 2020. Origin and cross-species transmission of bat
+coronaviruses in China. - bioRxiv: The Preprint Server for Biology:
+2020.05.31.116061.
 
@@
 
-\leavevmode\hypertarget{ref-Muylaert2021PreFut}{}%
-Muylaert, R. L. et al. 2021. Present and future distribution of bat
+\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
+Muylaert, R. L. et al. 2022. Present and future distribution of bat
 hosts of sarbecoviruses: Implications for conservation and public
-health.
+health. - Proceedings of the Royal Society B: Biological Sciences 289:
+20220397.
+
+\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
+Nguyen, L.-T. et al. 2015. IQ-TREE: A Fast and Effective Stochastic
+Algorithm for Estimating Maximum-Likelihood Phylogenies. - Molecular
+Biology and Evolution 32: 268--274.
 
@@
 
+\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
+Villalobos, F. and Arita, H. T. 2010. The diversity field of New World
+leaf-nosed bats (Phyllostomidae). - Global Ecology and Biogeography 19:
+200--211.
+
 \leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%

### 2022-06-17 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@fa58cdffbc410db83637c21cff76d58ca3bd6ac3 🚀

@@
 \noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
-bat-betacoronaviruses spillover risk}}}
+\emph{Betacoronavirus} spillover risk from bat hosts}}}
 

### 2022-06-17 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@7018c2fbc6439ce51f68b1d0eeb77170b87fd7c8 🚀

@@
 %
-\,\textsuperscript{5}\quad %
+\,\textsuperscript{6,7,8}\quad %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
@@
 Sciences\quad \textsuperscript{3}\,Molecular Epidemiology and Public
-Health Laboratory, Hopkirk Research Institute, Massey University,
-Palmerston North, New Zealand\quad \textsuperscript{4}\,Paul G. Allen
-School for Global Health, Washington State University, Pullman, WA,
-United States\quad \textsuperscript{5}\,???
+Health Laboratory, School of Veterinary Science, Massey University, New
+Zealand\quad \textsuperscript{4}\,Paul G. Allen School for Global
+Health, Washington State University, Pullman, WA, United
+States\quad \textsuperscript{5}\,Department of Biology, Georgetown
+University, Washington, DC, USA\quad \textsuperscript{6}\,Department of
+Biology, Georgetown University, Washington,
+DC,\quad \textsuperscript{7}\,Center for Global Health Science and
+Security, Georgetown University Medical Center, Washington, DC,
+USA\quad \textsuperscript{8}\,Department of Microbiology and Immunology,
+Georgetown University Medical Center, Washington, DC, USA
 

### 2022-06-17 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@7018c2fbc6439ce51f68b1d0eeb77170b87fd7c8 🚀

@@
 %
-\,\textsuperscript{5}, %
+\,\textsuperscript{6,7,8}, %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
@@
 Centre for Biodiversity Sciences; \textsuperscript{3}\,Molecular
-Epidemiology and Public Health Laboratory, Hopkirk Research Institute,
-Massey University, Palmerston North, New
-Zealand; \textsuperscript{4}\,Paul G. Allen School for Global Health,
-Washington State University, Pullman, WA, United
-States; \textsuperscript{5}\,???\\
+Epidemiology and Public Health Laboratory, School of Veterinary Science,
+Massey University, New Zealand; \textsuperscript{4}\,Paul G. Allen
+School for Global Health, Washington State University, Pullman, WA,
+United States; \textsuperscript{5}\,Department of Biology, Georgetown
+University, Washington, DC, USA; \textsuperscript{6}\,Department of
+Biology, Georgetown University, Washington,
+DC,; \textsuperscript{7}\,Center for Global Health Science and Security,
+Georgetown University Medical Center, Washington, DC,
+USA; \textsuperscript{8}\,Department of Microbiology and Immunology,
+Georgetown University Medical Center, Washington, DC, USA\\
 \textsuperscript{‡}\,These authors contributed equally to the work\\

### 2022-06-18 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@47e9da202a84e537f493b7741affa594f76d6fa6 🚀

@@
 %
-\,\textsuperscript{1,2}\quad %
+\,\textsuperscript{1,2,‡}\quad %
 Renata L.\,Muylaert%
@@
 \,\textsuperscript{5}\quad %
+Daniel J.\,Becker%
+%
+\,\textsuperscript{6}\quad %
 Colin J.\,Carlson%
 %
-\,\textsuperscript{6,7,8}\quad %
+\,\textsuperscript{7,8,9,‡}\quad %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
@@
 University, Washington, DC, USA\quad \textsuperscript{6}\,Department of
-Biology, Georgetown University, Washington,
-DC,\quad \textsuperscript{7}\,Center for Global Health Science and
-Security, Georgetown University Medical Center, Washington, DC,
-USA\quad \textsuperscript{8}\,Department of Microbiology and Immunology,
-Georgetown University Medical Center, Washington, DC, USA
+Biology, University of Oklahoma, Norman, OK,
+USA\quad \textsuperscript{7}\,Department of Biology, Georgetown
+University, Washington, DC,\quad \textsuperscript{8}\,Center for Global
+Health Science and Security, Georgetown University Medical Center,
+Washington, DC, USA\quad \textsuperscript{9}\,Department of Microbiology
+and Immunology, Georgetown University Medical Center, Washington, DC,
+USA
 
@@
 Driven by the need to understand the ecological factors involved in the
-emergence of betacoronavirus (the genus causing the SARS and MERS
-disease in human) through bat hosts, we develop an approach to the
+emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
+and COVID-19 in human) through bat hosts, we develop an approach to the
 assesment of spillover risk based on the Geographic Mosaic Theory of
 Coevolution. In doing so, we provide a global mapping of the spillover
-risk posed by betacoronaviruses, reflecting the fact that this risk has
-many ecological and evolutionary origins. Our framework reveals that
-reservoir richness alone, although a component of viral hazard, is not a
-sufficiently integrative predictor of risk. We offer alternative
-insights based on viral sharing, host compositional uniqueness, and host
-phylogenetic diversity. By comparing our aggregated measure of risk to a
-proxy for human density, namely the proportion of each pixel that is
-covered by urban or built land, we provide a synthetic risk map,
-allowing the identification of hotspots where the bat-betacoronavirus
-system may originate novel viruses in humans.
+risk posed by betacoronaviruses, reflecting the fact that this risk is
+best understood through the multi-faceted prism of ecological and
+evolutionary mechanisms. Our framework reveals that reservoir richness
+alone, although a component of viral hazard, is not a sufficiently
+integrative predictor of risk. We offer alternative insights based on
+viral sharing, host compositional uniqueness, and host phylogenetic
+diversity and phylogeographic regions. By comparing our aggregated
+measure of risk to a proxy for human density, namely the proportion of
+each pixel that is covered by urban or built land, we provide a
+synthetic risk map, allowing the identification of hotspots where the
+bat-betacoronavirus system may originate novel viruses in humans.
 
@@
 competent hosts, it is driven by a multiplicity of factors (Plowright et
-al. 2017). Although proxies for the local richness of hosts is commonly
-analysed (see \emph{e.g.} Anthony et al. 2017 for coronaviruses), there
-is an argument to be made that species who are not (known as) competent
-hosts of a specific virus genus may not factor into this (Plowright et
-al. 2015), calling for species-level information. This is especially
-true as competence data increases predictive power when the taxonomic
-scope of hosts of a viral family increases (Becker et al. 2020, Mull et
-al. 2022). Similarly, host species who share viruses at different rates
+al. 2017). Although proxies for the local richness of hosts are commonly
+analysed (Anthony et al. 2017, Ruiz-Aravena et al. 2022), there is an
+argument to be made that species who are not (known as) competent hosts
+of a specific virus genus may not factor into this (Plowright et al.
+2015), calling for species-level information. This is especially true as
+competence data increases predictive power when the taxonomic scope of
+hosts of a viral family increases (Becker et al. 2020, Mull et al.
+2022). Similarly, host species who share viruses at different rates
 should be weighted accordingly (Albery et al. 2020). In mammals, key
 functional traits (for which phylogeny is a reasonable proxy) are
-determinants of the spillover potential (Olival et al. 2017); these
-include, notably, body mass, and affinity for urban environments (Albery
-et al. 2022). Finally, especially when the pool of potential hosts spans
-the entire globe, there may be local host pools that are highly unique;
-not having been observed in other locations, these can act on the
-overall risk either by providing novel contact opportunities, reflecting
-unique host-environment combinations (Engering et al. 2013), or
-facilitating rapid evolutionary changes in specialism of their pathogens
-(Agosta et al. 2010). In the specific case of generalist pathogens
-(which is the case many viruses in the \emph{Betacoronavirus} genus, see
-\emph{e.g.} MacLean et al. 2021), there is conceptual and empirical
-support to the idea that these community-level mechanisms are even more
-important in driving the overall risk (Power and Mitchell 2004).
-
-In this paper, we examine the biogeographic structure of
-bats-betacoronaviruses associations, based on a curated dataset of all
-confirmed bat hosts of betacoronaviruses. By drawing on concepts from
-the Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005), we
-turn these associations into a spatially explicit additive mapping of
-zoonotic risk components, revealing the extreme heterogeneity of risk at
-the global scale. Explicitely framing the notion of spillover risk based
-on propositions from the GMTC (which is to say, based on a framework
-linking interactions between species to change within species) is a
-novel idea, that should be relatively general. Indeed, it only assumes
-the action of well described evolutionary mechanisms. The benefit of
-this approach is to provide the potential for a more dynamic and nuanced
-understanding of risk: not only on ecological timescales, but also by
-providing clues about which areas can change over micro-evolutionary
-timescales. This provides a way to look at spatial structure by
-accounting for more notions than species richness/similarity, but also a
-way to identify spatial areas of higher risk.
-
-We identify the Amazon and South-Eastern Asia as hotspots where the
-phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
-(Anthony et al. 2017); surprisingly, current data suggest that viral
-sharing between hosts is high in the Amazon and low in South-Eastern
-Asia, which has the potential to result in different evolutionary
-dynamics between these two regions, hinting at different futures for
-their viral communities. This work is important both as a description of
-the bats-betacoronaviruses complex, but also because more broadly, bats
-are known reservoirs for a variety of emerging viruses and other
-pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
-the needs for bat conservation and disease prevention most likely very
-difficult and a source of human-wildlife conflicts, especially in more
-densely populated areas (Stone et al. 2015, Rego et al. 2015).
+determinants of the spillover potential of their viruses (Olival et al.
+2017); these include, notably, body mass, life history, diet, and use
+for human-occupied spaces (Albery et al. 2022). Finally, especially when
+the pool of potential hosts spans the entire globe, there may be local
+host pools that are highly unique; not having been observed in other
+locations, these can act on the overall risk either by providing novel
+contact opportunities, reflecting unique host-environment combinations
+(Engering et al. 2013), or facilitating rapid evolutionary changes in
+specialism of their pathogens (Agosta et al. 2010). In the specific case
+of generalist pathogens (which is the case for many viruses in the
+\emph{Betacoronavirus} genus, see \emph{e.g.} MacLean et al. 2021),
+there is conceptual and empirical support to the idea that these
+community-level mechanisms are even more important in driving the
+overall risk (Power and Mitchell 2004). In this paper, we examine the
+biogeographic structure of bats-betacoronaviruses associations, based on
+a curated dataset of all confirmed bat hosts of betacoronaviruses, by
+adopting an analysis framework grounded in concepts from the Geographic
+Mosaic Theory of Coevolution (GMTC; Thompson 2005). The GMTC posits that
+coevolutionary processes between pairs (Thompson 1994) or complexes
+(Janzen 1980) of species are structured in space by the rippling effects
+of abiotic conditions onto evolutionary mechansism, resulting in
+spatially fragmented evolutionary dynamics, coupled only by
+dispersal-related processes (Gomulkiewicz et al. 2000). In turn, these
+spatially fragmented processes can lead taxonomically homogeneous
+systems (in our case, the bats-betacoronaviruses complex) to have
+different structure and dynamics of large spatial extents (see
+\emph{e.g.} Price 2002).
+
+The GMTC predicts a spatial fragmentation of coevolutionary dynamics
+under the joint action of three processes (see notably Gomulkiewicz et
+al. 2007), which all have the potential to act on outbreak potential,
+pathogen tramission, and disease virulence (Parratt et al. 2016, Turner
+et al. 2021). First hot/cold spots of coevolution can appear when the
+intensity of \emph{interaction} (in terms of reciprocal fitness
+consequences) varies spatially, because of \emph{e.g.} partial range
+overlap between organisms (Nuismer et al. 2003); hot/cold spots are
+notoriously difficult to identify in nature (Laine 2005), especially
+when the structure of range overlap is complex, and the reciprocal
+fitness effets are relatively low. Both of these conditions are met in
+the bats-coronaviruses complex. Second, GMTC supposes the existence of
+selection mosaics, wherein the intensity of \emph{selection} (in terms
+of selective consequences of fitness effects) varies across space; the
+strength of reciprocal selection responds to the biotic complexity of
+the community (locally diverse hosts and viruses are more biotically
+complex; Thrall et al. 2007) \emph{and} to the local favorability of the
+environment (Hochberg and Baalen 1998). In a system with near global
+distribution and variations in relative richness, like the
+bats-betacoronaviruses complex, the GMTC predicts the establishment of
+many different coevolutionary trajectories. The final GMCT process is
+trait remixing, under which changes in coevolutionary dynamics happen
+due to the arrival (or departure) of functional traits, through changes
+in community composition due to invasions, meta-community dynamics, and
+disperal; adopting an host-centric perspective, high viral sharing would
+expose the host immune system to multiple new viral traits, thereby
+changing the selection landscape for all viruses.
+
+We turn the processes on the GMTC into definitions of spillover risk
+from viruses to hosts (focusing on the bats-betacoronavirus complex),
+with a specific emphasis on the potential to create independant
+coevolutionary dynamics (and therefore spatial fragmentation in the
+risk) through heterogeneity. Our components of risk are (i) the
+phylogenetic diversity of hosts, representing the diversification
+potential of the system (hotspots, selection mosaics), under the
+assumption that bat diversification is not primarily driven by viruses
+(Gorbunova et al. 2020); (ii) the local uniqueness of the bat community,
+representing the potential for viruses to be exposed to novel host
+traits in unusual combinations (cold spots, trait remixing); finally,
+(iii) viral sharing among hosts, representing the potential for hosts to
+be exposed to complex and functionaly diverse communities (trait
+remixing, biotic complexity). We turn these predicates into a spatially
+explicit additive mapping of zoonotic risk components, revealing the
+extreme heterogeneity of risk at the global scale. Explicitely framing
+the notion of spillover risk based on propositions from the GMTC (which
+is to say, based on a framework linking interactions between species to
+change within species) is a novel idea, that should be relatively
+general. Indeed, it only assumes the action of well described
+evolutionary mechanisms. The benefit of this approach is to provide the
+potential for a more dynamic and nuanced understanding of risk: not only
+on ecological timescales, but also by providing clues about which areas
+can change over micro-evolutionary timescales. This provides a way to
+look at spatial structure by accounting for more notions than species
+richness/similarity, but also a way to identify spatial areas of higher
+risk.
 
@@
 SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
-competence assays since the emergence of SARS-CoV-2). The original
+competence assays since the initial data collection). The original
 database was assembled by a combination of data mining and literature
@@
 slightly different portrait than bat richness alone. Chiropterans can be
-classified, from a macro-evolutionary standpoint, as Yangochiroptera in
-the ``New World,'' and Yinpterochiroptera elsewhere (Teeling et al.
-2005, Springer 2013). Specifically, we would expect that the so-called
-``New World'' group of bats, being more evolutionary distinct, would
-also have evolutionary distinct viruses. Indeed fig.~\ref{fig:richness}
-(bottom) reveals it to be the case, and this region harbors a distinct
-bat-betacoronaviruses complex. By contrast, South-Eastern Asia has a lot
-of non-evolutionary distinct bats, but evolutionary-distinct viruses.
+classified, from a macro-evolutionary standpoint, as Yangochiroptera and
+Yinpterochiroptera elsewhere (Teeling et al. 2005, Springer 2013).
+Specifically, we would expect that the so-called ``New World'' group of
+bats, being more evolutionary distinct, would also have evolutionary
+distinct viruses. Indeed fig.~\ref{fig:richness} (bottom) reveals it to
+be the case, and this region harbors a distinct bat-betacoronaviruses
+complex. This can be explained by the fact that Yangochiroptera,
+although not limited to the western hemisphere, contain the highly
+diverse adaptive radiation in the Phyllostomidae (Villalobos and Arita
+2010), which is restricted to the western hemisphere. By contrast,
+South-Eastern Asia has a lot of non-evolutionary distinct bats, who
+nevertheless hosted evolutionary-distinct viruses.
 
@@
 
-Our study focuses largely on the biogeography of hosts. Yet, we know
-that viruses with high host plasticity, that is, the ability of a given
-virus to adapt to various taxonomic orders and ecological groups
-(Kreuder Johnson et al. 2015), are more likely to amplify viral
-spillover, followed by secondary human-to-human transmission, and
-geographical spread (Hazarie et al. 2021). High viral host plasticity is
-an especially important trait for RNA viruses like betacoronaviruses
-(Haddad et al. 2021). Indeed, our analysis of viral sequences reveals
-that Latin America is a hotspot of viral distinctiveness, suggesting
-that this part of the bats-betacoronaviruses complex may be undergoing
-independent evolutionary dynamics (related species sharing viruses that
-are different from the rest of the global pool). The other hotspot of
-viral distinctiveness is S.E. Asia, in which richness is high but
-sharing is low; this suggests a different type of evolutionary dynamics
-(unrelated viruses coevolving with evolutionarily distinct hosts,
-generating high diversity locally, see \emph{e.g.} Latinne et al. 2020).
-Both of these areas should be priority areas for sampling, especially
-since Becker et al. (2022) advance that they harbor undiscovered hosts
-of beta-coronaviruses. This diversity of hosts, and the mechanisms by
-which the exchange of viruses occurs between species, is largely
-affected by the local environmental conditions and environmental change.
+Our study focuses largely on the biogeography of hosts. Specifically, we
+identify the Amazon and South-Eastern Asia as hotspots where the
+phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
+(Anthony et al. 2017); surprisingly, current data suggest that viral
+sharing between hosts is high in the Amazon and low in South-Eastern
+Asia, which has the potential to result in different evolutionary
+dynamics between these two regions, hinting at different futures for
+their viral communities. This work is important both as a description of
+the bats-betacoronaviruses complex, but also because more broadly, bats
+are known reservoirs for a variety of emerging viruses and other
+pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
+the needs for bat conservation and disease prevention most likely very
+difficult and a source of human-wildlife conflicts, especially in more
+densely populated areas (Stone et al. 2015, Rego et al. 2015).
+
+Yet, we know that viruses with high host plasticity, that is, the
+ability of a given virus to adapt to various taxonomic orders and
+ecological groups (Kreuder Johnson et al. 2015), are more likely to
+amplify viral spillover, followed by secondary human-to-human
+transmission, and geographical spread (Hazarie et al. 2021). High viral
+host plasticity is an especially important trait for RNA viruses like
+betacoronaviruses (Haddad et al. 2021). Indeed, our analysis of viral
+sequences reveals that Latin America is a hotspot of viral
+distinctiveness, suggesting that this part of the bats-betacoronaviruses
+complex may be undergoing independent evolutionary dynamics (related
+species sharing viruses that are different from the rest of the global
+pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
+richness is high but sharing is low; this suggests a different type of
+evolutionary dynamics (unrelated viruses coevolving with evolutionarily
+distinct hosts, generating high diversity locally, see \emph{e.g.}
+Latinne et al. 2020). Both of these areas should be priority areas for
+sampling, especially since Becker et al. (2022) advance that they harbor
+undiscovered hosts of beta-coronaviruses. This diversity of hosts, and
+the mechanisms by which the exchange of viruses occurs between species,
+is largely affected by the local environmental conditions and
+environmental change.
 
@@
 Albery, G. F. et al. 2022. Urban-adapted mammal species have more known
-pathogens. in press.
+pathogens. - Nature Ecology \& Evolution 6: 794--801.
 
@@
 
+\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
+Gomulkiewicz, R. et al. 2007. Dos and don'ts of testing the geographic
+mosaic theory of coevolution. - Heredity 98: 249--258.
+
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
@@
 
+\leavevmode\hypertarget{ref-Hochberg1998AntCoe}{}%
+Hochberg, M. E. and Baalen, M. 1998. Antagonistic coevolution over
+productivity gradients. - The American Naturalist 152: 620--634.
+
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
@@
 IUCN 2021. The IUCN Red List of Threatened Species.
+
+\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
+Janzen, D. H. 1980. When is it Coevolution? - Evolution 34: 611--612.
 
@@
 14830.
+
+\leavevmode\hypertarget{ref-Laine2005SpaSca}{}%
+Laine, A.-L. 2005. Spatial scale of local adaptation in a plant-pathogen
+metapopulation. - Journal of Evolutionary Biology 18: 930--938.
 
@@
 
+\leavevmode\hypertarget{ref-Nuismer2003CoeHos}{}%
+Nuismer, S. L. et al. 2003. Coevolution between hosts and parasites with
+partially overlapping geographic ranges. - Journal of Evolutionary
+Biology 16: 1337--1345.
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
@@
 
+\leavevmode\hypertarget{ref-Parratt2016InfDis}{}%
+Parratt, S. R. et al. 2016. Infectious Disease Dynamics in Heterogeneous
+Landscapes. - Annual Review of Ecology, Evolution, and Systematics 47:
+283--306.
+
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
@@
 
+\leavevmode\hypertarget{ref-Price2002MacThe}{}%
+Price, P. W. 2002. Macroevolutionary Theory on Macroecological Patterns.
+- Cambridge University Press.
+
 \leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
@@
 
+\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
+Ruiz-Aravena, M. et al. 2022. Ecology, evolution and spillover of
+coronaviruses from bats. - Nature Reviews Microbiology 20: 299--314.
+
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
@@
 
+\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
+Thompson, J. N. 1994. The Coevolutionary Process. - University of
+Chicago Press.
+
 \leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
@@
 Of Chicago Press.
+
+\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
+Thrall, P. H. et al. 2007. Coevolution of symbiotic mutualists and
+parasites in a community context. - Trends in Ecology \& Evolution 22:
+120--126.
+
+\leavevmode\hypertarget{ref-Turner2021RolEnv}{}%
+Turner, W. C. et al. 2021. The roles of environmental variation and
+parasite survival in virulence--transmission relationships. - Royal
+Society open science 8: 210088.
 

### 2022-06-18 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@47e9da202a84e537f493b7741affa594f76d6fa6 🚀

@@
 %
-\,\textsuperscript{1,2}, %
+\,\textsuperscript{1,2,‡}, %
 Renata L.\,Muylaert%
@@
 \,\textsuperscript{5}, %
+Daniel J.\,Becker%
+%
+\,\textsuperscript{6}, %
 Colin J.\,Carlson%
 %
-\,\textsuperscript{6,7,8}, %
+\,\textsuperscript{7,8,9,‡}, %
 \href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
@@
 University, Washington, DC, USA; \textsuperscript{6}\,Department of
-Biology, Georgetown University, Washington,
-DC,; \textsuperscript{7}\,Center for Global Health Science and Security,
-Georgetown University Medical Center, Washington, DC,
-USA; \textsuperscript{8}\,Department of Microbiology and Immunology,
+Biology, University of Oklahoma, Norman, OK,
+USA; \textsuperscript{7}\,Department of Biology, Georgetown University,
+Washington, DC,; \textsuperscript{8}\,Center for Global Health Science
+and Security, Georgetown University Medical Center, Washington, DC,
+USA; \textsuperscript{9}\,Department of Microbiology and Immunology,
 Georgetown University Medical Center, Washington, DC, USA\\

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@63a51b8ed8b53a85690ceaf73d894e2b7187b884 🚀

@@
 
-Spillover risk is multidimensional and complicated. Within a pool of
-competent hosts, it is driven by a multiplicity of factors (Plowright et
-al. 2017). Although proxies for the local richness of hosts are commonly
-analysed (Anthony et al. 2017, Ruiz-Aravena et al. 2022), there is an
-argument to be made that species who are not (known as) competent hosts
-of a specific virus genus may not factor into this (Plowright et al.
-2015), calling for species-level information. This is especially true as
-competence data increases predictive power when the taxonomic scope of
-hosts of a viral family increases (Becker et al. 2020, Mull et al.
-2022). Similarly, host species who share viruses at different rates
-should be weighted accordingly (Albery et al. 2020). In mammals, key
-functional traits (for which phylogeny is a reasonable proxy) are
-determinants of the spillover potential of their viruses (Olival et al.
-2017); these include, notably, body mass, life history, diet, and use
-for human-occupied spaces (Albery et al. 2022). Finally, especially when
-the pool of potential hosts spans the entire globe, there may be local
-host pools that are highly unique; not having been observed in other
-locations, these can act on the overall risk either by providing novel
-contact opportunities, reflecting unique host-environment combinations
-(Engering et al. 2013), or facilitating rapid evolutionary changes in
-specialism of their pathogens (Agosta et al. 2010). In the specific case
-of generalist pathogens (which is the case for many viruses in the
-\emph{Betacoronavirus} genus, see \emph{e.g.} MacLean et al. 2021),
-there is conceptual and empirical support to the idea that these
+Spillover risk is complex and, within a pool of susceptible wildlife
+hosts, is driven by a multiplicity of factors (Plowright et al. 2017).
+Although proxies for the local richness of hosts are commonly analysed
+(Anthony et al. 2017, Ruiz-Aravena et al. 2022), these approaches
+deliberately oversimplify interspecific heterogeneity in immunity,
+behavior, and contact with humans. At the coarsest level, some species
+play a more central role in the host-virus network because of key
+functional traits (for which phylogeny is a reasonable proxy) (Albery et
+al. 2020); these include, notably, body mass, life history, diet, and
+use for human-occupied spaces Olival et al. (2017). Global maps of
+spillover risk often struggle to distill these features into
+interpretable risk maps, and overlook highly-unique host pools that
+allow for the rapid evolution of highly divergent viruses (Agosta et al.
+2010). In the case of generalist pathogens like betacoronaviruses, there
+is conceptual and empirical support to the idea that these
 community-level mechanisms are even more important in driving the
 overall risk (Power and Mitchell 2004). In this paper, we examine the
-biogeographic structure of bats-betacoronaviruses associations, based on
-a curated dataset of all confirmed bat hosts of betacoronaviruses, by
-adopting an analysis framework grounded in concepts from the Geographic
-Mosaic Theory of Coevolution (GMTC; Thompson 2005). The GMTC posits that
+biogeographic structure of bat-betacoronavirus associations by adopting
+an analysis framework grounded in concepts from the Geographic Mosaic
+Theory of Coevolution (GMTC; Thompson 2005). The GMTC posits that
 coevolutionary processes between pairs (Thompson 1994) or complexes
 (Janzen 1980) of species are structured in space by the rippling effects
-of abiotic conditions onto evolutionary mechansism, resulting in
+of abiotic conditions onto evolutionary mechanism, resulting in
 spatially fragmented evolutionary dynamics, coupled only by
@@
 spatially fragmented processes can lead taxonomically homogeneous
-systems (in our case, the bats-betacoronaviruses complex) to have
-different structure and dynamics of large spatial extents (see
-\emph{e.g.} Price 2002).
+systems to have different structure and dynamics over large spatial
+extents (see \emph{e.g.} Price 2002).
 
@@
 pathogen tramission, and disease virulence (Parratt et al. 2016, Turner
-et al. 2021). First hot/cold spots of coevolution can appear when the
-intensity of \emph{interaction} (in terms of reciprocal fitness
+et al. 2021). First hot- and coldspots of coevolution can appear when
+the intensity of \emph{interaction} (in terms of reciprocal fitness
 consequences) varies spatially, because of \emph{e.g.} partial range
-overlap between organisms (Nuismer et al. 2003); hot/cold spots are
+overlap between organisms (Nuismer et al. 2003). Hot- and coldspots are
 notoriously difficult to identify in nature (Laine 2005), especially
 when the structure of range overlap is complex, and the reciprocal
-fitness effets are relatively low. Both of these conditions are met in
-the bats-coronaviruses complex. Second, GMTC supposes the existence of
-selection mosaics, wherein the intensity of \emph{selection} (in terms
-of selective consequences of fitness effects) varies across space; the
-strength of reciprocal selection responds to the biotic complexity of
-the community (locally diverse hosts and viruses are more biotically
-complex; Thrall et al. 2007) \emph{and} to the local favorability of the
-environment (Hochberg and Baalen 1998). In a system with near global
-distribution and variations in relative richness, like the
-bats-betacoronaviruses complex, the GMTC predicts the establishment of
-many different coevolutionary trajectories. The final GMCT process is
-trait remixing, under which changes in coevolutionary dynamics happen
-due to the arrival (or departure) of functional traits, through changes
-in community composition due to invasions, meta-community dynamics, and
-disperal; adopting an host-centric perspective, high viral sharing would
-expose the host immune system to multiple new viral traits, thereby
-changing the selection landscape for all viruses.
+fitness effects are relatively low---both of which conditions are met in
+the bat-betacoronavirus complex. Second, the GMTC supposes the existence
+of selection mosaics, wherein the intensity of \emph{selection} varies
+across space; the strength of reciprocal selection responds to the
+biotic complexity of the community (locally diverse hosts and viruses
+are more biotically complex; Thrall et al. 2007) \emph{and} to the local
+favorability of the environment (Hochberg and Baalen 1998). In a system
+with near global distribution and variations in relative richness, like
+the bat-betacoronavirus complex, the GMTC predicts the establishment of
+many different coevolutionary trajectories (and resulting viral
+regimes). The final GMCT process is trait remixing, under which changes
+in coevolutionary dynamics happen due to the arrival (or departure) of
+functional traits, through changes in community composition due to
+invasions, meta-community dynamics, and disperal; these changes are
+likely to happen in overlap zones between ``long branches'' of
+coevolution.
 
@@
 assumption that bat diversification is not primarily driven by viruses
-(Gorbunova et al. 2020); (ii) the local uniqueness of the bat community,
+(Gorbunova et al. 2020); (ii) viral sharing among hosts, representing
+levels of viral exchange within host communities that may act as a
+buffer against viral specialization in any given host (i.e., high
+sharing creates coldspots of selective pressures from any given host
+species); and (iii) the local uniqueness of the bat community,
 representing the potential for viruses to be exposed to novel host
-traits in unusual combinations (cold spots, trait remixing); finally,
-(iii) viral sharing among hosts, representing the potential for hosts to
-be exposed to complex and functionaly diverse communities (trait
-remixing, biotic complexity). We turn these predicates into a spatially
-explicit additive mapping of zoonotic risk components, revealing the
-extreme heterogeneity of risk at the global scale. Explicitely framing
-the notion of spillover risk based on propositions from the GMTC (which
-is to say, based on a framework linking interactions between species to
-change within species) is a novel idea, that should be relatively
-general. Indeed, it only assumes the action of well described
-evolutionary mechanisms. The benefit of this approach is to provide the
-potential for a more dynamic and nuanced understanding of risk: not only
-on ecological timescales, but also by providing clues about which areas
-can change over micro-evolutionary timescales. This provides a way to
-look at spatial structure by accounting for more notions than species
-richness/similarity, but also a way to identify spatial areas of higher
-risk.
+traits in unusual combinations (trait remixing). We turn these features
+into a spatially-explicit additive mapping of zoonotic risk components,
+revealing the extreme heterogeneity of risk at the global scale.
+Explicitely framing the notion of spillover risk based on propositions
+from the GMTC (which is to say, based on a framework linking
+interactions between species to change within species) is a novel idea,
+that should be relatively general. Indeed, it only assumes the action of
+well described evolutionary mechanisms. The benefit of this approach is
+to provide the potential for a more dynamic and nuanced understanding of
+risk: not only on ecological timescales, but also by providing clues
+about which areas can change over micro-evolutionary timescales. This
+provides a way to look at spatial structure by accounting for more
+notions than species richness/similarity, but also a way to identify
+spatial areas of higher risk.
 
@@
 
-\leavevmode\hypertarget{ref-Becker2020InfInt}{}%
-Becker, D. J. et al. 2020. Beyond Infection: Integrating Competence into
-Reservoir Host Prediction. - Trends in Ecology \& Evolution 35:
-1062--1065.
-
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
@@
 45--56.
-
-\leavevmode\hypertarget{ref-Engering2013PatHos}{}%
-Engering, A. et al. 2013. Pathogen--host--environment interplay and
-disease emergence. - Emerging Microbes \& Infections 2: e5.
 
@@
 
-\leavevmode\hypertarget{ref-MacLean2021NatSel}{}%
-MacLean, O. A. et al. 2021. Natural selection in the evolution of
-SARS-CoV-2 in bats created a generalist virus and highly capable human
-pathogen. - PLOS Biology 19: e3001115.
-
 \leavevmode\hypertarget{ref-Melaun2014BatPot}{}%
@@
 
-\leavevmode\hypertarget{ref-Mull2022VirIso}{}%
-Mull, N. et al. 2022. Virus isolation data improve host predictions for
-New World rodent orthohantaviruses. - Journal of Animal Ecology in
-press.
-
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
@@
 
-\leavevmode\hypertarget{ref-Plowright2015EcoDyn}{}%
-Plowright, R. K. et al. 2015. Ecological dynamics of emerging bat virus
-spillover. - Proceedings of the Royal Society B: Biological Sciences
-282: 20142124.
-
 \leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@bdd862abac2f96426c34cd878e05be6d7808831f 🚀

@@
 
-Spillover risk is complex and, within a pool of susceptible wildlife
-hosts, is driven by a multiplicity of factors (Plowright et al. 2017).
-Although proxies for the local richness of hosts are commonly analysed
-(Anthony et al. 2017, Ruiz-Aravena et al. 2022), these approaches
-deliberately oversimplify interspecific heterogeneity in immunity,
-behavior, and contact with humans. At the coarsest level, some species
-play a more central role in the host-virus network because of key
-functional traits (for which phylogeny is a reasonable proxy) (Albery et
-al. 2020); these include, notably, body mass, life history, diet, and
-use for human-occupied spaces Olival et al. (2017). Global maps of
+Spillover risk is complex and, even within a pool of susceptible
+wildlife hosts, driven by a multiplicity of factors (Plowright et al.
+2017). Although host richness is often used as a coarse proxy for
+hotspots of emergence risk (Anthony et al. 2017, Ruiz-Aravena et al.
+2022), these approaches deliberately oversimplify interspecific
+heterogeneity in immunology, behavior, and other traits. Global maps of
 spillover risk often struggle to distill these features into
@@
 is conceptual and empirical support to the idea that these
-community-level mechanisms are even more important in driving the
-overall risk (Power and Mitchell 2004). In this paper, we examine the
-biogeographic structure of bat-betacoronavirus associations by adopting
-an analysis framework grounded in concepts from the Geographic Mosaic
-Theory of Coevolution (GMTC; Thompson 2005). The GMTC posits that
-coevolutionary processes between pairs (Thompson 1994) or complexes
-(Janzen 1980) of species are structured in space by the rippling effects
-of abiotic conditions onto evolutionary mechanism, resulting in
-spatially fragmented evolutionary dynamics, coupled only by
+community-level mechanisms are even more important in driving hotspots
+of overall risk (Power and Mitchell 2004).
+
+These kinds of dynamics lend themselves to interpretation through the
+lens of the Geographic Mosaic Theory of Coevolution (GMTC; Thompson
+2005), which connects microevolutionary dynamics in symbiotic
+interactions to macroecological dynamics in host communities. The GMTC
+posits that coevolutionary processes between pairs (Thompson 1994) or
+complexes (Janzen 1980) of species are structured in space by the
+rippling effects of abiotic conditions onto evolutionary mechanism,
+resulting in spatially fragmented evolutionary dynamics, coupled only by
 dispersal-related processes (Gomulkiewicz et al. 2000). In turn, these
@@
 systems to have different structure and dynamics over large spatial
-extents (see \emph{e.g.} Price 2002).
-
-The GMTC predicts a spatial fragmentation of coevolutionary dynamics
-under the joint action of three processes (see notably Gomulkiewicz et
-al. 2007), which all have the potential to act on outbreak potential,
-pathogen tramission, and disease virulence (Parratt et al. 2016, Turner
-et al. 2021). First hot- and coldspots of coevolution can appear when
-the intensity of \emph{interaction} (in terms of reciprocal fitness
-consequences) varies spatially, because of \emph{e.g.} partial range
-overlap between organisms (Nuismer et al. 2003). Hot- and coldspots are
-notoriously difficult to identify in nature (Laine 2005), especially
-when the structure of range overlap is complex, and the reciprocal
-fitness effects are relatively low---both of which conditions are met in
-the bat-betacoronavirus complex. Second, the GMTC supposes the existence
+extents (see \emph{e.g.} Price 2002). The GMTC predicts a spatial
+fragmentation of coevolutionary dynamics under the joint action of three
+processes (see notably Gomulkiewicz et al. 2007), which all have the
+potential to act on outbreak potential, pathogen tramission, and disease
+virulence (Parratt et al. 2016, Turner et al. 2021). First hot- and
+coldspots of coevolution can appear when the intensity of
+\emph{interaction} (in terms of reciprocal fitness consequences) varies
+spatially, because of \emph{e.g.} partial range overlap between
+organisms (Nuismer et al. 2003). Second, the GMTC supposes the existence
 of selection mosaics, wherein the intensity of \emph{selection} varies
@@
 are more biotically complex; Thrall et al. 2007) \emph{and} to the local
-favorability of the environment (Hochberg and Baalen 1998). In a system
-with near global distribution and variations in relative richness, like
-the bat-betacoronavirus complex, the GMTC predicts the establishment of
-many different coevolutionary trajectories (and resulting viral
-regimes). The final GMCT process is trait remixing, under which changes
-in coevolutionary dynamics happen due to the arrival (or departure) of
-functional traits, through changes in community composition due to
-invasions, meta-community dynamics, and disperal; these changes are
-likely to happen in overlap zones between ``long branches'' of
-coevolution.
+favorability of the environment (Hochberg and Baalen 1998). Finally,
+trait remixing occurs when coevolutionary dynamics are driven by the
+arrival (or departure) of functional traits, through changes in
+community composition due to invasions, meta-community dynamics, and
+disperal.
+
+Each of these elements can be applied to a relatively quantifiable
+aspect of host-virus ecology: (i) viral sharing among hosts,
+representing the strength of potential interaction between viruses and
+any one host (i.e., places where viruses undergo constant host switching
+may be coevolutionary coldspots); (ii) the phylogenetic diversity of
+hosts, as a proxy for variation in the immunological mechanisms that
+antagonize viruses (i.e., the selection mosaic); and (iii) the local
+uniqueness of the bat community, representing the potential for viruses
+to be exposed to novel host traits (e.g., variation in receptor
+sequences). Together, we argue that these can be used to identify the
+evolutionary drivers that---in conjunction with transmission processes
+(e.g., viral prevalence in reservoirs and animal-human contact rates)---
+determine disease emergence risk.
+
+Here,
 
@@
 from viruses to hosts (focusing on the bats-betacoronavirus complex),
-with a specific emphasis on the potential to create independant
+with a specific emphasis on the potential to create independent
 coevolutionary dynamics (and therefore spatial fragmentation in the
-risk) through heterogeneity. Our components of risk are (i) the
-phylogenetic diversity of hosts, representing the diversification
-potential of the system (hotspots, selection mosaics), under the
-assumption that bat diversification is not primarily driven by viruses
-(Gorbunova et al. 2020); (ii) viral sharing among hosts, representing
-levels of viral exchange within host communities that may act as a
-buffer against viral specialization in any given host (i.e., high
-sharing creates coldspots of selective pressures from any given host
-species); and (iii) the local uniqueness of the bat community,
-representing the potential for viruses to be exposed to novel host
-traits in unusual combinations (trait remixing). We turn these features
-into a spatially-explicit additive mapping of zoonotic risk components,
-revealing the extreme heterogeneity of risk at the global scale.
-Explicitely framing the notion of spillover risk based on propositions
-from the GMTC (which is to say, based on a framework linking
-interactions between species to change within species) is a novel idea,
-that should be relatively general. Indeed, it only assumes the action of
-well described evolutionary mechanisms. The benefit of this approach is
-to provide the potential for a more dynamic and nuanced understanding of
-risk: not only on ecological timescales, but also by providing clues
-about which areas can change over micro-evolutionary timescales. This
-provides a way to look at spatial structure by accounting for more
-notions than species richness/similarity, but also a way to identify
-spatial areas of higher risk.
+risk) through heterogeneity.
 
@@
 
+RELOCATE: Indeed, it only assumes the action of well described
+evolutionary mechanisms. The benefit of this approach is to provide the
+potential for a more dynamic and nuanced understanding of risk: not only
+on ecological timescales, but also by providing clues about which areas
+can change over micro-evolutionary timescales.
+
 Our study focuses largely on the biogeography of hosts. Specifically, we
@@
 
-\leavevmode\hypertarget{ref-Albery2022UrbMam}{}%
-Albery, G. F. et al. 2022. Urban-adapted mammal species have more known
-pathogens. - Nature Ecology \& Evolution 6: 794--801.
-
 \leavevmode\hypertarget{ref-Allen2017GloHot}{}%
@@
 14830.
-
-\leavevmode\hypertarget{ref-Laine2005SpaSca}{}%
-Laine, A.-L. 2005. Spatial scale of local adaptation in a plant-pathogen
-metapopulation. - Journal of Evolutionary Biology 18: 930--938.
 

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6e59aed62281889e728b148dd8371076deca2901 🚀

@@
 
-Spillover risk is complex and, even within a pool of susceptible
-wildlife hosts, driven by a multiplicity of factors (Plowright et al.
-2017). Although host richness is often used as a coarse proxy for
-hotspots of emergence risk (Anthony et al. 2017, Ruiz-Aravena et al.
-2022), these approaches deliberately oversimplify interspecific
-heterogeneity in immunology, behavior, and other traits. Global maps of
-spillover risk often struggle to distill these features into
-interpretable risk maps, and overlook highly-unique host pools that
-allow for the rapid evolution of highly divergent viruses (Agosta et al.
-2010). In the case of generalist pathogens like betacoronaviruses, there
-is conceptual and empirical support to the idea that these
-community-level mechanisms are even more important in driving hotspots
-of overall risk (Power and Mitchell 2004).
-
-These kinds of dynamics lend themselves to interpretation through the
-lens of the Geographic Mosaic Theory of Coevolution (GMTC; Thompson
-2005), which connects microevolutionary dynamics in symbiotic
-interactions to macroecological dynamics in host communities. The GMTC
-posits that coevolutionary processes between pairs (Thompson 1994) or
-complexes (Janzen 1980) of species are structured in space by the
-rippling effects of abiotic conditions onto evolutionary mechanism,
-resulting in spatially fragmented evolutionary dynamics, coupled only by
-dispersal-related processes (Gomulkiewicz et al. 2000). In turn, these
-spatially fragmented processes can lead taxonomically homogeneous
-systems to have different structure and dynamics over large spatial
-extents (see \emph{e.g.} Price 2002). The GMTC predicts a spatial
-fragmentation of coevolutionary dynamics under the joint action of three
-processes (see notably Gomulkiewicz et al. 2007), which all have the
-potential to act on outbreak potential, pathogen tramission, and disease
-virulence (Parratt et al. 2016, Turner et al. 2021). First hot- and
-coldspots of coevolution can appear when the intensity of
+Disease emergence is complex, and is driven not only by animal-human
+contact, but also by the underlying evolutionary dynamics in viral
+reservoirs (Plowright et al. 2017). Although host richness is often used
+as a superficial proxy for spillover risk (Anthony et al. 2017,
+Ruiz-Aravena et al. 2022), these approaches oversimplify the relevant
+interspecific heterogeneity in immunology, behavior, and other traits,
+and therefore overlook unique host pools that allow for the rapid
+evolution of highly divergent viruses (Agosta et al. 2010). In the case
+of generalist pathogens like betacoronaviruses, there is conceptual and
+empirical support to the idea that these community-level mechanisms are
+even more important (Power and Mitchell 2004). This creates a disconnect
+between coevolutionary theory and ecological studies of spillover risk.
+
+The Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005)
+attempts to explicitly connect microevolutionary dynamics to the
+macroecology and biogeography of symbiotic interactions. The GMTC posits
+that coevolutionary processes between pairs (Thompson 1994) or complexes
+(Janzen 1980) of species are structured in space by the rippling effects
+of abiotic conditions onto evolutionary mechanism, giving rise to
+fragmented systems with different structure and ecologically dynamics
+over large spatial extents (see \emph{e.g.} Price 2002). The GMTC
+predicts a spatial fragmentation of coevolutionary dynamics under the
+joint action of three processes (see notably Gomulkiewicz et al. 2007):
+coevolutionary hot- and coldspots, which appear when the intensity of
 \emph{interaction} (in terms of reciprocal fitness consequences) varies
-spatially, because of \emph{e.g.} partial range overlap between
-organisms (Nuismer et al. 2003). Second, the GMTC supposes the existence
-of selection mosaics, wherein the intensity of \emph{selection} varies
-across space; the strength of reciprocal selection responds to the
-biotic complexity of the community (locally diverse hosts and viruses
-are more biotically complex; Thrall et al. 2007) \emph{and} to the local
-favorability of the environment (Hochberg and Baalen 1998). Finally,
-trait remixing occurs when coevolutionary dynamics are driven by the
-arrival (or departure) of functional traits, through changes in
-community composition due to invasions, meta-community dynamics, and
-disperal.
-
-Each of these elements can be applied to a relatively quantifiable
-aspect of host-virus ecology: (i) viral sharing among hosts,
-representing the strength of potential interaction between viruses and
-any one host (i.e., places where viruses undergo constant host switching
-may be coevolutionary coldspots); (ii) the phylogenetic diversity of
-hosts, as a proxy for variation in the immunological mechanisms that
-antagonize viruses (i.e., the selection mosaic); and (iii) the local
-uniqueness of the bat community, representing the potential for viruses
-to be exposed to novel host traits (e.g., variation in receptor
-sequences). Together, we argue that these can be used to identify the
-evolutionary drivers that---in conjunction with transmission processes
-(e.g., viral prevalence in reservoirs and animal-human contact rates)---
-determine disease emergence risk.
-
-Here,
-
-We turn the processes on the GMTC into definitions of spillover risk
-from viruses to hosts (focusing on the bats-betacoronavirus complex),
-with a specific emphasis on the potential to create independent
-coevolutionary dynamics (and therefore spatial fragmentation in the
-risk) through heterogeneity.
+spatially; selection mosaics, wherein the intensity of \emph{selection}
+varies across space, driven by both the biotic complexity of the
+community (locally diverse hosts and viruses are more biotically
+complex) and the local favorability of the environment
+(\textbf{Thrall2007?}); and trait remixing, which occurs when
+coevolutionary dynamics are driven by by the arrival (or departure) of
+functional traits, through changes in community composition due to
+invasions, meta-community dynamics, and disperal.
+
+Here, we develop a framework that quantifies each of these elements in
+light of a measurable aspect of host-virus interactions. To do so, we
+explore the global biogeography of betacoronaviruses, the group that
+includes SARS-CoV, MERS-CoV, and SARS-CoV-2. Working from a recently
+published database of bat hosts of betacoronaviruses, we develop the
+first global maps of both host and virus evolutionary distinctiveness
+and biogeographic regions for this system. Aiming to explain these
+patterns, we develop a generalized framework for applying the GMTC to
+host-virus interactions, with a specific emphasis on the potential to
+create independent coevolutionary dynamics (and therefore spatial
+fragmentation in risk) through heterogeneity. We develop a trivariate
+risk assessment system that connects each GMTC mechanism o a
+quantifiable aspect of host-virus interactions: (i) viral sharing rates
+in host communities, representing the strength of potential interaction
+between viruses and any one host (i.e., places where viruses undergo
+constant host switching may be coevolutionary coldspots); (ii) the
+phylogenetic diversity of hosts, as a proxy for variation in the
+immunological mechanisms that antagonize viruses (i.e., the selection
+mosaic); and (iii) the local uniqueness of the bat community,
+representing the potential for viruses to be exposed to novel host
+traits (e.g., variation in receptor sequences). Together, we argue that
+these can be used to identify and map the evolutionary drivers that---in
+conjunction with transmission processes (e.g., viral prevalence in
+reservoirs and animal-human contact rates)--- determine disease
+emergence risk.
 
@@
 
-\leavevmode\hypertarget{ref-Hochberg1998AntCoe}{}%
-Hochberg, M. E. and Baalen, M. 1998. Antagonistic coevolution over
-productivity gradients. - The American Naturalist 152: 620--634.
-
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
@@
 
-\leavevmode\hypertarget{ref-Nuismer2003CoeHos}{}%
-Nuismer, S. L. et al. 2003. Coevolution between hosts and parasites with
-partially overlapping geographic ranges. - Journal of Evolutionary
-Biology 16: 1337--1345.
-
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
@@
 
-\leavevmode\hypertarget{ref-Parratt2016InfDis}{}%
-Parratt, S. R. et al. 2016. Infectious Disease Dynamics in Heterogeneous
-Landscapes. - Annual Review of Ecology, Evolution, and Systematics 47:
-283--306.
-
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
@@
 
-\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
-Thrall, P. H. et al. 2007. Coevolution of symbiotic mutualists and
-parasites in a community context. - Trends in Ecology \& Evolution 22:
-120--126.
-
-\leavevmode\hypertarget{ref-Turner2021RolEnv}{}%
-Turner, W. C. et al. 2021. The roles of environmental variation and
-parasite survival in virulence--transmission relationships. - Royal
-Society open science 8: 210088.
-
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a215c933433f5547874781355eed948badd49b0e 🚀

@@
 even more important (Power and Mitchell 2004). This creates a disconnect
-between coevolutionary theory and ecological studies of spillover risk.
+between coevolutionary theory and existing ecological frameworks for
+mapping spillover risk.
 
@@
 macroecology and biogeography of symbiotic interactions. The GMTC posits
-that coevolutionary processes between pairs (Thompson 1994) or complexes
+that coevolutionary processes among pairs (Thompson 1994) or complexes
 (Janzen 1980) of species are structured in space by the rippling effects
@@
 coevolutionary dynamics are driven by by the arrival (or departure) of
-functional traits, through changes in community composition due to
-invasions, meta-community dynamics, and disperal.
-
-Here, we develop a framework that quantifies each of these elements in
-light of a measurable aspect of host-virus interactions. To do so, we
-explore the global biogeography of betacoronaviruses, the group that
-includes SARS-CoV, MERS-CoV, and SARS-CoV-2. Working from a recently
-published database of bat hosts of betacoronaviruses, we develop the
-first global maps of both host and virus evolutionary distinctiveness
-and biogeographic regions for this system. Aiming to explain these
-patterns, we develop a generalized framework for applying the GMTC to
-host-virus interactions, with a specific emphasis on the potential to
-create independent coevolutionary dynamics (and therefore spatial
-fragmentation in risk) through heterogeneity. We develop a trivariate
-risk assessment system that connects each GMTC mechanism o a
-quantifiable aspect of host-virus interactions: (i) viral sharing rates
-in host communities, representing the strength of potential interaction
-between viruses and any one host (i.e., places where viruses undergo
-constant host switching may be coevolutionary coldspots); (ii) the
-phylogenetic diversity of hosts, as a proxy for variation in the
-immunological mechanisms that antagonize viruses (i.e., the selection
-mosaic); and (iii) the local uniqueness of the bat community,
-representing the potential for viruses to be exposed to novel host
-traits (e.g., variation in receptor sequences). Together, we argue that
-these can be used to identify and map the evolutionary drivers that---in
-conjunction with transmission processes (e.g., viral prevalence in
-reservoirs and animal-human contact rates)--- determine disease
-emergence risk.
+\emph{functional traits}, through changes in community composition due
+to invasions, meta-community dynamics, and disperal.
+
+Here, we apply the GMTC to explore and explain the global biogeography
+of betacoronaviruses, the group that includes SARS-CoV, MERS-CoV, and
+SARS-CoV-2. Working from a recently published database of bat hosts of
+betacoronaviruses, we develop the first global maps of both host and
+virus evolutionary distinctiveness and biogeographic regions for this
+system. Aiming to explain these patterns, we develop a generalized
+framework for applying the GMTC to host-virus interactions, with a
+specific emphasis on the potential to create independent coevolutionary
+dynamics (and therefore spatial fragmentation in risk) through
+heterogeneity. We develop a trivariate risk assessment system that
+connects each GMTC mechanism to a quantifiable aspect of host-virus
+interactions: (i) viral sharing rates in host communities, representing
+the strength of potential interaction between viruses and any one host
+(i.e., places where viruses undergo constant host switching may be
+coevolutionary coldspots); (ii) the phylogenetic diversity of hosts, as
+a proxy for variation in the immunological mechanisms that antagonize
+viruses (i.e., the selection mosaic); and (iii) the local uniqueness of
+the bat community, representing the potential for viruses to be exposed
+to novel host traits (e.g., variation in receptor sequences). Together,
+we argue that these can be used to identify and map the evolutionary
+drivers that---in conjunction with transmission processes (e.g., viral
+prevalence in reservoirs and animal-human contact rates)--- determine
+disease emergence risk.
 

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@acf0c1ba63d6594b25eaf195d286af60757f1bb2 🚀

@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-\hypertarget{known-betacoronavirus-hosts}{%
-\subsection{\texorpdfstring{Known \emph{Betacoronavirus}
-hosts}{Known Betacoronavirus hosts}}\label{known-betacoronavirus-hosts}}
-
-We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
-by Becker et al. (2022) from
-\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
-filtered it to ``known'' hosts (established before the emergence of
-SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
-competence assays since the initial data collection). The original
-database was assembled by a combination of data mining and literature
-surveys, including automated alerts on the ``bats'' and ``coronavirus''
-keywords to identify novel empirical evidence of bats-betacoronaviruses
-associations; this yielded a total of 126 known hosts, 47 of which were
-novel hosts.
-
-\hypertarget{bats-occurrences}{%
-\subsection{Bats occurrences}\label{bats-occurrences}}
-
-We downloaded the rangemap of every current bat species that was
-classified as an empirically documented host of \emph{Betacoronavirus}
-from the previous step, according to recent IUCN data (IUCN 2021). The
-range maps were subsequently rasterized using the \texttt{rasterize}
-function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
-approximately 100kmx100km. For every pixel in the resulting raster where
-at least one bat host of \emph{Betacoronavirus} was present, we extract
-the species pool (list of all competent bat hosts), which was used to
-calculate the following risk assessment components: bat phylogenetic
-diversity, bat compositional uniqueness, and predicted viral sharing
-risk.
-
-\hypertarget{bats-phylogenetic-diversity}{%
-\subsection{Bats phylogenetic
-diversity}\label{bats-phylogenetic-diversity}}
-
-For every pixel, we measured Faith's Phylogenetic Diversity (Faith 1992)
-based on a recent synthetic tree with robust time calibration, covering
-about 6000 mammalian species (Upham et al. 2019). Faith's PD measures
-the sum of unique branches from an arbitrary root to a set of tips, and
-comparatively larger values indicate a more phylogenetic diverse species
-pool. We measured phylogenetic diversity starting from the root of the
-entire tree (and not from Chiroptera); this bears no consequences on the
-resulting values, since all branches leading up to Chiroptera are only
-counted one per species pool, and (as we explain when describing the
-assembly of the composite risk map), all individual risk components are
-ranged in {[}0,1{]}. This measure incorporates a richness component,
-which we chose not to correct for; the interpretation of the
-phylogenetic diversity is therefore a weighted species richness, that
-accounts for phylogenetic over/under-dispersal in some places.
-
-\hypertarget{bats-compositional-uniqueness}{%
-\subsection{Bats compositional
-uniqueness}\label{bats-compositional-uniqueness}}
-
-For every species pool, we measured its Local Contribution to
-Beta-Diversity (Legendre and De Cáceres 2013); LCBD works from a
-species-data matrix (traditionally noted as \(\mathbf{Y}\)), where
-species are rows and sites are columns, and a value of 1 indicates
-occurrence. We extracted the Y matrix assuming that every pixel
-represents a unique location, and following best practices (Legendre and
-Condit 2019) transformed it using Hellinger's distance to account for
-unequal bat richness at different pixels. The correction of raw
-community data is particularly important for two reasons: first, it
-prevents the artifact of richer sites having higher importance; second,
-it removes the effect of overall species richness, which is already
-incorporated in the phylogenetic diversity component. High values of
-LCBD indicate that the pixel has a community that is on average more
-dissimilar in species composition than what is expected knowing the
-entire matrix, i.e.~a more unique community. Recent results by Dansereau
-et al. (2022) shows that LCBD measures are robust with regards to
-spatial scale, and are therefore applicable at the global scale.
-
-\hypertarget{viral-sharing-between-hosts}{%
-\subsection{Viral sharing between
-hosts}\label{viral-sharing-between-hosts}}
-
-For all bat hosts of \emph{Betacoronavirus}, we extracted their
-predicted viral sharing network, generated from a previously published
-generalized additive mixed model of virus sharing by a tensor function
-of phylogenetic distance and geographic range overlap (Albery et al.
-2020). This network stores pairwise values of viral community
-similarity. To project viral sharing values into a single value for
-every pixel, we averaged the pairwise scores. High values of the average
-sharing propensity means that this specific extant bat assemblage is
-likely to be proficient at exchanging viruses.
-
-\hypertarget{composite-risk-map}{%
-\subsection{Composite risk map}\label{composite-risk-map}}
-
-To visualize the aggregated risk at the global scale, we combine the
-three individual risk components (phylogenetic diversity, compositional
-uniqueness, and viral sharing) using an additive color model (Seekell et
-al. 2018). In this approach, every risk component gets assigned a
-component in the RGB color model (phylogenetic diversity is green,
-compositional uniqueness is red, and viral sharing is blue). In order to
-achieve a valid RGB measure, all components are re-scaled to the
-{[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
-diversity, and no compositional uniqueness is black, and a pixel with
-maximal values for each is white. This additive model conveys both the
-intensity of the overall risk, but also the nature of the risk as colors
-diverge towards combinations of values for three risk components. Out of
-the possible combinations, the most risky in terms or rapid
-diversification and spillover potential is high phylogenetic diversity
-and low viral sharing (Gomulkiewicz et al. 2000, Cavender-Bares et al.
-2009), in that this allows multiple independent host-virus
-coevolutionary dynamics to take place in the same location. In the
-colorimetric space, this correspond to yellow -- because the HSV space
-is more amenable to calculations for feature extraction (see \emph{e.g.}
-Keke et al. 2010), we measured the risk level by calculating the angular
-distance of the hue of each pixel to a reference value of 60, and
-weighted this risk level by the value component. Specifically, given a
-pixel with colorimetric coordinates \((h,s,v)\), its ranged weighted
-risk value is
-
-\[
-v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right|}{2\pi}\right]\,,
-\]
-
-where X is
-\(\text{atan}\left(\text{cos}(\text{rad}(60)), \text{sin}(\text{rad}(60))\right)\),
-a constant approximately equal to \(0.5235\).
-
-\hypertarget{viral-phylogeography-and-evolutionary-diversification}{%
-\subsection{Viral phylogeography and evolutionary
-diversification}\label{viral-phylogeography-and-evolutionary-diversification}}
-
-To next represent phylogeography of betacoronaviruses in bats, we
-aggregated and analyzed betacoronavirus sequence data. We used the
-following query to pull all \emph{Betacoronavirus} sequence data from
-the GenBank Nucleotide database except SARS-CoV-2;
-(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
-NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
-OR sars-cov-2{[}All Fields{]}). We added a single representative
-sequence for SARS-CoV-2 and manually curated to remove sequences without
-the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
-indicating recombinant or laboratory strains including ``patent,''
-``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
-taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
-syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
-betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
-{[}Katoh and Standley (2013); Algorithm FFT-NS-2, Scoring matrix 200PAM
-/ k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
-likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al. 2015)
-with ModelFinder (Kalyaanamoorthy et al. 2017) ultrafast bootstrap
-approximation (Hoang et al. 2018) with a general time reversible model
-with empirical base frequencies and the 5-discrete-rate-category
-FreeRaye model of nucleotide substitution (GTR+F+R5).
-
-We first tested the hypothesis that hotspots of viral diversification
-would track hotspots of bat diversification. To do so, we plotted the
-number of known bat hosts (specifically only those included in the
-phylogeny, so there was a 1:1 correspondence between data sources)
-against the ``mean evolutionary distinctiveness'' of the associated
-viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness (Isaac et al. 2007) for each of the viruses in the tree,
-then averaged these at the bat species level, projected these values
-onto their geographic distributions, and averaged across every bat found
-in a given pixel. As such, this can be thought of as a map of the mean
-evolutionary distinctiveness of the known viral community believed to be
-associated with a particular subset of bats present.
-
-\hypertarget{co-distribution-of-hosts-and-viral-hotspots}{%
-\subsection{Co-distribution of hosts and viral
-hotspots}\label{co-distribution-of-hosts-and-viral-hotspots}}
-
-Subsequently, we tested the hypothesis that the biogeography of bat
-betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from (Kreft and Jetz 2007, 2010),
-who proposed a phylogenetic method for the delineation of animal
-biogeographic regions. In their original method, a distance matrix -
-where each row or column represents a geographic raster's grid cell, and
-the dissimilarity values are the ``beta diversity similarity'' of their
-community assemble - undergoes non-metric multidimensional scaling
-(NMDS); the first two axes of the NMDS are projected geographically
-using a four-color bivariate map. Here, we build on this idea with an
-entirely novel methodology. First, we measure the phylogenetic distance
-between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
-subsequently, we take a principal components analysis of that distance
-matrix (readily interchangeable for NMDS in this case) to project the
-viral tree into an n-dimensional space. We then take the first two
-principal components and, as with the evolutionary distinctiveness
-analysis, aggregated these to a mean host value and projected them using
-a four-color bivariate map.
-
-\hypertarget{results-and-discussion}{%
-\section{Results and discussion}\label{results-and-discussion}}
-
 \hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
@@
 
-RELOCATE: Indeed, it only assumes the action of well described
-evolutionary mechanisms. The benefit of this approach is to provide the
-potential for a more dynamic and nuanced understanding of risk: not only
-on ecological timescales, but also by providing clues about which areas
-can change over micro-evolutionary timescales.
-
-Our study focuses largely on the biogeography of hosts. Specifically, we
-identify the Amazon and South-Eastern Asia as hotspots where the
-phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
-(Anthony et al. 2017); surprisingly, current data suggest that viral
-sharing between hosts is high in the Amazon and low in South-Eastern
-Asia, which has the potential to result in different evolutionary
-dynamics between these two regions, hinting at different futures for
-their viral communities. This work is important both as a description of
-the bats-betacoronaviruses complex, but also because more broadly, bats
-are known reservoirs for a variety of emerging viruses and other
-pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
-the needs for bat conservation and disease prevention most likely very
-difficult and a source of human-wildlife conflicts, especially in more
-densely populated areas (Stone et al. 2015, Rego et al. 2015).
-
-Yet, we know that viruses with high host plasticity, that is, the
-ability of a given virus to adapt to various taxonomic orders and
-ecological groups (Kreuder Johnson et al. 2015), are more likely to
-amplify viral spillover, followed by secondary human-to-human
-transmission, and geographical spread (Hazarie et al. 2021). High viral
-host plasticity is an especially important trait for RNA viruses like
-betacoronaviruses (Haddad et al. 2021). Indeed, our analysis of viral
-sequences reveals that Latin America is a hotspot of viral
-distinctiveness, suggesting that this part of the bats-betacoronaviruses
-complex may be undergoing independent evolutionary dynamics (related
-species sharing viruses that are different from the rest of the global
-pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
-richness is high but sharing is low; this suggests a different type of
-evolutionary dynamics (unrelated viruses coevolving with evolutionarily
-distinct hosts, generating high diversity locally, see \emph{e.g.}
-Latinne et al. 2020). Both of these areas should be priority areas for
-sampling, especially since Becker et al. (2022) advance that they harbor
-undiscovered hosts of beta-coronaviruses. This diversity of hosts, and
-the mechanisms by which the exchange of viruses occurs between species,
-is largely affected by the local environmental conditions and
-environmental change.
-
 Bats are important reservoir hosts for different classes of
@@
 
+RELOCATE: Indeed, it only assumes the action of well described
+evolutionary mechanisms. The benefit of this approach is to provide the
+potential for a more dynamic and nuanced understanding of risk: not only
+on ecological timescales, but also by providing clues about which areas
+can change over micro-evolutionary timescales.
+
+Our study focuses largely on the biogeography of hosts. Specifically, we
+identify the Amazon and South-Eastern Asia as hotspots where the
+phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
+(Anthony et al. 2017); surprisingly, current data suggest that viral
+sharing between hosts is high in the Amazon and low in South-Eastern
+Asia, which has the potential to result in different evolutionary
+dynamics between these two regions, hinting at different futures for
+their viral communities. This work is important both as a description of
+the bats-betacoronaviruses complex, but also because more broadly, bats
+are known reservoirs for a variety of emerging viruses and other
+pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
+the needs for bat conservation and disease prevention most likely very
+difficult and a source of human-wildlife conflicts, especially in more
+densely populated areas (Stone et al. 2015, Rego et al. 2015).
+
+Yet, we know that viruses with high host plasticity, that is, the
+ability of a given virus to adapt to various taxonomic orders and
+ecological groups (Kreuder Johnson et al. 2015), are more likely to
+amplify viral spillover, followed by secondary human-to-human
+transmission, and geographical spread (Hazarie et al. 2021). High viral
+host plasticity is an especially important trait for RNA viruses like
+betacoronaviruses (Haddad et al. 2021). Indeed, our analysis of viral
+sequences reveals that Latin America is a hotspot of viral
+distinctiveness, suggesting that this part of the bats-betacoronaviruses
+complex may be undergoing independent evolutionary dynamics (related
+species sharing viruses that are different from the rest of the global
+pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
+richness is high but sharing is low; this suggests a different type of
+evolutionary dynamics (unrelated viruses coevolving with evolutionarily
+distinct hosts, generating high diversity locally, see \emph{e.g.}
+Latinne et al. 2020). Both of these areas should be priority areas for
+sampling, especially since Becker et al. (2022) advance that they harbor
+undiscovered hosts of beta-coronaviruses. This diversity of hosts, and
+the mechanisms by which the exchange of viruses occurs between species,
+is largely affected by the local environmental conditions and
+environmental change.
+
 There are several factors that drive changes in the diversity of bats
@@
 
+\newpage
+
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{known-betacoronavirus-hosts}{%
+\subsection{\texorpdfstring{Known \emph{Betacoronavirus}
+hosts}{Known Betacoronavirus hosts}}\label{known-betacoronavirus-hosts}}
+
+We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
+by Becker et al. (2022) from
+\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
+filtered it to ``known'' hosts (established before the emergence of
+SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
+competence assays since the initial data collection). The original
+database was assembled by a combination of data mining and literature
+surveys, including automated alerts on the ``bats'' and ``coronavirus''
+keywords to identify novel empirical evidence of bats-betacoronaviruses
+associations; this yielded a total of 126 known hosts, 47 of which were
+novel hosts.
+
+\hypertarget{bat-occurrences}{%
+\subsection{Bat occurrences}\label{bat-occurrences}}
+
+We downloaded the rangemap of every current bat species that was
+classified as an empirically documented host of \emph{Betacoronavirus}
+from the previous step, according to recent IUCN data (IUCN 2021). The
+range maps were subsequently rasterized using the \texttt{rasterize}
+function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
+approximately 100kmx100km. For every pixel in the resulting raster where
+at least one bat host of \emph{Betacoronavirus} was present, we extract
+the species pool (list of all competent bat hosts), which was used to
+calculate the following risk assessment components: bat phylogenetic
+diversity, bat compositional uniqueness, and predicted viral sharing
+risk.
+
+\hypertarget{bat-phylogenetic-diversity}{%
+\subsection{Bat phylogenetic
+diversity}\label{bat-phylogenetic-diversity}}
+
+For every pixel, we measured Faith's Phylogenetic Diversity (Faith 1992)
+based on a recent synthetic tree with robust time calibration, covering
+about 6000 mammalian species (Upham et al. 2019). Faith's PD measures
+the sum of unique branches from an arbitrary root to a set of tips, and
+comparatively larger values indicate a more phylogenetic diverse species
+pool. We measured phylogenetic diversity starting from the root of the
+entire tree (and not from Chiroptera); this bears no consequences on the
+resulting values, since all branches leading up to Chiroptera are only
+counted one per species pool, and (as we explain when describing the
+assembly of the composite risk map), all individual risk components are
+ranged in {[}0,1{]}. This measure incorporates a richness component,
+which we chose not to correct for; the interpretation of the
+phylogenetic diversity is therefore a weighted species richness, that
+accounts for phylogenetic over/under-dispersal in some places.
+
+\hypertarget{bat-compositional-uniqueness}{%
+\subsection{Bat compositional
+uniqueness}\label{bat-compositional-uniqueness}}
+
+For every species pool, we measured its Local Contribution to
+Beta-Diversity (Legendre and De Cáceres 2013); LCBD works from a
+species-data matrix (traditionally noted as \(\mathbf{Y}\)), where
+species are rows and sites are columns, and a value of 1 indicates
+occurrence. We extracted the Y matrix assuming that every pixel
+represents a unique location, and following best practices (Legendre and
+Condit 2019) transformed it using Hellinger's distance to account for
+unequal bat richness at different pixels. The correction of raw
+community data is particularly important for two reasons: first, it
+prevents the artifact of richer sites having higher importance; second,
+it removes the effect of overall species richness, which is already
+incorporated in the phylogenetic diversity component. High values of
+LCBD indicate that the pixel has a community that is on average more
+dissimilar in species composition than what is expected knowing the
+entire matrix, i.e.~a more unique community. Recent results by Dansereau
+et al. (2022) shows that LCBD measures are robust with regards to
+spatial scale, and are therefore applicable at the global scale.
+
+\hypertarget{viral-sharing-between-hosts}{%
+\subsection{Viral sharing between
+hosts}\label{viral-sharing-between-hosts}}
+
+For all bat hosts of \emph{Betacoronavirus}, we extracted their
+predicted viral sharing network, generated from a previously published
+generalized additive mixed model of virus sharing by a tensor function
+of phylogenetic distance and geographic range overlap (Albery et al.
+2020). This network stores pairwise values of viral community
+similarity. To project viral sharing values into a single value for
+every pixel, we averaged the pairwise scores. High values of the average
+sharing propensity means that this specific extant bat assemblage is
+likely to be proficient at exchanging viruses.
+
+\hypertarget{composite-risk-map}{%
+\subsection{Composite risk map}\label{composite-risk-map}}
+
+To visualize the aggregated risk at the global scale, we combine the
+three individual risk components (phylogenetic diversity, compositional
+uniqueness, and viral sharing) using an additive color model (Seekell et
+al. 2018). In this approach, every risk component gets assigned a
+component in the RGB color model (phylogenetic diversity is green,
+compositional uniqueness is red, and viral sharing is blue). In order to
+achieve a valid RGB measure, all components are re-scaled to the
+{[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
+diversity, and no compositional uniqueness is black, and a pixel with
+maximal values for each is white. This additive model conveys both the
+intensity of the overall risk, but also the nature of the risk as colors
+diverge towards combinations of values for three risk components. Out of
+the possible combinations, the most risky in terms or rapid
+diversification and spillover potential is high phylogenetic diversity
+and low viral sharing (Gomulkiewicz et al. 2000, Cavender-Bares et al.
+2009), in that this allows multiple independent host-virus
+coevolutionary dynamics to take place in the same location. In the
+colorimetric space, this correspond to yellow -- because the HSV space
+is more amenable to calculations for feature extraction (see \emph{e.g.}
+Keke et al. 2010), we measured the risk level by calculating the angular
+distance of the hue of each pixel to a reference value of 60, and
+weighted this risk level by the value component. Specifically, given a
+pixel with colorimetric coordinates \((h,s,v)\), its ranged weighted
+risk value is
+
+\[
+v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right|}{2\pi}\right]\,,
+\]
+
+where X is
+\(\text{atan}\left(\text{cos}(\text{rad}(60)), \text{sin}(\text{rad}(60))\right)\),
+a constant approximately equal to \(0.5235\).
+
+\hypertarget{viral-phylogeography-and-evolutionary-diversification}{%
+\subsection{Viral phylogeography and evolutionary
+diversification}\label{viral-phylogeography-and-evolutionary-diversification}}
+
+To next represent phylogeography of betacoronaviruses in bats, we
+aggregated and analyzed betacoronavirus sequence data. We used the
+following query to pull all \emph{Betacoronavirus} sequence data from
+the GenBank Nucleotide database except SARS-CoV-2;
+(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
+NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
+OR sars-cov-2{[}All Fields{]}). We added a single representative
+sequence for SARS-CoV-2 and manually curated to remove sequences without
+the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
+indicating recombinant or laboratory strains including ``patent,''
+``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
+taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
+syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
+betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
+{[}Katoh and Standley (2013); Algorithm FFT-NS-2, Scoring matrix 200PAM
+/ k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
+likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al. 2015)
+with ModelFinder (Kalyaanamoorthy et al. 2017) ultrafast bootstrap
+approximation (Hoang et al. 2018) with a general time reversible model
+with empirical base frequencies and the 5-discrete-rate-category
+FreeRaye model of nucleotide substitution (GTR+F+R5).
+
+We first tested the hypothesis that hotspots of viral diversification
+would track hotspots of bat diversification. To do so, we plotted the
+number of known bat hosts (specifically only those included in the
+phylogeny, so there was a 1:1 correspondence between data sources)
+against the ``mean evolutionary distinctiveness'' of the associated
+viruses. To calculate this, we derived the fair proportions evolutionary
+distinctiveness (Isaac et al. 2007) for each of the viruses in the tree,
+then averaged these at the bat species level, projected these values
+onto their geographic distributions, and averaged across every bat found
+in a given pixel. As such, this can be thought of as a map of the mean
+evolutionary distinctiveness of the known viral community believed to be
+associated with a particular subset of bats present.
+
+\hypertarget{co-distribution-of-hosts-and-viral-hotspots}{%
+\subsection{Co-distribution of hosts and viral
+hotspots}\label{co-distribution-of-hosts-and-viral-hotspots}}
+
+Subsequently, we tested the hypothesis that the biogeography of bat
+betacoronaviruses should track the biogeography of their hosts. To test
+this idea, we loosely adapted a method from (Kreft and Jetz 2007, 2010),
+who proposed a phylogenetic method for the delineation of animal
+biogeographic regions. In their original method, a distance matrix -
+where each row or column represents a geographic raster's grid cell, and
+the dissimilarity values are the ``beta diversity similarity'' of their
+community assemble - undergoes non-metric multidimensional scaling
+(NMDS); the first two axes of the NMDS are projected geographically
+using a four-color bivariate map. Here, we build on this idea with an
+entirely novel methodology. First, we measure the phylogenetic distance
+between the different viruses in the betacoronaviruses tree by using the
+cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
+subsequently, we take a principal components analysis of that distance
+matrix (readily interchangeable for NMDS in this case) to project the
+viral tree into an n-dimensional space. We then take the first two
+principal components and, as with the evolutionary distinctiveness
+analysis, aggregated these to a mean host value and projected them using
+a four-color bivariate map.
+
 \hypertarget{references}{%

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@985d7c950248429a9991142d4c0691e94823181b 🚀

@@
 
-RELOCATE: Indeed, it only assumes the action of well described
-evolutionary mechanisms. The benefit of this approach is to provide the
-potential for a more dynamic and nuanced understanding of risk: not only
-on ecological timescales, but also by providing clues about which areas
-can change over micro-evolutionary timescales.
-
-Our study focuses largely on the biogeography of hosts. Specifically, we
-identify the Amazon and South-Eastern Asia as hotspots where the
-phylogenetic distinctiveness of \emph{Betacoronavirus} is the highest
-(Anthony et al. 2017); surprisingly, current data suggest that viral
-sharing between hosts is high in the Amazon and low in South-Eastern
-Asia, which has the potential to result in different evolutionary
-dynamics between these two regions, hinting at different futures for
-their viral communities. This work is important both as a description of
-the bats-betacoronaviruses complex, but also because more broadly, bats
-are known reservoirs for a variety of emerging viruses and other
-pathogens (Calisher et al. 2006, Melaun et al. 2014), making balancing
-the needs for bat conservation and disease prevention most likely very
-difficult and a source of human-wildlife conflicts, especially in more
-densely populated areas (Stone et al. 2015, Rego et al. 2015).
-
-Yet, we know that viruses with high host plasticity, that is, the
-ability of a given virus to adapt to various taxonomic orders and
-ecological groups (Kreuder Johnson et al. 2015), are more likely to
-amplify viral spillover, followed by secondary human-to-human
-transmission, and geographical spread (Hazarie et al. 2021). High viral
-host plasticity is an especially important trait for RNA viruses like
-betacoronaviruses (Haddad et al. 2021). Indeed, our analysis of viral
-sequences reveals that Latin America is a hotspot of viral
-distinctiveness, suggesting that this part of the bats-betacoronaviruses
-complex may be undergoing independent evolutionary dynamics (related
-species sharing viruses that are different from the rest of the global
-pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
-richness is high but sharing is low; this suggests a different type of
-evolutionary dynamics (unrelated viruses coevolving with evolutionarily
-distinct hosts, generating high diversity locally, see \emph{e.g.}
-Latinne et al. 2020). Both of these areas should be priority areas for
-sampling, especially since Becker et al. (2022) advance that they harbor
-undiscovered hosts of beta-coronaviruses. This diversity of hosts, and
-the mechanisms by which the exchange of viruses occurs between species,
-is largely affected by the local environmental conditions and
-environmental change.
-
-There are several factors that drive changes in the diversity of bats
-(Alves et al. 2018), but human activities' effects on the ecosystem
-(like modifications of land use) could significantly decrease it.
-Therefore, it can be suggested that changes in the diversity of betacovs
-in bats are linked to their biogeographic variation, and human
-population density and other anthropogenic factors are decisive
-moderators for its implications in public health. With the increase of
-contact between humans and potential hosts, we also increase the risk of
-emergence of novel diseases (Johnson et al. 2020), as previous studies
-on RNA viruses suggest the importance of host phylogeography at the time
-of virus dispersal (Gryseels et al. 2017). One of these scenarios where
-interaction between bats and humans can occur can be seed dispersal in
-tropical agroecosystems. It opens the discussion of whether the fruits
-thrown by bats not only disperse seeds but could also be a source of
-indirect interaction between viruses of bat origin and humans (Deshpande
-et al. 2022). This represents a challenge for conservation strategies
-and disease ecology since some areas can haveboth potential for the
-acquisition of zoonotic viruses and bat-human interactions; in
-particular, the challenge lies in the fact that actual exposure must
-then be quantified accounting for several transmission scenarios,
-including both direct and indirect bat - human interaction.
+Previous studies that have used host richness as proxy have predicted a
+high diversity of unsampled bat viruses {[}Olival{]}, bat coronaviruses
+{[}Anthony{]}, and even specifically betacoronaviruses {[}Becker{]} in
+both the Amazon and southeast Asia. While we find that both regions are
+characterized by highly divergent host and viral communities, our
+framework identifies key differences between the regions. We find that
+Latin America is a hotspot of both host and viral distinctiveness,
+suggesting that this branch of the bat-betacoronavirus complex may be
+undergoing independent evolutionary dynamics from the rest of the global
+pool, but with limited potential for viral diversification--- a finding
+that is supported by previous work indicating a higher rate of
+codivergence in Latin America {[}Anthony{]}. In contrast, in southeast
+Asia, host richness and viral distinctiveness are high but sharing is
+low; this suggests a different type of evolutionary dynamics that could
+generate high viral diversity locally (see \emph{e.g.} (Latinne et al.
+2020), as well as the discovery of recombinant viruses that share
+genetic material from both the SARS-CoV and SARS-CoV-2 branches of the
+Sarbecovirus lineage {[}Wu{]}). Both of these regions are priority areas
+for sampling, especially given predictions by Becker et al. (2022) that
+they contain many bat hosts of undiscovered betacoronaviruses. However,
+both the evolutionary and ecological aspects of emergence risk are
+likely higher in southeast Asia---a fact that will only become more
+relevant, as bats track shifting climates and exchange viruses with
+other species, creating a hotspot of cross-species transmission unique
+to the region {[}Carlson{]}.
+
+{[}para about conservation{]} There are several factors that drive
+changes in the diversity of bats (Alves et al. 2018), but human
+activities' effects on the ecosystem (like modifications of land use)
+could significantly decrease it. Therefore, it can be suggested that
+changes in the diversity of betacoronaviruses in bats are linked to
+their biogeographic variation, and human population density and other
+anthropogenic factors are decisive moderators for its implications in
+public health. With the increase of contact between humans and potential
+hosts, we also increase the risk of emergence of novel diseases (Johnson
+et al. 2020), as previous studies on RNA viruses suggest the importance
+of host phylogeography at the time of virus dispersal (Gryseels et al.
+2017). One of these scenarios where interaction between bats and humans
+can occur can be seed dispersal in tropical agroecosystems. It opens the
+discussion of whether the fruits thrown by bats not only disperse seeds
+but could also be a source of indirect interaction between viruses of
+bat origin and humans (Deshpande et al. 2022). This represents a
+challenge for conservation strategies and disease ecology since some
+areas can haveboth potential for the acquisition of zoonotic viruses and
+bat-human interactions; in particular, the challenge lies in the fact
+that actual exposure must then be quantified accounting for several
+transmission scenarios, including both direct and indirect bat - human
+interaction.
 
@@
 
-\leavevmode\hypertarget{ref-Haddad2021SarPos}{}%
-Haddad, D. et al. 2021. SARS-CoV-2: Possible recombination and emergence
-of potentially more virulent strains (H Attoui, Ed.). - PLOS ONE 16:
-e0251368.
-
 \leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
@@
 2029--2053.
-
-\leavevmode\hypertarget{ref-KreuderJohnson2015SpiPan}{}%
-Kreuder Johnson, C. et al. 2015. Spillover and pandemic properties of
-zoonotic viruses with high host plasticity. - Scientific Reports 5:
-14830.
 
@@
 
-\leavevmode\hypertarget{ref-Melaun2014BatPot}{}%
-Melaun, C. et al. 2014. Bats as Potential Reservoir Hosts for
-Vector-Borne Diseases. - In: Klimpel, S. and Mehlhorn, H. (eds), Bats
-(Chiroptera) as Vectors of Diseases and Parasites: Facts and Myths.
-Parasitology Research Monographs. Springer, pp. 25--61.
-
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
@@
 
-\leavevmode\hypertarget{ref-Rego2015AssHum}{}%
-Rego, K. M. da C. et al. 2015. Assessing human-bat interactions around a
-protected area in northeastern Brazil. - Journal of Ethnobiology and
-Ethnomedicine 11: 80.
-
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
@@
 
-\leavevmode\hypertarget{ref-Stone2015ManCon}{}%
-Stone, E. et al. 2015. Managing Conflict between Bats and Humans: The
-Response of Soprano Pipistrelles (Pipistrellus pygmaeus) to Exclusion
-from Roosts in Houses. - PLoS ONE 10: e0131825.
-
 \leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@85ea305586fa11eef39b9fb32a52100fd76a1565 🚀

@@
 
-Previous studies that have used host richness as proxy have predicted a
-high diversity of unsampled bat viruses {[}Olival{]}, bat coronaviruses
-{[}Anthony{]}, and even specifically betacoronaviruses {[}Becker{]} in
-both the Amazon and southeast Asia. While we find that both regions are
-characterized by highly divergent host and viral communities, our
-framework identifies key differences between the regions. We find that
-Latin America is a hotspot of both host and viral distinctiveness,
-suggesting that this branch of the bat-betacoronavirus complex may be
-undergoing independent evolutionary dynamics from the rest of the global
-pool, but with limited potential for viral diversification--- a finding
-that is supported by previous work indicating a higher rate of
-codivergence in Latin America {[}Anthony{]}. In contrast, in southeast
-Asia, host richness and viral distinctiveness are high but sharing is
-low; this suggests a different type of evolutionary dynamics that could
-generate high viral diversity locally (see \emph{e.g.} (Latinne et al.
-2020), as well as the discovery of recombinant viruses that share
-genetic material from both the SARS-CoV and SARS-CoV-2 branches of the
-Sarbecovirus lineage {[}Wu{]}). Both of these regions are priority areas
-for sampling, especially given predictions by Becker et al. (2022) that
-they contain many bat hosts of undiscovered betacoronaviruses. However,
-both the evolutionary and ecological aspects of emergence risk are
-likely higher in southeast Asia---a fact that will only become more
-relevant, as bats track shifting climates and exchange viruses with
-other species, creating a hotspot of cross-species transmission unique
-to the region {[}Carlson{]}.
-
-{[}para about conservation{]} There are several factors that drive
-changes in the diversity of bats (Alves et al. 2018), but human
-activities' effects on the ecosystem (like modifications of land use)
-could significantly decrease it. Therefore, it can be suggested that
-changes in the diversity of betacoronaviruses in bats are linked to
+Here, we propose a simple freamework with broad explanatory power that
+helps contextualize discoveries like highly divergent nobecoviruses in
+Madagascar and the previously-neglected adaptive radiation of
+sarbecoviruses outside of southern China and throughout southeast Asia.
+In doing so, it advances ecological theory beyond the current state of
+the art for global maps of emergence risk. For example, previous studies
+that have used host richness as proxy have predicted a high diversity of
+unsampled bat viruses {[}Olival{]}, bat coronaviruses {[}Anthony{]}, and
+even specifically betacoronaviruses {[}Becker{]} in both the Amazon and
+southeast Asia. While we find that both regions are characterized by
+highly divergent host and viral communities, our framework identifies
+key differences between the regions. We find that Latin America is a
+hotspot of both host and viral distinctiveness, suggesting that this
+branch of the bat-betacoronavirus complex may be undergoing independent
+evolutionary dynamics from the rest of the global pool, but with limited
+potential for viral diversification--- a finding that is supported by
+previous work indicating a higher rate of codivergence in Latin America
+{[}Anthony{]}. In contrast, in southeast Asia, host richness and viral
+distinctiveness are high but sharing is low; this suggests a different
+type of evolutionary dynamics that could generate high local diversity
+of viruses through host switching and viral recombination (see
+\emph{e.g.} (Latinne et al. 2020), as well as the discovery of
+recombinant viruses that share genetic material from both the SARS-CoV
+and SARS-CoV-2 branches of the Sarbecovirus lineage {[}Wu{]}). Both of
+these regions are priority areas for sampling, especially given
+predictions by Becker et al. (2022) that they contain many bat hosts of
+undiscovered betacoronaviruses. However, both the evolutionary and
+ecological aspects of emergence risk are likely higher in southeast
+Asia---a fact that will only become more relevant, as bats track
+shifting climates and exchange viruses with other species, creating a
+hotspot of cross-species transmission unique to the region
+{[}Carlson{]}.
+
+{[}Tim tinker with para about conservation{]} There are several factors
+that drive changes in the diversity of bats (Alves et al. 2018), but
+human activities' effects on the ecosystem (like modifications of land
+use) could significantly decrease it. Therefore, it can be suggested
+that changes in the diversity of betacoronaviruses in bats are linked to
 their biogeographic variation, and human population density and other
@@
 at least one bat host of \emph{Betacoronavirus} was present, we extract
-the species pool (list of all competent bat hosts), which was used to
+the species pool (list of all known bat hosts), which was used to
 calculate the following risk assessment components: bat phylogenetic

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@33a1b0ced81c54a6379e039b524e45bfb8aff78d 🚀

@@
 empirical support to the idea that these community-level mechanisms are
-even more important (Power and Mitchell 2004). This creates a disconnect
-between coevolutionary theory and existing ecological frameworks for
-mapping spillover risk.
+even more important (Power and Mitchell 2004), particularly given that
+cross-species transmission may, as a rule, structure viral evolution
+more than co-divergence with hosts {[}Geoghegan{]}. This creates a
+disconnect between coevolutionary theory (including empirical evidence
+from virology) and most existing ecological frameworks for mapping
+spillover risk.
 
@@
 of betacoronaviruses, the group that includes SARS-CoV, MERS-CoV, and
-SARS-CoV-2. Working from a recently published database of bat hosts of
-betacoronaviruses, we develop the first global maps of both host and
-virus evolutionary distinctiveness and biogeographic regions for this
-system. Aiming to explain these patterns, we develop a generalized
-framework for applying the GMTC to host-virus interactions, with a
-specific emphasis on the potential to create independent coevolutionary
-dynamics (and therefore spatial fragmentation in risk) through
-heterogeneity. We develop a trivariate risk assessment system that
-connects each GMTC mechanism to a quantifiable aspect of host-virus
-interactions: (i) viral sharing rates in host communities, representing
-the strength of potential interaction between viruses and any one host
-(i.e., places where viruses undergo constant host switching may be
-coevolutionary coldspots); (ii) the phylogenetic diversity of hosts, as
-a proxy for variation in the immunological mechanisms that antagonize
-viruses (i.e., the selection mosaic); and (iii) the local uniqueness of
-the bat community, representing the potential for viruses to be exposed
-to novel host traits (e.g., variation in receptor sequences). Together,
-we argue that these can be used to identify and map the evolutionary
-drivers that---in conjunction with transmission processes (e.g., viral
-prevalence in reservoirs and animal-human contact rates)--- determine
-disease emergence risk.
+SARS-CoV-2. In their bat reservoirs, coronaviruses evolve through a mix
+of host jumps, recombination among disparate lineages, and, to a lesser
+degree, co-divergence with their hosts {[}Anthony{]}---a mix of
+mechanisms that creates a complex and nonlinear relationship between
+host diversity and viral emergence. Working from a recently published
+database of bat hosts of betacoronaviruses, we develop the first global
+maps of both host and virus evolutionary distinctiveness and
+biogeographic regions for this system. Aiming to explain these patterns,
+we develop a generalized framework for applying the GMTC to host-virus
+interactions, with a specific emphasis on the potential to create
+independent coevolutionary dynamics (and therefore spatial fragmentation
+in risk) through heterogeneity. We develop a trivariate risk assessment
+system that connects each GMTC mechanism to a quantifiable aspect of
+host-virus interactions: (i) viral sharing rates in host communities,
+representing the strength of potential interaction between viruses and
+any one host (i.e., places where viruses undergo constant host switching
+may be coevolutionary coldspots); (ii) the phylogenetic diversity of
+hosts, as a proxy for variation in the immunological mechanisms that
+antagonize viruses (i.e., the selection mosaic); and (iii) the local
+uniqueness of the bat community, representing the potential for viruses
+to be exposed to novel host traits (e.g., variation in receptor
+sequences). Together, we argue that these can be used to identify and
+map the evolutionary drivers that---in conjunction with transmission
+processes (e.g., viral prevalence in reservoirs and animal-human contact
+rates)--- determine disease emergence risk.
 

### 2022-06-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c3b033a4e022255700ed689fe367e4000ac0daba 🚀

@@
 
-\hypertarget{coevolution-informed-spillover-risk-is-different-in-space}{%
-\subsection{Coevolution-informed spillover risk is different in
-space}\label{coevolution-informed-spillover-risk-is-different-in-space}}
+\hypertarget{coevolution-informed-emergence-risk-is-different-in-space}{%
+\subsection{Coevolution-informed emergence risk is different in
+space}\label{coevolution-informed-emergence-risk-is-different-in-space}}
 

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1b4ddd7d8db0b6e031f2dc59c93240bedff693b2 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@16e453a1bff6288438ff77acaba93a8c85556ae7 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily The coevolutionary mosaic of
-\emph{Betacoronavirus} spillover risk from bat hosts}
+{\Large\bfseries\sffamily The coevolutionary mosaic of betacoronavirus
+emergence risk}
 \vskip 5em
@@
 contact, but also by the underlying evolutionary dynamics in viral
-reservoirs (Plowright et al. 2017). Although host richness is often used
-as a superficial proxy for spillover risk (Anthony et al. 2017,
-Ruiz-Aravena et al. 2022), these approaches oversimplify the relevant
-interspecific heterogeneity in immunology, behavior, and other traits,
-and therefore overlook unique host pools that allow for the rapid
-evolution of highly divergent viruses (Agosta et al. 2010). In the case
-of generalist pathogens like betacoronaviruses, there is conceptual and
-empirical support to the idea that these community-level mechanisms are
-even more important (Power and Mitchell 2004), particularly given that
-cross-species transmission may, as a rule, structure viral evolution
-more than co-divergence with hosts {[}Geoghegan{]}. This creates a
-disconnect between coevolutionary theory (including empirical evidence
-from virology) and most existing ecological frameworks for mapping
-spillover risk.
-
-The Geographic Mosaic Theory of Coevolution (GMTC; Thompson 2005)
+reservoirs.\textsuperscript{1} Although host richness is often used as a
+superficial proxy for spillover risk,\textsuperscript{2,3} these
+approaches oversimplify the relevant interspecific heterogeneity in
+immunology, behavior, and other traits, and therefore overlook unique
+host pools that allow for the rapid evolution of highly divergent
+viruses.\textsuperscript{4} In the case of generalist pathogens like
+betacoronaviruses, there is conceptual and empirical support to the idea
+that these community-level mechanisms are even more
+important,\textsuperscript{5} particularly given that cross-species
+transmission may, as a rule, structure viral evolution more than
+co-divergence with hosts {[}Geoghegan{]}. This creates a disconnect
+between coevolutionary theory (including empirical evidence from
+virology) and most existing ecological frameworks for mapping spillover
+risk.
+
+The Geographic Mosaic Theory of Coevolution\textsuperscript{GMTC; 6}
 attempts to explicitly connect microevolutionary dynamics to the
 macroecology and biogeography of symbiotic interactions. The GMTC posits
-that coevolutionary processes among pairs (Thompson 1994) or complexes
-(Janzen 1980) of species are structured in space by the rippling effects
-of abiotic conditions onto evolutionary mechanism, giving rise to
-fragmented systems with different structure and ecologically dynamics
-over large spatial extents (see \emph{e.g.} Price 2002). The GMTC
-predicts a spatial fragmentation of coevolutionary dynamics under the
-joint action of three processes (see notably Gomulkiewicz et al. 2007):
-coevolutionary hot- and coldspots, which appear when the intensity of
-\emph{interaction} (in terms of reciprocal fitness consequences) varies
-spatially; selection mosaics, wherein the intensity of \emph{selection}
-varies across space, driven by both the biotic complexity of the
-community (locally diverse hosts and viruses are more biotically
-complex) and the local favorability of the environment
-(\textbf{Thrall2007?}); and trait remixing, which occurs when
-coevolutionary dynamics are driven by by the arrival (or departure) of
-\emph{functional traits}, through changes in community composition due
-to invasions, meta-community dynamics, and disperal.
+that coevolutionary processes among pairs\textsuperscript{7} or
+complexes\textsuperscript{8} of species are structured in space by the
+rippling effects of abiotic conditions onto evolutionary mechanism,
+giving rise to fragmented systems with different structure and
+ecologically dynamics over large spatial extents.\textsuperscript{see
+\emph{e.g.} 9} The GMTC predicts a spatial fragmentation of
+coevolutionary dynamics under the joint action of three
+processes:\textsuperscript{see notably 10} coevolutionary hot- and
+coldspots, which appear when the intensity of \emph{interaction} (in
+terms of reciprocal fitness consequences) varies spatially; selection
+mosaics, wherein the intensity of \emph{selection} varies across space,
+driven by both the biotic complexity of the community (locally diverse
+hosts and viruses are more biotically complex) and the local
+favorability of the environment;\textsuperscript{11} and trait remixing,
+which occurs when coevolutionary dynamics are driven by by the arrival
+(or departure) of \emph{functional traits}, through changes in community
+composition due to invasions, meta-community dynamics, and disperal.
 
@@
 of host jumps, recombination among disparate lineages, and, to a lesser
-degree, co-divergence with their hosts {[}Anthony{]}---a mix of
+degree, co-divergence with their hosts---\textsuperscript{2}a mix of
 mechanisms that creates a complex and nonlinear relationship between
@@
 serve as the main animal reservoir for different strains of
-betacoronaviruses (Drexler et al. 2014). This has attracted attention to
+betacoronaviruses.\textsuperscript{12} This has attracted attention to
 areas where high diversity of bats, and therefore presumably high
 diversity of betacoronaviruses, can be an important issue for human
-health (Calisher et al. 2006, Moratelli and Calisher 2015). By
-overlaying the IUCN range maps for confirmed bat hosts of
-betacoronaviruses {[}fig.~\ref{fig:richness}; top{]}, we see that the
-the main hotspots (both in terms of size and higher values) of host
-richness are primarily South-Eastern Asia, parts of Southern Europe, and
-to a lesser extent parts of Africa in the -25-0 range of latitudes. The
-description of host richness is an important first step towards
-understanding risk, as previous research (Anthony et al. 2017, Mollentze
-and Streicker 2020) states that locally diverse bat communities could
-maintain more viruses and hence, a higher probability of having a
-pathogen that could represent a risk for human health.
+health.\textsuperscript{13,14} By overlaying the IUCN range maps for
+confirmed bat hosts of betacoronaviruses {[}fig.~\ref{fig:richness};
+top{]}, we see that the the main hotspots (both in terms of size and
+higher values) of host richness are primarily South-Eastern Asia, parts
+of Southern Europe, and to a lesser extent parts of Africa in the -25-0
+range of latitudes. The description of host richness is an important
+first step towards understanding risk, as previous
+research\textsuperscript{2,15} states that locally diverse bat
+communities could maintain more viruses and hence, a higher probability
+of having a pathogen that could represent a risk for human health.
 
@@
 represent an increased risk of spillover under climate change through
-the creation of novel interactions (Carlson et al. 2022), and therefore
+the creation of novel interactions,\textsuperscript{16} and therefore
 the diversity of \emph{Betacoronavirus} strains should similarly be
@@
 classified, from a macro-evolutionary standpoint, as Yangochiroptera and
-Yinpterochiroptera elsewhere (Teeling et al. 2005, Springer 2013).
-Specifically, we would expect that the so-called ``New World'' group of
-bats, being more evolutionary distinct, would also have evolutionary
-distinct viruses. Indeed fig.~\ref{fig:richness} (bottom) reveals it to
-be the case, and this region harbors a distinct bat-betacoronaviruses
-complex. This can be explained by the fact that Yangochiroptera,
-although not limited to the western hemisphere, contain the highly
-diverse adaptive radiation in the Phyllostomidae (Villalobos and Arita
-2010), which is restricted to the western hemisphere. By contrast,
-South-Eastern Asia has a lot of non-evolutionary distinct bats, who
-nevertheless hosted evolutionary-distinct viruses.
+Yinpterochiroptera elsewhere.\textsuperscript{17,18} Specifically, we
+would expect that the so-called ``New World'' group of bats, being more
+evolutionary distinct, would also have evolutionary distinct viruses.
+Indeed fig.~\ref{fig:richness} (bottom) reveals it to be the case, and
+this region harbors a distinct bat-betacoronaviruses complex. This can
+be explained by the fact that Yangochiroptera, although not limited to
+the western hemisphere, contain the highly diverse adaptive radiation in
+the Phyllostomidae,\textsuperscript{19} which is restricted to the
+western hemisphere. By contrast, South-Eastern Asia has a lot of
+non-evolutionary distinct bats, who nevertheless hosted
+evolutionary-distinct viruses.
 
@@
 betacoronaviruses likely remain to be discovered in this region. Indeed,
-global predictions by Becker et al. (2022) highlight that southeast Asia
+global predictions by\textsuperscript{20} highlight that southeast Asia
 is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
@@
 be a product of under-sampling: South-America is one of the places where
-the fewest \emph{Betacoronavirus} sequences have been discovered
-(Anthony et al. 2017, Olival et al. 2017, Allen et al. 2017), resulting
-in sparser phylogenetic tree, thereby artificially inflating
-distinctiveness. Adding more viruses would bring the distinctiveness of
-known sequences down. Finally, South-America is the range of
-Phyllostomidae, a group of bats that underwent explosive diversification
-events (Villalobos and Arita 2010), which may drive the emergence of
-multiple viral lineages.
+the fewest \emph{Betacoronavirus} sequences have been
+discovered,\textsuperscript{2,21,22} resulting in sparser phylogenetic
+tree, thereby artificially inflating distinctiveness. Adding more
+viruses would bring the distinctiveness of known sequences down.
+Finally, South-America is the range of Phyllostomidae, a group of bats
+that underwent explosive diversification events,\textsuperscript{19}
+which may drive the emergence of multiple viral lineages.
 
@@
 are reasons to expect that the phylogeography of bats and
-betacoronaviruses should show some degree of congruence (Letko et al.
-2020, Van Brussel and Holmes 2022). In particular, this should be the
+betacoronaviruses should show some degree of
+congruence.\textsuperscript{23,24} In particular, this should be the
 case if viruses can circulate among hosts and co-evolve with local hosts
@@
 bat antiviral immune system, resulting in spatially distinct
-immunological responses, as evidenced in several bat species (Banerjee
-et al. 2020). Immune characteristics that allow bats to be better
-adapted to infection by emerging viruses (Gorbunova et al. 2020, Irving
-et al. 2021), in addition to being hardcoded in their genome (Jebb et
-al. 2020), may be related to a wide variety of diets (Banerjee et al.
-2020, Moreno Santillán et al. 2021), themselves likely to be driven by
-spatial effects, especially at the local scale -- bats, indeed, occupy a
-variety of environments, and therefore display a variety of adaptations
-to these environments (Muylaert et al. 2022).
+immunological responses, as evidenced in several bat
+species.\textsuperscript{25} Immune characteristics that allow bats to
+be better adapted to infection by emerging
+viruses,\textsuperscript{26,27} in addition to being hardcoded in their
+genome,\textsuperscript{28} may be related to a wide variety of
+diets,\textsuperscript{25,29} themselves likely to be driven by spatial
+effects, especially at the local scale -- bats, indeed, occupy a variety
+of environments, and therefore display a variety of adaptations to these
+environments.\textsuperscript{30}
 
@@
 nearby location where serological (antibody) evidence has indicated
-human exposure to SARS-like viruses (Wang et al. 2018); MERS-CoV in
+human exposure to SARS-like viruses;\textsuperscript{31} MERS-CoV in
 Saudi Arabia based on index cases available from a recently-published
-compendium of cases (Ramshaw et al. 2019). For the latest event, most if
+compendium of cases.\textsuperscript{32} For the latest event, most if
 not all index cases are presumed to be camel-to-human transmission, and
@@
 uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti (Lednicky et al. 2021) and Europe (Vlasova et al. 2022) are not
+Haiti\textsuperscript{33} and Europe\textsuperscript{34} are not
 relevant here, as bats' involvement in these cycles of transmission have
@@
 co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
-Coevolution (Thompson 2005) to provide a measure of risk accounting for
-multiple processes. In fig.~\ref{fig:trivariate}, we overlapped three
-components of spillover risk: viral sharing, \emph{i.e.} the chance that
-two bats will share viruses overall; Local Contribution to Beta
-Diversity, \emph{i.e.} the fact that a bat community is compositionally
-unique compared to the average compositional similarity across the
-entire system; finally, host phylogenetic diversity, \emph{i.e.} how
-dispersed the bats in a location are within the tree of life. This
-approach leads to the definition of broad biogeographic regions of risk,
-where the same color represents the same type of risk. By way of
-constrat to figures fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo},
-these regions do not necessarilly overlap with previous spatial
-partitions of the bat-betacoronaviruses complex.
+Coevolution\textsuperscript{6} to provide a measure of risk accounting
+for multiple processes. In fig.~\ref{fig:trivariate}, we overlapped
+three components of spillover risk: viral sharing, \emph{i.e.} the
+chance that two bats will share viruses overall; Local Contribution to
+Beta Diversity, \emph{i.e.} the fact that a bat community is
+compositionally unique compared to the average compositional similarity
+across the entire system; finally, host phylogenetic diversity,
+\emph{i.e.} how dispersed the bats in a location are within the tree of
+life. This approach leads to the definition of broad biogeographic
+regions of risk, where the same color represents the same type of risk.
+By way of constrat to figures fig.~\ref{fig:richness} and
+fig.~\ref{fig:biogeo}, these regions do not necessarilly overlap with
+previous spatial partitions of the bat-betacoronaviruses complex.
 
@@
 (inter-genus or higher levels of cross-species transmission) and, to a
-lesser degree, cospeciation and sharing (intra-genus cross-species
-transmission; Anthony et al. 2017). This diversification is not an
-actual risk factor for spillover itself, but acts downstream of a
-spillover event by increasing the random chance of the emergence of a
-virus with the raw genomic components required for the potential to
-infect humans.
+lesser degree, cospeciation and sharing.\textsuperscript{intra-genus
+cross-species transmission; 2} This diversification is not an actual
+risk factor for spillover itself, but acts downstream of a spillover
+event by increasing the random chance of the emergence of a virus with
+the raw genomic components required for the potential to infect humans.
 
@@
 Madagascar, where most bat species are endemic following evolutionary
-divergence from sister species in both African and Asian continents
-(\emph{e.g.} Shi et al. 2014). Recent surveillance (Kettenburg et al.
-2022) has identified a novel \emph{Betacoronavirus} (in the subgenus
-\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
-(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
-strong proof of principle in model predictions.
+divergence from sister species in both African and Asian
+continents.\textsuperscript{\emph{e.g.} 35} Recent
+surveillance\textsuperscript{36} has identified a novel
+\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
+Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
+\emph{Rousettus madagascariensis}), emphasizing strong proof of
+principle in model predictions.
 
@@
 reasonable proxy for the density of humans per unit area, which
-increases the probability of pathogen spread more widely (Hazarie et al.
-2021). Since human activity is required to amplify the frequency of
-virus encounters and thus create areas of viral amplification, mapping
-the potential risk against measures of land use is required to generate
-a more actionable assessment of risk. This map is presented in
-fig.~\ref{fig:compound}. Most of South America and Europe are at
-comparatively lower risk, as although densely populated, settlements
-tend to be in areas with lower potential risk. Regions like Malaysia and
-the North coast of Australia have a high risk component, but should
-represent a relatively lower effective risk due to low human density.
-However, this mapping reveals that South-East Asia, the Indian
+increases the probability of pathogen spread more
+widely.\textsuperscript{37} Since human activity is required to amplify
+the frequency of virus encounters and thus create areas of viral
+amplification, mapping the potential risk against measures of land use
+is required to generate a more actionable assessment of risk. This map
+is presented in fig.~\ref{fig:compound}. Most of South America and
+Europe are at comparatively lower risk, as although densely populated,
+settlements tend to be in areas with lower potential risk. Regions like
+Malaysia and the North coast of Australia have a high risk component,
+but should represent a relatively lower effective risk due to low human
+density. However, this mapping reveals that South-East Asia, the Indian
 subcontinent, and parts of sub-Saharan Africa, are at high risk due to
@@
 opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China, Xu et al. (2004) present
+looking for the origins of SARS in China,\textsuperscript{38} present
 serological evidence that strongest human-animal contact results in
@@
 amount of risk represented by each activity. Our map of risk overlays
-with recent results from Cohen et al. (2022) -- areas of purported high
+with recent results from\textsuperscript{39} -- areas of purported high
 risk/diversitifcation potential (Madagascar, South-America) overlay with
@@
 Bats are important reservoir hosts for different classes of
-microorganisms, many of which a threat to human health (Letko et al.
-2020, Van Brussel and Holmes 2022). Chiropterans emerged around 64
-million years ago and are one of the most diverse mammalian orders, with
-an estimated richness of more than 1400 species (Peixoto et al. 2018,
-Simmons and Cirranello 2020). They exhibit a broad variety of habitat
-use, behaviour, and feeding strategies, putting them at key positions in
-the delivery and provisioning of several ecosystem services, tied to
-important ecosystem-derived benefits to human (Kasso and Balakrishnan
-2013). For example, bats are an essential component of many
-seed-dispersal networks (Mello et al. 2011). Over two-thirds of bats are
-know to be either obligate or facultative insectivores, therefore
-actively contributing for agricultural pest control (Williams-Guillén et
-al. 2008, Voigt and Kingston 2016), and vectors of pathogens that put a
-risk on human health (Gonsalves et al. 2013a, b). Because bats are
-globally distributed and have a long evolutionary history,
-phylogeographic and biogeographic approaches are required to shed light
-on the contemporary distribution of coevolutionary processes between
-bats and the pathogens they host. Not all areas in which bats, viruses,
-and human are co-occuring are facing a risk of spillover towards human
-populations, and the areas in which this risk exist may not be facing
-risks of the same nature and magnitude.
+microorganisms, many of which a threat to human
+health.\textsuperscript{23,24} Chiropterans emerged around 64 million
+years ago and are one of the most diverse mammalian orders, with an
+estimated richness of more than 1400 species.\textsuperscript{40,41}
+They exhibit a broad variety of habitat use, behaviour, and feeding
+strategies, putting them at key positions in the delivery and
+provisioning of several ecosystem services, tied to important
+ecosystem-derived benefits to human.\textsuperscript{42} For example,
+bats are an essential component of many seed-dispersal
+networks.\textsuperscript{43} Over two-thirds of bats are know to be
+either obligate or facultative insectivores, therefore actively
+contributing for agricultural pest control,\textsuperscript{44,45} and
+vectors of pathogens that put a risk on human
+health.\textsuperscript{46,47} Because bats are globally distributed and
+have a long evolutionary history, phylogeographic and biogeographic
+approaches are required to shed light on the contemporary distribution
+of coevolutionary processes between bats and the pathogens they host.
+Not all areas in which bats, viruses, and human are co-occuring are
+facing a risk of spillover towards human populations, and the areas in
+which this risk exist may not be facing risks of the same nature and
+magnitude.
 
@@
 that have used host richness as proxy have predicted a high diversity of
-unsampled bat viruses {[}Olival{]}, bat coronaviruses {[}Anthony{]}, and
-even specifically betacoronaviruses {[}Becker{]} in both the Amazon and
-southeast Asia. While we find that both regions are characterized by
-highly divergent host and viral communities, our framework identifies
-key differences between the regions. We find that Latin America is a
-hotspot of both host and viral distinctiveness, suggesting that this
-branch of the bat-betacoronavirus complex may be undergoing independent
+unsampled bat viruses,\textsuperscript{22} bat
+coronaviruses,\textsuperscript{2} and even specifically
+betacoronaviruses\textsuperscript{20} in both the Amazon and southeast
+Asia. While we find that both regions are characterized by highly
+divergent host and viral communities, our framework identifies key
+differences between the regions. We find that Latin America is a hotspot
+of both host and viral distinctiveness, suggesting that this branch of
+the bat-betacoronavirus complex may be undergoing independent
 evolutionary dynamics from the rest of the global pool, but with limited
 potential for viral diversification--- a finding that is supported by
-previous work indicating a higher rate of codivergence in Latin America
-{[}Anthony{]}. In contrast, in southeast Asia, host richness and viral
-distinctiveness are high but sharing is low; this suggests a different
-type of evolutionary dynamics that could generate high local diversity
-of viruses through host switching and viral recombination (see
-\emph{e.g.} (Latinne et al. 2020), as well as the discovery of
-recombinant viruses that share genetic material from both the SARS-CoV
-and SARS-CoV-2 branches of the Sarbecovirus lineage {[}Wu{]}). Both of
-these regions are priority areas for sampling, especially given
-predictions by Becker et al. (2022) that they contain many bat hosts of
-undiscovered betacoronaviruses. However, both the evolutionary and
-ecological aspects of emergence risk are likely higher in southeast
-Asia---a fact that will only become more relevant, as bats track
-shifting climates and exchange viruses with other species, creating a
-hotspot of cross-species transmission unique to the region
-{[}Carlson{]}.
+previous work indicating a higher rate of codivergence in Latin
+America.\textsuperscript{2} In contrast, in southeast Asia, host
+richness and viral distinctiveness are high but sharing is low; this
+suggests a different type of evolutionary dynamics that could generate
+high local diversity of viruses through host switching and viral
+recombination (see \emph{e.g.},\textsuperscript{48} as well as the
+discovery of recombinant viruses that share genetic material from both
+the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
+lineage).\textsuperscript{49} Both of these regions are priority areas
+for sampling, especially given predictions that they contain many bat
+hosts of undiscovered betacoronaviruses.\textsuperscript{20,39} However,
+both the evolutionary and ecological aspects of emergence risk are
+likely higher in southeast Asia---a fact that will only become more
+relevant, as bats track shifting climates and exchange viruses with
+other species, creating a hotspot of cross-species transmission unique
+to the region.\textsuperscript{16}
 
 {[}Tim tinker with para about conservation{]} There are several factors
-that drive changes in the diversity of bats (Alves et al. 2018), but
+that drive changes in the diversity of bats,\textsuperscript{50} but
 human activities' effects on the ecosystem (like modifications of land
@@
 public health. With the increase of contact between humans and potential
-hosts, we also increase the risk of emergence of novel diseases (Johnson
-et al. 2020), as previous studies on RNA viruses suggest the importance
-of host phylogeography at the time of virus dispersal (Gryseels et al.
-2017). One of these scenarios where interaction between bats and humans
-can occur can be seed dispersal in tropical agroecosystems. It opens the
-discussion of whether the fruits thrown by bats not only disperse seeds
-but could also be a source of indirect interaction between viruses of
-bat origin and humans (Deshpande et al. 2022). This represents a
-challenge for conservation strategies and disease ecology since some
-areas can haveboth potential for the acquisition of zoonotic viruses and
-bat-human interactions; in particular, the challenge lies in the fact
-that actual exposure must then be quantified accounting for several
-transmission scenarios, including both direct and indirect bat - human
-interaction.
+hosts, we also increase the risk of emergence of novel
+diseases,\textsuperscript{51} as previous studies on RNA viruses suggest
+the importance of host phylogeography at the time of virus
+dispersal.\textsuperscript{52} One of these scenarios where interaction
+between bats and humans can occur can be seed dispersal in tropical
+agroecosystems. It opens the discussion of whether the fruits thrown by
+bats not only disperse seeds but could also be a source of indirect
+interaction between viruses of bat origin and
+humans.\textsuperscript{53} This represents a challenge for conservation
+strategies and disease ecology since some areas can haveboth potential
+for the acquisition of zoonotic viruses and bat-human interactions; in
+particular, the challenge lies in the fact that actual exposure must
+then be quantified accounting for several transmission scenarios,
+including both direct and indirect bat - human interaction.
 
@@
 We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
-by Becker et al. (2022) from
+by\textsuperscript{20} from
 \texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
@@
 classified as an empirically documented host of \emph{Betacoronavirus}
-from the previous step, according to recent IUCN data (IUCN 2021). The
-range maps were subsequently rasterized using the \texttt{rasterize}
-function from \texttt{GDAL} (Rouault et al. 2022) at a resolution of
-approximately 100kmx100km. For every pixel in the resulting raster where
-at least one bat host of \emph{Betacoronavirus} was present, we extract
-the species pool (list of all known bat hosts), which was used to
-calculate the following risk assessment components: bat phylogenetic
-diversity, bat compositional uniqueness, and predicted viral sharing
-risk.
+from the previous step, according to recent IUCN
+data.\textsuperscript{54} The range maps were subsequently rasterized
+using the \texttt{rasterize} function from
+\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
+100kmx100km. For every pixel in the resulting raster where at least one
+bat host of \emph{Betacoronavirus} was present, we extract the species
+pool (list of all known bat hosts), which was used to calculate the
+following risk assessment components: bat phylogenetic diversity, bat
+compositional uniqueness, and predicted viral sharing risk.
 
@@
 
-For every pixel, we measured Faith's Phylogenetic Diversity (Faith 1992)
-based on a recent synthetic tree with robust time calibration, covering
-about 6000 mammalian species (Upham et al. 2019). Faith's PD measures
-the sum of unique branches from an arbitrary root to a set of tips, and
-comparatively larger values indicate a more phylogenetic diverse species
-pool. We measured phylogenetic diversity starting from the root of the
-entire tree (and not from Chiroptera); this bears no consequences on the
+For every pixel, we measured Faith's Phylogenetic
+Diversity\textsuperscript{56} based on a recent synthetic tree with
+robust time calibration, covering about 6000 mammalian
+species.\textsuperscript{57} Faith's PD measures the sum of unique
+branches from an arbitrary root to a set of tips, and comparatively
+larger values indicate a more phylogenetic diverse species pool. We
+measured phylogenetic diversity starting from the root of the entire
+tree (and not from Chiroptera); this bears no consequences on the
 resulting values, since all branches leading up to Chiroptera are only
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity (Legendre and De Cáceres 2013); LCBD works from a
-species-data matrix (traditionally noted as \(\mathbf{Y}\)), where
-species are rows and sites are columns, and a value of 1 indicates
-occurrence. We extracted the Y matrix assuming that every pixel
-represents a unique location, and following best practices (Legendre and
-Condit 2019) transformed it using Hellinger's distance to account for
-unequal bat richness at different pixels. The correction of raw
-community data is particularly important for two reasons: first, it
-prevents the artifact of richer sites having higher importance; second,
-it removes the effect of overall species richness, which is already
-incorporated in the phylogenetic diversity component. High values of
-LCBD indicate that the pixel has a community that is on average more
-dissimilar in species composition than what is expected knowing the
-entire matrix, i.e.~a more unique community. Recent results by Dansereau
-et al. (2022) shows that LCBD measures are robust with regards to
-spatial scale, and are therefore applicable at the global scale.
+Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
+matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
+and sites are columns, and a value of 1 indicates occurrence. We
+extracted the Y matrix assuming that every pixel represents a unique
+location, and following best practices\textsuperscript{59} transformed
+it using Hellinger's distance to account for unequal bat richness at
+different pixels. The correction of raw community data is particularly
+important for two reasons: first, it prevents the artifact of richer
+sites having higher importance; second, it removes the effect of overall
+species richness, which is already incorporated in the phylogenetic
+diversity component. High values of LCBD indicate that the pixel has a
+community that is on average more dissimilar in species composition than
+what is expected knowing the entire matrix, i.e.~a more unique
+community. Recent results by\textsuperscript{60} shows that LCBD
+measures are robust with regards to spatial scale, and are therefore
+applicable at the global scale.
 
@@
 generalized additive mixed model of virus sharing by a tensor function
-of phylogenetic distance and geographic range overlap (Albery et al.
-2020). This network stores pairwise values of viral community
-similarity. To project viral sharing values into a single value for
-every pixel, we averaged the pairwise scores. High values of the average
-sharing propensity means that this specific extant bat assemblage is
-likely to be proficient at exchanging viruses.
+of phylogenetic distance and geographic range
+overlap.\textsuperscript{61} This network stores pairwise values of
+viral community similarity. To project viral sharing values into a
+single value for every pixel, we averaged the pairwise scores. High
+values of the average sharing propensity means that this specific extant
+bat assemblage is likely to be proficient at exchanging viruses.
 
@@
 three individual risk components (phylogenetic diversity, compositional
-uniqueness, and viral sharing) using an additive color model (Seekell et
-al. 2018). In this approach, every risk component gets assigned a
-component in the RGB color model (phylogenetic diversity is green,
-compositional uniqueness is red, and viral sharing is blue). In order to
-achieve a valid RGB measure, all components are re-scaled to the
-{[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
+uniqueness, and viral sharing) using an additive color
+model.\textsuperscript{62} In this approach, every risk component gets
+assigned a component in the RGB color model (phylogenetic diversity is
+green, compositional uniqueness is red, and viral sharing is blue). In
+order to achieve a valid RGB measure, all components are re-scaled to
+the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
 diversity, and no compositional uniqueness is black, and a pixel with
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing (Gomulkiewicz et al. 2000, Cavender-Bares et al.
-2009), in that this allows multiple independent host-virus
-coevolutionary dynamics to take place in the same location. In the
-colorimetric space, this correspond to yellow -- because the HSV space
-is more amenable to calculations for feature extraction (see \emph{e.g.}
-Keke et al. 2010), we measured the risk level by calculating the angular
-distance of the hue of each pixel to a reference value of 60, and
-weighted this risk level by the value component. Specifically, given a
-pixel with colorimetric coordinates \((h,s,v)\), its ranged weighted
-risk value is
+and low viral sharing,\textsuperscript{63,64} in that this allows
+multiple independent host-virus coevolutionary dynamics to take place in
+the same location. In the colorimetric space, this correspond to yellow
+-- because the HSV space is more amenable to calculations for feature
+extraction,\textsuperscript{see \emph{e.g.} 65} we measured the risk
+level by calculating the angular distance of the hue of each pixel to a
+reference value of 60, and weighted this risk level by the value
+component. Specifically, given a pixel with colorimetric coordinates
+\((h,s,v)\), its ranged weighted risk value is
 
@@
 the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
-indicating recombinant or laboratory strains including ``patent,''
-``mutant,'' ``GFP,'' and ``recombinant.'' We filtered over-represented
+indicating recombinant or laboratory strains including ``patent'',
+``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
 taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
@@
 betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
-{[}Katoh and Standley (2013); Algorithm FFT-NS-2, Scoring matrix 200PAM
-/ k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
-likelihood tree reconstructed in IQ-TREE v 1.6.12 (Nguyen et al. 2015)
-with ModelFinder (Kalyaanamoorthy et al. 2017) ultrafast bootstrap
-approximation (Hoang et al. 2018) with a general time reversible model
+{[}\textsuperscript{66}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
+k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
+likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{67}
+with ModelFinder\textsuperscript{68} ultrafast bootstrap
+approximation\textsuperscript{69} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness (Isaac et al. 2007) for each of the viruses in the tree,
+distinctiveness\textsuperscript{70} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from (Kreft and Jetz 2007, 2010),
-who proposed a phylogenetic method for the delineation of animal
+this idea, we loosely adapted a method from,\textsuperscript{71,72} who
+proposed a phylogenetic method for the delineation of animal
 biogeographic regions. In their original method, a distance matrix -
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
-subsequently, we take a principal components analysis of that distance
-matrix (readily interchangeable for NMDS in this case) to project the
-viral tree into an n-dimensional space. We then take the first two
-principal components and, as with the evolutionary distinctiveness
-analysis, aggregated these to a mean host value and projected them using
-a four-color bivariate map.
+cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
+we take a principal components analysis of that distance matrix (readily
+interchangeable for NMDS in this case) to project the viral tree into an
+n-dimensional space. We then take the first two principal components
+and, as with the evolutionary distinctiveness analysis, aggregated these
+to a mean host value and projected them using a four-color bivariate
+map.
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
+\begin{CSLReferences}{0}{0}
+\leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%
+\CSLLeftMargin{1. }
+\CSLRightInline{Plowright, R. K. \emph{et al.} Pathways to zoonotic
+spillover. \emph{Nature Reviews Microbiology} \textbf{15}, 502--510
+(2017).}
+
+\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
+\CSLLeftMargin{2. }
+\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
+coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}
+
+\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
+\CSLLeftMargin{3. }
+\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
+spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
+\textbf{20}, 299--314 (2022).}
+
 \leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
-Agosta, S. J. et al. 2010. How specialists can be generalists: Resolving
-the {``parasite paradox''} and implications for emerging infectious
-disease. - Zoologia (Curitiba) 27: 151--162.
+\CSLLeftMargin{4. }
+\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
+can be generalists: Resolving the {`parasite paradox'} and implications
+for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
+151--162 (2010).}
+
+\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
+\CSLLeftMargin{5. }
+\CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
+Disease Epidemics. \emph{The American Naturalist} \textbf{164}, S79--S89
+(2004).}
+
+\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
+\CSLLeftMargin{6. }
+\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
+Coevolution}. (University Of Chicago Press, 2005).}
+
+\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
+\CSLLeftMargin{7. }
+\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
+(University of Chicago Press, 1994).}
+
+\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
+\CSLLeftMargin{8. }
+\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
+\textbf{34}, 611--612 (1980).}
+
+\leavevmode\hypertarget{ref-Price2002MacThe}{}%
+\CSLLeftMargin{9. }
+\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
+Macroecological Patterns}. (Cambridge University Press, 2002).
+doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}
+
+\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
+\CSLLeftMargin{10. }
+\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
+the geographic mosaic theory of coevolution. \emph{Heredity}
+\textbf{98}, 249--258 (2007).}
+
+\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
+\CSLLeftMargin{11. }
+\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
+J. D. Coevolution of symbiotic mutualists and parasites in a community
+context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
+(2007).}
+
+\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
+\CSLLeftMargin{12. }
+\CSLRightInline{Drexler, J. F., Corman, V. M. \& Drosten, C. Ecology,
+evolution and classification of bat coronaviruses in the aftermath of
+SARS. \emph{Antiviral Research} \textbf{101}, 45--56 (2014).}
+
+\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
+\CSLLeftMargin{13. }
+\CSLRightInline{Calisher, C. H., Childs, J. E., Field, H. E., Holmes, K.
+V. \& Schountz, T. Bats: Important Reservoir Hosts of Emerging Viruses.
+\emph{Clinical Microbiology Reviews} \textbf{19}, 531--545 (2006).}
+
+\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
+\CSLLeftMargin{14. }
+\CSLRightInline{Moratelli, R. \& Calisher, C. H. Bats and zoonotic
+viruses: Can we confidently link bats with emerging deadly viruses?
+\emph{Memórias do Instituto Oswaldo Cruz} \textbf{110}, 1--22 (2015).}
+
+\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
+\CSLLeftMargin{15. }
+\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
+homogenous among taxonomic orders of mammalian and avian reservoir
+hosts. \emph{Proceedings of the National Academy of Sciences}
+\textbf{117}, 9423--9430 (2020).}
+
+\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
+\CSLLeftMargin{16. }
+\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
+cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
+doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
+
+\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
+\CSLLeftMargin{17. }
+\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
+Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}
+
+\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
+\CSLLeftMargin{18. }
+\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
+Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
+York, N.Y.)} \textbf{307}, 580--584 (2005).}
+
+\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
+\CSLLeftMargin{19. }
+\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
+New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
+Biogeography} \textbf{19}, 200--211 (2010).}
+
+\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
+\CSLLeftMargin{20. }
+\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
+to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
+Microbe} (2022)
+doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}
+
+\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
+\CSLLeftMargin{21. }
+\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
+of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
+(2017).}
+
+\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
+\CSLLeftMargin{22. }
+\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
+predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
+646--650 (2017).}
+
+\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
+\CSLLeftMargin{23. }
+\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
+virome diversity in bats. \emph{Current Opinion in Virology}
+\textbf{52}, 192--202 (2022).}
+
+\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
+\CSLLeftMargin{24. }
+\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
+K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
+\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}
+
+\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
+\CSLLeftMargin{25. }
+\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
+Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}
+
+\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
+\CSLLeftMargin{26. }
+\CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
+Goes Bats: Living Longer and Tolerating Viruses. \emph{Cell Metabolism}
+\textbf{32}, 31--43 (2020).}
+
+\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
+\CSLLeftMargin{27. }
+\CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
+Wang, L.-F. Lessons from the host defences of bats, a unique viral
+reservoir. \emph{Nature} \textbf{589}, 363--370 (2021).}
+
+\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
+\CSLLeftMargin{28. }
+\CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
+reveal evolution of bat adaptations. \emph{Nature} \textbf{583},
+578--584 (2020).}
+
+\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
+\CSLLeftMargin{29. }
+\CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
+sampling reveals unique immunity and metabolic adaptations in bats.
+\emph{Molecular Ecology} mec.16027 (2021)
+doi:\href{https://doi.org/10.1111/mec.16027}{10.1111/mec.16027}.}
+
+\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
+\CSLLeftMargin{30. }
+\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
+distribution of bat hosts of sarbecoviruses: Implications for
+conservation and public health. \emph{Proceedings of the Royal Society
+B: Biological Sciences} \textbf{289}, 20220397 (2022).}
+
+\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
+\CSLLeftMargin{31. }
+\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
+SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
+Sinica} \textbf{33}, 104--107 (2018).}
+
+\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
+\CSLLeftMargin{32. }
+\CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
+Middle East Respiratory Syndrome Coronavirus occurrences.
+\emph{Scientific Data} \textbf{6}, 318 (2019).}
+
+\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
+\CSLLeftMargin{33. }
+\CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
+Recombinant Canine Coronavirus from a Visitor to Haiti: Further Evidence
+of Transmission of Coronaviruses of Zoonotic Origin to Humans.
+\emph{Clinical Infectious Diseases: An Official Publication of the
+Infectious Diseases Society of America} ciab924 (2021)
+doi:\href{https://doi.org/10.1093/cid/ciab924}{10.1093/cid/ciab924}.}
+
+\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
+\CSLLeftMargin{34. }
+\CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
+found in human patients with acute respiratory illness in different
+countries. \emph{Emerging Microbes \& Infections} \textbf{11}, 699--702
+(2022).}
+
+\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
+\CSLLeftMargin{35. }
+\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
+Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
+Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}
+
+\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
+\CSLLeftMargin{36. }
+\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
+Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
+for This Coronavirus Clade. \emph{Frontiers in Public Health}
+\textbf{10}, (2022).}
+
+\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
+\CSLLeftMargin{37. }
+\CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
+Gómez-Gardeñes, J. \& Ghoshal, G. Interplay between population density
+and mobility in determining the spread of epidemics in cities.
+\emph{Communications Physics} \textbf{4}, 191 (2021).}
+
+\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
+\CSLLeftMargin{38. }
+\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
+Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
+1030--1037 (2004).}
+
+\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
+\CSLLeftMargin{39. }
+\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
+Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
+bat coronaviruses. 2022.06.15.496296 (2022)
+doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}
+
+\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
+\CSLLeftMargin{40. }
+\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
+synthesis of ecological and evolutionary determinants of bat diversity
+across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}
+
+\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
+\CSLLeftMargin{41. }
+\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
+World: A taxonomic and geographic database. \url{https://batnames.org/}
+(2020).}
+
+\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
+\CSLLeftMargin{42. }
+\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
+Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
+\textbf{2013}, e187415 (2013).}
+
+\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
+\CSLLeftMargin{43. }
+\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
+Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
+\emph{PLOS ONE} \textbf{6}, e17395 (2011).}
+
+\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
+\CSLLeftMargin{44. }
+\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
+a Changing World}. (Springer International Publishing, 2016).
+doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}
+
+\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
+\CSLLeftMargin{45. }
+\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
+Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
+\textbf{320}, 70--70 (2008).}
+
+\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
+\CSLLeftMargin{46. }
+\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
+Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
+\emph{PLOS ONE} \textbf{8}, e77183 (2013).}
+
+\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
+\CSLLeftMargin{47. }
+\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
+Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
+Research} \textbf{40}, 10--24 (2013).}
+
+\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
+\CSLLeftMargin{48. }
+\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
+transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
+Server for Biology} 2020.05.31.116061 (2020)
+doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
+
+\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
+\CSLLeftMargin{49. }
+\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
+sarbecoviruses across China for the origin tracing of SARS-CoV and
+SARS-CoV-2}. (2021)
+doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}
+
+\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
+\CSLLeftMargin{50. }
+\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
+Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
+relationship between large-scale environmental determinants and bat
+species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
+(2018).}
+
+\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
+\CSLLeftMargin{51. }
+\CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
+population trends reveal key predictors of virus spillover risk.
+\emph{Proceedings of the Royal Society B: Biological Sciences}
+\textbf{287}, 20192736 (2020).}
+
+\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
+\CSLLeftMargin{52. }
+\CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
+The Importance of Host Phylogeographic Structure in the Spatial Spread
+of Arenaviruses. \emph{PLOS Pathogens} \textbf{13}, e1006073 (2017).}
+
+\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
+\CSLLeftMargin{53. }
+\CSLRightInline{Deshpande, K., Vanak, A. T., Devy, M. S. \&
+Krishnaswamy, J. Forbidden fruits? Ecosystem services from seed
+dispersal by fruit bats in the context of latent zoonotic risk.
+\emph{Oikos (Copenhagen, Denmark)} oik.08359 (2022)
+doi:\href{https://doi.org/10.1111/oik.08359}{10.1111/oik.08359}.}
+
+\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
+\CSLLeftMargin{54. }
+\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
+\url{https://www.iucnredlist.org} (2021).}
+
+\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
+\CSLLeftMargin{55. }
+\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
+Abstraction software Library}. (Zenodo, 2022).
+doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}
+
+\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
+\CSLLeftMargin{56. }
+\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
+diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}
+
+\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
+\CSLLeftMargin{57. }
+\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
+mammal tree: Species-level sets of phylogenies for questions in ecology,
+evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
+(2019).}
+
+\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
+\CSLLeftMargin{58. }
+\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
+variance of community data: Dissimilarity coefficients and partitioning.
+\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}
+
+\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
+\CSLLeftMargin{59. }
+\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
+of beta diversity in the Barro Colorado Island forest dynamics plot,
+Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}
+
+\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
+\CSLLeftMargin{60. }
+\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
+ecological uniqueness over broad spatial extents using species
+distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}
 
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-Albery, G. F. et al. 2020. Predicting the global mammalian viral sharing
-network using phylogeography. - Nature Communications 11: 2260.
-
-\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
-Allen, T. et al. 2017. Global hotspots and correlates of emerging
-zoonotic diseases. - Nature Communications in press.
-
-\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-Alves, D. M. C. C. et al. 2018. Geographic variation in the relationship
-between large-scale environmental determinants and bat species richness.
-- Basic and Applied Ecology 27: 1--8.
-
-\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
-Anthony, S. J. et al. 2017. Global patterns in coronavirus diversity. -
-Virus Evolution in press.
-
-\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-Banerjee, A. et al. 2020. Novel Insights Into Immune Systems of Bats. -
-Frontiers in Immunology 11: 26.
-
-\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-Becker, D. J. et al. 2022. Optimising predictive models to prioritise
-viral discovery in zoonotic reservoirs. - The Lancet Microbe in press.
-
-\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
-Calisher, C. H. et al. 2006. Bats: Important Reservoir Hosts of Emerging
-Viruses. - Clinical Microbiology Reviews 19: 531--545.
-
-\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-Carlson, C. J. et al. 2022. Climate change increases cross-species viral
-transmission risk. - Nature: 1--1.
+\CSLLeftMargin{61. }
+\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
+Predicting the global mammalian viral sharing network using
+phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}
+
+\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
+\CSLLeftMargin{62. }
+\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
+geography of lake carbon cycling. \emph{Limnology and Oceanography
+Letters} \textbf{3}, 49--56 (2018).}
 
 \leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
-Cavender-Bares, J. et al. 2009. The merging of community ecology and
-phylogenetic biology. - Ecol. Lett. 12: 693--715.
-
-\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-Cohen, L. E. et al. 2022. Sampling strategies and pre-pandemic
-surveillance gaps for bat coronaviruses.: 2022.06.15.496296.
-
-\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-Dansereau, G. et al. 2022. Evaluating ecological uniqueness over broad
-spatial extents using species distribution modelling. - Oikos n/a:
-e09063.
-
-\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
-Deshpande, K. et al. 2022. Forbidden fruits? Ecosystem services from
-seed dispersal by fruit bats in the context of latent zoonotic risk. -
-Oikos (Copenhagen, Denmark): oik.08359.
-
-\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
-Drexler, J. F. et al. 2014. Ecology, evolution and classification of bat
-coronaviruses in the aftermath of SARS. - Antiviral Research 101:
-45--56.
-
-\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-Faith, D. P. 1992. Conservation evaluation and phylogenetic diversity. -
-Biological Conservation 61: 1--10.
+\CSLLeftMargin{63. }
+\CSLRightInline{Cavender-Bares, J., Kozak, K. H., Fine, P. V. A. \&
+Kembel, S. W. The merging of community ecology and phylogenetic biology.
+\emph{Ecol. Lett.} \textbf{12}, 693--715 (2009).}
 
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-Gomulkiewicz, R. et al. 2000. Hot Spots, Cold Spots, and the Geographic
-Mosaic Theory of Coevolution. - The American Naturalist 156: 156--174.
-
-\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
-Gomulkiewicz, R. et al. 2007. Dos and don'ts of testing the geographic
-mosaic theory of coevolution. - Heredity 98: 249--258.
-
-\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-Gonsalves, L. et al. 2013a. Do mosquitoes influence bat activity in
-coastal habitats? - Wildlife Research 40: 10--24.
-
-\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-Gonsalves, L. et al. 2013b. Mosquito Consumption by Insectivorous Bats:
-Does Size Matter? - PLOS ONE 8: e77183.
-
-\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
-Gorbunova, V. et al. 2020. The World Goes Bats: Living Longer and
-Tolerating Viruses. - Cell Metabolism 32: 31--43.
-
-\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-Gryseels, S. et al. 2017. When Viruses Don't Go Viral: The Importance of
-Host Phylogeographic Structure in the Spatial Spread of Arenaviruses (JH
-Kuhn, Ed.). - PLOS Pathogens 13: e1006073.
-
-\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
-Hazarie, S. et al. 2021. Interplay between population density and
-mobility in determining the spread of epidemics in cities. -
-Communications Physics 4: 191.
+\CSLLeftMargin{64. }
+\CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
+S. L. \& Hochberg, M. E. Hot Spots, Cold Spots, and the Geographic
+Mosaic Theory of Coevolution. \emph{The American Naturalist}
+\textbf{156}, 156--174 (2000).}
+
+\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
+\CSLLeftMargin{65. }
+\CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
+image segmentation used by Fuzzy-c-means arithmetic. in \emph{2010
+Seventh International Conference on Fuzzy Systems and Knowledge
+Discovery} vol. 2 612--615 (2010).}
+
+\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
+\CSLLeftMargin{66. }
+\CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
+Alignment Software Version 7: Improvements in Performance and Usability.
+\emph{Molecular Biology and Evolution} \textbf{30}, 772--780 (2013).}
+
+\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
+\CSLLeftMargin{67. }
+\CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
+B. Q. IQ-TREE: A Fast and Effective Stochastic Algorithm for Estimating
+Maximum-Likelihood Phylogenies. \emph{Molecular Biology and Evolution}
+\textbf{32}, 268--274 (2015).}
+
+\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
+\CSLLeftMargin{68. }
+\CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
+Haeseler, A. \& Jermiin, L. S. ModelFinder: Fast model selection for
+accurate phylogenetic estimates. \emph{Nature Methods} \textbf{14},
+587--589 (2017).}
 
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-Hoang, D. T. et al. 2018. UFBoot2: Improving the Ultrafast Bootstrap
-Approximation. - Molecular Biology and Evolution 35: 518--522.
-
-\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
-Irving, A. T. et al. 2021. Lessons from the host defences of bats, a
-unique viral reservoir. - Nature 589: 363--370.
+\CSLLeftMargin{69. }
+\CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
+Q. \& Vinh, L. S. UFBoot2: Improving the Ultrafast Bootstrap
+Approximation. \emph{Molecular Biology and Evolution} \textbf{35},
+518--522 (2018).}
 
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-Isaac, N. J. B. et al. 2007. Mammals on the EDGE: Conservation
-Priorities Based on Threat and Phylogeny. - PLOS ONE 2: e296.
-
-\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-IUCN 2021. The IUCN Red List of Threatened Species.
-
-\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
-Janzen, D. H. 1980. When is it Coevolution? - Evolution 34: 611--612.
-
-\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
-Jebb, D. et al. 2020. Six reference-quality genomes reveal evolution of
-bat adaptations. - Nature 583: 578--584.
-
-\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-Johnson, C. K. et al. 2020. Global shifts in mammalian population trends
-reveal key predictors of virus spillover risk. - Proceedings of the
-Royal Society B: Biological Sciences 287: 20192736.
-
-\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-Kalyaanamoorthy, S. et al. 2017. ModelFinder: Fast model selection for
-accurate phylogenetic estimates. - Nature Methods 14: 587--589.
-
-\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-Kasso, M. and Balakrishnan, M. 2013. Ecological and Economic Importance
-of Bats (Order Chiroptera). - ISRN Biodiversity 2013: e187415.
-
-\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-Katoh, K. and Standley, D. M. 2013. MAFFT Multiple Sequence Alignment
-Software Version 7: Improvements in Performance and Usability. -
-Molecular Biology and Evolution 30: 772--780.
-
-\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-Keke, S. et al. 2010. Study on skin color image segmentation used by
-Fuzzy-c-means arithmetic. - 2010 Seventh International Conference on
-Fuzzy Systems and Knowledge Discovery 2: 612--615.
-
-\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-Kettenburg, G. et al. 2022. Full Genome Nobecovirus Sequences From
-Malagasy Fruit Bats Define a Unique Evolutionary History for This
-Coronavirus Clade. - Frontiers in Public Health in press.
+\CSLLeftMargin{70. }
+\CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
+\& Baillie, J. E. M. Mammals on the EDGE: Conservation Priorities Based
+on Threat and Phylogeny. \emph{PLOS ONE} \textbf{2}, e296 (2007).}
 
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-Kreft, H. and Jetz, W. 2007. Global patterns and determinants of
-vascular plant diversity. - Proceedings of the National Academy of
-Sciences 104: 5925--5930.
+\CSLLeftMargin{71. }
+\CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
+of vascular plant diversity. \emph{Proceedings of the National Academy
+of Sciences} \textbf{104}, 5925--5930 (2007).}
 
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-Kreft, H. and Jetz, W. 2010. A framework for delineating biogeographical
-regions based on species distributions. - Journal of Biogeography 37:
-2029--2053.
-
-\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-Latinne, A. et al. 2020. Origin and cross-species transmission of bat
-coronaviruses in China. - bioRxiv: The Preprint Server for Biology:
-2020.05.31.116061.
-
-\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-Lednicky, J. A. et al. 2021. Isolation of a Novel Recombinant Canine
-Coronavirus from a Visitor to Haiti: Further Evidence of Transmission of
-Coronaviruses of Zoonotic Origin to Humans. - Clinical Infectious
-Diseases: An Official Publication of the Infectious Diseases Society of
-America: ciab924.
-
-\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-Legendre, P. and De Cáceres, M. 2013. Beta diversity as the variance of
-community data: Dissimilarity coefficients and partitioning (H Morlon,
-Ed.). - Ecology Letters 16: 951--963.
-
-\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-Legendre, P. and Condit, R. 2019. Spatial and temporal analysis of beta
-diversity in the Barro Colorado Island forest dynamics plot, Panama. -
-Forest Ecosystems 6: 7.
-
-\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-Letko, M. et al. 2020. Bat-borne virus diversity, spillover and
-emergence. - Nature Reviews Microbiology 18: 461--471.
-
-\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-Mello, M. A. R. et al. 2011. The Missing Part of Seed Dispersal
-Networks: Structure and Robustness of Bat-Fruit Interactions. - PLOS ONE
-6: e17395.
-
-\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
-Mollentze, N. and Streicker, D. G. 2020. Viral zoonotic risk is
-homogenous among taxonomic orders of mammalian and avian reservoir
-hosts. - Proceedings of the National Academy of Sciences 117:
-9423--9430.
-
-\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
-Moratelli, R. and Calisher, C. H. 2015. Bats and zoonotic viruses: Can
-we confidently link bats with emerging deadly viruses? - Memórias do
-Instituto Oswaldo Cruz 110: 1--22.
-
-\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
-Moreno Santillán, D. D. et al. 2021. Large-scale genome sampling reveals
-unique immunity and metabolic adaptations in bats. - Molecular Ecology:
-mec.16027.
-
-\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-Muylaert, R. L. et al. 2022. Present and future distribution of bat
-hosts of sarbecoviruses: Implications for conservation and public
-health. - Proceedings of the Royal Society B: Biological Sciences 289:
-20220397.
-
-\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-Nguyen, L.-T. et al. 2015. IQ-TREE: A Fast and Effective Stochastic
-Algorithm for Estimating Maximum-Likelihood Phylogenies. - Molecular
-Biology and Evolution 32: 268--274.
-
-\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-Olival, K. J. et al. 2017. Host and viral traits predict zoonotic
-spillover from mammals. - Nature 546: 646--650.
+\CSLLeftMargin{72. }
+\CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
+biogeographical regions based on species distributions. \emph{Journal of
+Biogeography} \textbf{37}, 2029--2053 (2010).}
 
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-Paradis, E. and Schliep, K. 2019. Ape 5.0: An environment for modern
-phylogenetics and evolutionary analyses in R. - Bioinformatics 35:
-526--528.
-
-\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-Peixoto, F. P. et al. 2018. A synthesis of ecological and evolutionary
-determinants of bat diversity across spatial scales. - BMC ecology 18:
-18.
-
-\leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%
-Plowright, R. K. et al. 2017. Pathways to zoonotic spillover. - Nature
-Reviews Microbiology 15: 502--510.
-
-\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
-Power, A. G. and Mitchell, C. E. 2004. Pathogen Spillover in Disease
-Epidemics. - The American Naturalist 164: S79--S89.
-
-\leavevmode\hypertarget{ref-Price2002MacThe}{}%
-Price, P. W. 2002. Macroevolutionary Theory on Macroecological Patterns.
-- Cambridge University Press.
-
-\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
-Ramshaw, R. E. et al. 2019. A database of geopositioned Middle East
-Respiratory Syndrome Coronavirus occurrences. - Scientific Data 6: 318.
-
-\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-Rouault, E. et al. 2022. GDAL/OGR Geospatial Data Abstraction software
-Library. - Zenodo.
-
-\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
-Ruiz-Aravena, M. et al. 2022. Ecology, evolution and spillover of
-coronaviruses from bats. - Nature Reviews Microbiology 20: 299--314.
-
-\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-Seekell, D. A. et al. 2018. A geography of lake carbon cycling. -
-Limnology and Oceanography Letters 3: 49--56.
-
-\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-Shi, J. J. et al. 2014. A Deep Divergence Time between Sister Species of
-Eidolon (Pteropodidae) with Evidence for Widespread Panmixia. - Acta
-Chiropterologica 16: 279--292.
-
-\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-Simmons, N. B. and Cirranello, A. L. 2020. Bat Species of the World: A
-taxonomic and geographic database.
-
-\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
-Springer, M. S. 2013. Phylogenetics: Bats United, Microbats Divided. -
-Current Biology 23: R999--R1001.
-
-\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
-Teeling, E. C. et al. 2005. A Molecular Phylogeny for Bats Illuminates
-Biogeography and the Fossil Record. - Science (New York, N.Y.) 307:
-580--584.
-
-\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
-Thompson, J. N. 1994. The Coevolutionary Process. - University of
-Chicago Press.
-
-\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
-Thompson, J. N. 2005. The Geographic Mosaic of Coevolution. - University
-Of Chicago Press.
-
-\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-Upham, N. S. et al. 2019. Inferring the mammal tree: Species-level sets
-of phylogenies for questions in ecology, evolution, and conservation. -
-PLOS Biology 17: e3000494.
-
-\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-Van Brussel, K. and Holmes, E. C. 2022. Zoonotic disease and virome
-diversity in bats. - Current Opinion in Virology 52: 192--202.
-
-\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
-Villalobos, F. and Arita, H. T. 2010. The diversity field of New World
-leaf-nosed bats (Phyllostomidae). - Global Ecology and Biogeography 19:
-200--211.
-
-\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-Vlasova, A. N. et al. 2022. Animal alphacoronaviruses found in human
-patients with acute respiratory illness in different countries. -
-Emerging Microbes \& Infections 11: 699--702.
-
-\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-2016. Bats in the Anthropocene: Conservation of Bats in a Changing World
-(CC Voigt and T Kingston, Eds.). - Springer International Publishing.
-
-\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-Wang, N. et al. 2018. Serological Evidence of Bat SARS-Related
-Coronavirus Infection in Humans, China. - Virologica Sinica 33:
-104--107.
-
-\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-Williams-Guillén, K. et al. 2008. Bats Limit Insects in a Neotropical
-Agroforestry System. - Science 320: 70--70.
-
-\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-Xu, R.-H. et al. 2004. Epidemiologic Clues to SARS Origin in China. -
-Emerging Infectious Diseases 10: 1030--1037.
+\CSLLeftMargin{73. }
+\CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for
+modern phylogenetics and evolutionary analyses in R.
+\emph{Bioinformatics} \textbf{35}, 526--528 (2019).}
 

### 2022-06-21 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1b4ddd7d8db0b6e031f2dc59c93240bedff693b2 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@16e453a1bff6288438ff77acaba93a8c85556ae7 🚀

@@
 \noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
-\emph{Betacoronavirus} spillover risk from bat hosts}}}
+betacoronavirus emergence risk}}}
 

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@eea6e201f3438c201940b84fa06464c6f96cb8a2 🚀

@@
 transmission may, as a rule, structure viral evolution more than
-co-divergence with hosts {[}Geoghegan{]}. This creates a disconnect
+co-divergence with hosts.\textsuperscript{6} This creates a disconnect
 between coevolutionary theory (including empirical evidence from
@@
 
-The Geographic Mosaic Theory of Coevolution\textsuperscript{GMTC; 6}
-attempts to explicitly connect microevolutionary dynamics to the
-macroecology and biogeography of symbiotic interactions. The GMTC posits
-that coevolutionary processes among pairs\textsuperscript{7} or
-complexes\textsuperscript{8} of species are structured in space by the
+The Geographic Mosaic Theory of Coevolution (GMTC) attempts to
+explicitly connect microevolutionary dynamics to the macroecology and
+biogeography of symbiotic interactions.\textsuperscript{7} The GMTC
+posits that coevolutionary processes among pairs\textsuperscript{8} or
+complexes\textsuperscript{9} of species are structured in space by the
 rippling effects of abiotic conditions onto evolutionary mechanism,
 giving rise to fragmented systems with different structure and
-ecologically dynamics over large spatial extents.\textsuperscript{see
-\emph{e.g.} 9} The GMTC predicts a spatial fragmentation of
-coevolutionary dynamics under the joint action of three
-processes:\textsuperscript{see notably 10} coevolutionary hot- and
-coldspots, which appear when the intensity of \emph{interaction} (in
-terms of reciprocal fitness consequences) varies spatially; selection
-mosaics, wherein the intensity of \emph{selection} varies across space,
-driven by both the biotic complexity of the community (locally diverse
-hosts and viruses are more biotically complex) and the local
-favorability of the environment;\textsuperscript{11} and trait remixing,
-which occurs when coevolutionary dynamics are driven by by the arrival
-(or departure) of \emph{functional traits}, through changes in community
-composition due to invasions, meta-community dynamics, and disperal.
+ecologically dynamics over large spatial extents.\textsuperscript{10}
+The GMTC predicts a spatial fragmentation of coevolutionary dynamics
+under the joint action of three processes:\textsuperscript{11}
+coevolutionary hot- and coldspots, which appear when the intensity of
+\emph{interaction} (in terms of reciprocal fitness consequences) varies
+spatially; selection mosaics, wherein the intensity of \emph{selection}
+varies across space, driven by both the biotic complexity of the
+community (locally diverse hosts and viruses are more biotically
+complex) and the local favorability of the
+environment;\textsuperscript{12} and trait remixing, which occurs when
+coevolutionary dynamics are driven by by the arrival (or departure) of
+\emph{functional traits}, through changes in community composition due
+to invasions, meta-community dynamics, and disperal.
 
@@
 serve as the main animal reservoir for different strains of
-betacoronaviruses.\textsuperscript{12} This has attracted attention to
+betacoronaviruses.\textsuperscript{13} This has attracted attention to
 areas where high diversity of bats, and therefore presumably high
 diversity of betacoronaviruses, can be an important issue for human
-health.\textsuperscript{13,14} By overlaying the IUCN range maps for
+health.\textsuperscript{14,15} By overlaying the IUCN range maps for
 confirmed bat hosts of betacoronaviruses {[}fig.~\ref{fig:richness};
@@
 first step towards understanding risk, as previous
-research\textsuperscript{2,15} states that locally diverse bat
+research\textsuperscript{2,16} states that locally diverse bat
 communities could maintain more viruses and hence, a higher probability
@@
 represent an increased risk of spillover under climate change through
-the creation of novel interactions,\textsuperscript{16} and therefore
+the creation of novel interactions,\textsuperscript{17} and therefore
 the diversity of \emph{Betacoronavirus} strains should similarly be
@@
 classified, from a macro-evolutionary standpoint, as Yangochiroptera and
-Yinpterochiroptera elsewhere.\textsuperscript{17,18} Specifically, we
+Yinpterochiroptera elsewhere.\textsuperscript{18,19} Specifically, we
 would expect that the so-called ``New World'' group of bats, being more
@@
 the western hemisphere, contain the highly diverse adaptive radiation in
-the Phyllostomidae,\textsuperscript{19} which is restricted to the
+the Phyllostomidae,\textsuperscript{20} which is restricted to the
 western hemisphere. By contrast, South-Eastern Asia has a lot of
@@
 betacoronaviruses likely remain to be discovered in this region. Indeed,
-global predictions by\textsuperscript{20} highlight that southeast Asia
-is a likely hostpot of unconfirmed hosts of betacoronaviruses, which
-would likely result in additional viral discoveries. This idea is
-unsurprising given the growing realization, especially since the
-emergence of SARS-CoV-2, that unique lineages of similar viruses are
-widespread in bats but still mostly undescribed. The most distinct
+global predictions highlight that southeast Asia is a likely hostpot of
+unconfirmed hosts of betacoronaviruses,\textsuperscript{21} which would
+likely result in additional viral discoveries. This idea is unsurprising
+given the growing realization, especially since the emergence of
+SARS-CoV-2, that unique lineages of similar viruses are widespread in
+bats but still mostly undescribed. The most distinct
 bats-betacoronaviruses complex is found in South America, a region with
@@
 the fewest \emph{Betacoronavirus} sequences have been
-discovered,\textsuperscript{2,21,22} resulting in sparser phylogenetic
+discovered,\textsuperscript{2,22,23} resulting in sparser phylogenetic
 tree, thereby artificially inflating distinctiveness. Adding more
@@
 Finally, South-America is the range of Phyllostomidae, a group of bats
-that underwent explosive diversification events,\textsuperscript{19}
+that underwent explosive diversification events,\textsuperscript{20}
 which may drive the emergence of multiple viral lineages.
@@
 betacoronaviruses should show some degree of
-congruence.\textsuperscript{23,24} In particular, this should be the
+congruence.\textsuperscript{24,25} In particular, this should be the
 case if viruses can circulate among hosts and co-evolve with local hosts
@@
 immunological responses, as evidenced in several bat
-species.\textsuperscript{25} Immune characteristics that allow bats to
+species.\textsuperscript{26} Immune characteristics that allow bats to
 be better adapted to infection by emerging
-viruses,\textsuperscript{26,27} in addition to being hardcoded in their
-genome,\textsuperscript{28} may be related to a wide variety of
-diets,\textsuperscript{25,29} themselves likely to be driven by spatial
+viruses,\textsuperscript{27,28} in addition to being hardcoded in their
+genome,\textsuperscript{29} may be related to a wide variety of
+diets,\textsuperscript{26,30} themselves likely to be driven by spatial
 effects, especially at the local scale -- bats, indeed, occupy a variety
 of environments, and therefore display a variety of adaptations to these
-environments.\textsuperscript{30}
+environments.\textsuperscript{31}
 
@@
 nearby location where serological (antibody) evidence has indicated
-human exposure to SARS-like viruses;\textsuperscript{31} MERS-CoV in
+human exposure to SARS-like viruses;\textsuperscript{32} MERS-CoV in
 Saudi Arabia based on index cases available from a recently-published
-compendium of cases.\textsuperscript{32} For the latest event, most if
+compendium of cases.\textsuperscript{33} For the latest event, most if
 not all index cases are presumed to be camel-to-human transmission, and
@@
 uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti\textsuperscript{33} and Europe\textsuperscript{34} are not
+Haiti\textsuperscript{34} and Europe\textsuperscript{35} are not
 relevant here, as bats' involvement in these cycles of transmission have
@@
 co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
-Coevolution\textsuperscript{6} to provide a measure of risk accounting
+Coevolution\textsuperscript{7} to provide a measure of risk accounting
 for multiple processes. In fig.~\ref{fig:trivariate}, we overlapped
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{\emph{e.g.} 35} Recent
-surveillance\textsuperscript{36} has identified a novel
+continents.\textsuperscript{\emph{e.g.} 36} Recent
+surveillance\textsuperscript{37} has identified a novel
 \emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
@@
 increases the probability of pathogen spread more
-widely.\textsuperscript{37} Since human activity is required to amplify
+widely.\textsuperscript{38} Since human activity is required to amplify
 the frequency of virus encounters and thus create areas of viral
@@
 opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China,\textsuperscript{38} present
+looking for the origins of SARS in China,\textsuperscript{39} present
 serological evidence that strongest human-animal contact results in
@@
 amount of risk represented by each activity. Our map of risk overlays
-with recent results from\textsuperscript{39} -- areas of purported high
+with recent results from\textsuperscript{40} -- areas of purported high
 risk/diversitifcation potential (Madagascar, South-America) overlay with
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{23,24} Chiropterans emerged around 64 million
+health.\textsuperscript{24,25} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{40,41}
+estimated richness of more than 1400 species.\textsuperscript{41,42}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{42} For example,
+ecosystem-derived benefits to human.\textsuperscript{43} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{43} Over two-thirds of bats are know to be
+networks.\textsuperscript{44} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{44,45} and
+contributing for agricultural pest control,\textsuperscript{45,46} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{46,47} Because bats are globally distributed and
+health.\textsuperscript{47,48} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 that have used host richness as proxy have predicted a high diversity of
-unsampled bat viruses,\textsuperscript{22} bat
+unsampled bat viruses,\textsuperscript{23} bat
 coronaviruses,\textsuperscript{2} and even specifically
-betacoronaviruses\textsuperscript{20} in both the Amazon and southeast
+betacoronaviruses\textsuperscript{21} in both the Amazon and southeast
 Asia. While we find that both regions are characterized by highly
@@
 high local diversity of viruses through host switching and viral
-recombination (see \emph{e.g.},\textsuperscript{48} as well as the
+recombination (see \emph{e.g.},\textsuperscript{49} as well as the
 discovery of recombinant viruses that share genetic material from both
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{49} Both of these regions are priority areas
+lineage).\textsuperscript{50} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{20,39} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{21,40} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{16}
+to the region.\textsuperscript{17}
 
 {[}Tim tinker with para about conservation{]} There are several factors
-that drive changes in the diversity of bats,\textsuperscript{50} but
+that drive changes in the diversity of bats,\textsuperscript{51} but
 human activities' effects on the ecosystem (like modifications of land
@@
 hosts, we also increase the risk of emergence of novel
-diseases,\textsuperscript{51} as previous studies on RNA viruses suggest
+diseases,\textsuperscript{52} as previous studies on RNA viruses suggest
 the importance of host phylogeography at the time of virus
-dispersal.\textsuperscript{52} One of these scenarios where interaction
+dispersal.\textsuperscript{53} One of these scenarios where interaction
 between bats and humans can occur can be seed dispersal in tropical
@@
 interaction between viruses of bat origin and
-humans.\textsuperscript{53} This represents a challenge for conservation
+humans.\textsuperscript{54} This represents a challenge for conservation
 strategies and disease ecology since some areas can haveboth potential
@@
 We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
-by\textsuperscript{20} from
+by\textsuperscript{21} from
 \texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{54} The range maps were subsequently rasterized
+data.\textsuperscript{55} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{56} based on a recent synthetic tree with
+Diversity\textsuperscript{57} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{57} Faith's PD measures the sum of unique
+species.\textsuperscript{58} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{59} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{59} transformed
+location, and following best practices\textsuperscript{60} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{60} shows that LCBD
+community. Recent results by\textsuperscript{61} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range
-overlap.\textsuperscript{61} This network stores pairwise values of
+overlap.\textsuperscript{62} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{62} In this approach, every risk component gets
+model.\textsuperscript{63} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{63,64} in that this allows
+and low viral sharing,\textsuperscript{64,65} in that this allows
 multiple independent host-virus coevolutionary dynamics to take place in
@@
 -- because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{see \emph{e.g.} 65} we measured the risk
+extraction,\textsuperscript{see \emph{e.g.} 66} we measured the risk
 level by calculating the angular distance of the hue of each pixel to a
@@
 betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
-{[}\textsuperscript{66}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
+{[}\textsuperscript{67}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
 k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
-likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{67}
-with ModelFinder\textsuperscript{68} ultrafast bootstrap
-approximation\textsuperscript{69} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{68}
+with ModelFinder\textsuperscript{69} ultrafast bootstrap
+approximation\textsuperscript{70} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{70} for each of the viruses in the tree,
+distinctiveness\textsuperscript{71} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{71,72} who
+this idea, we loosely adapted a method from,\textsuperscript{72,73} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{74} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Geoghegan2017ComAna}{}%
+\CSLLeftMargin{6. }
+\CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
+Comparative analysis estimates the relative frequencies of co-divergence
+and cross-species transmission within viral families. \emph{PLOS
+Pathogens} \textbf{13}, e1006215 (2017).}
+
 \leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
-\CSLLeftMargin{6. }
+\CSLLeftMargin{7. }
 \CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
@@
 \leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
-\CSLLeftMargin{7. }
+\CSLLeftMargin{8. }
 \CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
@@
 \leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
-\CSLLeftMargin{8. }
+\CSLLeftMargin{9. }
 \CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-Price2002MacThe}{}%
-\CSLLeftMargin{9. }
+\CSLLeftMargin{10. }
 \CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
-\CSLLeftMargin{10. }
+\CSLLeftMargin{11. }
 \CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
@@
 \leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
-\CSLLeftMargin{11. }
+\CSLLeftMargin{12. }
 \CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
@@
 \leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
-\CSLLeftMargin{12. }
+\CSLLeftMargin{13. }
 \CSLRightInline{Drexler, J. F., Corman, V. M. \& Drosten, C. Ecology,
@@
 \leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
-\CSLLeftMargin{13. }
+\CSLLeftMargin{14. }
 \CSLRightInline{Calisher, C. H., Childs, J. E., Field, H. E., Holmes, K.
@@
 \leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
-\CSLLeftMargin{14. }
+\CSLLeftMargin{15. }
 \CSLRightInline{Moratelli, R. \& Calisher, C. H. Bats and zoonotic
@@
 \leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
-\CSLLeftMargin{15. }
+\CSLLeftMargin{16. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{16. }
+\CSLLeftMargin{17. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
-\CSLLeftMargin{17. }
+\CSLLeftMargin{18. }
 \CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
@@
 \leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
@@
 \leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
@@
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
@@
 \leavevmode\hypertarget{ref-Allen2017GloHot}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
@@
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
@@
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
@@
 \leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
@@
 \leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
@@
 \leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
@@
 \leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{45. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Deshpande, K., Vanak, A. T., Devy, M. S. \&
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{55. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Cavender-Bares, J., Kozak, K. H., Fine, P. V. A. \&
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{73. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{73. }
+\CSLLeftMargin{74. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6922fbcfc88b17b9b42b7d43e59e81efebaa73cb 🚀

@@
 health.\textsuperscript{14,15} By overlaying the IUCN range maps for
-confirmed bat hosts of betacoronaviruses {[}fig.~\ref{fig:richness};
-top{]}, we see that the the main hotspots (both in terms of size and
-higher values) of host richness are primarily South-Eastern Asia, parts
-of Southern Europe, and to a lesser extent parts of Africa in the -25-0
+confirmed bat hosts of betacoronaviruses (fig.~\ref{fig:richness}; top),
+we see that the the main hotspots (both in terms of size and higher
+values) of host richness are primarily South-Eastern Asia, parts of
+Southern Europe, and to a lesser extent parts of Africa in the -25-0
 range of latitudes. The description of host richness is an important
@@
 co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
-Coevolution\textsuperscript{7} to provide a measure of risk accounting
-for multiple processes. In fig.~\ref{fig:trivariate}, we overlapped
-three components of spillover risk: viral sharing, \emph{i.e.} the
-chance that two bats will share viruses overall; Local Contribution to
-Beta Diversity, \emph{i.e.} the fact that a bat community is
-compositionally unique compared to the average compositional similarity
-across the entire system; finally, host phylogenetic diversity,
-\emph{i.e.} how dispersed the bats in a location are within the tree of
-life. This approach leads to the definition of broad biogeographic
-regions of risk, where the same color represents the same type of risk.
-By way of constrat to figures fig.~\ref{fig:richness} and
-fig.~\ref{fig:biogeo}, these regions do not necessarilly overlap with
-previous spatial partitions of the bat-betacoronaviruses complex.
+Coevolution to provide a measure of risk accounting for multiple
+processes. In fig.~\ref{fig:trivariate}, we overlapped three components
+of spillover risk: viral sharing, \emph{i.e.} the chance that two bats
+will share viruses overall; Local Contribution to Beta Diversity,
+\emph{i.e.} the fact that a bat community is compositionally unique
+compared to the average compositional similarity across the entire
+system; finally, host phylogenetic diversity, \emph{i.e.} how dispersed
+the bats in a location are within the tree of life. This approach leads
+to the definition of broad biogeographic regions of risk, where the same
+color represents the same type of risk. By way of constrat to figures
+fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo}, these regions do not
+necessarilly overlap with previous spatial partitions of the
+bat-betacoronaviruses complex.
 
@@
 perspective on risk, other regions of the world are of lesser concern
-fig.~\ref{fig:risk}. Our risk decomposition does not account for viral
+(fig.~\ref{fig:risk}). Our risk decomposition does not account for viral
 diversity or distinctiveness. The simple rationale behind it is that the
@@
 (inter-genus or higher levels of cross-species transmission) and, to a
-lesser degree, cospeciation and sharing.\textsuperscript{intra-genus
-cross-species transmission; 2} This diversification is not an actual
-risk factor for spillover itself, but acts downstream of a spillover
-event by increasing the random chance of the emergence of a virus with
-the raw genomic components required for the potential to infect humans.
+lesser degree, cospeciation and sharing, representing intra-genus
+cross-species transmission.\textsuperscript{2} This diversification is
+not an actual risk factor for spillover itself, but acts downstream of a
+spillover event by increasing the random chance of the emergence of a
+virus with the raw genomic components required for the potential to
+infect humans.
 
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{\emph{e.g.} 36} Recent
-surveillance\textsuperscript{37} has identified a novel
-\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
-Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
-\emph{Rousettus madagascariensis}), emphasizing strong proof of
-principle in model predictions.
+continents.\textsuperscript{36} Recent surveillance\textsuperscript{37}
+has identified a novel \emph{Betacoronavirus} (in the subgenus
+\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
+(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
+strong proof of principle in model predictions.
 
@@
 further, alebit at the cost of more hypotheses required to estimate the
-amount of risk represented by each activity. Our map of risk overlays
-with recent results from\textsuperscript{40} -- areas of purported high
+amount of risk represented by each activity. Our map of purported high
 risk/diversitifcation potential (Madagascar, South-America) overlay with
-sampling gaps for \emph{Betacoronavirus}, stressing the need for
-spatially targeted monitoring and discovery.
+sampling gaps for \emph{Betacoronavirus},\textsuperscript{40} stressing
+the need for spatially targeted monitoring and discovery.
 

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1aa4d1699319cabba9f9b73cd8b67632f8a2cd89 🚀

@@
 syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
-betacoronavirus RdRp sequences were then aligned using MAFFT v 1.4.0
-{[}\textsuperscript{67}; Algorithm FFT-NS-2, Scoring matrix 200PAM /
-k=2, gap open penalty 1.53m offset value 0.123{]} and a maximum
-likelihood tree reconstructed in IQ-TREE v 1.6.12\textsuperscript{68}
+betacoronavirus RdRp sequences were then aligned using
+MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
+likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
 with ModelFinder\textsuperscript{69} ultrafast bootstrap
@@
 
+\newpage
+
 \hypertarget{references}{%

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@be5dbd63ba5ee717e5879239f1b41feb1bd1e6e7 🚀

@@
 
-We downloaded the data on bats hosts of \emph{Betacoronavirus} assembled
-by\textsuperscript{21} from
-\texttt{https://www.viralemergence.org/betacov} on Apr.~2022, and
-filtered it to ``known'' hosts (established before the emergence of
-SARS-CoV-2) and ``novel'' hosts (confirmed through sampling and
-competence assays since the initial data collection). The original
-database was assembled by a combination of data mining and literature
-surveys, including automated alerts on the ``bats'' and ``coronavirus''
-keywords to identify novel empirical evidence of bats-betacoronaviruses
-associations; this yielded a total of 126 known hosts, 47 of which were
-novel hosts.
+We downloaded the data on bats hosts of \emph{Betacoronavirus} from
+\texttt{https://www.viralemergence.org/betacov} on
+Apr.~2022,\textsuperscript{21} and filtered it to ``known'' hosts
+(established before the emergence of SARS-CoV-2) and ``novel'' hosts
+(confirmed through sampling and competence assays since the initial data
+collection). The original database was assembled by a combination of
+data mining and literature surveys, including automated alerts on the
+``bats'' and ``coronavirus'' keywords to identify novel empirical
+evidence of bats-betacoronaviruses associations; this yielded a total of
+126 known hosts, 47 of which were novel hosts.
 
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{64,65} in that this allows
-multiple independent host-virus coevolutionary dynamics to take place in
-the same location. In the colorimetric space, this correspond to yellow
--- because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{see \emph{e.g.} 66} we measured the risk
-level by calculating the angular distance of the hue of each pixel to a
-reference value of 60, and weighted this risk level by the value
+and low viral sharing,\textsuperscript{64} in that this allows multiple
+independent host-virus coevolutionary dynamics to take place in the same
+location. In the colorimetric space, this correspond to yellow --
+because the HSV space is more amenable to calculations for feature
+extraction,\textsuperscript{65} we measured the risk level by
+calculating the angular distance of the hue of each pixel to a reference
+value of 60 (yellow), and weighted this risk level by the value
 component. Specifically, given a pixel with colorimetric coordinates
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
-with ModelFinder\textsuperscript{69} ultrafast bootstrap
-approximation\textsuperscript{70} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{67} v1.6.12
+with ModelFinder\textsuperscript{68} ultrafast bootstrap
+approximation\textsuperscript{69} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{71} for each of the viruses in the tree,
+distinctiveness\textsuperscript{70} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{72,73} who
+this idea, we loosely adapted a method from,\textsuperscript{71,72} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{74} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
+\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
 \CSLLeftMargin{64. }
-\CSLRightInline{Cavender-Bares, J., Kozak, K. H., Fine, P. V. A. \&
-Kembel, S. W. The merging of community ecology and phylogenetic biology.
-\emph{Ecol. Lett.} \textbf{12}, 693--715 (2009).}
-
-\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{65. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{73. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{74. }
+\CSLLeftMargin{73. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1397fb2684d2e8fb22048453717ca8cf3a8c3112 🚀

@@
 \hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
-\subsection{Host richness does not predict virus
+\section{Host richness does not predict virus
 distinctiveness}\label{host-richness-does-not-predict-virus-distinctiveness}}
@@
 \hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
-\subsection{The phylogeographic regions of hosts and their viruses
+\section{The phylogeographic regions of hosts and their viruses
 overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}
@@
 \hypertarget{coevolution-informed-emergence-risk-is-different-in-space}{%
-\subsection{Coevolution-informed emergence risk is different in
+\section{Coevolution-informed emergence risk is different in
 space}\label{coevolution-informed-emergence-risk-is-different-in-space}}
@@
 \hypertarget{human-occupancy-drives-different-levels-of-effective-risk-globally}{%
-\subsection{Human occupancy drives different levels of effective risk
+\section{Human occupancy drives different levels of effective risk
 globally}\label{human-occupancy-drives-different-levels-of-effective-risk-globally}}

### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f716db62355123121d22ac5466a04778c04fea61 🚀

--- 2022-06-21 draft.tex
@@
 best understood through the multi-faceted prism of ecological and
-evolutionary mechanisms. Our framework reveals that reservoir richness
-alone, although a component of viral hazard, is not a sufficiently
-integrative predictor of risk. We offer alternative insights based on
-viral sharing, host compositional uniqueness, and host phylogenetic
-diversity and phylogeographic regions. By comparing our aggregated
-measure of risk to a proxy for human density, namely the proportion of
-each pixel that is covered by urban or built land, we provide a
-synthetic risk map, allowing the identification of hotspots where the
-bat-betacoronavirus system may originate novel viruses in humans.
+evolutionary mechanisms. Our framework reveals that host richness alone,
+although a component of viral hazard, is not a sufficiently integrative
+predictor of risk. We offer alternative insights based on viral sharing,
+host compositional uniqueness, and host phylogenetic diversity and
+phylogeographic regions. By comparing our aggregated measure of risk to
+a proxy for human density, namely the proportion of each pixel that is
+covered by urban or built land, we provide a synthetic risk map,
+allowing the identification of hotspots where the bat-betacoronavirus
+interaction network may facilitate the emergence of novel viruses and
+their spillover into human populations.
 
@@
 bats-betacoronaviruses complex is found in South America, a region with
-a comparatively lower number of hosts; this matches with the isolation
-through viariance of the host group, and may highlight a different
-co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
-be a product of under-sampling: South-America is one of the places where
-the fewest \emph{Betacoronavirus} sequences have been
+a comparatively lower number of hosts; this matches with isolation
+through host vicariance, and may highlight a different co-evolutionary
+dynamic. Alternatively, this distinctiveness hostpot may be a product of
+under-sampling: South-America is one of the places where the fewest
+\emph{Betacoronavirus} sequences have been
 discovered,\textsuperscript{2,22,23} resulting in sparser phylogenetic
@@
 
-{[}Tim tinker with para about conservation{]} There are several factors
-that drive changes in the diversity of bats,\textsuperscript{51} but
-human activities' effects on the ecosystem (like modifications of land
-use) could significantly decrease it. Therefore, it can be suggested
-that changes in the diversity of betacoronaviruses in bats are linked to
-their biogeographic variation, and human population density and other
-anthropogenic factors are decisive moderators for its implications in
-public health. With the increase of contact between humans and potential
-hosts, we also increase the risk of emergence of novel
-diseases,\textsuperscript{52} as previous studies on RNA viruses suggest
-the importance of host phylogeography at the time of virus
-dispersal.\textsuperscript{53} One of these scenarios where interaction
-between bats and humans can occur can be seed dispersal in tropical
-agroecosystems. It opens the discussion of whether the fruits thrown by
-bats not only disperse seeds but could also be a source of indirect
-interaction between viruses of bat origin and
-humans.\textsuperscript{54} This represents a challenge for conservation
-strategies and disease ecology since some areas can haveboth potential
-for the acquisition of zoonotic viruses and bat-human interactions; in
-particular, the challenge lies in the fact that actual exposure must
-then be quantified accounting for several transmission scenarios,
-including both direct and indirect bat - human interaction.
+The diversity and diversification potential of bats responds to
+anthropogenic factors others than shifting climates.\textsuperscript{51}
+Land use changes could significantly decrease bat suitability, notably
+through effects on diet and availability of
+habitats.\textsuperscript{52} As our results establish that the
+diversification of bats betacoronaviruses happens on top of processes
+affecting hosts, biogeographic variation in human population density and
+anthropogenic disturbances may feed into co-evolutionary dynamics.
+Increase in humans-hosts contacts also increase the risk of emergence of
+novel diseases,\textsuperscript{53} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{54} This
+represents a challenge for both conservation strategies and disease
+ecology: some areas can a high emergence risk and more potential for the
+acquisition of zoonotic viruses through bat-human
+encounters.\textsuperscript{55} In particular, the challenge ahead lies
+in the need to quantify actual exposure (and risk) accounting for
+several transmission scenarios, including both direct and indirect bat -
+human interactions, and feeding back into the provision of ecosystem
+services by bats.
 
@@
 Courtois Foundation. RLM was supported by Bryce Carmine and Anne Carmine
-(née Percival), through the Massey University Foundation.
+(née Percival), through the Massey University Foundation. DJB was
+supported by the National Institute of General Medical Sciences of the
+National Institutes of Health (P20GM134973).
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{55} The range maps were subsequently rasterized
+data.\textsuperscript{56} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{57} based on a recent synthetic tree with
+Diversity\textsuperscript{58} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{58} Faith's PD measures the sum of unique
+species.\textsuperscript{59} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{59} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{60} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{60} transformed
+location, and following best practices\textsuperscript{61} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{61} shows that LCBD
+community. Recent results by\textsuperscript{62} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 generalized additive mixed model of virus sharing by a tensor function
-of phylogenetic distance and geographic range
-overlap.\textsuperscript{62} This network stores pairwise values of
+of phylogenetic distance and geographic range overlap across
+mammals.\textsuperscript{63} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{63} In this approach, every risk component gets
+model.\textsuperscript{64} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{64,65} in that this allows
-multiple independent host-virus coevolutionary dynamics to take place in
-the same location. In the colorimetric space, this correspond to yellow
--- because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{see \emph{e.g.} 66} we measured the risk
-level by calculating the angular distance of the hue of each pixel to a
-reference value of 60, and weighted this risk level by the value
+and low viral sharing,\textsuperscript{65} in that this allows multiple
+independent host-virus coevolutionary dynamics to take place in the same
+location. In the colorimetric space, this correspond to yellow --
+because the HSV space is more amenable to calculations for feature
+extraction,\textsuperscript{66} we measured the risk level by
+calculating the angular distance of the hue of each pixel to a reference
+value of 60 (yellow), and weighted this risk level by the value
 component. Specifically, given a pixel with colorimetric coordinates
@@
 
+\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
+\CSLLeftMargin{52. }
+\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
+effect of local land use and loss of forests on bats and nocturnal
+insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}
+
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 
-\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
-\CSLLeftMargin{54. }
-\CSLRightInline{Deshpande, K., Vanak, A. T., Devy, M. S. \&
-Krishnaswamy, J. Forbidden fruits? Ecosystem services from seed
-dispersal by fruit bats in the context of latent zoonotic risk.
-\emph{Oikos (Copenhagen, Denmark)} oik.08359 (2022)
-doi:\href{https://doi.org/10.1111/oik.08359}{10.1111/oik.08359}.}
+\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
+\CSLLeftMargin{55. }
+\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
+of bats in emerging zoonoses: Balancing ecology, conservation and public
+health interest}. (FAO, 2011).}
 
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{56. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 Letters} \textbf{3}, 49--56 (2018).}
-
-\leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
-\CSLLeftMargin{64. }
-\CSLRightInline{Cavender-Bares, J., Kozak, K. H., Fine, P. V. A. \&
-Kembel, S. W. The merging of community ecology and phylogenetic biology.
-\emph{Ecol. Lett.} \textbf{12}, 693--715 (2009).}
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6e1ff4016222589bcf665d30237cd734d28d5598 🚀

@@
 
+\hypertarget{results-and-discussion}{%
+\section{Results and Discussion}\label{results-and-discussion}}
+
 \hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
-\section{Host richness does not predict virus
+\subsection{Host richness does not predict virus
 distinctiveness}\label{host-richness-does-not-predict-virus-distinctiveness}}
@@
 \hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
-\section{The phylogeographic regions of hosts and their viruses
+\subsection{The phylogeographic regions of hosts and their viruses
 overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}
@@
 \hypertarget{coevolution-informed-emergence-risk-is-different-in-space}{%
-\section{Coevolution-informed emergence risk is different in
+\subsection{Coevolution-informed emergence risk is different in
 space}\label{coevolution-informed-emergence-risk-is-different-in-space}}
@@
 \hypertarget{human-occupancy-drives-different-levels-of-effective-risk-globally}{%
-\section{Human occupancy drives different levels of effective risk
+\subsection{Human occupancy drives different levels of effective risk
 globally}\label{human-occupancy-drives-different-levels-of-effective-risk-globally}}

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@890e7b3f2b398279abf2c9d6ada1b51469df2d43 🚀

@@
 
-Bats, the second most diverse groups of mammals, are found worldwide and
-serve as the main animal reservoir for different strains of
-betacoronaviruses.\textsuperscript{13} This has attracted attention to
-areas where high diversity of bats, and therefore presumably high
-diversity of betacoronaviruses, can be an important issue for human
-health.\textsuperscript{14,15} By overlaying the IUCN range maps for
-confirmed bat hosts of betacoronaviruses (fig.~\ref{fig:richness}; top),
-we see that the the main hotspots (both in terms of size and higher
-values) of host richness are primarily South-Eastern Asia, parts of
-Southern Europe, and to a lesser extent parts of Africa in the -25-0
-range of latitudes. The description of host richness is an important
-first step towards understanding risk, as previous
-research\textsuperscript{2,16} states that locally diverse bat
-communities could maintain more viruses and hence, a higher probability
-of having a pathogen that could represent a risk for human health.
+Bats, the second most diverse groups of mammals, are found worldwide;
+gradients in bat species richness generally track broader patterns of
+mammal diversity, with a striking Neotropical hotspot (especially in the
+Amazon basin) and a secondary hotspot centered in the southeast Asian
+peninsula. These hotspots of bat diversity are generally presumed to be
+hotspots of viral diversity, and therefore areas of concern for human
+health.\textsuperscript{2,13} However, the hotspots of bat
+betacoronavirus reservoirs show a distinct pattern, with primary
+hotspots (both in terms of size and higher values) of host richness
+situated primarily South-Eastern Asia, parts of Southern Europe, and to
+a lesser extent parts of Africa in the -25-0 range of latitudes (Fig. 1;
+top). Although hundreds of species likely host undiscovered
+betacoronaviruses, machine learning predictions have suggested that
+these undiscovered reservoirs should follow the same diversity
+gradient.\textsuperscript{14}
 
@@
 represent an increased risk of spillover under climate change through
-the creation of novel interactions,\textsuperscript{17} and therefore
+the creation of novel interactions,\textsuperscript{15} and therefore
 the diversity of \emph{Betacoronavirus} strains should similarly be
@@
 classified, from a macro-evolutionary standpoint, as Yangochiroptera and
-Yinpterochiroptera elsewhere.\textsuperscript{18,19} Specifically, we
+Yinpterochiroptera elsewhere.\textsuperscript{16,17} Specifically, we
 would expect that the so-called ``New World'' group of bats, being more
@@
 the western hemisphere, contain the highly diverse adaptive radiation in
-the Phyllostomidae,\textsuperscript{20} which is restricted to the
+the Phyllostomidae,\textsuperscript{18} which is restricted to the
 western hemisphere. By contrast, South-Eastern Asia has a lot of
@@
 global predictions highlight that southeast Asia is a likely hostpot of
-unconfirmed hosts of betacoronaviruses,\textsuperscript{21} which would
+unconfirmed hosts of betacoronaviruses,\textsuperscript{14} which would
 likely result in additional viral discoveries. This idea is unsurprising
@@
 \emph{Betacoronavirus} sequences have been
-discovered,\textsuperscript{2,22,23} resulting in sparser phylogenetic
+discovered,\textsuperscript{2,13,19} resulting in sparser phylogenetic
 tree, thereby artificially inflating distinctiveness. Adding more
@@
 Finally, South-America is the range of Phyllostomidae, a group of bats
-that underwent explosive diversification events,\textsuperscript{20}
+that underwent explosive diversification events,\textsuperscript{18}
 which may drive the emergence of multiple viral lineages.
@@
 
-Despite the difference in evolutionary distinctiveness globally, there
+Despite differences in hotspots of evolutionary distinctiveness, there
 are reasons to expect that the phylogeography of bats and
 betacoronaviruses should show some degree of
-congruence.\textsuperscript{24,25} In particular, this should be the
+congruence.\textsuperscript{20,21} In particular, this should be the
 case if viruses can circulate among hosts and co-evolve with local hosts
@@
 immunological responses, as evidenced in several bat
-species.\textsuperscript{26} Immune characteristics that allow bats to
+species.\textsuperscript{22} Immune characteristics that allow bats to
 be better adapted to infection by emerging
-viruses,\textsuperscript{27,28} in addition to being hardcoded in their
-genome,\textsuperscript{29} may be related to a wide variety of
-diets,\textsuperscript{26,30} themselves likely to be driven by spatial
+viruses,\textsuperscript{23,24} in addition to being hardcoded in their
+genome,\textsuperscript{25} may be related to a wide variety of
+diets,\textsuperscript{22,26} themselves likely to be driven by spatial
 effects, especially at the local scale -- bats, indeed, occupy a variety
 of environments, and therefore display a variety of adaptations to these
-environments.\textsuperscript{31}
+environments.\textsuperscript{27}
 
@@
 nearby location where serological (antibody) evidence has indicated
-human exposure to SARS-like viruses;\textsuperscript{32} MERS-CoV in
+human exposure to SARS-like viruses;\textsuperscript{28} MERS-CoV in
 Saudi Arabia based on index cases available from a recently-published
-compendium of cases.\textsuperscript{33} For the latest event, most if
+compendium of cases.\textsuperscript{29} For the latest event, most if
 not all index cases are presumed to be camel-to-human transmission, and
@@
 uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti\textsuperscript{34} and Europe\textsuperscript{35} are not
+Haiti\textsuperscript{30} and Europe\textsuperscript{31} are not
 relevant here, as bats' involvement in these cycles of transmission have
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{36} Recent surveillance\textsuperscript{37}
+continents.\textsuperscript{32} Recent surveillance\textsuperscript{33}
 has identified a novel \emph{Betacoronavirus} (in the subgenus
@@
 increases the probability of pathogen spread more
-widely.\textsuperscript{38} Since human activity is required to amplify
+widely.\textsuperscript{34} Since human activity is required to amplify
 the frequency of virus encounters and thus create areas of viral
@@
 opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China,\textsuperscript{39} present
+looking for the origins of SARS in China,\textsuperscript{35} present
 serological evidence that strongest human-animal contact results in
@@
 risk/diversitifcation potential (Madagascar, South-America) overlay with
-sampling gaps for \emph{Betacoronavirus},\textsuperscript{40} stressing
+sampling gaps for \emph{Betacoronavirus},\textsuperscript{36} stressing
 the need for spatially targeted monitoring and discovery.
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{24,25} Chiropterans emerged around 64 million
+health.\textsuperscript{20,21} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{41,42}
+estimated richness of more than 1400 species.\textsuperscript{37,38}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{43} For example,
+ecosystem-derived benefits to human.\textsuperscript{39} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{44} Over two-thirds of bats are know to be
+networks.\textsuperscript{40} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{45,46} and
+contributing for agricultural pest control,\textsuperscript{41,42} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{47,48} Because bats are globally distributed and
+health.\textsuperscript{43,44} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 that have used host richness as proxy have predicted a high diversity of
-unsampled bat viruses,\textsuperscript{23} bat
+unsampled bat viruses,\textsuperscript{13} bat
 coronaviruses,\textsuperscript{2} and even specifically
-betacoronaviruses\textsuperscript{21} in both the Amazon and southeast
+betacoronaviruses\textsuperscript{14} in both the Amazon and southeast
 Asia. While we find that both regions are characterized by highly
@@
 high local diversity of viruses through host switching and viral
-recombination (see \emph{e.g.},\textsuperscript{49} as well as the
+recombination (see \emph{e.g.},\textsuperscript{45} as well as the
 discovery of recombinant viruses that share genetic material from both
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{50} Both of these regions are priority areas
+lineage).\textsuperscript{46} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{21,40} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{14,36} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{17}
+to the region.\textsuperscript{15}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{51}
+anthropogenic factors others than shifting climates.\textsuperscript{47}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{52} As our results establish that the
+habitats.\textsuperscript{48} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{53} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{54} This
+novel diseases,\textsuperscript{49} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{50} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{55} In particular, the challenge ahead lies
+encounters.\textsuperscript{51} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 \texttt{https://www.viralemergence.org/betacov} on
-Apr.~2022,\textsuperscript{21} and filtered it to ``known'' hosts
+Apr.~2022,\textsuperscript{14} and filtered it to ``known'' hosts
 (established before the emergence of SARS-CoV-2) and ``novel'' hosts
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{56} The range maps were subsequently rasterized
+data.\textsuperscript{52} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{53} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{58} based on a recent synthetic tree with
+Diversity\textsuperscript{54} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{59} Faith's PD measures the sum of unique
+species.\textsuperscript{55} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{60} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{56} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{61} transformed
+location, and following best practices\textsuperscript{57} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{62} shows that LCBD
+community. Recent results by\textsuperscript{58} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{63} This network stores pairwise values of
+mammals.\textsuperscript{59} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{64} In this approach, every risk component gets
+model.\textsuperscript{60} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{65} in that this allows multiple
+and low viral sharing,\textsuperscript{61} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{66} we measured the risk level by
+extraction,\textsuperscript{62} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{63} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
-with ModelFinder\textsuperscript{69} ultrafast bootstrap
-approximation\textsuperscript{70} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{64} v1.6.12
+with ModelFinder\textsuperscript{65} ultrafast bootstrap
+approximation\textsuperscript{66} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{71} for each of the viruses in the tree,
+distinctiveness\textsuperscript{67} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{72,73} who
+this idea, we loosely adapted a method from,\textsuperscript{68,69} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{74} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{70} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Drexler2014EcoEvo}{}%
+\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
 \CSLLeftMargin{13. }
-\CSLRightInline{Drexler, J. F., Corman, V. M. \& Drosten, C. Ecology,
-evolution and classification of bat coronaviruses in the aftermath of
-SARS. \emph{Antiviral Research} \textbf{101}, 45--56 (2014).}
-
-\leavevmode\hypertarget{ref-Calisher2006BatImp}{}%
+\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
+predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
+646--650 (2017).}
+
+\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
 \CSLLeftMargin{14. }
-\CSLRightInline{Calisher, C. H., Childs, J. E., Field, H. E., Holmes, K.
-V. \& Schountz, T. Bats: Important Reservoir Hosts of Emerging Viruses.
-\emph{Clinical Microbiology Reviews} \textbf{19}, 531--545 (2006).}
-
-\leavevmode\hypertarget{ref-Moratelli2015BatZoo}{}%
-\CSLLeftMargin{15. }
-\CSLRightInline{Moratelli, R. \& Calisher, C. H. Bats and zoonotic
-viruses: Can we confidently link bats with emerging deadly viruses?
-\emph{Memórias do Instituto Oswaldo Cruz} \textbf{110}, 1--22 (2015).}
-
-\leavevmode\hypertarget{ref-Mollentze2020VirZoo}{}%
-\CSLLeftMargin{16. }
-\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
-homogenous among taxonomic orders of mammalian and avian reservoir
-hosts. \emph{Proceedings of the National Academy of Sciences}
-\textbf{117}, 9423--9430 (2020).}
-
-\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{17. }
-\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
-cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
-doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
-
-\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
-\CSLLeftMargin{18. }
-\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
-Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}
-
-\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
-\CSLLeftMargin{19. }
-\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
-Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
-York, N.Y.)} \textbf{307}, 580--584 (2005).}
-
-\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
-\CSLLeftMargin{20. }
-\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
-New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
-Biogeography} \textbf{19}, 200--211 (2010).}
-
-\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-\CSLLeftMargin{21. }
 \CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
@@
 
+\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
+\CSLLeftMargin{15. }
+\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
+cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
+doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
+
+\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
+\CSLLeftMargin{16. }
+\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
+Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}
+
+\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
+\CSLLeftMargin{17. }
+\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
+Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
+York, N.Y.)} \textbf{307}, 580--584 (2005).}
+
+\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
+\CSLLeftMargin{18. }
+\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
+New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
+Biogeography} \textbf{19}, 200--211 (2010).}
+
 \leavevmode\hypertarget{ref-Allen2017GloHot}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
@@
 
-\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-\CSLLeftMargin{23. }
-\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
-predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
-646--650 (2017).}
-
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
@@
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
@@
 \leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
@@
 \leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
@@
 \leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
@@
 \leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{41. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{52. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{73. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{74. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@3f97d39ccff49b8819786ddb750ed9dd1c424697 🚀

@@
 
-\hypertarget{host-richness-does-not-predict-virus-distinctiveness}{%
-\subsection{Host richness does not predict virus
-distinctiveness}\label{host-richness-does-not-predict-virus-distinctiveness}}
+\hypertarget{hotspots-of-host-richness-and-viral-diversification-are-distinct}{%
+\subsection{Hotspots of host richness and viral diversification are
+distinct}\label{hotspots-of-host-richness-and-viral-diversification-are-distinct}}
 
 Bats, the second most diverse groups of mammals, are found worldwide;
-gradients in bat species richness generally track broader patterns of
+gradients in their species richness generally track broader patterns of
 mammal diversity, with a striking Neotropical hotspot (especially in the
@@
 peninsula. These hotspots of bat diversity are generally presumed to be
-hotspots of viral diversity, and therefore areas of concern for human
-health.\textsuperscript{2,13} However, the hotspots of bat
+hotspots of viral adaptive radiation, and therefore areas of concern for
+human health.\textsuperscript{2,13} However, the hotspots of bat
 betacoronavirus reservoirs show a distinct pattern, with primary
 hotspots (both in terms of size and higher values) of host richness
-situated primarily South-Eastern Asia, parts of Southern Europe, and to
-a lesser extent parts of Africa in the -25-0 range of latitudes (Fig. 1;
-top). Although hundreds of species likely host undiscovered
-betacoronaviruses, machine learning predictions have suggested that
-these undiscovered reservoirs should follow the same diversity
-gradient.\textsuperscript{14}
+situated primarily southeast Asia, parts of southern Europe, and to a
+lesser extent parts of Africa in the -25-0 range of latitudes
+(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
+undiscovered betacoronaviruses, machine learning predictions have
+suggested that these undiscovered reservoirs should follow the same
+diversity gradient.\textsuperscript{14} In principle, these hotspots of
+locally-diverse, virus-rich bat communities should drive more adaptive
+diversification in their viruses.
 
@@
 between the evolutionary distinctiveness of the hosts (grey to blue) and
-the viruses (grey to red). By contrast to the richness map, this reveals
-that South America has the most evolutionary distinct hosts \emph{and}
-viruses, whereas South-Eastern Asia and the Rift Valley region have
-mostly distinct viruses. This is congruent with know results about New
-World bats being evolutionary distinct, and suggests that they similarly
-have distinct viruses.}\label{fig:richness}
+the viruses (grey to red).}\label{fig:richness}
 }
@@
 
-Nevertheless, locally diverse and virus-rich bat communities could
-represent an increased risk of spillover under climate change through
-the creation of novel interactions,\textsuperscript{15} and therefore
-the diversity of \emph{Betacoronavirus} strains should similarly be
-ccounted for. In fig.~\ref{fig:richness} (bottom), we contrast the
-evolutionary distinctiveness of bats and viruses -- this reveals a
-slightly different portrait than bat richness alone. Chiropterans can be
-classified, from a macro-evolutionary standpoint, as Yangochiroptera and
-Yinpterochiroptera elsewhere.\textsuperscript{16,17} Specifically, we
-would expect that the so-called ``New World'' group of bats, being more
-evolutionary distinct, would also have evolutionary distinct viruses.
-Indeed fig.~\ref{fig:richness} (bottom) reveals it to be the case, and
-this region harbors a distinct bat-betacoronaviruses complex. This can
-be explained by the fact that Yangochiroptera, although not limited to
-the western hemisphere, contain the highly diverse adaptive radiation in
-the Phyllostomidae,\textsuperscript{18} which is restricted to the
-western hemisphere. By contrast, South-Eastern Asia has a lot of
-non-evolutionary distinct bats, who nevertheless hosted
-evolutionary-distinct viruses.
-
-It is noteworthy that outside of South America, viral evolutionary
-distinctiveness does not accurately track host diversity, with some
-areas having over-distinct viruses (eastern China but, oddly, not the
-rest of southeast Asia). There are a number of likely explanations.
-First, given the richness of bats in southeast Asia, many
-betacoronaviruses likely remain to be discovered in this region. Indeed,
-global predictions highlight that southeast Asia is a likely hostpot of
-unconfirmed hosts of betacoronaviruses,\textsuperscript{14} which would
-likely result in additional viral discoveries. This idea is unsurprising
-given the growing realization, especially since the emergence of
-SARS-CoV-2, that unique lineages of similar viruses are widespread in
-bats but still mostly undescribed. The most distinct
-bats-betacoronaviruses complex is found in South America, a region with
-a comparatively lower number of hosts; this matches with isolation
-through host vicariance, and may highlight a different co-evolutionary
-dynamic. Alternatively, this distinctiveness hostpot may be a product of
-under-sampling: South-America is one of the places where the fewest
-\emph{Betacoronavirus} sequences have been
-discovered,\textsuperscript{2,13,19} resulting in sparser phylogenetic
-tree, thereby artificially inflating distinctiveness. Adding more
-viruses would bring the distinctiveness of known sequences down.
-Finally, South-America is the range of Phyllostomidae, a group of bats
-that underwent explosive diversification events,\textsuperscript{18}
-which may drive the emergence of multiple viral lineages.
+However, we find that the global pattern of betacoronavirus phylogenetic
+distinctiveness is quite distinct from both bat host richness and
+phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
+contrast to the sparsity of Neotropical betacoronavirus hosts, South
+America has the most evolutionary distinct hosts \emph{and} viruses,
+followed by secondary hotspots in southeast Asia and the Rift Valley
+region have mostly distinct viruses. Some degree of sampling bias may
+contribute to these patterns: for example, South-America is one of the
+places where the fewest bat betacoronavirus sequences have been
+generated,\textsuperscript{2,13,15} resulting in a sparser phylogenetic
+tree, and artificially inflating distinctiveness; conversely,
+disproportionate research effort in eastern China\textsuperscript{16}
+may have led to a more complete inventory of the local diversity of
+coronaviruses, again inflating these metrics relative to underlying
+patterns. Even accounting for these potential biases, though, there are
+obvious and distinctive regional regimes of bat-betacoronavirus
+diversification.
+
+These global patterns recapitulate the evolutionary history of both the
+order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe bats
+(Rhinolophidae) are both the reservoirs of the SARS-like viruses
+(subgenus \emph{Sarbecovirus}) and the presumed ancestral host of
+\emph{Betacoronavirus}.\textsuperscript{17} The hotspot of host richness
+and viral diversity in southeast Asia---both of which are
+disproportionately high considering the global landscape of bat species
+richness---is almost entirely driven by dynamics in this
+clade.\textsuperscript{3,14} In contrast, the Neotropical hotspot is
+driven by a combination of distinct bat and virus adaptive radiation.
+The hotspot of bat diversity in the Americas is driven by the New World
+leaf-nosed bats (Phyllostomidae),\textsuperscript{18} a group of bats
+that have undergone multiple explosive diversification events--- but
+that are more often hosts of alphacoronaviruses. Out of the four main
+groups of betacoronaviruses, only the subgenus \emph{Merbecovirus}---the
+outgroup of the betacoronavirus tree---has been found in New World bats;
+while found worldwide, the merbecoviruses are almost entirely restricted
+to vesper bats (Vespertilionidae) {[}{]}. In this light, the hotspot of
+virus distinctiveness in the Americas is the result of isolation by
+vicariance--- a unique branch of coronavirus-bat coevolution, separate
+from the rest of the globe.
 
@@
 betacoronaviruses should show some degree of
-congruence.\textsuperscript{20,21} In particular, this should be the
+congruence.\textsuperscript{19,20} In particular, this should be the
 case if viruses can circulate among hosts and co-evolve with local hosts
@@
 immunological responses, as evidenced in several bat
-species.\textsuperscript{22} Immune characteristics that allow bats to
+species.\textsuperscript{21} Immune characteristics that allow bats to
 be better adapted to infection by emerging
-viruses,\textsuperscript{23,24} in addition to being hardcoded in their
-genome,\textsuperscript{25} may be related to a wide variety of
-diets,\textsuperscript{22,26} themselves likely to be driven by spatial
+viruses,\textsuperscript{22,23} in addition to being hardcoded in their
+genome,\textsuperscript{24} may be related to a wide variety of
+diets,\textsuperscript{21,25} themselves likely to be driven by spatial
 effects, especially at the local scale -- bats, indeed, occupy a variety
 of environments, and therefore display a variety of adaptations to these
-environments.\textsuperscript{27}
+environments.\textsuperscript{26}
 
@@
 nearby location where serological (antibody) evidence has indicated
-human exposure to SARS-like viruses;\textsuperscript{28} MERS-CoV in
+human exposure to SARS-like viruses;\textsuperscript{27} MERS-CoV in
 Saudi Arabia based on index cases available from a recently-published
-compendium of cases.\textsuperscript{29} For the latest event, most if
+compendium of cases.\textsuperscript{28} For the latest event, most if
 not all index cases are presumed to be camel-to-human transmission, and
@@
 uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti\textsuperscript{30} and Europe\textsuperscript{31} are not
+Haiti\textsuperscript{29} and Europe\textsuperscript{30} are not
 relevant here, as bats' involvement in these cycles of transmission have
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{32} Recent surveillance\textsuperscript{33}
+continents.\textsuperscript{31} Recent surveillance\textsuperscript{32}
 has identified a novel \emph{Betacoronavirus} (in the subgenus
@@
 increases the probability of pathogen spread more
-widely.\textsuperscript{34} Since human activity is required to amplify
+widely.\textsuperscript{33} Since human activity is required to amplify
 the frequency of virus encounters and thus create areas of viral
@@
 opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China,\textsuperscript{35} present
+looking for the origins of SARS in China,\textsuperscript{34} present
 serological evidence that strongest human-animal contact results in
@@
 risk/diversitifcation potential (Madagascar, South-America) overlay with
-sampling gaps for \emph{Betacoronavirus},\textsuperscript{36} stressing
+sampling gaps for \emph{Betacoronavirus},\textsuperscript{16} stressing
 the need for spatially targeted monitoring and discovery.
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{20,21} Chiropterans emerged around 64 million
+health.\textsuperscript{19,20} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{37,38}
+estimated richness of more than 1400 species.\textsuperscript{35,36}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{39} For example,
+ecosystem-derived benefits to human.\textsuperscript{37} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{40} Over two-thirds of bats are know to be
+networks.\textsuperscript{38} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{41,42} and
+contributing for agricultural pest control,\textsuperscript{39,40} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{43,44} Because bats are globally distributed and
+health.\textsuperscript{41,42} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 high local diversity of viruses through host switching and viral
-recombination (see \emph{e.g.},\textsuperscript{45} as well as the
+recombination (see \emph{e.g.},\textsuperscript{17} as well as the
 discovery of recombinant viruses that share genetic material from both
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{46} Both of these regions are priority areas
+lineage).\textsuperscript{43} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{14,36} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{14,16} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{15}
+to the region.\textsuperscript{44}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{47}
+anthropogenic factors others than shifting climates.\textsuperscript{45}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{48} As our results establish that the
+habitats.\textsuperscript{46} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{49} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{50} This
+novel diseases,\textsuperscript{47} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{48} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{51} In particular, the challenge ahead lies
+encounters.\textsuperscript{49} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{52} The range maps were subsequently rasterized
+data.\textsuperscript{50} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{53} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{51} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{54} based on a recent synthetic tree with
+Diversity\textsuperscript{52} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{55} Faith's PD measures the sum of unique
+species.\textsuperscript{53} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{56} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{54} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{57} transformed
+location, and following best practices\textsuperscript{55} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{58} shows that LCBD
+community. Recent results by\textsuperscript{56} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{59} This network stores pairwise values of
+mammals.\textsuperscript{57} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{60} In this approach, every risk component gets
+model.\textsuperscript{58} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{61} in that this allows multiple
+and low viral sharing,\textsuperscript{59} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{62} we measured the risk level by
+extraction,\textsuperscript{60} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{63} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{61} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{64} v1.6.12
-with ModelFinder\textsuperscript{65} ultrafast bootstrap
-approximation\textsuperscript{66} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{62} v1.6.12
+with ModelFinder\textsuperscript{63} ultrafast bootstrap
+approximation\textsuperscript{64} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{67} for each of the viruses in the tree,
+distinctiveness\textsuperscript{65} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{68,69} who
+this idea, we loosely adapted a method from,\textsuperscript{66,67} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{70} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{68} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
+\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
 \CSLLeftMargin{15. }
-\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
-cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
-doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
-
-\leavevmode\hypertarget{ref-Springer2013PhyBat}{}%
+\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
+of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
+(2017).}
+
+\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
 \CSLLeftMargin{16. }
-\CSLRightInline{Springer, M. S. Phylogenetics: Bats United, Microbats
-Divided. \emph{Current Biology} \textbf{23}, R999--R1001 (2013).}
-
-\leavevmode\hypertarget{ref-Teeling2005MolPhy}{}%
+\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
+Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
+bat coronaviruses. 2022.06.15.496296 (2022)
+doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}
+
+\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
 \CSLLeftMargin{17. }
-\CSLRightInline{Teeling, E. C. \emph{et al.} A Molecular Phylogeny for
-Bats Illuminates Biogeography and the Fossil Record. \emph{Science (New
-York, N.Y.)} \textbf{307}, 580--584 (2005).}
+\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
+transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
+Server for Biology} 2020.05.31.116061 (2020)
+doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
 
@@
 
-\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
+\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
 \CSLLeftMargin{19. }
-\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
-of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
-(2017).}
-
-\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{20. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
@@
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
@@
 \leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
@@
 \leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
@@
 \leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
@@
 \leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 
-\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{36. }
-\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
-Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
-bat coronaviruses. 2022.06.15.496296 (2022)
-doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}
-
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{39. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 
-\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-\CSLLeftMargin{45. }
-\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
-transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
-Server for Biology} 2020.05.31.116061 (2020)
-doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
-
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 
+\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
+\CSLLeftMargin{44. }
+\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
+cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
+doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
+
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{50. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@7fbeb69803b427454cd7d1bde2843f3d17331183 🚀

@@
 coronaviruses, again inflating these metrics relative to underlying
-patterns. Even accounting for these potential biases, though, there are
-obvious and distinctive regional regimes of bat-betacoronavirus
-diversification.
-
-These global patterns recapitulate the evolutionary history of both the
-order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe bats
+patterns. Even accounting for these potential biases, though, there is
+obvious heterogeneity in betacoronavirus evolutionary distinctiveness.
+
+These patterns recapitulate the evolutionary history of both the order
+Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe bats
 (Rhinolophidae) are both the reservoirs of the SARS-like viruses
@@
 and viral diversity in southeast Asia---both of which are
-disproportionately high considering the global landscape of bat species
-richness---is almost entirely driven by dynamics in this
-clade.\textsuperscript{3,14} In contrast, the Neotropical hotspot is
-driven by a combination of distinct bat and virus adaptive radiation.
-The hotspot of bat diversity in the Americas is driven by the New World
-leaf-nosed bats (Phyllostomidae),\textsuperscript{18} a group of bats
-that have undergone multiple explosive diversification events--- but
-that are more often hosts of alphacoronaviruses. Out of the four main
-groups of betacoronaviruses, only the subgenus \emph{Merbecovirus}---the
-outgroup of the betacoronavirus tree---has been found in New World bats;
-while found worldwide, the merbecoviruses are almost entirely restricted
-to vesper bats (Vespertilionidae) {[}{]}. In this light, the hotspot of
-virus distinctiveness in the Americas is the result of isolation by
-vicariance--- a unique branch of coronavirus-bat coevolution, separate
-from the rest of the globe.
+disproportionately high, considering the global landscape of bat species
+richness---is almost entirely driven by adaptive radiation in this
+clade\textsuperscript{3,14}. In contrast, the Neotropical hotspot of
+viral distinctiveness is the result of isolation by vicariance for both
+bats and their viruses. Out of the four main groups of
+betacoronaviruses, only the subgenus \emph{Merbecovirus} has been found
+in animals in the Americas---an introduction that is generally presumed
+to be ancient. While comparatively understudied, New World
+merbecoviruses have been found in the ghost-faced bats (Mormoopidae),
+New World leaf-nosed bats (Phyllostomidae), and free-tailed bats
+(Molossidae) (add cite: Olival 2020 PLoS Pathogens). The former two
+groups are endemic to the Neotropics, while the latter two and
+particularly the phyllostomids are responsible for the hotspot of bat
+diversity in the Amazon. Together, these groups play host to a distinct
+regime of bat-betacoronavirus coevolution in the Americas.
 
@@
 betacoronaviruses should show some degree of
-congruence.\textsuperscript{19,20} In particular, this should be the
+congruence.\textsuperscript{18,19} In particular, this should be the
 case if viruses can circulate among hosts and co-evolve with local hosts
@@
 immunological responses, as evidenced in several bat
-species.\textsuperscript{21} Immune characteristics that allow bats to
+species.\textsuperscript{20} Immune characteristics that allow bats to
 be better adapted to infection by emerging
-viruses,\textsuperscript{22,23} in addition to being hardcoded in their
-genome,\textsuperscript{24} may be related to a wide variety of
-diets,\textsuperscript{21,25} themselves likely to be driven by spatial
+viruses,\textsuperscript{21,22} in addition to being hardcoded in their
+genome,\textsuperscript{23} may be related to a wide variety of
+diets,\textsuperscript{20,24} themselves likely to be driven by spatial
 effects, especially at the local scale -- bats, indeed, occupy a variety
 of environments, and therefore display a variety of adaptations to these
-environments.\textsuperscript{26}
+environments.\textsuperscript{25}
 
@@
 nearby location where serological (antibody) evidence has indicated
-human exposure to SARS-like viruses;\textsuperscript{27} MERS-CoV in
+human exposure to SARS-like viruses;\textsuperscript{26} MERS-CoV in
 Saudi Arabia based on index cases available from a recently-published
-compendium of cases.\textsuperscript{28} For the latest event, most if
+compendium of cases.\textsuperscript{27} For the latest event, most if
 not all index cases are presumed to be camel-to-human transmission, and
@@
 uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti\textsuperscript{29} and Europe\textsuperscript{30} are not
+Haiti\textsuperscript{28} and Europe\textsuperscript{29} are not
 relevant here, as bats' involvement in these cycles of transmission have
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{31} Recent surveillance\textsuperscript{32}
+continents.\textsuperscript{30} Recent surveillance\textsuperscript{31}
 has identified a novel \emph{Betacoronavirus} (in the subgenus
@@
 increases the probability of pathogen spread more
-widely.\textsuperscript{33} Since human activity is required to amplify
+widely.\textsuperscript{32} Since human activity is required to amplify
 the frequency of virus encounters and thus create areas of viral
@@
 opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China,\textsuperscript{34} present
+looking for the origins of SARS in China,\textsuperscript{33} present
 serological evidence that strongest human-animal contact results in
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{19,20} Chiropterans emerged around 64 million
+health.\textsuperscript{18,19} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{35,36}
+estimated richness of more than 1400 species.\textsuperscript{34,35}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{37} For example,
+ecosystem-derived benefits to human.\textsuperscript{36} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{38} Over two-thirds of bats are know to be
+networks.\textsuperscript{37} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{39,40} and
+contributing for agricultural pest control,\textsuperscript{38,39} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{41,42} Because bats are globally distributed and
+health.\textsuperscript{40,41} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{43} Both of these regions are priority areas
+lineage).\textsuperscript{42} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{44}
+to the region.\textsuperscript{43}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{45}
+anthropogenic factors others than shifting climates.\textsuperscript{44}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{46} As our results establish that the
+habitats.\textsuperscript{45} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{47} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{48} This
+novel diseases,\textsuperscript{46} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{47} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{49} In particular, the challenge ahead lies
+encounters.\textsuperscript{48} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{50} The range maps were subsequently rasterized
+data.\textsuperscript{49} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{51} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{52} based on a recent synthetic tree with
+Diversity\textsuperscript{51} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{53} Faith's PD measures the sum of unique
+species.\textsuperscript{52} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{54} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{53} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{55} transformed
+location, and following best practices\textsuperscript{54} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{56} shows that LCBD
+community. Recent results by\textsuperscript{55} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{57} This network stores pairwise values of
+mammals.\textsuperscript{56} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{58} In this approach, every risk component gets
+model.\textsuperscript{57} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{59} in that this allows multiple
+and low viral sharing,\textsuperscript{58} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{60} we measured the risk level by
+extraction,\textsuperscript{59} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{61} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{60} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{62} v1.6.12
-with ModelFinder\textsuperscript{63} ultrafast bootstrap
-approximation\textsuperscript{64} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{61} v1.6.12
+with ModelFinder\textsuperscript{62} ultrafast bootstrap
+approximation\textsuperscript{63} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{65} for each of the viruses in the tree,
+distinctiveness\textsuperscript{64} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{66,67} who
+this idea, we loosely adapted a method from,\textsuperscript{65,66} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{68} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{67} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Villalobos2010DivFie}{}%
+\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
 \CSLLeftMargin{18. }
-\CSLRightInline{Villalobos, F. \& Arita, H. T. The diversity field of
-New World leaf-nosed bats (Phyllostomidae). \emph{Global Ecology and
-Biogeography} \textbf{19}, 200--211 (2010).}
-
-\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{19. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
@@
 \leavevmode\hypertarget{ref-Irving2021LesHos}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
@@
 \leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
@@
 \leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
@@
 \leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
@@
 \leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{38. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{49. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f7b24e3dc31af379be75123d05d7c99049d263fe 🚀

@@
 patterns. Even accounting for these potential biases, though, there is
-obvious heterogeneity in betacoronavirus evolutionary distinctiveness.
-
-These patterns recapitulate the evolutionary history of both the order
-Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe bats
-(Rhinolophidae) are both the reservoirs of the SARS-like viruses
-(subgenus \emph{Sarbecovirus}) and the presumed ancestral host of
+obvious heterogeneity in betacoronavirus evolutionary distinctiveness
+that is distinct from overall bat diversity.
+
+On closer inspection, these patterns recapitulate the evolutionary
+history of both the order Chiroptera and the genus
+\emph{Betacoronavirus}. Horseshoe bats (Rhinolophidae) are both the
+reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}) and
+the presumed ancestral host of
 \emph{Betacoronavirus}.\textsuperscript{17} The hotspot of host richness

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@cf5941efd35908a02185710410fc5afdf838af52 🚀

@@
 disproportionately high, considering the global landscape of bat species
-richness---is almost entirely driven by adaptive radiation in this
-clade\textsuperscript{3,14}. In contrast, the Neotropical hotspot of
-viral distinctiveness is the result of isolation by vicariance for both
-bats and their viruses. Out of the four main groups of
-betacoronaviruses, only the subgenus \emph{Merbecovirus} has been found
-in animals in the Americas---an introduction that is generally presumed
-to be ancient. While comparatively understudied, New World
-merbecoviruses have been found in the ghost-faced bats (Mormoopidae),
-New World leaf-nosed bats (Phyllostomidae), and free-tailed bats
-(Molossidae) (add cite: Olival 2020 PLoS Pathogens). The former two
-groups are endemic to the Neotropics, while the latter two and
-particularly the phyllostomids are responsible for the hotspot of bat
-diversity in the Amazon. Together, these groups play host to a distinct
-regime of bat-betacoronavirus coevolution in the Americas.
+richness---is almost entirely driven by viral adaptive radiation through
+host switching within this clade\textsuperscript{3,14}. In contrast, the
+Neotropical hotspot of viral distinctiveness is driven by isolation by
+host vicariance. Out of the four main groups of betacoronaviruses, only
+the subgenus \emph{Merbecovirus} (MERS-like viruses) has been found in
+animals in the Americas---an introduction that is generally presumed to
+be ancient.\textsuperscript{3} While comparatively understudied, New
+World merbecoviruses have been found in the ghost-faced bats
+(Mormoopidae), New World leaf-nosed bats (Phyllostomidae), and
+free-tailed bats (Molossidae) (add cite: Olival 2020 PLoS Pathogens).
+The former two groups are endemic to the Neotropics, while the explosive
+adaptive radiations of the latter two (and particularly the
+phyllostomids) are responsible for the hotspot of bat diversity in the
+Amazon. Together, these clades of New World bats play host to a distinct
+regime of betacoronavirus coevolution.
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e9d427c180600ce8759a5f59777d64c33b08a09b 🚀

@@
 host diversity and viral emergence. Working from a recently published
-database of bat hosts of betacoronaviruses, we develop the first global
-maps of both host and virus evolutionary distinctiveness and
-biogeographic regions for this system. Aiming to explain these patterns,
-we develop a generalized framework for applying the GMTC to host-virus
-interactions, with a specific emphasis on the potential to create
-independent coevolutionary dynamics (and therefore spatial fragmentation
-in risk) through heterogeneity. We develop a trivariate risk assessment
-system that connects each GMTC mechanism to a quantifiable aspect of
-host-virus interactions: (i) viral sharing rates in host communities,
-representing the strength of potential interaction between viruses and
-any one host (i.e., places where viruses undergo constant host switching
-may be coevolutionary coldspots); (ii) the phylogenetic diversity of
-hosts, as a proxy for variation in the immunological mechanisms that
-antagonize viruses (i.e., the selection mosaic); and (iii) the local
-uniqueness of the bat community, representing the potential for viruses
-to be exposed to novel host traits (e.g., variation in receptor
-sequences). Together, we argue that these can be used to identify and
-map the evolutionary drivers that---in conjunction with transmission
-processes (e.g., viral prevalence in reservoirs and animal-human contact
-rates)--- determine disease emergence risk.
+database of bat hosts of betacoronaviruses, we test whether spatial
+structure in bat-betacoronavirus coevolution is identifiable at a global
+scale. Aiming to explain these patterns, we develop a generalized
+framework for applying the GMTC to host-virus interactions, with a
+specific emphasis on the potential to create independent coevolutionary
+dynamics (and therefore spatial fragmentation in risk) through
+heterogeneity. We develop a trivariate risk assessment system that
+connects each GMTC mechanism to a quantifiable aspect of host-virus
+interactions: (i) viral sharing rates in host communities, representing
+the strength of potential interaction between viruses and any one host
+(i.e., places where viruses undergo constant host switching may be
+coevolutionary coldspots); (ii) the phylogenetic diversity of hosts, as
+a proxy for variation in the immunological mechanisms that antagonize
+viruses (i.e., the selection mosaic); and (iii) the local uniqueness of
+the bat community, representing the potential for viruses to be exposed
+to novel host traits (e.g., variation in receptor sequences). Together,
+we argue that these can be used to identify and map the evolutionary
+drivers that---in conjunction with transmission processes (e.g., viral
+prevalence in reservoirs and animal-human contact rates)--- determine
+disease emergence risk.
 
@@
 \emph{Betacoronavirus}. Horseshoe bats (Rhinolophidae) are both the
-reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}) and
-the presumed ancestral host of
-\emph{Betacoronavirus}.\textsuperscript{17} The hotspot of host richness
-and viral diversity in southeast Asia---both of which are
-disproportionately high, considering the global landscape of bat species
-richness---is almost entirely driven by viral adaptive radiation through
-host switching within this clade\textsuperscript{3,14}. In contrast, the
-Neotropical hotspot of viral distinctiveness is driven by isolation by
-host vicariance. Out of the four main groups of betacoronaviruses, only
-the subgenus \emph{Merbecovirus} (MERS-like viruses) has been found in
-animals in the Americas---an introduction that is generally presumed to
-be ancient.\textsuperscript{3} While comparatively understudied, New
-World merbecoviruses have been found in the ghost-faced bats
-(Mormoopidae), New World leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae) (add cite: Olival 2020 PLoS Pathogens).
-The former two groups are endemic to the Neotropics, while the explosive
-adaptive radiations of the latter two (and particularly the
-phyllostomids) are responsible for the hotspot of bat diversity in the
-Amazon. Together, these clades of New World bats play host to a distinct
-regime of betacoronavirus coevolution.
-
-\hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
-\subsection{The phylogeographic regions of hosts and their viruses
-overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}
-
-Despite differences in hotspots of evolutionary distinctiveness, there
-are reasons to expect that the phylogeography of bats and
-betacoronaviruses should show some degree of
-congruence.\textsuperscript{18,19} In particular, this should be the
-case if viruses can circulate among hosts and co-evolve with local hosts
-communities, making their evolutionary process more than a byproduct of
-host evolution. High density of hosts sharing the same virus (albeit
-possibly different strains) can drive or result from evolution of the
-bat antiviral immune system, resulting in spatially distinct
-immunological responses, as evidenced in several bat
-species.\textsuperscript{20} Immune characteristics that allow bats to
-be better adapted to infection by emerging
-viruses,\textsuperscript{21,22} in addition to being hardcoded in their
-genome,\textsuperscript{23} may be related to a wide variety of
-diets,\textsuperscript{20,24} themselves likely to be driven by spatial
-effects, especially at the local scale -- bats, indeed, occupy a variety
-of environments, and therefore display a variety of adaptations to these
-environments.\textsuperscript{25}
+reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}) and a
+possible ancestral host of \emph{Betacoronavirus}.\textsuperscript{17}
+The hotspots of host richness and viral diversity in southeast
+Asia---both of which are disproportionately high, considering the global
+landscape of bat species richness---are almost entirely driven by viral
+adaptive radiation through host switching within this
+clade\textsuperscript{3,14}. In contrast, the Neotropical hotspot of
+viral distinctiveness is driven by isolation by host vicariance. Out of
+the four main groups of betacoronaviruses, only the subgenus
+\emph{Merbecovirus} (MERS-like viruses) has been found in animals in the
+Americas---an introduction that is generally presumed to be
+ancient.\textsuperscript{3} While comparatively understudied, New World
+merbecoviruses have been found in the ghost-faced bats (Mormoopidae),
+New World leaf-nosed bats (Phyllostomidae), and free-tailed bats
+(Molossidae) (add cite: Olival 2020 PLoS Pathogens). The former two
+groups are endemic to the Neotropics, while the explosive adaptive
+radiations of the latter two (and particularly the phyllostomids) are
+responsible for the hotspot of bat diversity in the Amazon. Together,
+these clades of New World bats play host to a distinct regime of
+betacoronavirus coevolution.
+
+\hypertarget{global-biogeographic-regions-are-consistent-for-bats-and-betacoronaviruses}{%
+\subsection{Global biogeographic regions are consistent for bats and
+betacoronaviruses}\label{global-biogeographic-regions-are-consistent-for-bats-and-betacoronaviruses}}
+
+Most previous work has assumed that coronavirus biogeography is driven
+by coevolutionary regimes that form at finer taxonomic scales, treating
+the presence or richness of key bat host groups as predictive of these
+viruses' distribution.\textsuperscript{2,3} Projecting bat and
+betacoronavirus phylogeny over space (fig.~\ref{fig:biogeo}), we find
+further support for the idea that bat community assembly is directly
+responsible for spatial heterogeneity in viral coevolutionary regimes.
+The distinct groupings (represented by different colors, symbolizing
+positions in a subspace formed by the first two phylogenetic principal
+components) are essentially equivalent between the two groups, and can
+be coarsely delineated as (1) south and southeast Asia, (2) east Asia
+(including northern China), west Asia, and the Mediterranean coast; (3)
+Eurasia above a northing of 40; and (4) Africa and south America. These
+results suggest that, although the evolutionary distinctiveness of the
+bat-betacoronavirus complex varies spatially, biogeographic regions are
+consistent between bats and their viruses. In some cases, this may
+diverge from expectations about coronavirus biogeography: for example,
+previous work has rarely flagged India as a region of concern, but for
+both bats and betacoronaviruses, the subcontinent falls into the same
+phylogeographic regions as the southeast Asian peninsula (and indeed,
+the region is home to known bat hosts of nobecoviruses, sarbecoviruses,
+and merbecoviruses {[}Ruiz-Aravena2022EcoEvo{]}).
 
@@
 \end{figure}
-
-In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
-signal of bats (top) and viruses (bottom) in space; the distinct
-groupings (represented by different colors symbolizing positions in the
-subspace formed by the first two axes of the PCoA) are essentially
-equivalent between the two groups, and can be coarsely delineated as
-southeast Asia, Eurasia above a northing of 25, and Africa and south
-America. These results suggest that, although the evolutionary
-distinctiveness of the bat-betacoronaviruses complex varies spatially,
-the system shows an important degree of spatial consistency, with a
-reduced number of bioregions. Available information describing the
-spillover of zoonotic betacoronaviruses of bat origin where data was
-available before and up through the COVID-19 pandemic puts spillover
-events of SARS-CoV-2 in Wuhan, China; SARS-CoV in Guangdong, China based
-on the presence of closest known viruses circulating in nature, and a
-nearby location where serological (antibody) evidence has indicated
-human exposure to SARS-like viruses;\textsuperscript{26} MERS-CoV in
-Saudi Arabia based on index cases available from a recently-published
-compendium of cases.\textsuperscript{27} For the latest event, most if
-not all index cases are presumed to be camel-to-human transmission, and
-the precise origin point (if it exists) of MERS-CoV in bats is
-uncertain. Recent recombinant canine coronavirus spillover events in
-Haiti\textsuperscript{28} and Europe\textsuperscript{29} are not
-relevant here, as bats' involvement in these cycles of transmission have
-been supposed to be non-existent. These index cases fall within
-different phylogeographic bioregions (fig.~\ref{fig:biogeo}), which
-further highlight the issue that different host-virus sub-systems may
-lead to widespread emergence.
 
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{30} Recent surveillance\textsuperscript{31}
+continents.\textsuperscript{18} Recent surveillance\textsuperscript{19}
 has identified a novel \emph{Betacoronavirus} (in the subgenus
@@
 increases the probability of pathogen spread more
-widely.\textsuperscript{32} Since human activity is required to amplify
+widely.\textsuperscript{20} Since human activity is required to amplify
 the frequency of virus encounters and thus create areas of viral
@@
 opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China,\textsuperscript{33} present
+looking for the origins of SARS in China,\textsuperscript{21} present
 serological evidence that strongest human-animal contact results in
@@
 
+PUT THIS SOMEWHERE: Available information describing the spillover of
+zoonotic betacoronaviruses of bat origin where data was available before
+and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
+in Wuhan, China; SARS-CoV in Guangdong, China based on the presence of
+closest known viruses circulating in nature, and a nearby location where
+serological (antibody) evidence has indicated human exposure to
+SARS-like viruses;\textsuperscript{22} MERS-CoV in Saudi Arabia based on
+index cases available from a recently-published compendium of
+cases.\textsuperscript{23} For the latest event, most if not all index
+cases are presumed to be camel-to-human transmission, and the precise
+origin point (if it exists) of MERS-CoV in bats is uncertain. Recent
+recombinant canine coronavirus spillover events in
+Haiti\textsuperscript{24} and Europe\textsuperscript{25} are not
+relevant here, as bats' involvement in these cycles of transmission have
+been supposed to be non-existent. These index cases fall within
+different phylogeographic bioregions (fig.~\ref{fig:biogeo}), which
+further highlight the issue that different host-virus sub-systems may
+lead to widespread emergence.
+
 \hypertarget{conclusion}{%
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{18,19} Chiropterans emerged around 64 million
+health.\textsuperscript{26,27} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{34,35}
+estimated richness of more than 1400 species.\textsuperscript{28,29}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{36} For example,
+ecosystem-derived benefits to human.\textsuperscript{30} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{37} Over two-thirds of bats are know to be
+networks.\textsuperscript{31} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{38,39} and
+contributing for agricultural pest control,\textsuperscript{32,33} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{40,41} Because bats are globally distributed and
+health.\textsuperscript{34,35} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{42} Both of these regions are priority areas
+lineage).\textsuperscript{36} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{43}
+to the region.\textsuperscript{37}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{44}
+anthropogenic factors others than shifting climates.\textsuperscript{38}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{45} As our results establish that the
+habitats.\textsuperscript{39} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{46} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{47} This
+novel diseases,\textsuperscript{40} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{41} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{48} In particular, the challenge ahead lies
+encounters.\textsuperscript{42} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{49} The range maps were subsequently rasterized
+data.\textsuperscript{43} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{44} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{51} based on a recent synthetic tree with
+Diversity\textsuperscript{45} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{52} Faith's PD measures the sum of unique
+species.\textsuperscript{46} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{53} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{47} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{54} transformed
+location, and following best practices\textsuperscript{48} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{55} shows that LCBD
+community. Recent results by\textsuperscript{49} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{56} This network stores pairwise values of
+mammals.\textsuperscript{50} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{57} In this approach, every risk component gets
+model.\textsuperscript{51} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{58} in that this allows multiple
+and low viral sharing,\textsuperscript{52} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{59} we measured the risk level by
+extraction,\textsuperscript{53} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{60} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{54} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{61} v1.6.12
-with ModelFinder\textsuperscript{62} ultrafast bootstrap
-approximation\textsuperscript{63} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{55} v1.6.12
+with ModelFinder\textsuperscript{56} ultrafast bootstrap
+approximation\textsuperscript{57} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{64} for each of the viruses in the tree,
+distinctiveness\textsuperscript{58} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{65,66} who
+this idea, we loosely adapted a method from,\textsuperscript{59,60} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{67} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{61} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
+\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
 \CSLLeftMargin{18. }
-\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
-virome diversity in bats. \emph{Current Opinion in Virology}
-\textbf{52}, 192--202 (2022).}
-
-\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
+\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
+Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
+Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}
+
+\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
 \CSLLeftMargin{19. }
-\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
-K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
-\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}
-
-\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
+\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
+Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
+for This Coronavirus Clade. \emph{Frontiers in Public Health}
+\textbf{10}, (2022).}
+
+\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
 \CSLLeftMargin{20. }
-\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
-Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}
-
-\leavevmode\hypertarget{ref-Gorbunova2020WorGoe}{}%
+\CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
+Gómez-Gardeñes, J. \& Ghoshal, G. Interplay between population density
+and mobility in determining the spread of epidemics in cities.
+\emph{Communications Physics} \textbf{4}, 191 (2021).}
+
+\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
 \CSLLeftMargin{21. }
-\CSLRightInline{Gorbunova, V., Seluanov, A. \& Kennedy, B. K. The World
-Goes Bats: Living Longer and Tolerating Viruses. \emph{Cell Metabolism}
-\textbf{32}, 31--43 (2020).}
-
-\leavevmode\hypertarget{ref-Irving2021LesHos}{}%
+\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
+Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
+1030--1037 (2004).}
+
+\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
 \CSLLeftMargin{22. }
-\CSLRightInline{Irving, A. T., Ahn, M., Goh, G., Anderson, D. E. \&
-Wang, L.-F. Lessons from the host defences of bats, a unique viral
-reservoir. \emph{Nature} \textbf{589}, 363--370 (2021).}
-
-\leavevmode\hypertarget{ref-Jebb2020SixRef}{}%
-\CSLLeftMargin{23. }
-\CSLRightInline{Jebb, D. \emph{et al.} Six reference-quality genomes
-reveal evolution of bat adaptations. \emph{Nature} \textbf{583},
-578--584 (2020).}
-
-\leavevmode\hypertarget{ref-MorenoSantillan2021LarGen}{}%
-\CSLLeftMargin{24. }
-\CSLRightInline{Moreno Santillán, D. D. \emph{et al.} Large-scale genome
-sampling reveals unique immunity and metabolic adaptations in bats.
-\emph{Molecular Ecology} mec.16027 (2021)
-doi:\href{https://doi.org/10.1111/mec.16027}{10.1111/mec.16027}.}
-
-\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{25. }
-\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
-distribution of bat hosts of sarbecoviruses: Implications for
-conservation and public health. \emph{Proceedings of the Royal Society
-B: Biological Sciences} \textbf{289}, 20220397 (2022).}
-
-\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{26. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
@@
 \leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
@@
 \leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
@@
 
-\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{30. }
-\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
-Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
-Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}
-
-\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{31. }
-\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
-Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
-for This Coronavirus Clade. \emph{Frontiers in Public Health}
-\textbf{10}, (2022).}
-
-\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
-\CSLLeftMargin{32. }
-\CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
-Gómez-Gardeñes, J. \& Ghoshal, G. Interplay between population density
-and mobility in determining the spread of epidemics in cities.
-\emph{Communications Physics} \textbf{4}, 191 (2021).}
-
-\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{33. }
-\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
-Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
-1030--1037 (2004).}
+\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
+\CSLLeftMargin{26. }
+\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
+K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
+\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}
+
+\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
+\CSLLeftMargin{27. }
+\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
+virome diversity in bats. \emph{Current Opinion in Virology}
+\textbf{52}, 192--202 (2022).}
 
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{32. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{43. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@bd8dbb2df687e024451db49cf6e4be7914267245 🚀

@@
 (including northern China), west Asia, and the Mediterranean coast; (3)
-Eurasia above a northing of 40; and (4) Africa and south America. These
-results suggest that, although the evolutionary distinctiveness of the
-bat-betacoronavirus complex varies spatially, biogeographic regions are
-consistent between bats and their viruses. In some cases, this may
-diverge from expectations about coronavirus biogeography: for example,
-previous work has rarely flagged India as a region of concern, but for
-both bats and betacoronaviruses, the subcontinent falls into the same
-phylogeographic regions as the southeast Asian peninsula (and indeed,
-the region is home to known bat hosts of nobecoviruses, sarbecoviruses,
-and merbecoviruses {[}Ruiz-Aravena2022EcoEvo{]}).
+Eurasia above a northing of 40; and (4) Africa and south America. In
+some cases, this diverges from expectations about coronavirus
+biogeography: for example, previous work has rarely flagged India as a
+region of concern, but for both bats and betacoronaviruses, the
+subcontinent falls into the same phylogeographic regions as the
+southeast Asian peninsula (and indeed, the region is home to known bat
+hosts of nobecoviruses, sarbecoviruses, and merbecoviruses
+{[}Ruiz-Aravena2022EcoEvo{]}).
 
@@
 \end{figure}
+
+Overall, these results suggest that, although there are unique hotspots
+of viral adaptive radiation, biogeographic regions are mostly consistent
+between bats and their viruses. This may be surprising, given that
+cospeciation plays a minor role in coronavirus
+diversification,\textsuperscript{2} which could theoretically allow for
+broad divergence in their biogeography. However, host jumps at the
+family level or higher are relatively rare and significant events in
+coronavirus evolutionary history;\textsuperscript{2,17} as a result, we
+suggest the mosaic of betacoronavirus phylogeography is assembled from a
+set of overlapping smaller coevolutionary systems, superimposed in space
+and filtered by the importance of different subgroups in local host
+communities. For example, the most speciose and cosmopolitan family of
+bats, the vesper bats (Vespertilionidae), are considered the primary
+reservoir of merbecoviruses;\textsuperscript{3,17} but in the Americas,
+where merbecoviruses are the only lineage present, they have only been
+found in other bat taxa. At the coarsest scale, these heterogeneities
+are lost, and betacoronavirus biogeography tracks the deep rifts in bat
+evolutionary history---but within these broad regions, the component
+coevolutionary systems can have very different dynamics.
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@19dcaf28f6830ab06f98bf10806dff89566caacf 🚀

@@
 southeast Asian peninsula (and indeed, the region is home to known bat
-hosts of nobecoviruses, sarbecoviruses, and merbecoviruses
-{[}Ruiz-Aravena2022EcoEvo{]}).
+hosts of nobecoviruses, sarbecoviruses, and
+merbecoviruses).\textsuperscript{3}
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@05560fa1bb7abf6736a3912e8ae484ca2c47d35e 🚀

@@
 \emph{Betacoronavirus}. Horseshoe bats (Rhinolophidae) are both the
-reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}) and a
-possible ancestral host of \emph{Betacoronavirus}.\textsuperscript{17}
-The hotspots of host richness and viral diversity in southeast
-Asia---both of which are disproportionately high, considering the global
-landscape of bat species richness---are almost entirely driven by viral
-adaptive radiation through host switching within this
-clade\textsuperscript{3,14}. In contrast, the Neotropical hotspot of
-viral distinctiveness is driven by isolation by host vicariance. Out of
-the four main groups of betacoronaviruses, only the subgenus
-\emph{Merbecovirus} (MERS-like viruses) has been found in animals in the
-Americas---an introduction that is generally presumed to be
-ancient.\textsuperscript{3} While comparatively understudied, New World
-merbecoviruses have been found in the ghost-faced bats (Mormoopidae),
-New World leaf-nosed bats (Phyllostomidae), and free-tailed bats
-(Molossidae) (add cite: Olival 2020 PLoS Pathogens). The former two
-groups are endemic to the Neotropics, while the explosive adaptive
-radiations of the latter two (and particularly the phyllostomids) are
-responsible for the hotspot of bat diversity in the Amazon. Together,
-these clades of New World bats play host to a distinct regime of
-betacoronavirus coevolution.
+reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}), the
+group of pandemic threats that have been of the greatest interest to
+researchers\textsuperscript{17} (and so have been sampled most
+intensively).\textsuperscript{16} The hotspots of host richness and
+viral diversity in southeast Asia---both of which are disproportionately
+high, considering the global landscape of bat species richness---are
+almost entirely driven by viral adaptive radiation through host
+switching within this clade\textsuperscript{3,14}. In contrast, the
+Neotropical hotspot of viral distinctiveness is driven by isolation by
+host vicariance. Out of the four main groups of betacoronaviruses, only
+the subgenus \emph{Merbecovirus} (MERS-like viruses) has been found in
+animals in the Americas---an introduction that is generally presumed to
+be ancient.\textsuperscript{3} While comparatively understudied, New
+World merbecoviruses have been found in the ghost-faced bats
+(Mormoopidae), New World leaf-nosed bats (Phyllostomidae), and
+free-tailed bats (Molossidae) (add cite: Olival 2020 PLoS Pathogens).
+The former two groups are endemic to the Neotropics, while the explosive
+adaptive radiations of the latter two (and particularly the
+phyllostomids) are responsible for the hotspot of bat diversity in the
+Amazon. Together, these clades of New World bats play host to a distinct
+regime of betacoronavirus coevolution.
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@8d14abcf17db4e1500203637797d46cbe3a3af37 🚀

@@
 
-On closer inspection, these patterns recapitulate the evolutionary
-history of both the order Chiroptera and the genus
-\emph{Betacoronavirus}. Horseshoe bats (Rhinolophidae) are both the
-reservoirs of the SARS-like viruses (subgenus \emph{Sarbecovirus}), the
-group of pandemic threats that have been of the greatest interest to
-researchers\textsuperscript{17} (and so have been sampled most
-intensively).\textsuperscript{16} The hotspots of host richness and
-viral diversity in southeast Asia---both of which are disproportionately
-high, considering the global landscape of bat species richness---are
-almost entirely driven by viral adaptive radiation through host
-switching within this clade\textsuperscript{3,14}. In contrast, the
-Neotropical hotspot of viral distinctiveness is driven by isolation by
-host vicariance. Out of the four main groups of betacoronaviruses, only
-the subgenus \emph{Merbecovirus} (MERS-like viruses) has been found in
-animals in the Americas---an introduction that is generally presumed to
-be ancient.\textsuperscript{3} While comparatively understudied, New
-World merbecoviruses have been found in the ghost-faced bats
-(Mormoopidae), New World leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae) (add cite: Olival 2020 PLoS Pathogens).
-The former two groups are endemic to the Neotropics, while the explosive
-adaptive radiations of the latter two (and particularly the
-phyllostomids) are responsible for the hotspot of bat diversity in the
-Amazon. Together, these clades of New World bats play host to a distinct
-regime of betacoronavirus coevolution.
+Overall, these patterns recapitulate the evolutionary history of both
+the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
+bats (Rhinolophidae) are both the reservoirs of the SARS-like viruses
+(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
+been of the greatest interest to researchers\textsuperscript{17} (and so
+have been sampled most intensively).\textsuperscript{16} The hotspots of
+host richness and viral diversity in southeast Asia---both of which are
+disproportionately high, considering the global landscape of bat species
+richness---are almost entirely driven by viral adaptive radiation
+through host switching within this clade\textsuperscript{3,14}. In
+contrast, the Neotropical hotspot of viral distinctiveness is driven by
+isolation by host vicariance. Out of the four main groups of
+betacoronaviruses, only the subgenus \emph{Merbecovirus} (MERS-like
+viruses) has been found in animals in the Americas---an introduction
+that is generally presumed to be ancient.\textsuperscript{3} While
+comparatively understudied, New World merbecoviruses have been found in
+the ghost-faced bats (Mormoopidae), New World leaf-nosed bats
+(Phyllostomidae), and free-tailed bats (Molossidae) (add cite: Olival
+2020 PLoS Pathogens). The former two groups are endemic to the
+Neotropics, while the explosive adaptive radiations of the latter two
+(and particularly the phyllostomids) are responsible for the hotspot of
+bat diversity in the Amazon. Together, these clades of New World bats
+play host to a distinct regime of betacoronavirus coevolution.
 
@@
 family level or higher are relatively rare and significant events in
-coronavirus evolutionary history;\textsuperscript{2,17} as a result, we
-suggest the mosaic of betacoronavirus phylogeography is assembled from a
-set of overlapping smaller coevolutionary systems, superimposed in space
-and filtered by the importance of different subgroups in local host
+coronavirus evolutionary history;\textsuperscript{2,17} as a result, the
+mosaic of betacoronavirus phylogeography is assembled from a set of
+overlapping smaller coevolutionary systems, superimposed in space and
+filtered by the importance of different subgroups in local host
 communities. For example, the most speciose and cosmopolitan family of

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@0bfcf11d2028cedf01a37ba47a7b6385fe6e60cd 🚀

@@
 co-divergence with hosts.\textsuperscript{6} This creates a disconnect
-between coevolutionary theory (including empirical evidence from
-virology) and most existing ecological frameworks for mapping spillover
-risk.
+between coevolutionary theory and most existing ecological frameworks
+for mapping spillover risk.
 
@@
 biogeography: for example, previous work has rarely flagged India as a
-region of concern, but for both bats and betacoronaviruses, the
-subcontinent falls into the same phylogeographic regions as the
-southeast Asian peninsula (and indeed, the region is home to known bat
-hosts of nobecoviruses, sarbecoviruses, and
-merbecoviruses).\textsuperscript{3}
+region of interest, but for both bats and betacoronaviruses, the
+subcontinent falls into the same regions as the southeast Asian
+peninsula (and indeed, the region is home to known bat hosts of
+nobecoviruses, sarbecoviruses, and merbecoviruses).\textsuperscript{3}
 
@@
 are lost, and betacoronavirus biogeography tracks the deep rifts in bat
-evolutionary history---but within these broad regions, the component
-coevolutionary systems can have very different dynamics.
+evolutionary history---but within broad regions, the component
+coevolutionary systems may have very different dynamics.
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@79b20b62fd5d3b81ec57b105439336f545a184b0 🚀

@@
 
+\hypertarget{bat-and-betacoronavirus-biogeography-are-broadly-consistent}{%
+\subsection{Bat and betacoronavirus biogeography are broadly
+consistent}\label{bat-and-betacoronavirus-biogeography-are-broadly-consistent}}
+
+Most previous work has assumed that the presence or richness of key
+groups of bat hosts are predictive of coronavirus
+diversity.\textsuperscript{2,3} Projecting bat and betacoronavirus
+phylogeny over space (fig.~\ref{fig:biogeo}), we find support for the
+idea that bat community assembly is directly responsible for a global
+mosaic of viral evolution. The distinct groupings (represented by
+different colors, symbolizing positions in a subspace formed by the
+first two phylogenetic principal components) are essentially equivalent
+between the two groups, and can be coarsely delineated as (1) south and
+southeast Asia, (2) east Asia (including northern China), west Asia, and
+the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
+Africa and south America. In some cases, this diverges from expectations
+about coronavirus biogeography: for example, previous work has rarely
+flagged India as a region of interest, but for both bats and
+betacoronaviruses, the subcontinent falls into the same regions as the
+southeast Asian peninsula (and indeed, the region is home to known bat
+hosts of nobecoviruses, sarbecoviruses, and
+merbecoviruses).\textsuperscript{3}
+
+\begin{figure}
+\hypertarget{fig:biogeo}{%
+\centering
+\includegraphics{figures/combined_biogeo.png}
+\caption{Phylogeographic regions of bats (top) and viruses (bottom)
+based on the joint analysis of their occurrence and phylogenetic
+relatedness. The different colors show tendencies to separate alongside
+the first two components of a PCoA. Note that the PCoA for the bats and
+viruses are independent, and so cannot be compared directly -- that
+being said, the regions can be compared across maps.}\label{fig:biogeo}
+}
+\end{figure}
+
+Overall, these results suggest that the boundaries of bat and
+betacoronavirus biogeographic regions are largely consistent. This may
+be surprising, given that cospeciation plays a minor role in coronavirus
+diversification,\textsuperscript{2} a property that would theoretically
+allow for substantial broad divergence in their biogeography. However,
+host jumps at the family level or higher are relatively rare and
+significant events in coronavirus evolutionary
+history;\textsuperscript{2,13} as a result, the mosaic of
+betacoronavirus phylogeography is assembled from a set of overlapping
+smaller coevolutionary systems, superimposed in space and filtered by
+the importance of different subgroups in local host communities. For
+example, the most speciose and cosmopolitan family of bats, the vesper
+bats (Vespertilionidae), are considered the primary reservoir of
+merbecoviruses;\textsuperscript{3,13} but in the Americas, where
+merbecoviruses are the only lineage present, they have only been found
+in other bat taxa. At the coarsest scale, these heterogeneities are
+lost, and betacoronavirus biogeography tracks the deep rifts in bat
+evolutionary history---but within broad regions, the component
+coevolutionary systems may have very different dynamics.
+
 \hypertarget{hotspots-of-host-richness-and-viral-diversification-are-distinct}{%
@@
 hotspots of viral adaptive radiation, and therefore areas of concern for
-human health.\textsuperscript{2,13} However, the hotspots of bat
+human health.\textsuperscript{2,14} However, the hotspots of bat
 betacoronavirus reservoirs show a distinct pattern, with primary
@@
 suggested that these undiscovered reservoirs should follow the same
-diversity gradient.\textsuperscript{14} In principle, these hotspots of
+diversity gradient.\textsuperscript{15} In principle, these hotspots of
 locally-diverse, virus-rich bat communities should drive more adaptive
@@
 places where the fewest bat betacoronavirus sequences have been
-generated,\textsuperscript{2,13,15} resulting in a sparser phylogenetic
+generated,\textsuperscript{2,14,16} resulting in a sparser phylogenetic
 tree, and artificially inflating distinctiveness; conversely,
-disproportionate research effort in eastern China\textsuperscript{16}
+disproportionate research effort in eastern China\textsuperscript{17}
 may have led to a more complete inventory of the local diversity of
@@
 (subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
-been of the greatest interest to researchers\textsuperscript{17} (and so
-have been sampled most intensively).\textsuperscript{16} The hotspots of
+been of the greatest interest to researchers\textsuperscript{13} (and so
+have been sampled most intensively).\textsuperscript{17} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 richness---are almost entirely driven by viral adaptive radiation
-through host switching within this clade\textsuperscript{3,14}. In
+through host switching within this clade\textsuperscript{3,15}. In
 contrast, the Neotropical hotspot of viral distinctiveness is driven by
@@
 play host to a distinct regime of betacoronavirus coevolution.
-
-\hypertarget{global-biogeographic-regions-are-consistent-for-bats-and-betacoronaviruses}{%
-\subsection{Global biogeographic regions are consistent for bats and
-betacoronaviruses}\label{global-biogeographic-regions-are-consistent-for-bats-and-betacoronaviruses}}
-
-Most previous work has assumed that coronavirus biogeography is driven
-by coevolutionary regimes that form at finer taxonomic scales, treating
-the presence or richness of key bat host groups as predictive of these
-viruses' distribution.\textsuperscript{2,3} Projecting bat and
-betacoronavirus phylogeny over space (fig.~\ref{fig:biogeo}), we find
-further support for the idea that bat community assembly is directly
-responsible for spatial heterogeneity in viral coevolutionary regimes.
-The distinct groupings (represented by different colors, symbolizing
-positions in a subspace formed by the first two phylogenetic principal
-components) are essentially equivalent between the two groups, and can
-be coarsely delineated as (1) south and southeast Asia, (2) east Asia
-(including northern China), west Asia, and the Mediterranean coast; (3)
-Eurasia above a northing of 40; and (4) Africa and south America. In
-some cases, this diverges from expectations about coronavirus
-biogeography: for example, previous work has rarely flagged India as a
-region of interest, but for both bats and betacoronaviruses, the
-subcontinent falls into the same regions as the southeast Asian
-peninsula (and indeed, the region is home to known bat hosts of
-nobecoviruses, sarbecoviruses, and merbecoviruses).\textsuperscript{3}
-
-\begin{figure}
-\hypertarget{fig:biogeo}{%
-\centering
-\includegraphics{figures/combined_biogeo.png}
-\caption{Phylogeographic regions of bats (top) and viruses (bottom)
-based on the joint analysis of their occurrence and phylogenetic
-relatedness. The different colors show tendencies to separate alongside
-the first two components of a PCoA. Note that the PCoA for the bats and
-viruses are independent, and so cannot be compared directly -- that
-being said, the regions can be compared across maps.}\label{fig:biogeo}
-}
-\end{figure}
-
-Overall, these results suggest that, although there are unique hotspots
-of viral adaptive radiation, biogeographic regions are mostly consistent
-between bats and their viruses. This may be surprising, given that
-cospeciation plays a minor role in coronavirus
-diversification,\textsuperscript{2} which could theoretically allow for
-broad divergence in their biogeography. However, host jumps at the
-family level or higher are relatively rare and significant events in
-coronavirus evolutionary history;\textsuperscript{2,17} as a result, the
-mosaic of betacoronavirus phylogeography is assembled from a set of
-overlapping smaller coevolutionary systems, superimposed in space and
-filtered by the importance of different subgroups in local host
-communities. For example, the most speciose and cosmopolitan family of
-bats, the vesper bats (Vespertilionidae), are considered the primary
-reservoir of merbecoviruses;\textsuperscript{3,17} but in the Americas,
-where merbecoviruses are the only lineage present, they have only been
-found in other bat taxa. At the coarsest scale, these heterogeneities
-are lost, and betacoronavirus biogeography tracks the deep rifts in bat
-evolutionary history---but within broad regions, the component
-coevolutionary systems may have very different dynamics.
 
@@
 risk/diversitifcation potential (Madagascar, South-America) overlay with
-sampling gaps for \emph{Betacoronavirus},\textsuperscript{16} stressing
+sampling gaps for \emph{Betacoronavirus},\textsuperscript{17} stressing
 the need for spatially targeted monitoring and discovery.
@@
 that have used host richness as proxy have predicted a high diversity of
-unsampled bat viruses,\textsuperscript{13} bat
+unsampled bat viruses,\textsuperscript{14} bat
 coronaviruses,\textsuperscript{2} and even specifically
-betacoronaviruses\textsuperscript{14} in both the Amazon and southeast
+betacoronaviruses\textsuperscript{15} in both the Amazon and southeast
 Asia. While we find that both regions are characterized by highly
@@
 high local diversity of viruses through host switching and viral
-recombination (see \emph{e.g.},\textsuperscript{17} as well as the
+recombination (see \emph{e.g.},\textsuperscript{13} as well as the
 discovery of recombinant viruses that share genetic material from both
@@
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{14,16} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{15,17} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 \texttt{https://www.viralemergence.org/betacov} on
-Apr.~2022,\textsuperscript{14} and filtered it to ``known'' hosts
+Apr.~2022,\textsuperscript{15} and filtered it to ``known'' hosts
 (established before the emergence of SARS-CoV-2) and ``novel'' hosts
@@
 
+\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
+\CSLLeftMargin{13. }
+\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
+transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
+Server for Biology} 2020.05.31.116061 (2020)
+doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-\CSLLeftMargin{13. }
+\CSLLeftMargin{14. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
@@
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-\CSLLeftMargin{14. }
+\CSLLeftMargin{15. }
 \CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
@@
 \leavevmode\hypertarget{ref-Allen2017GloHot}{}%
-\CSLLeftMargin{15. }
+\CSLLeftMargin{16. }
 \CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
@@
 \leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{16. }
+\CSLLeftMargin{17. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}
-
-\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-\CSLLeftMargin{17. }
-\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
-transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
-Server for Biology} 2020.05.31.116061 (2020)
-doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f5918d40bb552fe4245e75ff3f52c2a7506f06ac 🚀

@@
 
-\hypertarget{hotspots-of-host-richness-and-viral-diversification-are-distinct}{%
-\subsection{Hotspots of host richness and viral diversification are
-distinct}\label{hotspots-of-host-richness-and-viral-diversification-are-distinct}}
+\hypertarget{coevolutionary-regimes-create-distinct-hotspots-of-betacoronavirus-biodiversity}{%
+\subsection{Coevolutionary regimes create distinct hotspots of
+betacoronavirus
+biodiversity}\label{coevolutionary-regimes-create-distinct-hotspots-of-betacoronavirus-biodiversity}}
 
@@
 hotspots of viral adaptive radiation, and therefore areas of concern for
-human health.\textsuperscript{2,14} However, the hotspots of bat
-betacoronavirus reservoirs show a distinct pattern, with primary
-hotspots (both in terms of size and higher values) of host richness
-situated primarily southeast Asia, parts of southern Europe, and to a
-lesser extent parts of Africa in the -25-0 range of latitudes
-(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
-undiscovered betacoronaviruses, machine learning predictions have
-suggested that these undiscovered reservoirs should follow the same
-diversity gradient.\textsuperscript{15} In principle, these hotspots of
+human health.\textsuperscript{2,14} However, the hotspots of known bat
+betacoronavirus hosts show a distinct pattern, with primary hotspots
+(both in terms of size and higher values) of host richness situated in
+southeast Asia, parts of southern Europe, and to a lesser extent parts
+of Africa in the -25-0 range of latitudes (fig.~\ref{fig:richness};
+top). Although hundreds of species likely host undiscovered
+betacoronaviruses, machine learning predictions have suggested that
+these undiscovered reservoirs should follow the same diversity
+gradient.\textsuperscript{15} In principle, these hotspots of
 locally-diverse, virus-rich bat communities should drive more adaptive

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@348041f6d355b99d641733ea19c6c0e367cf761b 🚀

@@
 
-\hypertarget{coevolutionary-regimes-create-distinct-hotspots-of-betacoronavirus-biodiversity}{%
-\subsection{Coevolutionary regimes create distinct hotspots of
-betacoronavirus
-biodiversity}\label{coevolutionary-regimes-create-distinct-hotspots-of-betacoronavirus-biodiversity}}
+\hypertarget{hotspots-of-bat-and-betacoronavirus-biodiversity-are-distinct}{%
+\subsection{Hotspots of bat and betacoronavirus biodiversity are
+distinct}\label{hotspots-of-bat-and-betacoronavirus-biodiversity-are-distinct}}
 

### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@bcbe5051cbe2512efc17e7645574e9bd042b9fa8 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@96da5ab50d0162ebdf6d6a957760c7b868229a6a 🚀

@@
 \thispagestyle{empty}
-{\Large\bfseries\sffamily The coevolutionary mosaic of betacoronavirus
-emergence risk}
+{\Large\bfseries\sffamily The coevolutionary mosaic of bat
+betacoronavirus emergence risk}
 \vskip 5em
@@
 
-\hypertarget{human-occupancy-drives-different-levels-of-effective-risk-globally}{%
-\subsection{Human occupancy drives different levels of effective risk
-globally}\label{human-occupancy-drives-different-levels-of-effective-risk-globally}}
-
-Based on the previous result, we extracted the risk component from the
-composite map (see Methods), to provide a single measure of risk varying
-between 0 and 1. This measure is presented in fig.~\ref{fig:risk}. As
-this map represents the potential risk, it must be weighed by the
-potential for contacts with humans. As a proxy for this measure, we used
-the proportion of build/urban land from the EarthEnv dataset: this is a
-reasonable proxy for the density of humans per unit area, which
-increases the probability of pathogen spread more
-widely.\textsuperscript{20} Since human activity is required to amplify
-the frequency of virus encounters and thus create areas of viral
-amplification, mapping the potential risk against measures of land use
-is required to generate a more actionable assessment of risk. This map
-is presented in fig.~\ref{fig:compound}. Most of South America and
-Europe are at comparatively lower risk, as although densely populated,
-settlements tend to be in areas with lower potential risk. Regions like
-Malaysia and the North coast of Australia have a high risk component,
-but should represent a relatively lower effective risk due to low human
-density. However, this mapping reveals that South-East Asia, the Indian
-subcontinent, and parts of sub-Saharan Africa, are at high risk due to
-the overlap between built areas and bat communities representing more
-opportunities for cross-species transmission of betacoronaviruses. In
-looking for the origins of SARS in China,\textsuperscript{21} present
-serological evidence that strongest human-animal contact results in
-higher risk of virus exposure, regardless of the animal species, but
-that different types of contact had different impacts. Ideally,
-finer-grained information about human activity (rather than human
-presence through anthropisation) could allow to partition this risk
-further, alebit at the cost of more hypotheses required to estimate the
-amount of risk represented by each activity. Our map of purported high
-risk/diversitifcation potential (Madagascar, South-America) overlay with
-sampling gaps for \emph{Betacoronavirus},\textsuperscript{17} stressing
-the need for spatially targeted monitoring and discovery.
+\hypertarget{human-landscapes-filter-the-geography-of-emergence-risk}{%
+\subsection{Human landscapes filter the geography of emergence
+risk}\label{human-landscapes-filter-the-geography-of-emergence-risk}}
+
+The relationship between the underlying pathogen pool and emergence risk
+is mediated by human-wildlife contact and the risk that spillovers
+become epidemics. As a proxy for both, we finally overlaid the risk
+component from the composite map (see above) with the proportion of
+built land, as a proxy for a mix of habitat disturbance, potential for
+bat synanthropy, and human population density and
+connectivity\textsuperscript{1,20} (add: Muylaert 2022 Proc B; Hassell
+2017 TREE) (fig.~\ref{fig:compound}).
+
+Most of South America and Europe are at comparatively lower risk, as
+although densely populated, settlements tend to be in areas with lower
+potential risk. Regions like Malaysia and the North coast of Australia
+have a high risk component, but should represent a relatively lower
+effective risk due to low human density. However, this mapping reveals
+that South-East Asia, the Indian subcontinent, and parts of sub-Saharan
+Africa, are at high risk due to the overlap between built areas and bat
+communities representing more opportunities for cross-species
+transmission of betacoronaviruses. In looking for the origins of SARS in
+China,\textsuperscript{21} present serological evidence that strongest
+human-animal contact results in higher risk of virus exposure,
+regardless of the animal species, but that different types of contact
+had different impacts. Ideally, finer-grained information about human
+activity (rather than human presence through anthropisation) could allow
+to partition this risk further, alebit at the cost of more hypotheses
+required to estimate the amount of risk represented by each activity.
+Our map of purported high risk/diversitifcation potential (Madagascar,
+South-America) overlay with sampling gaps for
+\emph{Betacoronavirus},\textsuperscript{17} stressing the need for
+spatially targeted monitoring and discovery.
 

### 2022-06-27 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@bcbe5051cbe2512efc17e7645574e9bd042b9fa8 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@96da5ab50d0162ebdf6d6a957760c7b868229a6a 🚀

@@
 
-\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of
+\noindent {\LARGE \textbf{\textsf{The coevolutionary mosaic of bat
 betacoronavirus emergence risk}}}

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@921fd2c197768a12420e4cda39e410d51d50afcd 🚀

@@
 The relationship between the underlying pathogen pool and emergence risk
-is mediated by human-wildlife contact and the risk that spillovers
-become epidemics. As a proxy for both, we finally overlaid the risk
-component from the composite map (see above) with the proportion of
-built land, as a proxy for a mix of habitat disturbance, potential for
-bat synanthropy, and human population density and
+is mediated by both human-wildlife interfaces (the probability of
+spillover) and opportunities for onward transmission (the probability
+that spillovers become epidemics). As a proxy for both, we finally
+overlaid the risk component from the composite map (see above) with the
+proportion of built land, as a proxy for a mix of habitat disturbance,
+potential for bat synanthropy or contact with bridge hosts like
+livestock (cite Rulli et al 2021 Nature Food; cite Cui et al 2019 Nature
+Reviews Microbiology), and human population density and
 connectivity\textsuperscript{1,20} (add: Muylaert 2022 Proc B; Hassell
-2017 TREE) (fig.~\ref{fig:compound}).
-
-Most of South America and Europe are at comparatively lower risk, as
-although densely populated, settlements tend to be in areas with lower
-potential risk. Regions like Malaysia and the North coast of Australia
-have a high risk component, but should represent a relatively lower
-effective risk due to low human density. However, this mapping reveals
-that South-East Asia, the Indian subcontinent, and parts of sub-Saharan
-Africa, are at high risk due to the overlap between built areas and bat
-communities representing more opportunities for cross-species
-transmission of betacoronaviruses. In looking for the origins of SARS in
-China,\textsuperscript{21} present serological evidence that strongest
-human-animal contact results in higher risk of virus exposure,
-regardless of the animal species, but that different types of contact
-had different impacts. Ideally, finer-grained information about human
-activity (rather than human presence through anthropisation) could allow
-to partition this risk further, alebit at the cost of more hypotheses
-required to estimate the amount of risk represented by each activity.
-Our map of purported high risk/diversitifcation potential (Madagascar,
-South-America) overlay with sampling gaps for
-\emph{Betacoronavirus},\textsuperscript{17} stressing the need for
-spatially targeted monitoring and discovery.
+2017 TREE) (fig.~\ref{fig:compound}). Accounting for these factors, most
+of South America and Europe are at comparatively lower risk,
+as--although densely populated--settlements tend to be in areas with
+lower potential risk. Conversely, regions like Malaysia and the northern
+coast of Australia have a high evolutionary risk component, but should
+represent a relatively lower effective risk due to low human density.
+However, southeast Asia, the Indian subcontinent, and scattered hotspots
+in sub-Saharan Africa are at high risk due to the overlap between human
+populations and natural opportunities for cross-species transmission of
+betacoronaviruses.
 
@@
 
-PUT THIS SOMEWHERE: Available information describing the spillover of
-zoonotic betacoronaviruses of bat origin where data was available before
-and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
-in Wuhan, China; SARS-CoV in Guangdong, China based on the presence of
-closest known viruses circulating in nature, and a nearby location where
-serological (antibody) evidence has indicated human exposure to
-SARS-like viruses;\textsuperscript{22} MERS-CoV in Saudi Arabia based on
-index cases available from a recently-published compendium of
-cases.\textsuperscript{23} For the latest event, most if not all index
-cases are presumed to be camel-to-human transmission, and the precise
-origin point (if it exists) of MERS-CoV in bats is uncertain. Recent
-recombinant canine coronavirus spillover events in
-Haiti\textsuperscript{24} and Europe\textsuperscript{25} are not
-relevant here, as bats' involvement in these cycles of transmission have
-been supposed to be non-existent. These index cases fall within
-different phylogeographic bioregions (fig.~\ref{fig:biogeo}), which
-further highlight the issue that different host-virus sub-systems may
-lead to widespread emergence.
+Reassuringly, these predictions are validated by the sites where the
+three highly pathogenic coronavirus are believed to have recently
+emerged. While available information puts the spillover of SARS-CoV-2 in
+a live animal market in Wuhan, China, the ultimate origin of the virus
+is almost certainly in a Lineage 2 sarbecovirus from a southeast Asian
+horseshoe bat. Similarly, the SARS-CoV outbreak began in Guangdong
+province in 2002, reaching humans through small carnivore bridge hosts,
+but was eventually traced back to a set of likely progenitor viruses
+found in cave-dwelling horseshoe bats in Yunnan province (cite Hu et
+al.~2017 PLoS Pathogens); nearby, antibody evidence has indicated human
+exposure to SARS-like viruses.\textsuperscript{21} MERS-CoV was
+originally detected in Saudi Arabia, accompanied by a nearly identical
+virus sequenced from an Egyptian tomb bat (\emph{Taphozous perforatus})
+(cite Memish et al.~2013 EID), but is widespread in camels in East
+Africa and the Middle East, and may have reached its bridge host decades
+earlier than originally supposed (cite Muller et al 2014 EID); as a
+result, the geography of the original bat-to-camel transmission is still
+widely regarded as uncertain. All of these are broadly consistent with
+the risk factors we identify. Notably, India and west Africa are
+additional hotspots that have yet to experience the emergence of a bat
+coronavirus into human populations, but may still be at
+risk---particularly given known gaps in bat
+surveillance,\textsuperscript{17} and a dense population in both regions
+with global connectivity. In any of these regions, surveillance on viral
+reservoirs can be paired with targeted monitoring of high-risk human
+populations (i.e., those with regular wildlife
+contact)\textsuperscript{22} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{26,27} Chiropterans emerged around 64 million
+health.\textsuperscript{23,24} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{28,29}
+estimated richness of more than 1400 species.\textsuperscript{25,26}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{30} For example,
+ecosystem-derived benefits to human.\textsuperscript{27} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{31} Over two-thirds of bats are know to be
+networks.\textsuperscript{28} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{32,33} and
+contributing for agricultural pest control,\textsuperscript{29,30} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{34,35} Because bats are globally distributed and
+health.\textsuperscript{31,32} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{36} Both of these regions are priority areas
+lineage).\textsuperscript{33} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{37}
+to the region.\textsuperscript{34}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{38}
+anthropogenic factors others than shifting climates.\textsuperscript{35}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{39} As our results establish that the
+habitats.\textsuperscript{36} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{40} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{41} This
+novel diseases,\textsuperscript{37} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{38} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{42} In particular, the challenge ahead lies
+encounters.\textsuperscript{39} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{43} The range maps were subsequently rasterized
+data.\textsuperscript{40} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{44} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{41} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{45} based on a recent synthetic tree with
+Diversity\textsuperscript{42} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{46} Faith's PD measures the sum of unique
+species.\textsuperscript{43} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{47} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{44} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{48} transformed
+location, and following best practices\textsuperscript{45} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{49} shows that LCBD
+community. Recent results by\textsuperscript{46} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{50} This network stores pairwise values of
+mammals.\textsuperscript{47} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{51} In this approach, every risk component gets
+model.\textsuperscript{48} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{52} in that this allows multiple
+and low viral sharing,\textsuperscript{49} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{53} we measured the risk level by
+extraction,\textsuperscript{50} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{54} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{51} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{55} v1.6.12
-with ModelFinder\textsuperscript{56} ultrafast bootstrap
-approximation\textsuperscript{57} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{52} v1.6.12
+with ModelFinder\textsuperscript{53} ultrafast bootstrap
+approximation\textsuperscript{54} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{58} for each of the viruses in the tree,
+distinctiveness\textsuperscript{55} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{59,60} who
+this idea, we loosely adapted a method from,\textsuperscript{56,57} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{61} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{58} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
+\CSLLeftMargin{21. }
+\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
+SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
+Sinica} \textbf{33}, 104--107 (2018).}
+
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 
-\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{22. }
-\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
-SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
-Sinica} \textbf{33}, 104--107 (2018).}
-
-\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
+\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
 \CSLLeftMargin{23. }
-\CSLRightInline{Ramshaw, R. E. \emph{et al.} A database of geopositioned
-Middle East Respiratory Syndrome Coronavirus occurrences.
-\emph{Scientific Data} \textbf{6}, 318 (2019).}
-
-\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
-\CSLLeftMargin{24. }
-\CSLRightInline{Lednicky, J. A. \emph{et al.} Isolation of a Novel
-Recombinant Canine Coronavirus from a Visitor to Haiti: Further Evidence
-of Transmission of Coronaviruses of Zoonotic Origin to Humans.
-\emph{Clinical Infectious Diseases: An Official Publication of the
-Infectious Diseases Society of America} ciab924 (2021)
-doi:\href{https://doi.org/10.1093/cid/ciab924}{10.1093/cid/ciab924}.}
-
-\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
-\CSLLeftMargin{25. }
-\CSLRightInline{Vlasova, A. N. \emph{et al.} Animal alphacoronaviruses
-found in human patients with acute respiratory illness in different
-countries. \emph{Emerging Microbes \& Infections} \textbf{11}, 699--702
-(2022).}
-
-\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{26. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{29. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{40. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a584545019c68fc3513be98c58545cb51310a121 🚀

@@
 a live animal market in Wuhan, China, the ultimate origin of the virus
-is almost certainly in a Lineage 2 sarbecovirus from a southeast Asian
-horseshoe bat. Similarly, the SARS-CoV outbreak began in Guangdong
-province in 2002, reaching humans through small carnivore bridge hosts,
-but was eventually traced back to a set of likely progenitor viruses
-found in cave-dwelling horseshoe bats in Yunnan province (cite Hu et
-al.~2017 PLoS Pathogens); nearby, antibody evidence has indicated human
-exposure to SARS-like viruses.\textsuperscript{21} MERS-CoV was
-originally detected in Saudi Arabia, accompanied by a nearly identical
-virus sequenced from an Egyptian tomb bat (\emph{Taphozous perforatus})
-(cite Memish et al.~2013 EID), but is widespread in camels in East
-Africa and the Middle East, and may have reached its bridge host decades
-earlier than originally supposed (cite Muller et al 2014 EID); as a
-result, the geography of the original bat-to-camel transmission is still
-widely regarded as uncertain. All of these are broadly consistent with
-the risk factors we identify. Notably, India and west Africa are
-additional hotspots that have yet to experience the emergence of a bat
-coronavirus into human populations, but may still be at
-risk---particularly given known gaps in bat
+is almost certainly in a divergent lineage of sarbecoviruses from the
+Indochinese peninsula that was poorly characterized prior to the
+pandemic {[}cite Worobey 2022 Zenodo; Temmam 2022 Nature; Boni 2020
+Nature Microbiology{]}. Similarly, the SARS-CoV outbreak began in
+Guangdong province in 2002, reaching humans through small carnivore
+bridge hosts, but was eventually traced back to a set of likely
+progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
+province (cite Hu et al.~2017 PLoS Pathogens); nearby, antibody evidence
+has indicated human exposure to SARS-like viruses.\textsuperscript{21}
+MERS-CoV was originally detected in Saudi Arabia, accompanied by a
+nearly identical virus sequenced from an Egyptian tomb bat
+(\emph{Taphozous perforatus}) (cite Memish et al.~2013 EID), but is
+widespread in camels in East Africa and the Middle East, and may have
+reached its bridge host decades earlier than originally supposed (cite
+Muller et al 2014 EID); as a result, the geography of the original
+bat-to-camel transmission is still widely regarded as uncertain. All of
+these are broadly consistent with the risk factors we identify. Notably,
+India and west Africa are additional hotspots that have yet to
+experience the emergence of a bat coronavirus into human populations,
+but may still be at risk---particularly given known gaps in bat
 surveillance,\textsuperscript{17} and a dense population in both regions

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d66d3a132cc144d2439e07bf1827a5dec0874916 🚀

@@
 spillover) and opportunities for onward transmission (the probability
-that spillovers become epidemics). As a proxy for both, we finally
-overlaid the risk component from the composite map (see above) with the
-proportion of built land, as a proxy for a mix of habitat disturbance,
-potential for bat synanthropy or contact with bridge hosts like
-livestock (cite Rulli et al 2021 Nature Food; cite Cui et al 2019 Nature
-Reviews Microbiology), and human population density and
-connectivity\textsuperscript{1,20} (add: Muylaert 2022 Proc B; Hassell
-2017 TREE) (fig.~\ref{fig:compound}). Accounting for these factors, most
-of South America and Europe are at comparatively lower risk,
-as--although densely populated--settlements tend to be in areas with
-lower potential risk. Conversely, regions like Malaysia and the northern
-coast of Australia have a high evolutionary risk component, but should
-represent a relatively lower effective risk due to low human density.
-However, southeast Asia, the Indian subcontinent, and scattered hotspots
-in sub-Saharan Africa are at high risk due to the overlap between human
-populations and natural opportunities for cross-species transmission of
-betacoronaviruses.
+that spillovers become epidemics)\textsuperscript{1}. As a proxy for
+both, we finally overlaid the risk component from the composite map (see
+above) with the proportion of built land, as a proxy for a mix of
+habitat disturbance, potential for bat synanthropy or contact with
+bridge hosts like livestock (cite Rulli et al 2021 Nature Food; cite Cui
+et al 2019 Nature Reviews Microbiology), and human population density
+and connectivity\textsuperscript{1,20} (add: Muylaert 2022 Proc B;
+Hassell 2017 TREE) (fig.~\ref{fig:compound}). Accounting for these
+factors, most of South America and Europe are at comparatively lower
+risk, as--although densely populated--settlements tend to be in areas
+with lower potential risk. Conversely, regions like Malaysia and the
+northern coast of Australia have a high evolutionary risk component, but
+should represent a relatively lower effective risk due to low human
+density. However, southeast Asia, the Indian subcontinent, and scattered
+hotspots in sub-Saharan Africa are at high risk due to the overlap
+between human populations and natural opportunities for cross-species
+transmission of betacoronaviruses.
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e1fd0c81afe4ddf4a3978ee73a21d0e6a0272d28 🚀

@@
 the ghost-faced bats (Mormoopidae), New World leaf-nosed bats
-(Phyllostomidae), and free-tailed bats (Molossidae) (add cite: Olival
-2020 PLoS Pathogens). The former two groups are endemic to the
-Neotropics, while the explosive adaptive radiations of the latter two
-(and particularly the phyllostomids) are responsible for the hotspot of
-bat diversity in the Amazon. Together, these clades of New World bats
-play host to a distinct regime of betacoronavirus coevolution.
+(Phyllostomidae), and free-tailed bats
+(Molossidae).\textsuperscript{\textbf{Olival2020PosZoo?}} The former two
+groups are endemic to the Neotropics, while the explosive adaptive
+radiations of the latter two (and particularly the phyllostomids) are
+responsible for the hotspot of bat diversity in the Amazon. Together,
+these clades of New World bats play host to a distinct regime of
+betacoronavirus coevolution.
 
@@
 habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock (cite Rulli et al 2021 Nature Food; cite Cui
-et al 2019 Nature Reviews Microbiology), and human population density
-and connectivity\textsuperscript{1,20} (add: Muylaert 2022 Proc B;
-Hassell 2017 TREE) (fig.~\ref{fig:compound}). Accounting for these
-factors, most of South America and Europe are at comparatively lower
-risk, as--although densely populated--settlements tend to be in areas
-with lower potential risk. Conversely, regions like Malaysia and the
-northern coast of Australia have a high evolutionary risk component, but
-should represent a relatively lower effective risk due to low human
-density. However, southeast Asia, the Indian subcontinent, and scattered
-hotspots in sub-Saharan Africa are at high risk due to the overlap
-between human populations and natural opportunities for cross-species
-transmission of betacoronaviruses.
+bridge hosts like livestock,\textsuperscript{20,21} and human population
+density and connectivity\textsuperscript{1,22,23}
+(fig.~\ref{fig:compound}). Accounting for these factors, most of South
+America and Europe are at comparatively lower risk, as--although densely
+populated--settlements tend to be in areas with lower potential risk.
+Conversely, regions like Malaysia and the northern coast of Australia
+have a high evolutionary risk component, but should represent a
+relatively lower effective risk due to low human density. However,
+southeast Asia, the Indian subcontinent, and scattered hotspots in
+sub-Saharan Africa are at high risk due to the overlap between human
+populations and natural opportunities for cross-species transmission of
+betacoronaviruses.
 
@@
 Indochinese peninsula that was poorly characterized prior to the
-pandemic {[}cite Worobey 2022 Zenodo; Temmam 2022 Nature; Boni 2020
-Nature Microbiology{]}. Similarly, the SARS-CoV outbreak began in
-Guangdong province in 2002, reaching humans through small carnivore
+pandemic.\textsuperscript{24--26} Similarly, the SARS-CoV outbreak began
+in Guangdong province in 2002, reaching humans through small carnivore
 bridge hosts, but was eventually traced back to a set of likely
 progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province (cite Hu et al.~2017 PLoS Pathogens); nearby, antibody evidence
-has indicated human exposure to SARS-like viruses.\textsuperscript{21}
-MERS-CoV was originally detected in Saudi Arabia, accompanied by a
-nearly identical virus sequenced from an Egyptian tomb bat
-(\emph{Taphozous perforatus}) (cite Memish et al.~2013 EID), but is
-widespread in camels in East Africa and the Middle East, and may have
-reached its bridge host decades earlier than originally supposed (cite
-Muller et al 2014 EID); as a result, the geography of the original
-bat-to-camel transmission is still widely regarded as uncertain. All of
-these are broadly consistent with the risk factors we identify. Notably,
-India and west Africa are additional hotspots that have yet to
-experience the emergence of a bat coronavirus into human populations,
-but may still be at risk---particularly given known gaps in bat
-surveillance,\textsuperscript{17} and a dense population in both regions
-with global connectivity. In any of these regions, surveillance on viral
-reservoirs can be paired with targeted monitoring of high-risk human
-populations (i.e., those with regular wildlife
-contact)\textsuperscript{22} for maximum impact.
+province;\textsuperscript{27} nearby, antibody evidence has indicated
+human exposure to SARS-like viruses.\textsuperscript{28} MERS-CoV was
+originally detected in Saudi Arabia, accompanied by a nearly identical
+virus sequenced from an Egyptian tomb bat (\emph{Taphozous
+perforatus})\textsuperscript{29}, but is widespread in camels in East
+Africa and the Middle East, and may have reached its bridge host decades
+earlier than originally supposed;\textsuperscript{30} as a result, the
+geography of the original bat-to-camel transmission is still widely
+regarded as uncertain. All of these are broadly consistent with the risk
+factors we identify. Notably, India and west Africa are additional
+hotspots that have yet to experience the emergence of a bat coronavirus
+into human populations, but may still be at risk---particularly given
+known gaps in bat surveillance,\textsuperscript{17} and a dense
+population in both regions with global connectivity. In any of these
+regions, surveillance on viral reservoirs can be paired with targeted
+monitoring of high-risk human populations (i.e., those with regular
+wildlife contact)\textsuperscript{31} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{23,24} Chiropterans emerged around 64 million
+health.\textsuperscript{32,33} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{25,26}
+estimated richness of more than 1400 species.\textsuperscript{34,35}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{27} For example,
+ecosystem-derived benefits to human.\textsuperscript{36} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{28} Over two-thirds of bats are know to be
+networks.\textsuperscript{37} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{29,30} and
+contributing for agricultural pest control,\textsuperscript{38,39} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{31,32} Because bats are globally distributed and
+health.\textsuperscript{40,41} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{33} Both of these regions are priority areas
+lineage).\textsuperscript{42} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{34}
+to the region.\textsuperscript{43}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{35}
+anthropogenic factors others than shifting climates.\textsuperscript{44}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{36} As our results establish that the
+habitats.\textsuperscript{45} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{37} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{38} This
+novel diseases,\textsuperscript{46} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{47} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{39} In particular, the challenge ahead lies
+encounters.\textsuperscript{48} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{40} The range maps were subsequently rasterized
+data.\textsuperscript{49} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{41} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{42} based on a recent synthetic tree with
+Diversity\textsuperscript{51} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{43} Faith's PD measures the sum of unique
+species.\textsuperscript{52} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{44} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{53} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{45} transformed
+location, and following best practices\textsuperscript{54} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{46} shows that LCBD
+community. Recent results by\textsuperscript{55} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{47} This network stores pairwise values of
+mammals.\textsuperscript{56} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{48} In this approach, every risk component gets
+model.\textsuperscript{57} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{49} in that this allows multiple
+and low viral sharing,\textsuperscript{58} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{50} we measured the risk level by
+extraction,\textsuperscript{59} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{51} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{60} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{52} v1.6.12
-with ModelFinder\textsuperscript{53} ultrafast bootstrap
-approximation\textsuperscript{54} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{61} v1.6.12
+with ModelFinder\textsuperscript{62} ultrafast bootstrap
+approximation\textsuperscript{63} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{55} for each of the viruses in the tree,
+distinctiveness\textsuperscript{64} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{56,57} who
+this idea, we loosely adapted a method from,\textsuperscript{65,66} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{58} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{67} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Hazarie2021IntPop}{}%
+\leavevmode\hypertarget{ref-Rulli2021LanUse}{}%
 \CSLLeftMargin{20. }
-\CSLRightInline{Hazarie, S., Soriano-Paños, D., Arenas, A.,
-Gómez-Gardeñes, J. \& Ghoshal, G. Interplay between population density
-and mobility in determining the spread of epidemics in cities.
-\emph{Communications Physics} \textbf{4}, 191 (2021).}
+\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
+Land-use change and the livestock revolution increase the risk of
+zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
+Food} \textbf{2}, 409--416 (2021).}
+
+\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
+\CSLLeftMargin{21. }
+\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
+pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
+\textbf{17}, 181--192 (2019).}
+
+\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
+\CSLLeftMargin{22. }
+\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
+distribution of bat hosts of sarbecoviruses: Implications for
+conservation and public health. \emph{Proceedings of the Royal Society
+B: Biological Sciences} \textbf{289}, 20220397 (2022).}
+
+\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
+\CSLLeftMargin{23. }
+\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
+Urbanization and disease emergence: Dynamics at the
+wildlife--livestock--human interface. \emph{Trends in ecology \&
+evolution} \textbf{32}, 55--67 (2017).}
+
+\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
+\CSLLeftMargin{24. }
+\CSLRightInline{Worobey, M. \emph{et al.} The huanan market was the
+epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022).}
+
+\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
+\CSLLeftMargin{25. }
+\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
+SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
+330--336 (2022).}
+
+\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
+\CSLLeftMargin{26. }
+\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
+SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
+\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}
+
+\leavevmode\hypertarget{ref-Hu2017DisRic}{}%
+\CSLLeftMargin{27. }
+\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
+bat SARS-related coronaviruses provides new insights into the origin of
+SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}
 
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 
+\leavevmode\hypertarget{ref-Memish2013MidEas}{}%
+\CSLLeftMargin{29. }
+\CSLRightInline{Memish, Z. A. \emph{et al.} Middle east respiratory
+syndrome coronavirus in bats, saudi arabia. \emph{Emerging infectious
+diseases} \textbf{19}, 1819 (2013).}
+
+\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
+\CSLLeftMargin{30. }
+\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
+neutralizing antibodies in camels, eastern africa, 1983--1997.
+\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}
+
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{38. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{49. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@77e13b501ab294fbcfda1097bf939bf1c5b301e0 🚀

@@
 the ghost-faced bats (Mormoopidae), New World leaf-nosed bats
-(Phyllostomidae), and free-tailed bats
-(Molossidae).\textsuperscript{\textbf{Olival2020PosZoo?}} The former two
-groups are endemic to the Neotropics, while the explosive adaptive
-radiations of the latter two (and particularly the phyllostomids) are
-responsible for the hotspot of bat diversity in the Amazon. Together,
-these clades of New World bats play host to a distinct regime of
-betacoronavirus coevolution.
+(Phyllostomidae), and free-tailed bats (Molossidae).\textsuperscript{18}
+The former two groups are endemic to the Neotropics, while the explosive
+adaptive radiations of the latter two (and particularly the
+phyllostomids) are responsible for the hotspot of bat diversity in the
+Amazon. Together, these clades of New World bats play host to a distinct
+regime of betacoronavirus coevolution.
 
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{18} Recent surveillance\textsuperscript{19}
+continents.\textsuperscript{19} Recent surveillance\textsuperscript{20}
 has identified a novel \emph{Betacoronavirus} (in the subgenus
@@
 habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock,\textsuperscript{20,21} and human population
-density and connectivity\textsuperscript{1,22,23}
+bridge hosts like livestock,\textsuperscript{21,22} and human population
+density and connectivity\textsuperscript{1,23,24}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 Indochinese peninsula that was poorly characterized prior to the
-pandemic.\textsuperscript{24--26} Similarly, the SARS-CoV outbreak began
+pandemic.\textsuperscript{25--27} Similarly, the SARS-CoV outbreak began
 in Guangdong province in 2002, reaching humans through small carnivore
@@
 progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province;\textsuperscript{27} nearby, antibody evidence has indicated
-human exposure to SARS-like viruses.\textsuperscript{28} MERS-CoV was
+province;\textsuperscript{28} nearby, antibody evidence has indicated
+human exposure to SARS-like viruses.\textsuperscript{29} MERS-CoV was
 originally detected in Saudi Arabia, accompanied by a nearly identical
 virus sequenced from an Egyptian tomb bat (\emph{Taphozous
-perforatus})\textsuperscript{29}, but is widespread in camels in East
+perforatus})\textsuperscript{30}, but is widespread in camels in East
 Africa and the Middle East, and may have reached its bridge host decades
-earlier than originally supposed;\textsuperscript{30} as a result, the
+earlier than originally supposed;\textsuperscript{31} as a result, the
 geography of the original bat-to-camel transmission is still widely
@@
 monitoring of high-risk human populations (i.e., those with regular
-wildlife contact)\textsuperscript{31} for maximum impact.
+wildlife contact)\textsuperscript{32} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{32,33} Chiropterans emerged around 64 million
+health.\textsuperscript{33,34} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{34,35}
+estimated richness of more than 1400 species.\textsuperscript{35,36}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{36} For example,
+ecosystem-derived benefits to human.\textsuperscript{37} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{37} Over two-thirds of bats are know to be
+networks.\textsuperscript{38} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{38,39} and
+contributing for agricultural pest control,\textsuperscript{39,40} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{40,41} Because bats are globally distributed and
+health.\textsuperscript{41,42} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{42} Both of these regions are priority areas
+lineage).\textsuperscript{43} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{43}
+to the region.\textsuperscript{44}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{44}
+anthropogenic factors others than shifting climates.\textsuperscript{45}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{45} As our results establish that the
+habitats.\textsuperscript{46} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{46} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{47} This
+novel diseases,\textsuperscript{47} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{48} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{48} In particular, the challenge ahead lies
+encounters.\textsuperscript{49} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{49} The range maps were subsequently rasterized
+data.\textsuperscript{50} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{51} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{51} based on a recent synthetic tree with
+Diversity\textsuperscript{52} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{52} Faith's PD measures the sum of unique
+species.\textsuperscript{53} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{53} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{54} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{54} transformed
+location, and following best practices\textsuperscript{55} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{55} shows that LCBD
+community. Recent results by\textsuperscript{56} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{56} This network stores pairwise values of
+mammals.\textsuperscript{57} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{57} In this approach, every risk component gets
+model.\textsuperscript{58} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{58} in that this allows multiple
+and low viral sharing,\textsuperscript{59} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{59} we measured the risk level by
+extraction,\textsuperscript{60} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{60} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{61} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{61} v1.6.12
-with ModelFinder\textsuperscript{62} ultrafast bootstrap
-approximation\textsuperscript{63} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{62} v1.6.12
+with ModelFinder\textsuperscript{63} ultrafast bootstrap
+approximation\textsuperscript{64} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{64} for each of the viruses in the tree,
+distinctiveness\textsuperscript{65} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{65,66} who
+this idea, we loosely adapted a method from,\textsuperscript{66,67} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{67} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{68} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Olival2020PosZoo}{}%
+\CSLLeftMargin{18. }
+\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
+zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
+study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}
+
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021LanUse}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Worobey, M. \emph{et al.} The huanan market was the
@@
 \leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Memish2013MidEas}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Memish, Z. A. \emph{et al.} Middle east respiratory
@@
 \leavevmode\hypertarget{ref-Muller2014MerCor}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{39. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{50. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@969464404b67c4a2d00f53e4c93c2eff97514bb0 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f08380c53e7fd9645cd7b52768acc1f057997063 🚀

--- 2022-06-28 draft.tex
@@
 habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock,\textsuperscript{20,21} and human population
-density and connectivity\textsuperscript{1,22,23}
+bridge hosts like
+livestock,\textsuperscript{20,\textbf{Rulli2021LanUse?}} and human
+population density and connectivity\textsuperscript{1,21,22}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 
-Reassuringly, these predictions are validated by the sites where the
-three highly pathogenic coronavirus are believed to have recently
-emerged. While available information puts the spillover of SARS-CoV-2 in
-a live animal market in Wuhan, China, the ultimate origin of the virus
-is almost certainly in a divergent lineage of sarbecoviruses from the
-Indochinese peninsula that was poorly characterized prior to the
-pandemic.\textsuperscript{24--26} Similarly, the SARS-CoV outbreak began
-in Guangdong province in 2002, reaching humans through small carnivore
-bridge hosts, but was eventually traced back to a set of likely
-progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province;\textsuperscript{27} nearby, antibody evidence has indicated
-human exposure to SARS-like viruses.\textsuperscript{28} MERS-CoV was
-originally detected in Saudi Arabia, accompanied by a nearly identical
-virus sequenced from an Egyptian tomb bat (\emph{Taphozous
-perforatus})\textsuperscript{29}, but is widespread in camels in East
-Africa and the Middle East, and may have reached its bridge host decades
-earlier than originally supposed;\textsuperscript{30} as a result, the
-geography of the original bat-to-camel transmission is still widely
-regarded as uncertain. All of these are broadly consistent with the risk
-factors we identify. Notably, India and west Africa are additional
-hotspots that have yet to experience the emergence of a bat coronavirus
-into human populations, but may still be at risk---particularly given
-known gaps in bat surveillance,\textsuperscript{17} and a dense
-population in both regions with global connectivity. In any of these
-regions, surveillance on viral reservoirs can be paired with targeted
-monitoring of high-risk human populations (i.e., those with regular
-wildlife contact)\textsuperscript{31} for maximum impact.
+Reassuringly, these predictions correspond to the geographic origins of
+the three bat-origin coronaviruses that have recently emerged in human
+populations. While available information puts the spillover of
+SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
+of the virus is almost certainly in a divergent lineage of
+sarbecoviruses from the Indochinese peninsula that was poorly
+characterized prior to the pandemic.\textsuperscript{23--25} Similarly,
+the SARS-CoV outbreak began in Guangdong province in 2002, reaching
+humans through small carnivore bridge hosts, but was eventually traced
+back to a set of likely progenitor viruses found in cave-dwelling
+horseshoe bats in Yunnan province;\textsuperscript{26} nearby, antibody
+evidence has indicated human exposure to SARS-like
+viruses.\textsuperscript{27} MERS-CoV was originally detected in Saudi
+Arabia, accompanied by a nearly identical virus sequenced from an
+Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{28}, but
+is widespread in camels in East Africa and the Middle East, and may have
+reached its bridge host decades earlier than originally
+supposed;\textsuperscript{29} as a result, the geography of the original
+bat-to-camel transmission is still widely regarded as uncertain. All of
+these are broadly consistent with the risk factors we identify. Notably,
+India and west Africa are additional hotspots that have yet to
+experience the emergence of a bat coronavirus into human populations,
+but may still be at risk---particularly given known gaps in bat
+surveillance,\textsuperscript{17} and a dense population in both regions
+with global connectivity. In any of these regions, surveillance on viral
+reservoirs can be paired with targeted monitoring of high-risk human
+populations (i.e., those with regular wildlife
+contact)\textsuperscript{30} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{32,33} Chiropterans emerged around 64 million
+health.\textsuperscript{31,32} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{34,35}
+estimated richness of more than 1400 species.\textsuperscript{33,34}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{36} For example,
+ecosystem-derived benefits to human.\textsuperscript{35} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{37} Over two-thirds of bats are know to be
+networks.\textsuperscript{36} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{38,39} and
+contributing for agricultural pest control,\textsuperscript{37,38} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{40,41} Because bats are globally distributed and
+health.\textsuperscript{39,40} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{42} Both of these regions are priority areas
+lineage).\textsuperscript{41} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{43}
+to the region.\textsuperscript{42}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{44}
+anthropogenic factors others than shifting climates.\textsuperscript{43}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{45} As our results establish that the
+habitats.\textsuperscript{44} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{46} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{47} This
+novel diseases,\textsuperscript{45} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{46} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{48} In particular, the challenge ahead lies
+encounters.\textsuperscript{47} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{49} The range maps were subsequently rasterized
+data.\textsuperscript{48} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{49} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{51} based on a recent synthetic tree with
+Diversity\textsuperscript{50} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{52} Faith's PD measures the sum of unique
+species.\textsuperscript{51} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{53} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{52} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{54} transformed
+location, and following best practices\textsuperscript{53} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{55} shows that LCBD
+community. Recent results by\textsuperscript{54} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{56} This network stores pairwise values of
+mammals.\textsuperscript{55} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{57} In this approach, every risk component gets
+model.\textsuperscript{56} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{58} in that this allows multiple
+and low viral sharing,\textsuperscript{57} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{59} we measured the risk level by
+extraction,\textsuperscript{58} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{60} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{59} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{61} v1.6.12
-with ModelFinder\textsuperscript{62} ultrafast bootstrap
-approximation\textsuperscript{63} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{60} v1.6.12
+with ModelFinder\textsuperscript{61} ultrafast bootstrap
+approximation\textsuperscript{62} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{64} for each of the viruses in the tree,
+distinctiveness\textsuperscript{63} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{65,66} who
+this idea, we loosely adapted a method from,\textsuperscript{64,65} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{67} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{66} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Rulli2021LanUse}{}%
+\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
 \CSLLeftMargin{20. }
-\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
-Land-use change and the livestock revolution increase the risk of
-zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
-Food} \textbf{2}, 409--416 (2021).}
-
-\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{21. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
+\CSLLeftMargin{23. }
+\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
+epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
+doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}
+
+\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
 \CSLLeftMargin{24. }
-\CSLRightInline{Worobey, M. \emph{et al.} The huanan market was the
-epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022).}
-
-\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{25. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Memish2013MidEas}{}%
+\CSLLeftMargin{28. }
+\CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
+syndrome coronavirus in bats, Saudi Arabia. \emph{Emerging infectious
+diseases} \textbf{19}, 1819 (2013).}
+
+\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
 \CSLLeftMargin{29. }
-\CSLRightInline{Memish, Z. A. \emph{et al.} Middle east respiratory
-syndrome coronavirus in bats, saudi arabia. \emph{Emerging infectious
-diseases} \textbf{19}, 1819 (2013).}
-
-\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
+\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
+neutralizing antibodies in camels, Eastern Africa, 1983--1997.
+\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}
+
+\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
 \CSLLeftMargin{30. }
-\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
-neutralizing antibodies in camels, eastern africa, 1983--1997.
-\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}
-
-\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{31. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{37. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{48. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@130106cead1e869da92e8484ce7935e60b6734ad 🚀

--- 2022-06-28 draft.tex
@@
 Driven by the need to understand the ecological factors involved in the
-emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
-and COVID-19 in human) through bat hosts, we develop an approach to the
+emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
+COVID-19 in human) through bat hosts, we develop an approach to the
 assesment of spillover risk based on the Geographic Mosaic Theory of
@@
 complexes\textsuperscript{9} of species are structured in space by the
-rippling effects of abiotic conditions onto evolutionary mechanism,
-giving rise to fragmented systems with different structure and
-ecologically dynamics over large spatial extents.\textsuperscript{10}
-The GMTC predicts a spatial fragmentation of coevolutionary dynamics
-under the joint action of three processes:\textsuperscript{11}
-coevolutionary hot- and coldspots, which appear when the intensity of
-\emph{interaction} (in terms of reciprocal fitness consequences) varies
-spatially; selection mosaics, wherein the intensity of \emph{selection}
-varies across space, driven by both the biotic complexity of the
-community (locally diverse hosts and viruses are more biotically
-complex) and the local favorability of the
+rippling effects of abiotic conditions onto evolutionary mechanisms,
+giving rise to fragmented systems with different ecologies over large
+spatial extents.\textsuperscript{10} The GMTC predicts a spatial
+fragmentation of coevolutionary dynamics under the joint action of three
+processes:\textsuperscript{11} coevolutionary hot- and coldspots, which
+appear when the intensity of \emph{interaction} (in terms of reciprocal
+fitness consequences) varies spatially; selection mosaics, wherein the
+intensity of \emph{selection} varies across space, driven by both the
+biotic complexity of the community (locally diverse hosts and viruses
+are more biotically complex) and the local favorability of the
 environment;\textsuperscript{12} and trait remixing, which occurs when
-coevolutionary dynamics are driven by by the arrival (or departure) of
-\emph{functional traits}, through changes in community composition due
-to invasions, meta-community dynamics, and disperal.
+coevolutionary dynamics change when community-level \emph{functional
+traits} change through meta-community dynamics.
 
@@
 
-Reassuringly, these predictions are validated by the sites where the
-three highly pathogenic coronavirus are believed to have recently
-emerged. While available information puts the spillover of SARS-CoV-2 in
-a live animal market in Wuhan, China, the ultimate origin of the virus
-is almost certainly in a divergent lineage of sarbecoviruses from the
-Indochinese peninsula that was poorly characterized prior to the
-pandemic.\textsuperscript{25--27} Similarly, the SARS-CoV outbreak began
-in Guangdong province in 2002, reaching humans through small carnivore
-bridge hosts, but was eventually traced back to a set of likely
-progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province;\textsuperscript{28} nearby, antibody evidence has indicated
-human exposure to SARS-like viruses.\textsuperscript{29} MERS-CoV was
-originally detected in Saudi Arabia, accompanied by a nearly identical
-virus sequenced from an Egyptian tomb bat (\emph{Taphozous
-perforatus})\textsuperscript{30}, but is widespread in camels in East
-Africa and the Middle East, and may have reached its bridge host decades
-earlier than originally supposed;\textsuperscript{31} as a result, the
-geography of the original bat-to-camel transmission is still widely
-regarded as uncertain. All of these are broadly consistent with the risk
-factors we identify. Notably, India and west Africa are additional
-hotspots that have yet to experience the emergence of a bat coronavirus
-into human populations, but may still be at risk---particularly given
-known gaps in bat surveillance,\textsuperscript{17} and a dense
-population in both regions with global connectivity. In any of these
-regions, surveillance on viral reservoirs can be paired with targeted
-monitoring of high-risk human populations (i.e., those with regular
-wildlife contact)\textsuperscript{32} for maximum impact.
+Reassuringly, these predictions correspond to the geographic origins of
+the three bat-origin coronaviruses that have recently emerged in human
+populations. While available information puts the spillover of
+SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
+of the virus is almost certainly in a divergent lineage of
+sarbecoviruses from the Indochinese peninsula that was poorly
+characterized prior to the pandemic.\textsuperscript{25--27} Similarly,
+the SARS-CoV outbreak began in Guangdong province in 2002, reaching
+humans through small carnivore bridge hosts, but was eventually traced
+back to a set of likely progenitor viruses found in cave-dwelling
+horseshoe bats in Yunnan province;\textsuperscript{28} nearby, antibody
+evidence has indicated human exposure to SARS-like
+viruses.\textsuperscript{29} MERS-CoV was originally detected in Saudi
+Arabia, accompanied by a nearly identical virus sequenced from an
+Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{30}, but
+is widespread in camels in East Africa and the Middle East, and may have
+reached its bridge host decades earlier than originally
+supposed;\textsuperscript{31} as a result, the geography of the original
+bat-to-camel transmission is still widely regarded as uncertain. All of
+these are broadly consistent with the risk factors we identify. Notably,
+India and west Africa are additional hotspots that have yet to
+experience the emergence of a bat coronavirus into human populations,
+but may still be at risk---particularly given known gaps in bat
+surveillance,\textsuperscript{17} and a dense population in both regions
+with global connectivity. In any of these regions, surveillance on viral
+reservoirs can be paired with targeted monitoring of high-risk human
+populations (i.e., those with regular wildlife
+contact)\textsuperscript{32} for maximum impact.
 
@@
 
-\leavevmode\hypertarget{ref-Olival2020PosZoo}{}%
+\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
 \CSLLeftMargin{18. }
@@
 
-\leavevmode\hypertarget{ref-Rulli2021LanUse}{}%
+\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
 \CSLLeftMargin{21. }
@@
 \CSLLeftMargin{25. }
-\CSLRightInline{Worobey, M. \emph{et al.} The huanan market was the
-epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022).}
+\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
+epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
+doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}
 
@@
 \CSLLeftMargin{30. }
-\CSLRightInline{Memish, Z. A. \emph{et al.} Middle east respiratory
-syndrome coronavirus in bats, saudi arabia. \emph{Emerging infectious
+\CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
+syndrome coronavirus in bats, Saudi Arabia. \emph{Emerging infectious
 diseases} \textbf{19}, 1819 (2013).}
@@
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
-neutralizing antibodies in camels, eastern africa, 1983--1997.
+neutralizing antibodies in camels, Eastern Africa, 1983--1997.
 \emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c17f262cb73c011919904ab00089570c6da12626 🚀

@@
 between the two groups, and can be coarsely delineated as (1) south and
-southeast Asia, (2) east Asia (including northern China), west Asia, and
+southeast Asia; (2) east Asia (including northern China), west Asia, and
 the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
-Africa and south America. In some cases, this diverges from expectations
+Africa and Latin America. In some cases, this diverges from expectations
 about coronavirus biogeography: for example, previous work has rarely
@@
 southeast Asian peninsula (and indeed, the region is home to known bat
-hosts of nobecoviruses, sarbecoviruses, and
-merbecoviruses).\textsuperscript{3}
+hosts of multiple betacoronavirus subgenera, including nobecoviruses,
+sarbecoviruses, and merbecoviruses).\textsuperscript{3}
 
@@
 betacoronavirus biogeographic regions are largely consistent. This may
-be surprising, given that cospeciation plays a minor role in coronavirus
-diversification,\textsuperscript{2} a property that would theoretically
+be surprising, given that cross-species transmission may play a stronger
+role in coronavirus diversification than
+cospeciation---\textsuperscript{2}a property that would theoretically
 allow for substantial broad divergence in their biogeography. However,
@@
 example, the most speciose and cosmopolitan family of bats, the vesper
-bats (Vespertilionidae), are considered the primary reservoir of
-merbecoviruses;\textsuperscript{3,13} but in the Americas, where
-merbecoviruses are the only lineage present, they have only been found
-in other bat taxa. At the coarsest scale, these heterogeneities are
-lost, and betacoronavirus biogeography tracks the deep rifts in bat
+bats (Vespertilionidae), are considered the primary hosts of the
+subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,13}
+but in the Americas, where merbecoviruses are the only lineage present,
+they have only been found in other bat taxa (e.g., Molossidae,
+Phyllostomidae) (refs1). At the coarsest scale, these heterogeneities
+are lost, and betacoronavirus biogeography tracks the deep rifts in bat
 evolutionary history---but within broad regions, the component
@@
 phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
-contrast to the sparsity of Neotropical betacoronavirus hosts, South
-America has the most evolutionary distinct hosts \emph{and} viruses,
-followed by secondary hotspots in southeast Asia and the Rift Valley
-region have mostly distinct viruses. Some degree of sampling bias may
-contribute to these patterns: for example, South-America is one of the
-places where the fewest bat betacoronavirus sequences have been
-generated,\textsuperscript{2,14,16} resulting in a sparser phylogenetic
-tree, and artificially inflating distinctiveness; conversely,
-disproportionate research effort in eastern China\textsuperscript{17}
-may have led to a more complete inventory of the local diversity of
-coronaviruses, again inflating these metrics relative to underlying
-patterns. Even accounting for these potential biases, though, there is
-obvious heterogeneity in betacoronavirus evolutionary distinctiveness
-that is distinct from overall bat diversity.
+contrast to the sparsity of Neotropical betacoronavirus hosts, South and
+Central America have the most evolutionary distinct hosts \emph{and}
+viruses, followed by secondary hotspots in southeast Asia and the Rift
+Valley region have mostly distinct viruses. Some degree of sampling bias
+may contribute to these patterns: for example, the Neotropics are one of
+the places where the fewest bat betacoronavirus sequences have been
+generated (cite2), resulting in a sparser phylogenetic tree, and
+artificially inflating distinctiveness; conversely, disproportionate
+research effort in eastern China\textsuperscript{16} may have led to a
+more complete inventory of the local diversity of coronaviruses, again
+inflating these metrics relative to underlying patterns. Even accounting
+for these potential biases, though, there is obvious heterogeneity in
+betacoronavirus evolutionary distinctiveness that is distinct from
+overall bat diversity.
 
@@
 the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
-bats (Rhinolophidae) are both the reservoirs of the SARS-like viruses
+bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
 (subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
 been of the greatest interest to researchers\textsuperscript{13} (and so
-have been sampled most intensively).\textsuperscript{17} The hotspots of
+have been sampled most intensively).\textsuperscript{16} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 isolation by host vicariance. Out of the four main groups of
-betacoronaviruses, only the subgenus \emph{Merbecovirus} (MERS-like
-viruses) has been found in animals in the Americas---an introduction
-that is generally presumed to be ancient.\textsuperscript{3} While
-comparatively understudied, New World merbecoviruses have been found in
-the ghost-faced bats (Mormoopidae), New World leaf-nosed bats
-(Phyllostomidae), and free-tailed bats (Molossidae).\textsuperscript{18}
-The former two groups are endemic to the Neotropics, while the explosive
-adaptive radiations of the latter two (and particularly the
-phyllostomids) are responsible for the hotspot of bat diversity in the
-Amazon. Together, these clades of New World bats play host to a distinct
-regime of betacoronavirus coevolution.
+betacoronaviruses, only merbecoviruses have been found in animals in the
+Americas--- an introduction that is generally presumed to be
+ancient.\textsuperscript{3,17} While comparatively understudied, New
+World merbecoviruses have been found in the ghost-faced bats
+(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
+free-tailed bats (Molossidae) (refs3). The former two groups and a clade
+of the latter are endemic to the Neotropics, while the explosive
+adaptive radiations of the phyllostomids are responsible for the hotspot
+of bat diversity in the Amazon (refs4). Together, these clades of New
+World bats play host to a distinct regime of betacoronavirus
+coevolution.
 
@@
 divergence from sister species in both African and Asian
-continents.\textsuperscript{19} Recent surveillance\textsuperscript{20}
+continents.\textsuperscript{18} Recent surveillance\textsuperscript{19}
 has identified a novel \emph{Betacoronavirus} (in the subgenus
@@
 habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock,\textsuperscript{21,22} and human population
-density and connectivity\textsuperscript{1,23,24}
+bridge hosts like livestock,\textsuperscript{20,21} and human population
+density and connectivity\textsuperscript{1,22,23}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 sarbecoviruses from the Indochinese peninsula that was poorly
-characterized prior to the pandemic.\textsuperscript{25--27} Similarly,
+characterized prior to the pandemic.\textsuperscript{24--26} Similarly,
 the SARS-CoV outbreak began in Guangdong province in 2002, reaching
@@
 back to a set of likely progenitor viruses found in cave-dwelling
-horseshoe bats in Yunnan province;\textsuperscript{28} nearby, antibody
+horseshoe bats in Yunnan province;\textsuperscript{27} nearby, antibody
 evidence has indicated human exposure to SARS-like
-viruses.\textsuperscript{29} MERS-CoV was originally detected in Saudi
+viruses.\textsuperscript{28} MERS-CoV was originally detected in Saudi
 Arabia, accompanied by a nearly identical virus sequenced from an
-Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{30}, but
+Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{29}, but
 is widespread in camels in East Africa and the Middle East, and may have
 reached its bridge host decades earlier than originally
-supposed;\textsuperscript{31} as a result, the geography of the original
+supposed;\textsuperscript{30} as a result, the geography of the original
 bat-to-camel transmission is still widely regarded as uncertain. All of
@@
 but may still be at risk---particularly given known gaps in bat
-surveillance,\textsuperscript{17} and a dense population in both regions
+surveillance,\textsuperscript{16} and a dense population in both regions
 with global connectivity. In any of these regions, surveillance on viral
@@
 populations (i.e., those with regular wildlife
-contact)\textsuperscript{32} for maximum impact.
+contact)\textsuperscript{31} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{33,34} Chiropterans emerged around 64 million
+health.\textsuperscript{32,33} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{35,36}
+estimated richness of more than 1400 species.\textsuperscript{34,35}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{43} Both of these regions are priority areas
+lineage).\textsuperscript{42} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{15,17} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{15,16} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 
-\leavevmode\hypertarget{ref-Allen2017GloHot}{}%
+\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
 \CSLLeftMargin{16. }
-\CSLRightInline{Allen, T. \emph{et al.} Global hotspots and correlates
-of emerging zoonotic diseases. \emph{Nature Communications} \textbf{8},
-(2017).}
-
-\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{17. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Olival2020PosRev}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{17. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{18. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
@@
 \leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Memish2013MidEas}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
@@
 \leavevmode\hypertarget{ref-Muller2014MerCor}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@abd9f48a0e70ba1f1619e6a9892a9eeb1e1aefc9 🚀

@@
 but in the Americas, where merbecoviruses are the only lineage present,
-they have only been found in other bat taxa (e.g., Molossidae,
-Phyllostomidae) (refs1). At the coarsest scale, these heterogeneities
-are lost, and betacoronavirus biogeography tracks the deep rifts in bat
-evolutionary history---but within broad regions, the component
-coevolutionary systems may have very different dynamics.
+they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
+Phyllostomidae).\textsuperscript{14--17} At the coarsest scale, these
+heterogeneities are lost, and betacoronavirus biogeography tracks the
+deep rifts in bat evolutionary history---but within broad regions, the
+component coevolutionary systems may have very different dynamics.
 
@@
 hotspots of viral adaptive radiation, and therefore areas of concern for
-human health.\textsuperscript{2,14} However, the hotspots of known bat
+human health.\textsuperscript{2,18} However, the hotspots of known bat
 betacoronavirus hosts show a distinct pattern, with primary hotspots
@@
 these undiscovered reservoirs should follow the same diversity
-gradient.\textsuperscript{15} In principle, these hotspots of
+gradient.\textsuperscript{19} In principle, these hotspots of
 locally-diverse, virus-rich bat communities should drive more adaptive
@@
 artificially inflating distinctiveness; conversely, disproportionate
-research effort in eastern China\textsuperscript{16} may have led to a
+research effort in eastern China\textsuperscript{20} may have led to a
 more complete inventory of the local diversity of coronaviruses, again
@@
 been of the greatest interest to researchers\textsuperscript{13} (and so
-have been sampled most intensively).\textsuperscript{16} The hotspots of
+have been sampled most intensively).\textsuperscript{20} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 richness---are almost entirely driven by viral adaptive radiation
-through host switching within this clade\textsuperscript{3,15}. In
+through host switching within this clade\textsuperscript{3,19}. In
 contrast, the Neotropical hotspot of viral distinctiveness is driven by
@@
 Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{3,17} While comparatively understudied, New
+ancient.\textsuperscript{3,21} While comparatively understudied, New
 World merbecoviruses have been found in the ghost-faced bats
 (Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae) (refs3). The former two groups and a clade
-of the latter are endemic to the Neotropics, while the explosive
-adaptive radiations of the phyllostomids are responsible for the hotspot
-of bat diversity in the Amazon (refs4). Together, these clades of New
-World bats play host to a distinct regime of betacoronavirus
-coevolution.
+free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
+groups and a clade of the latter are endemic to the Neotropics, while
+the explosive adaptive radiations of the phyllostomids are responsible
+for the hotspot of bat diversity in the Amazon.\textsuperscript{22}
+Together, these clades of New World bats play host to a distinct regime
+of betacoronavirus coevolution.
 
@@
 
-As host richness, joint distinctiveness, or phylogeographic structure
-suggest that the bat-betacoronaviruses complex is globally fragmented
-enough to give rise to both different levels of risk (as evidenced by
-the spatial location of spillover events) and different types of
-co-evolutionary dynamics, we turn to the Geographic Mosaic Theory of
-Coevolution to provide a measure of risk accounting for multiple
+Both the diverse origins of betacoronavirus spillover events and the
+existence of well defined co-phylogenetic regions suggest that the
+system is spatially fragmented enough to create risks of emergence that
+are different in their amplitude and in the coevolutionary processes
+underlying them. Therefore, we turn to the Geographic Mosaic Theory of
+Coevolution to provide an assesment of risk accounting for these
 processes. In fig.~\ref{fig:trivariate}, we overlapped three components
-of spillover risk: viral sharing, \emph{i.e.} the chance that two bats
-will share viruses overall; Local Contribution to Beta Diversity,
-\emph{i.e.} the fact that a bat community is compositionally unique
-compared to the average compositional similarity across the entire
-system; finally, host phylogenetic diversity, \emph{i.e.} how dispersed
-the bats in a location are within the tree of life. This approach leads
-to the definition of broad biogeographic regions of risk, where the same
-color represents the same type of risk. By way of constrat to figures
-fig.~\ref{fig:richness} and fig.~\ref{fig:biogeo}, these regions do not
-necessarilly overlap with previous spatial partitions of the
-bat-betacoronaviruses complex.
+of spillover risk: viral sharing (high viral sharing suggests slower,
+diffuse coevolution); host phylogenetic diversity (high diversity
+represents different evolutionary histories, suggesting more variation
+in immune strategies); host community uniqueness (high uniqueness
+suggests more potential for viruses to be exposed to novel host traits).
+This approach leads to the definition of broad biogeographic regions of
+risk, where close colors represent similar risks, and paler pixels
+represent overall higher risk (see Methods). These regions do not neatly
+overlap with those defined in fig.~\ref{fig:richness} or
+fig.~\ref{fig:biogeo}, reinforcing the notion that local-scale
+coevolutionary mosaics exist within co-phylogenetic regions.
 
@@
 host phylogenetic diversity in the green channel, and compositional
-uniqueness in the red channel. A pixel that would maximize all measures
-(highest possible risk) would be a pure white (specifically RGB(1.0,
-1.0. 1.0)), and a pixel with the lowest possible values would be pure
-black (specifically RGB(0.0, 0.0, 0.0)). Therefore, lighter values (the
-sum of the three channels gets closer to 3) indicate higher risk, and
-the color indicates the proportional distribution of the factors making
-up the total risk.}\label{fig:trivariate}
+uniqueness in the red channel. The main driver of emergence risk
+(possibility of spatially overlapped but coevolutionarily indepent
+host-viral dynamics) corresponds to low viral sharing, \emph{i.e.}
+pixels around yellow. Pixels in the yellow-green space (Africa)
+correspond to areas where, despite the potential for viral
+diversification, coevolutionary dynamics are likely to currently be
+dominated by the effect of host phylogenetic
+diversity.}\label{fig:trivariate}
 }
@@
 
-From the perspective of spillover risk, the most important combination
-of factors is a high phylogenetic diversity of hosts with low viral
-sharing; this, essentially, means that very different betacoronaviruses
-could co-exist within the same place. This is particularly the case
-given that betacoronaviruses often evolve and even achieve host shifts
-through recombination, which requires the co-occurrence of sufficiently
-distinct viruses to be a major driver of emergence. In
-fig.~\ref{fig:trivariate}, this corresponds to yellow to pale green
-areas, which are essentially limited to South-Eastern Asia, and to some
-part of Sub-Saharan Africa. Adopting a geographic mosaic theory
-perspective on risk, other regions of the world are of lesser concern
-(fig.~\ref{fig:risk}). Our risk decomposition does not account for viral
-diversity or distinctiveness. The simple rationale behind it is that the
-acquisition of viral data is rarely disconnected from the acquisition of
-host data. There are more sources of information on hosts than on
-viruses, allowing to develop a host-centric perspective on risk
-(although this estimate would more accurate with viral traits related to
-\emph{e.g.} ability to switch hosts or pathogenic potential). Areas with
-high bat diversity and high turnover \emph{may} facilitate the
-evolutionary radiation of viruses, matching previous findings that the
-diversification of bat coronaviruses is driven largely by host shifts
-(inter-genus or higher levels of cross-species transmission) and, to a
-lesser degree, cospeciation and sharing, representing intra-genus
+Emergence risk is maximized under low viral sharing (host-virus pairs
+coevolve independantly), high phylogenetic diversity (viruses are
+exposes to different host clades), and high host uniqueness (viruses are
+experiencing novel, heterogeneous host traits combinations). Under these
+conditions, very different betacoronaviruses could co-exist at the same
+place, yet evolve in independent ways. As betacoronaviruses often evolve
+(including host shifts) through recombination, the co-occurrence of
+sufficiently distinct viruses is a sufficient major driver of emergence,
+and the regions that meet these conditions therefore represent the
+higher risk. In fig.~\ref{fig:trivariate}, this corresponds to yellow
+areas (dynamics dominated by low viral sharing, with equal contributions
+of selection mosaics and trait remixing; South-Eastern Asia, and the
+Indian sub-continent), green-yellow areas (dynamics with low viral
+sharing but dominated by the selection mosaic effect of host diversity;
+Africa below the Sahara desert), and would correspond to red-yellow
+areas (dynamics with low viral sharing but dominated by trait remixing
+in host communities; Middle-East). Indeed, these regions are broadly
+reflected in the risk map fig.~\ref{fig:risk}.
+
+Under this framework, other regions (where high viral sharing dominates
+the dynamics; see \emph{e.g.} Latin America, Eurasia above a northing of
+30) represent a lower risk of emergence. Nevertheless, areas of high
+host uniqueness coupled with high viral sharing(red-to-pink in
+fig.~\ref{fig:trivariate}) could provide future hotspots of
+\emph{Betacoronavirus} emregence risk through the suddent divergence of
+currently diffuse coevolutionary dynamics; this is a likely scenario
+knowing that can be facilitated by codivergence followed by
+recombination. Madagascar, where most bat species are endemic following
+evolutionary divergence from sister species in both African and Asian
+continents,\textsuperscript{23} is one such potential future hotspot.
+Indeed, recent surveillance\textsuperscript{24} has identified a novel
+\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
+Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
+\emph{Rousettus madagascariensis}), emphasizing strong proof of
+principle in model predictions.
+
+Our risk decomposition does not account for viral diversity or
+distinctiveness. Viral data acquisition is rarely disconnected from the
+acquisition of host data. There are more sources of information on hosts
+than on viruses, allowing to develop a more robust host-centric
+perspective on risk. Any emergence risk estimate would benefit from
+viral traits related to \emph{e.g.} ability to switch hosts or
+pathogenic potential. This is particularly true under recent findings
+that the diversification of bat coronaviruses is driven largely by host
+shifts (inter-genus or higher levels of cross-species transmission) and,
+to a lesser degree, cospeciation and sharing, representing intra-genus
 cross-species transmission.\textsuperscript{2} This diversification is
-not an actual risk factor for spillover itself, but acts downstream of a
-spillover event by increasing the random chance of the emergence of a
-virus with the raw genomic components required for the potential to
-infect humans.
+not an actual risk factor for emergence itself, but acts downstream of
+an emergence event by increasing the random chance of the emergence of a
+virus with the raw genomic components required to eventually infect
+humans.
 
@@
 \includegraphics{figures/risk_map.png}
-\caption{Extraction of a measure of \emph{Betacoronavirus} spillover
+\caption{Extraction of a measure of \emph{Betacoronavirus} emergence
 risk from bat hosts based on the colorimetric space from
@@
 value and angular distance to the yellow hue, as defined in the methods,
-ranged in the unit space. Based on this analyses, regions at high risk
-of spillover are southeast Asia and Madagascar.}\label{fig:risk}
+ranged in the unit space. Based on these analyses, South-Eastern Asia,
+Madagascar, the Middle-East, and Africa below the Sahara desert have the
+highest relative risk of emergence.}\label{fig:risk}
 }
 \end{figure}
-
-From another perspective, areas of high host uniqueness and virus
-sharing (red-to-pink) could provide hotspots of \emph{Betacoronavirus}
-risk through mixing of unique viruses (via codivergence) and in turn
-recombination. Under our framework, such a hotspot was identified in
-Madagascar, where most bat species are endemic following evolutionary
-divergence from sister species in both African and Asian
-continents.\textsuperscript{18} Recent surveillance\textsuperscript{19}
-has identified a novel \emph{Betacoronavirus} (in the subgenus
-\emph{Nobecovirus}) in Madagascar-endemic pteropid bat species
-(\emph{Pteropus rufus}, \emph{Rousettus madagascariensis}), emphasizing
-strong proof of principle in model predictions.
 
@@
 habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock,\textsuperscript{20,21} and human population
-density and connectivity\textsuperscript{1,22,23}
+bridge hosts like livestock,\textsuperscript{25,26} and human population
+density and connectivity\textsuperscript{1,27,28}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 sarbecoviruses from the Indochinese peninsula that was poorly
-characterized prior to the pandemic.\textsuperscript{24--26} Similarly,
+characterized prior to the pandemic.\textsuperscript{29--31} Similarly,
 the SARS-CoV outbreak began in Guangdong province in 2002, reaching
@@
 back to a set of likely progenitor viruses found in cave-dwelling
-horseshoe bats in Yunnan province;\textsuperscript{27} nearby, antibody
+horseshoe bats in Yunnan province;\textsuperscript{32} nearby, antibody
 evidence has indicated human exposure to SARS-like
-viruses.\textsuperscript{28} MERS-CoV was originally detected in Saudi
+viruses.\textsuperscript{33} MERS-CoV was originally detected in Saudi
 Arabia, accompanied by a nearly identical virus sequenced from an
-Egyptian tomb bat (\emph{Taphozous perforatus})\textsuperscript{29}, but
+Egyptian tomb bat (\emph{Taphozous perforatus}),\textsuperscript{34} but
 is widespread in camels in East Africa and the Middle East, and may have
 reached its bridge host decades earlier than originally
-supposed;\textsuperscript{30} as a result, the geography of the original
+supposed;\textsuperscript{35} as a result, the geography of the original
 bat-to-camel transmission is still widely regarded as uncertain. All of
@@
 but may still be at risk---particularly given known gaps in bat
-surveillance,\textsuperscript{16} and a dense population in both regions
+surveillance,\textsuperscript{20} and a dense population in both regions
 with global connectivity. In any of these regions, surveillance on viral
@@
 populations (i.e., those with regular wildlife
-contact)\textsuperscript{31} for maximum impact.
+contact)\textsuperscript{36} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{32,33} Chiropterans emerged around 64 million
+health.\textsuperscript{37,38} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{34,35}
+estimated richness of more than 1400 species.\textsuperscript{39,40}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{36} For example,
+ecosystem-derived benefits to human.\textsuperscript{41} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{37} Over two-thirds of bats are know to be
+networks.\textsuperscript{42} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{38,39} and
+contributing for agricultural pest control,\textsuperscript{43,44} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{40,41} Because bats are globally distributed and
+health.\textsuperscript{45,46} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 that have used host richness as proxy have predicted a high diversity of
-unsampled bat viruses,\textsuperscript{14} bat
+unsampled bat viruses,\textsuperscript{18} bat
 coronaviruses,\textsuperscript{2} and even specifically
-betacoronaviruses\textsuperscript{15} in both the Amazon and southeast
+betacoronaviruses\textsuperscript{19} in both the Amazon and southeast
 Asia. While we find that both regions are characterized by highly
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{42} Both of these regions are priority areas
+lineage).\textsuperscript{47} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{15,16} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{19,20} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{43}
+to the region.\textsuperscript{48}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{44}
+anthropogenic factors others than shifting climates.\textsuperscript{49}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{45} As our results establish that the
+habitats.\textsuperscript{50} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{46} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{47} This
+novel diseases,\textsuperscript{51} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{52} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{48} In particular, the challenge ahead lies
+encounters.\textsuperscript{53} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 \texttt{https://www.viralemergence.org/betacov} on
-Apr.~2022,\textsuperscript{15} and filtered it to ``known'' hosts
+Apr.~2022,\textsuperscript{19} and filtered it to ``known'' hosts
 (established before the emergence of SARS-CoV-2) and ``novel'' hosts
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{49} The range maps were subsequently rasterized
+data.\textsuperscript{54} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{50} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{51} based on a recent synthetic tree with
+Diversity\textsuperscript{56} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{52} Faith's PD measures the sum of unique
+species.\textsuperscript{57} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{53} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{54} transformed
+location, and following best practices\textsuperscript{59} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{55} shows that LCBD
+community. Recent results by\textsuperscript{60} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{56} This network stores pairwise values of
+mammals.\textsuperscript{61} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{57} In this approach, every risk component gets
+model.\textsuperscript{62} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{58} in that this allows multiple
+and low viral sharing,\textsuperscript{63} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{59} we measured the risk level by
+extraction,\textsuperscript{64} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{60} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{65} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{61} v1.6.12
-with ModelFinder\textsuperscript{62} ultrafast bootstrap
-approximation\textsuperscript{63} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{66} v1.6.12
+with ModelFinder\textsuperscript{67} ultrafast bootstrap
+approximation\textsuperscript{68} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{64} for each of the viruses in the tree,
+distinctiveness\textsuperscript{69} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{65,66} who
+this idea, we loosely adapted a method from,\textsuperscript{70,71} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{67} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{72} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Anthony2013CorBat}{}%
+\CSLLeftMargin{14. }
+\CSLRightInline{Anthony, S. J. \emph{et al.} Coronaviruses in bats from
+Mexico. \emph{The Journal of General Virology} \textbf{94}, 1028--1038
+(2013).}
+
+\leavevmode\hypertarget{ref-Goes2013NovBat}{}%
+\CSLLeftMargin{15. }
+\CSLRightInline{Góes, L. G. B. \emph{et al.} Novel Bat Coronaviruses,
+Brazil and Mexico. \emph{Emerging Infectious Diseases} \textbf{19},
+1711--1713 (2013).}
+
+\leavevmode\hypertarget{ref-Goes2016GenDiv}{}%
+\CSLLeftMargin{16. }
+\CSLRightInline{Góes, L. G. B. \emph{et al.} Genetic diversity of bats
+coronaviruses in the Atlantic Forest hotspot biome, Brazil.
+\emph{Infection, Genetics and Evolution: Journal of Molecular
+Epidemiology and Evolutionary Genetics in Infectious Diseases}
+\textbf{44}, 510--513 (2016).}
+
+\leavevmode\hypertarget{ref-Brandao2008CorDet}{}%
+\CSLLeftMargin{17. }
+\CSLRightInline{Brandão, P. E. \emph{et al.} A coronavirus detected in
+the vampire bat Desmodus rotundus. \emph{Brazilian Journal of Infectious
+Diseases} \textbf{12}, 466--468 (2008).}
+
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-\CSLLeftMargin{14. }
+\CSLLeftMargin{18. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
@@
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-\CSLLeftMargin{15. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
@@
 \leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{16. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Olival2020PosRev}{}%
-\CSLLeftMargin{17. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
@@
 
+\leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
+\CSLLeftMargin{22. }
+\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
+molecular phylogenetic insights into the evolution of free-tailed bats
+in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
+Mammalogy} \textbf{93}, 12--28 (2012).}
+
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
@@
 \leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Memish2013MidEas}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
@@
 \leavevmode\hypertarget{ref-Muller2014MerCor}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{43. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
@@
 \leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{54. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ec3555b2e83eb8fec7839bfb58e33f91be22e17e 🚀

@@
 
-The Geographic Mosaic Theory of Coevolution (GMTC) attempts to
+The geographic mosaic theory of coevolution (GMTC) attempts to
 explicitly connect microevolutionary dynamics to the macroecology and
@@
 
-\hypertarget{coevolution-informed-emergence-risk-is-different-in-space}{%
-\subsection{Coevolution-informed emergence risk is different in
-space}\label{coevolution-informed-emergence-risk-is-different-in-space}}
-
-Both the diverse origins of betacoronavirus spillover events and the
-existence of well defined co-phylogenetic regions suggest that the
-system is spatially fragmented enough to create risks of emergence that
-are different in their amplitude and in the coevolutionary processes
-underlying them. Therefore, we turn to the Geographic Mosaic Theory of
-Coevolution to provide an assesment of risk accounting for these
-processes. In fig.~\ref{fig:trivariate}, we overlapped three components
-of spillover risk: viral sharing (high viral sharing suggests slower,
-diffuse coevolution); host phylogenetic diversity (high diversity
-represents different evolutionary histories, suggesting more variation
-in immune strategies); host community uniqueness (high uniqueness
-suggests more potential for viruses to be exposed to novel host traits).
-This approach leads to the definition of broad biogeographic regions of
-risk, where close colors represent similar risks, and paler pixels
-represent overall higher risk (see Methods). These regions do not neatly
-overlap with those defined in fig.~\ref{fig:richness} or
+\hypertarget{the-coevolutionary-mosaic-generates-emergence-regimes}{%
+\subsection{The coevolutionary mosaic generates emergence
+regimes}\label{the-coevolutionary-mosaic-generates-emergence-regimes}}
+
+The existence of well-defined cophylogenetic regions suggests that the
+bat-betacoronavirus system is spatially fragmented enough to create
+differential coevolutionary processes; in turn, the coevolutionary
+mosaic contributes to heterogeneity in emergence risk. These ideas are,
+respectively, supported by the existence of hotspots of viral uniqueness
+and the diverse origins of human betacoronaviruses. Together, these
+ideas point to a predictable relationship between host community
+structure and coevolutionary pressure: phylogeographic structure in bat
+hosts---and their diverse immune
+strategies---\textsuperscript{23}creates a landscape of selective
+pressure; the trajectory of viruses' coevolutionary response is, in
+turn, constrained by the opportunities they have for either
+coevolutionary specialization or diversification through host jumps and
+recombination.
+
+Based on the geographic mosaic theory of coevolution, we developed a
+trivariate map of three facets of coevolutionary pressure (see Methods):
+(1) \emph{host phylogenetic diversity}: a high diversity of evolutionary
+histories should expose viruses to more variation in host immune traits;
+(2) \emph{host community uniqueness}: coevolving with more unique host
+communities should create more unique branches of viral evolution; and
+(3) propensity for \emph{viral sharing}: frequent cross-species
+transmission may act as a buffer on selective pressure, while lower
+rates of exchange may enable more simultaneous trajectories of viral
+specialization to coexist within a given community. We combine global
+maps of all three to generate a map of coevolutionary regimes, where
+close colors represent similar risks, and paler pixels represent overall
+higher risk (fig.~\ref{fig:trivariate}). We find that these regions do
+not neatly overlap with those defined in fig.~\ref{fig:richness} or
 fig.~\ref{fig:biogeo}, reinforcing the notion that local-scale
-coevolutionary mosaics exist within co-phylogenetic regions.
+coevolutionary mosaics can form within cophylogenetic regions.
 
@@
 evolutionary divergence from sister species in both African and Asian
-continents,\textsuperscript{23} is one such potential future hotspot.
-Indeed, recent surveillance\textsuperscript{24} has identified a novel
+continents,\textsuperscript{24} is one such potential future hotspot.
+Indeed, recent surveillance\textsuperscript{25} has identified a novel
 \emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
@@
 habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock,\textsuperscript{25,26} and human population
-density and connectivity\textsuperscript{1,27,28}
+bridge hosts like livestock,\textsuperscript{26,27} and human population
+density and connectivity\textsuperscript{1,28,29}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 sarbecoviruses from the Indochinese peninsula that was poorly
-characterized prior to the pandemic.\textsuperscript{29--31} Similarly,
+characterized prior to the pandemic.\textsuperscript{30--32} Similarly,
 the SARS-CoV outbreak began in Guangdong province in 2002, reaching
@@
 back to a set of likely progenitor viruses found in cave-dwelling
-horseshoe bats in Yunnan province;\textsuperscript{32} nearby, antibody
+horseshoe bats in Yunnan province;\textsuperscript{33} nearby, antibody
 evidence has indicated human exposure to SARS-like
-viruses.\textsuperscript{33} MERS-CoV was originally detected in Saudi
+viruses.\textsuperscript{34} MERS-CoV was originally detected in Saudi
 Arabia, accompanied by a nearly identical virus sequenced from an
-Egyptian tomb bat (\emph{Taphozous perforatus}),\textsuperscript{34} but
+Egyptian tomb bat (\emph{Taphozous perforatus}),\textsuperscript{35} but
 is widespread in camels in East Africa and the Middle East, and may have
 reached its bridge host decades earlier than originally
-supposed;\textsuperscript{35} as a result, the geography of the original
+supposed;\textsuperscript{36} as a result, the geography of the original
 bat-to-camel transmission is still widely regarded as uncertain. All of
@@
 populations (i.e., those with regular wildlife
-contact)\textsuperscript{36} for maximum impact.
+contact)\textsuperscript{37} for maximum impact.
 
@@
 microorganisms, many of which a threat to human
-health.\textsuperscript{37,38} Chiropterans emerged around 64 million
+health.\textsuperscript{38,39} Chiropterans emerged around 64 million
 years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{39,40}
+estimated richness of more than 1400 species.\textsuperscript{40,41}
 They exhibit a broad variety of habitat use, behaviour, and feeding
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{41} For example,
+ecosystem-derived benefits to human.\textsuperscript{42} For example,
 bats are an essential component of many seed-dispersal
-networks.\textsuperscript{42} Over two-thirds of bats are know to be
+networks.\textsuperscript{43} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{43,44} and
+contributing for agricultural pest control,\textsuperscript{44,45} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{45,46} Because bats are globally distributed and
+health.\textsuperscript{46,47} Because bats are globally distributed and
 have a long evolutionary history, phylogeographic and biogeographic
@@
 the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{47} Both of these regions are priority areas
+lineage).\textsuperscript{48} Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{48}
+to the region.\textsuperscript{49}
 
 The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{49}
+anthropogenic factors others than shifting climates.\textsuperscript{50}
 Land use changes could significantly decrease bat suitability, notably
 through effects on diet and availability of
-habitats.\textsuperscript{50} As our results establish that the
+habitats.\textsuperscript{51} As our results establish that the
 diversification of bats betacoronaviruses happens on top of processes
@@
 Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{51} so does the changes in landscape
-connectivity at local/regional scales.\textsuperscript{52} This
+novel diseases,\textsuperscript{52} so does the changes in landscape
+connectivity at local/regional scales.\textsuperscript{53} This
 represents a challenge for both conservation strategies and disease
@@
 acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{53} In particular, the challenge ahead lies
+encounters.\textsuperscript{54} In particular, the challenge ahead lies
 in the need to quantify actual exposure (and risk) accounting for
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{61} This network stores pairwise values of
+mammals.\textsuperscript{62} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{63} in that this allows multiple
+and low viral sharing,\textsuperscript{64} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{64} we measured the risk level by
+extraction,\textsuperscript{65} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{65} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{66} v1.6.12
-with ModelFinder\textsuperscript{67} ultrafast bootstrap
-approximation\textsuperscript{68} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{67} v1.6.12
+with ModelFinder\textsuperscript{68} ultrafast bootstrap
+approximation\textsuperscript{69} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{69} for each of the viruses in the tree,
+distinctiveness\textsuperscript{70} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{70,71} who
+this idea, we loosely adapted a method from,\textsuperscript{71,72} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{72} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
+\CSLLeftMargin{23. }
+\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
+Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}
+
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
@@
 \leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Memish2013MidEas}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
@@
 \leavevmode\hypertarget{ref-Muller2014MerCor}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Letko2020BatVir}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
@@
 \leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{44. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{73. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b64f51898236a38c2da87a7aeaadb800649da1ab 🚀

@@
 
-\hypertarget{the-coevolutionary-mosaic-generates-emergence-regimes}{%
-\subsection{The coevolutionary mosaic generates emergence
-regimes}\label{the-coevolutionary-mosaic-generates-emergence-regimes}}
+\hypertarget{host-landscapes-generate-a-coevolutionary-mosaic}{%
+\subsection{Host landscapes generate a coevolutionary
+mosaic}\label{host-landscapes-generate-a-coevolutionary-mosaic}}
 
@@
 histories should expose viruses to more variation in host immune traits;
-(2) \emph{host community uniqueness}: coevolving with more unique host
-communities should create more unique branches of viral evolution; and
-(3) propensity for \emph{viral sharing}: frequent cross-species
-transmission may act as a buffer on selective pressure, while lower
-rates of exchange may enable more simultaneous trajectories of viral
-specialization to coexist within a given community. We combine global
-maps of all three to generate a map of coevolutionary regimes, where
-close colors represent similar risks, and paler pixels represent overall
-higher risk (fig.~\ref{fig:trivariate}). We find that these regions do
-not neatly overlap with those defined in fig.~\ref{fig:richness} or
-fig.~\ref{fig:biogeo}, reinforcing the notion that local-scale
-coevolutionary mosaics can form within cophylogenetic regions.
+(2) \emph{host community uniqueness}: exposure to greater host trait
+heterogeneity can drive viral diversification, and coevolving with more
+unique host communities should create more unique branches of viral
+evolution; and (3) propensity for \emph{viral sharing}: frequent
+cross-species transmission may act as a buffer on selective pressure,
+while lower rates of exchange may enable more simultaneous trajectories
+of viral specialization to coexist within a given community. We combine
+global maps of all three to generate a map of coevolutionary regimes,
+where close colors represent similar risks, and paler pixels represent
+overall higher risk (fig.~\ref{fig:trivariate}). We find that these
+regions do not neatly overlap with those defined in
+fig.~\ref{fig:biogeo} or fig.~\ref{fig:richness}, reinforcing the notion
+that local-scale coevolutionary mosaics can form within cophylogenetic
+regions.
 
@@
 host-viral dynamics) corresponds to low viral sharing, \emph{i.e.}
-pixels around yellow. Pixels in the yellow-green space (Africa)
-correspond to areas where, despite the potential for viral
-diversification, coevolutionary dynamics are likely to currently be
-dominated by the effect of host phylogenetic
-diversity.}\label{fig:trivariate}
+pixels around yellow.}\label{fig:trivariate}
 }
@@
 
-Emergence risk is maximized under low viral sharing (host-virus pairs
-coevolve independantly), high phylogenetic diversity (viruses are
-exposes to different host clades), and high host uniqueness (viruses are
-experiencing novel, heterogeneous host traits combinations). Under these
-conditions, very different betacoronaviruses could co-exist at the same
-place, yet evolve in independent ways. As betacoronaviruses often evolve
-(including host shifts) through recombination, the co-occurrence of
-sufficiently distinct viruses is a sufficient major driver of emergence,
-and the regions that meet these conditions therefore represent the
-higher risk. In fig.~\ref{fig:trivariate}, this corresponds to yellow
-areas (dynamics dominated by low viral sharing, with equal contributions
-of selection mosaics and trait remixing; South-Eastern Asia, and the
-Indian sub-continent), green-yellow areas (dynamics with low viral
-sharing but dominated by the selection mosaic effect of host diversity;
-Africa below the Sahara desert), and would correspond to red-yellow
-areas (dynamics with low viral sharing but dominated by trait remixing
-in host communities; Middle-East). Indeed, these regions are broadly
-reflected in the risk map fig.~\ref{fig:risk}.
-
-Under this framework, other regions (where high viral sharing dominates
-the dynamics; see \emph{e.g.} Latin America, Eurasia above a northing of
-30) represent a lower risk of emergence. Nevertheless, areas of high
-host uniqueness coupled with high viral sharing(red-to-pink in
-fig.~\ref{fig:trivariate}) could provide future hotspots of
-\emph{Betacoronavirus} emregence risk through the suddent divergence of
-currently diffuse coevolutionary dynamics; this is a likely scenario
-knowing that can be facilitated by codivergence followed by
-recombination. Madagascar, where most bat species are endemic following
-evolutionary divergence from sister species in both African and Asian
-continents,\textsuperscript{24} is one such potential future hotspot.
-Indeed, recent surveillance\textsuperscript{25} has identified a novel
-\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
-Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
-\emph{Rousettus madagascariensis}), emphasizing strong proof of
-principle in model predictions.
-
-Our risk decomposition does not account for viral diversity or
-distinctiveness. Viral data acquisition is rarely disconnected from the
-acquisition of host data. There are more sources of information on hosts
-than on viruses, allowing to develop a more robust host-centric
-perspective on risk. Any emergence risk estimate would benefit from
-viral traits related to \emph{e.g.} ability to switch hosts or
-pathogenic potential. This is particularly true under recent findings
-that the diversification of bat coronaviruses is driven largely by host
-shifts (inter-genus or higher levels of cross-species transmission) and,
-to a lesser degree, cospeciation and sharing, representing intra-genus
-cross-species transmission.\textsuperscript{2} This diversification is
-not an actual risk factor for emergence itself, but acts downstream of
-an emergence event by increasing the random chance of the emergence of a
-virus with the raw genomic components required to eventually infect
-humans.
+The greatest evolutionary potential for zoonotic emergence exists where
+pathogen pools have a high genetic diversity and high propensity for
+cross-species transmission. In our framework, emergence risk is
+therefore maximized under higher phylogenetic diversity (viruses are
+exposes to different host clades), higher host uniqueness (viruses are
+experiencing novel, heterogeneous host traits combinations), and low to
+medium viral sharing (host-virus pairs can coevolve independently, but
+divergent viruses may still have opportunities for recombination). In
+fig.~\ref{fig:trivariate}, this corresponds to yellow areas (dynamics
+dominated by low viral sharing, with equal contributions of selection
+mosaics and trait remixing; southeast Asia, and the Indian
+sub-continent), green-yellow areas (dynamics with low viral sharing but
+dominated by the selection mosaic effect of host diversity; Africa below
+the Sahara desert), and red-yellow areas (dynamics with low viral
+sharing but dominated by trait remixing in host communities; the Middle
+East). Translating this axis of variation back into a univariate risk
+map (fig.~\ref{fig:risk}) highlights that this evolutionary landscape
+has a striking correspondence to regions where zoonotic
+betacoronaviruses have previously emerged.
+
+Under this framework, other regions where high viral sharing dominates
+coevolutionary dynamics---such as Latin America, or Eurasia above a
+northing of 30---would pose less of a risk of zoonotic emergence.
+Nevertheless, areas of high host uniqueness coupled with high viral
+sharing (red-to-pink in fig.~\ref{fig:trivariate}) could create hotspots
+facilitated by viral codivergence. Our framework identifies Madagascar,
+where most bat species are endemic following evolutionary divergence
+from sister species in both African and Asian
+continents,\textsuperscript{24} as one such hotspot; interestingly, a
+recent study\textsuperscript{25} reported a novel and highly divergent
+lineage of nobecoviruses from Madagascar-endemic pteropid bat species
+(\emph{Pteropus rufus} and \emph{Rousettus madagascariensis}), again
+supporting the predictive power of the coevolutionary framework.
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6e83254033e0832aa8e4a280510cbc570cdcdc62 🚀

@@
 
-Under this framework, other regions where high viral sharing dominates
-coevolutionary dynamics---such as Latin America, or Eurasia above a
-northing of 30---would pose less of a risk of zoonotic emergence.
-Nevertheless, areas of high host uniqueness coupled with high viral
-sharing (red-to-pink in fig.~\ref{fig:trivariate}) could create hotspots
-facilitated by viral codivergence. Our framework identifies Madagascar,
-where most bat species are endemic following evolutionary divergence
-from sister species in both African and Asian
-continents,\textsuperscript{24} as one such hotspot; interestingly, a
-recent study\textsuperscript{25} reported a novel and highly divergent
-lineage of nobecoviruses from Madagascar-endemic pteropid bat species
-(\emph{Pteropus rufus} and \emph{Rousettus madagascariensis}), again
-supporting the predictive power of the coevolutionary framework.
+Compared to approaches that map emergence risk based only on the number
+of known bat hosts of betacoronaviruses, our framework suggests regions
+where high viral sharing dominates coevolutionary dynamics---such as
+Latin America, or Eurasia above a northing of 30---would pose less of a
+relative risk of zoonotic emergence. Nevertheless, areas of high host
+uniqueness coupled with high viral sharing (red-to-pink in
+fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
+codivergence. Our framework identifies Madagascar, where most bat
+species are endemic following evolutionary divergence from sister
+species in both African and Asian continents,\textsuperscript{24} as one
+such hotspot; interestingly, a recent study\textsuperscript{25} reported
+a novel and highly divergent lineage of nobecoviruses from
+Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
+\emph{Rousettus madagascariensis}), again supporting the predictive
+power of the coevolutionary framework.
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b73a0f667deef12e815dc9d6d434575272fd86b3 🚀

@@
 
-\hypertarget{host-landscapes-generate-a-coevolutionary-mosaic}{%
-\subsection{Host landscapes generate a coevolutionary
-mosaic}\label{host-landscapes-generate-a-coevolutionary-mosaic}}
+\hypertarget{coevolutionary-regimes-structure-evolutionary-risk-of-zoonotic-emergence}{%
+\subsection{Coevolutionary regimes structure evolutionary risk of
+zoonotic
+emergence}\label{coevolutionary-regimes-structure-evolutionary-risk-of-zoonotic-emergence}}
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@bfb6721e341ce678ea9848f21ba6695a41b0571f 🚀

@@
 sub-continent), green-yellow areas (dynamics with low viral sharing but
-dominated by the selection mosaic effect of host diversity; Africa below
-the Sahara desert), and red-yellow areas (dynamics with low viral
-sharing but dominated by trait remixing in host communities; the Middle
-East). Translating this axis of variation back into a univariate risk
-map (fig.~\ref{fig:risk}) highlights that this evolutionary landscape
-has a striking correspondence to regions where zoonotic
-betacoronaviruses have previously emerged.
+dominated by the selection mosaic effect of host diversity; sub-Saharan
+Africa), and red-yellow areas (dynamics with low viral sharing but
+dominated by trait remixing in host communities; the Middle East).
+Translating this axis of variation back into a univariate risk map
+(fig.~\ref{fig:risk}) highlights that this evolutionary landscape has a
+striking correspondence to regions where zoonotic betacoronaviruses have
+previously emerged.
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@dd96400fd33ea3c63c56143b8e9b6e484ba54afa 🚀

@@
 \includegraphics{figures/risk_trivariate.png}
-\caption{Trivariate additive mapping of the components of risk in the
-red/green/blue, where high virus sharing is encoded in the blue channel,
-host phylogenetic diversity in the green channel, and compositional
-uniqueness in the red channel. The main driver of emergence risk
-(possibility of spatially overlapped but coevolutionarily indepent
-host-viral dynamics) corresponds to low viral sharing, \emph{i.e.}
+\caption{Trivariate additive mapping of the components of risk. Viral
+sharing runs from yellow (low) to blue (high); host phylogenetic
+diversity runs from pink (low) to high (green) channel, and
+compositional uniqueness runs from cyan (low) to red (high). The GMTC
+suggests that the highest evolutionary potential for emergence exists in
+unique and diverse host communities with low viral sharing, \emph{i.e.}
 pixels around yellow.}\label{fig:trivariate}

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f552dd62fe1d0ee94574e0ff7a95536f59860deb 🚀

@@
 unique and diverse host communities with low viral sharing, \emph{i.e.}
-pixels around yellow.}\label{fig:trivariate}
+pixels around yellow. All components are scaled in brightness so that a
+pixel with no sharing, no phylogenetic diversity, and no compositional
+uniqueness would be black, and a pixel with maximal values for each
+would be white.}\label{fig:trivariate}
 }
@@
 \includegraphics{figures/risk_map.png}
-\caption{Extraction of a measure of \emph{Betacoronavirus} emergence
-risk from bat hosts based on the colorimetric space from
-fig.~\ref{fig:trivariate}. The risk is a composite measure of the color
-value and angular distance to the yellow hue, as defined in the methods,
-ranged in the unit space. Based on these analyses, South-Eastern Asia,
-Madagascar, the Middle-East, and Africa below the Sahara desert have the
-highest relative risk of emergence.}\label{fig:risk}
+\caption{Evolutionary potential for zoonotic emergence of bat-origin
+betacoronaviruses. Risk is a composite measure of the color value and
+angular distance to the yellow hue in fig.~\ref{fig:trivariate} (see
+Methods).}\label{fig:risk}
 }

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@34b8a1224cd30f556e77c6ce7ab7ff78d85122b4 🚀

@@
 \includegraphics{figures/combined_biogeo.png}
-\caption{Phylogeographic regions of bats (top) and viruses (bottom)
-based on the joint analysis of their occurrence and phylogenetic
-relatedness. The different colors show tendencies to separate alongside
-the first two components of a PCoA. Note that the PCoA for the bats and
-viruses are independent, and so cannot be compared directly -- that
-being said, the regions can be compared across maps.}\label{fig:biogeo}
+\caption{\textbf{Bat and betacoronavirus biogeographic regions.}
+Phylogeography of bats (top) and viruses (bottom) is categorized based
+on analysis of bat distributions paired with bat or virus phylogeny. The
+different colors show tendencies to separate alongside the first two
+components of a PCoA. Note that the PCoA for the bats and viruses are
+independent, and so cannot be compared directly -- that being said, the
+regions can be compared across maps.}\label{fig:biogeo}
 }
@@
 \includegraphics{figures/combined_richness.png}
-\caption{Top panel: relative diversity of known bat hosts of
-betacoronaviruses. This map shows that the region with the largest
-number of possible hosts is South-Eastern Asia. Bottom panel: congruence
-between the evolutionary distinctiveness of the hosts (grey to blue) and
-the viruses (grey to red).}\label{fig:richness}
+\caption{\textbf{Bat and betacoronavirus diversity.} Top panel: relative
+diversity of known bat hosts of betacoronaviruses. This map shows that
+the region with the largest number of possible hosts is South-Eastern
+Asia. Bottom panel: congruence between the evolutionary distinctiveness
+of the hosts (grey to blue) and the viruses (grey to
+red).}\label{fig:richness}
 }
@@
 bat-betacoronavirus system is spatially fragmented enough to create
-differential coevolutionary processes; in turn, the coevolutionary
-mosaic contributes to heterogeneity in emergence risk. These ideas are,
+divergent coevolutionary trajectories; in turn, the coevolutionary
+mosaic underlies the risk of zoonotic emergence. These ideas are,
 respectively, supported by the existence of hotspots of viral uniqueness
@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts---and their diverse immune
-strategies---\textsuperscript{23}creates a landscape of selective
-pressure; the trajectory of viruses' coevolutionary response is, in
-turn, constrained by the opportunities they have for either
-coevolutionary specialization or diversification through host jumps and
-recombination.
+hosts (and their diverse immune strategies)\textsuperscript{23} creates
+a landscape of selective pressure; the trajectory of viruses'
+coevolutionary response is, in turn, constrained by their opportunities
+for either coevolutionary specialization or diversification through host
+jumps and recombination.
 
 Based on the geographic mosaic theory of coevolution, we developed a
-trivariate map of three facets of coevolutionary pressure (see Methods):
-(1) \emph{host phylogenetic diversity}: a high diversity of evolutionary
-histories should expose viruses to more variation in host immune traits;
-(2) \emph{host community uniqueness}: exposure to greater host trait
-heterogeneity can drive viral diversification, and coevolving with more
-unique host communities should create more unique branches of viral
-evolution; and (3) propensity for \emph{viral sharing}: frequent
-cross-species transmission may act as a buffer on selective pressure,
-while lower rates of exchange may enable more simultaneous trajectories
-of viral specialization to coexist within a given community. We combine
-global maps of all three to generate a map of coevolutionary regimes,
-where close colors represent similar risks, and paler pixels represent
-overall higher risk (fig.~\ref{fig:trivariate}). We find that these
-regions do not neatly overlap with those defined in
+trivariate map of three facets of coevolutionary pressure
+(fig.~\ref{fig:trivariate}): (1) \emph{host phylogenetic diversity}: a
+high diversity of evolutionary histories should expose viruses to more
+variation in host immune traits; (2) \emph{host community uniqueness}:
+exposure to greater host trait heterogeneity can drive viral
+diversification, and coevolving with more unique host communities should
+create more unique branches of viral evolution; and (3) propensity for
+\emph{viral sharing}: frequent cross-species transmission may act as a
+buffer on selective pressure, while lower rates of exchange may enable
+more simultaneous trajectories of viral specialization to coexist within
+a given community. We combine global maps of all three to generate a map
+of coevolutionary regimes, where close colors represent similar risks,
+and paler pixels represent overall higher risk (see Methods). We find
+that these regions do not neatly overlap with those defined in
 fig.~\ref{fig:biogeo} or fig.~\ref{fig:richness}, reinforcing the notion
@@
 \includegraphics{figures/risk_trivariate.png}
-\caption{Trivariate additive mapping of the components of risk. Viral
-sharing runs from yellow (low) to blue (high); host phylogenetic
-diversity runs from pink (low) to high (green) channel, and
-compositional uniqueness runs from cyan (low) to red (high). The GMTC
-suggests that the highest evolutionary potential for emergence exists in
-unique and diverse host communities with low viral sharing, \emph{i.e.}
-pixels around yellow. All components are scaled in brightness so that a
-pixel with no sharing, no phylogenetic diversity, and no compositional
+\caption{\textbf{Trivariate additive mapping of the components of risk.}
+Viral sharing runs from yellow (low) to blue (high); host phylogenetic
+diversity runs from pink (low) to high (green); and host compositional
+uniqueness runs from cyan (low) to red (high). The GMTC suggests that
+the highest evolutionary potential for emergence exists in unique and
+diverse host communities with low viral sharing, \emph{i.e.} pixels
+around yellow. All components are scaled in brightness so that a pixel
+with no sharing, no phylogenetic diversity, and no compositional
 uniqueness would be black, and a pixel with maximal values for each
@@
 \includegraphics{figures/risk_map.png}
-\caption{Evolutionary potential for zoonotic emergence of bat-origin
-betacoronaviruses. Risk is a composite measure of the color value and
-angular distance to the yellow hue in fig.~\ref{fig:trivariate} (see
-Methods).}\label{fig:risk}
+\caption{\textbf{Evolutionary potential for zoonotic emergence of
+bat-origin betacoronaviruses.} Risk is a composite measure of the color
+value and angular distance to the yellow hue in
+fig.~\ref{fig:trivariate} (see Methods).}\label{fig:risk}
 }
@@
 \includegraphics{figures/risk_compounded.png}
-\caption{Overlap of the percent of each pixel occupied by urbanized
-structures, representing the degree of settlement, on the spillover risk
-map (where the risk comes only from wildlife, and ignores multi-hosts
-chains of transmissions including non-bats hosts). Darker pixels
-correspond to more risk, in that the GMTC-derived risk of
-fig.~\ref{fig:risk} is high \emph{and} the pixel is densely occupied by
-human populations. This approach increases the relative risk of several
-regions in Africa, and highlights the risk in India, southeast China,
-and the Arabian peninsula where areas of high to moderate risk overlap
-with areas of denser population.}\label{fig:compound}
+\caption{\textbf{Overlap between evolutionary potential and ecological
+opportunity for zoonotic emergence.} Overlap of the percent of each
+pixel occupied by urbanized structures, representing the degree of
+settlement, on the spillover risk map (where the risk comes only from
+wildlife, and ignores multi-hosts chains of transmissions including
+non-bats hosts). Darker pixels correspond to more risk, in that the
+GMTC-derived risk of fig.~\ref{fig:risk} is high \emph{and} the pixel is
+densely occupied by human populations.}\label{fig:compound}
 }
@@
 health.\textsuperscript{38,39} Chiropterans emerged around 64 million
-years ago and are one of the most diverse mammalian orders, with an
-estimated richness of more than 1400 species.\textsuperscript{40,41}
-They exhibit a broad variety of habitat use, behaviour, and feeding
-strategies, putting them at key positions in the delivery and
-provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to human.\textsuperscript{42} For example,
-bats are an essential component of many seed-dispersal
-networks.\textsuperscript{43} Over two-thirds of bats are know to be
+years ago and are one of the most diverse mammalian orders, with more
+than 1,400 estimated species.\textsuperscript{40,41} They exhibit a
+broad variety of habitat use, behaviour, and feeding strategies, putting
+them at key positions in the delivery and provisioning of several
+ecosystem services, tied to important ecosystem-derived benefits to
+humans.\textsuperscript{42} Over two-thirds of bats are know to be
 either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{44,45} and
+contributing for agricultural pest control,\textsuperscript{43,44} and
 vectors of pathogens that put a risk on human
-health.\textsuperscript{46,47} Because bats are globally distributed and
-have a long evolutionary history, phylogeographic and biogeographic
-approaches are required to shed light on the contemporary distribution
-of coevolutionary processes between bats and the pathogens they host.
-Not all areas in which bats, viruses, and human are co-occuring are
-facing a risk of spillover towards human populations, and the areas in
-which this risk exist may not be facing risks of the same nature and
-magnitude.
-
-Here, we propose a simple freamework with broad explanatory power that
+health;\textsuperscript{45,46} other species are essential links in many
+seed-dispersal networks.\textsuperscript{47} Areas where bats, viruses,
+and humans co-occur are not necessarily hotspots of risk for human
+heath; as such, developing more precise ways to map zoonotic hazards can
+help bats and humans coexist safely and support the conservation of
+these important and unique animals.
+
+Here, we propose a simple framework with broad explanatory power that
 helps contextualize discoveries like highly divergent nobecoviruses in
-Madagascar and the previously-neglected adaptive radiation of
-sarbecoviruses outside of southern China and throughout southeast Asia.
-In doing so, it advances ecological theory beyond the current state of
-the art for global maps of emergence risk. For example, previous studies
-that have used host richness as proxy have predicted a high diversity of
-unsampled bat viruses,\textsuperscript{18} bat
-coronaviruses,\textsuperscript{2} and even specifically
-betacoronaviruses\textsuperscript{19} in both the Amazon and southeast
-Asia. While we find that both regions are characterized by highly
-divergent host and viral communities, our framework identifies key
-differences between the regions. We find that Latin America is a hotspot
-of both host and viral distinctiveness, suggesting that this branch of
-the bat-betacoronavirus complex may be undergoing independent
-evolutionary dynamics from the rest of the global pool, but with limited
-potential for viral diversification--- a finding that is supported by
-previous work indicating a higher rate of codivergence in Latin
-America.\textsuperscript{2} In contrast, in southeast Asia, host
-richness and viral distinctiveness are high but sharing is low; this
-suggests a different type of evolutionary dynamics that could generate
-high local diversity of viruses through host switching and viral
-recombination (see \emph{e.g.},\textsuperscript{13} as well as the
-discovery of recombinant viruses that share genetic material from both
-the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{48} Both of these regions are priority areas
-for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{19,20} However,
-both the evolutionary and ecological aspects of emergence risk are
-likely higher in southeast Asia---a fact that will only become more
-relevant, as bats track shifting climates and exchange viruses with
-other species, creating a hotspot of cross-species transmission unique
-to the region.\textsuperscript{49}
+Madagascar and the once-neglected adaptive radiation of sarbecoviruses
+in the Indochinese peninsula. In doing so, it advances ecological theory
+beyond the current state of the art for global maps of emergence risk.
+For example, previous studies that have used host richness as a proxy
+have predicted a high diversity of unsampled bat
+viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
+even specifically betacoronaviruses\textsuperscript{19} in both the
+Amazon and southeast Asia. While we find that both regions are
+characterized by unique and diverse communities of both hosts and
+viruses, our framework is able to identify key differences between the
+two systems. We find that the merbecovirus complex in Latin America has
+been a unique branch of evolution separate from the rest of the global
+pool, but with limited potential for viral diversification--- a finding
+that is supported by previous work indicating a higher rate of
+codivergence in Latin America.\textsuperscript{2} In contrast, in
+southeast Asia, host richness and viral distinctiveness are high but
+sharing is low; this suggests a different type of evolutionary dynamics
+that could generate high local diversity of viruses through host
+switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
+as well as the discovery of recombinant viruses with genetic material
+from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
+lineage).\textsuperscript{48}
+
+Both of these regions are priority areas for sampling, especially given
+predictions that they contain many bat hosts of undiscovered
+betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
+and ecological aspects of emergence risk are higher in southeast
+Asia---a fact that will only become more relevant, as bats track
+shifting climates and exchange viruses with other species, creating a
+hotspot of elevated cross-species transmission unique to the
+region.\textsuperscript{49}
 
@@
 anthropogenic disturbances may feed into co-evolutionary dynamics.
-Increase in humans-hosts contacts also increase the risk of emergence of
-novel diseases,\textsuperscript{52} so does the changes in landscape
+Increase in human-wildlife contact also increase the risk of emergence
+of novel diseases,\textsuperscript{52} as do changes in landscape
 connectivity at local/regional scales.\textsuperscript{53} This
@@
 
+\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
+\CSLLeftMargin{43. }
+\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
+a Changing World}. (Springer International Publishing, 2016).
+doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}
+
+\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
+\CSLLeftMargin{44. }
+\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
+Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
+\textbf{320}, 70--70 (2008).}
+
+\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
+\CSLLeftMargin{45. }
+\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
+Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
+\emph{PLOS ONE} \textbf{8}, e77183 (2013).}
+
+\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
+\CSLLeftMargin{46. }
+\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
+Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
+Research} \textbf{40}, 10--24 (2013).}
+
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \emph{PLOS ONE} \textbf{6}, e17395 (2011).}
-
-\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{44. }
-\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
-a Changing World}. (Springer International Publishing, 2016).
-doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}
-
-\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{45. }
-\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
-Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
-\textbf{320}, 70--70 (2008).}
-
-\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{46. }
-\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
-Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
-\emph{PLOS ONE} \textbf{8}, e77183 (2013).}
-
-\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{47. }
-\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
-Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
-Research} \textbf{40}, 10--24 (2013).}
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f184848f9a50bc554a899660b728c3251b6f4044 🚀

@@
 hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{49}
-
-The diversity and diversification potential of bats responds to
-anthropogenic factors others than shifting climates.\textsuperscript{50}
-Land use changes could significantly decrease bat suitability, notably
-through effects on diet and availability of
-habitats.\textsuperscript{51} As our results establish that the
-diversification of bats betacoronaviruses happens on top of processes
-affecting hosts, biogeographic variation in human population density and
-anthropogenic disturbances may feed into co-evolutionary dynamics.
-Increase in human-wildlife contact also increase the risk of emergence
-of novel diseases,\textsuperscript{52} as do changes in landscape
-connectivity at local/regional scales.\textsuperscript{53} This
-represents a challenge for both conservation strategies and disease
-ecology: some areas can a high emergence risk and more potential for the
-acquisition of zoonotic viruses through bat-human
-encounters.\textsuperscript{54} In particular, the challenge ahead lies
-in the need to quantify actual exposure (and risk) accounting for
-several transmission scenarios, including both direct and indirect bat -
-human interactions, and feeding back into the provision of ecosystem
-services by bats.
+region.\textsuperscript{28,49} Bats---and the spillover of their
+viruses---are also sensitive responds to anthropogenic factors others
+than shifting climates, including deforestation and other kinds of
+habitat loss, increased stress, and greater contact with potential
+bridge hosts like domesticated species.\textsuperscript{26,50,51} This
+represents a challenge for both conservation strategies and pandemic
+prevention,\textsuperscript{52} but identifying areas at risk, and
+protecting the health of bats and ecosystems within those zones, can be
+a win-win intervention for both (add cites for DOIs:
+10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{55} The range maps were subsequently rasterized
+data.\textsuperscript{53} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{54} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{57} based on a recent synthetic tree with
+Diversity\textsuperscript{55} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{58} Faith's PD measures the sum of unique
+species.\textsuperscript{56} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{59} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{57} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{60} transformed
+location, and following best practices\textsuperscript{58} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{61} shows that LCBD
+community. Recent results by\textsuperscript{59} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{62} This network stores pairwise values of
+mammals.\textsuperscript{60} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{63} In this approach, every risk component gets
+model.\textsuperscript{61} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{64} in that this allows multiple
+and low viral sharing,\textsuperscript{62} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{65} we measured the risk level by
+extraction,\textsuperscript{63} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{64} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{67} v1.6.12
-with ModelFinder\textsuperscript{68} ultrafast bootstrap
-approximation\textsuperscript{69} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{65} v1.6.12
+with ModelFinder\textsuperscript{66} ultrafast bootstrap
+approximation\textsuperscript{67} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{70} for each of the viruses in the tree,
+distinctiveness\textsuperscript{68} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{71,72} who
+this idea, we loosely adapted a method from,\textsuperscript{69,70} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{73} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
+\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
 \CSLLeftMargin{52. }
-\CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
-population trends reveal key predictors of virus spillover risk.
-\emph{Proceedings of the Royal Society B: Biological Sciences}
-\textbf{287}, 20192736 (2020).}
-
-\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
-\CSLLeftMargin{53. }
-\CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
-The Importance of Host Phylogeographic Structure in the Spatial Spread
-of Arenaviruses. \emph{PLOS Pathogens} \textbf{13}, e1006073 (2017).}
-
-\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{54. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{53. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{73. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@1212f7fb3388b045df17ec4f4ef36d65913d0183 🚀

@@
 vectors of pathogens that put a risk on human
-health;\textsuperscript{45,46} other species are essential links in many
-seed-dispersal networks.\textsuperscript{47} Areas where bats, viruses,
-and humans co-occur are not necessarily hotspots of risk for human
-heath; as such, developing more precise ways to map zoonotic hazards can
-help bats and humans coexist safely and support the conservation of
-these important and unique animals.
+health;\textsuperscript{45,46} some other species are essential links in
+many seed-dispersal networks.\textsuperscript{47} However, many of these
+species face a high risk of extinction, particularly given persecution
+and killings that sometimes follows from messaging about their role in
+disease emergence. Areas where bats, viruses, and humans co-occur are
+not always hotspots of risk for human heath; as such, developing more
+precise ways to map zoonotic hazards can help bats and humans coexist
+safely, and support the conservation of these important and unique
+animals.
 
@@
 region.\textsuperscript{28,49} Bats---and the spillover of their
-viruses---are also sensitive responds to anthropogenic factors others
-than shifting climates, including deforestation and other kinds of
-habitat loss, increased stress, and greater contact with potential
-bridge hosts like domesticated species.\textsuperscript{26,50,51} This
-represents a challenge for both conservation strategies and pandemic
+viruses---are also sensitive to anthropogenic factors others than
+climate change, including deforestation and other kinds of habitat loss,
+increased stress, and greater contact with potential bridge hosts like
+domesticated species.\textsuperscript{26,50,51} This represents a
+challenge for both conservation strategies and pandemic
 prevention,\textsuperscript{52} but identifying areas at risk, and

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9896714d2bd4a62ccff14475e979edd2152b1a35 🚀

@@
 \vfill
-Driven by the need to understand the ecological factors involved in the
-emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
-COVID-19 in human) through bat hosts, we develop an approach to the
-assesment of spillover risk based on the Geographic Mosaic Theory of
-Coevolution. In doing so, we provide a global mapping of the spillover
-risk posed by betacoronaviruses, reflecting the fact that this risk is
-best understood through the multi-faceted prism of ecological and
-evolutionary mechanisms. Our framework reveals that host richness alone,
-although a component of viral hazard, is not a sufficiently integrative
-predictor of risk. We offer alternative insights based on viral sharing,
-host compositional uniqueness, and host phylogenetic diversity and
-phylogeographic regions. By comparing our aggregated measure of risk to
-a proxy for human density, namely the proportion of each pixel that is
-covered by urban or built land, we provide a synthetic risk map,
-allowing the identification of hotspots where the bat-betacoronavirus
-interaction network may facilitate the emergence of novel viruses and
-their spillover into human populations.
+Pathogen evolution is one of the least predictable components of disease
+emergence, particularly in zoonotic reservoirs. Here, building on
+principles established by the geographic mosaic theory of coevolution,
+we develop a quantitative, spatially-explicit framework for mapping the
+evolutionary risk of viral emergence. Driven by interest in diseases
+like SARS, MERS, and COVID-19, we examine the global biogeography of
+bat-origin betacoronaviruses, and apply our coevolutionary framework to
+develop a map of emergence risk. We find that our framework helps
+explain patterns like a unique pool of merbecoviruses in the Neotropics,
+a recently-discovered lineage of divergent nobecoviruses in Madagascar,
+and---most importantly---hotspots of diversification in southeast Asia,
+sub-Saharan Africa, and the Middle East that correspond to the site of
+previous zoonotic emergence events. Our framework may help identify
+hotspots of future risk that have also been previously overlooked, like
+west Africa and the Indian subcontinent, and may more broadly help
+researchers understand how reservoir ecology shapes the evolution of
+pandemic threats.
 

### 2022-06-28 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@9896714d2bd4a62ccff14475e979edd2152b1a35 🚀

@@
 
-Driven by the need to understand the ecological factors involved in the
-emergence of \emph{Betacoronavirus} (the genus causing SARS, MERS, and
-COVID-19 in human) through bat hosts, we develop an approach to the
-assesment of spillover risk based on the Geographic Mosaic Theory of
-Coevolution. In doing so, we provide a global mapping of the spillover
-risk posed by betacoronaviruses, reflecting the fact that this risk is
-best understood through the multi-faceted prism of ecological and
-evolutionary mechanisms. Our framework reveals that host richness alone,
-although a component of viral hazard, is not a sufficiently integrative
-predictor of risk. We offer alternative insights based on viral sharing,
-host compositional uniqueness, and host phylogenetic diversity and
-phylogeographic regions. By comparing our aggregated measure of risk to
-a proxy for human density, namely the proportion of each pixel that is
-covered by urban or built land, we provide a synthetic risk map,
-allowing the identification of hotspots where the bat-betacoronavirus
-interaction network may facilitate the emergence of novel viruses and
-their spillover into human populations.
+Pathogen evolution is one of the least predictable components of disease
+emergence, particularly in zoonotic reservoirs. Here, building on
+principles established by the geographic mosaic theory of coevolution,
+we develop a quantitative, spatially-explicit framework for mapping the
+evolutionary risk of viral emergence. Driven by interest in diseases
+like SARS, MERS, and COVID-19, we examine the global biogeography of
+bat-origin betacoronaviruses, and apply our coevolutionary framework to
+develop a map of emergence risk. We find that our framework helps
+explain patterns like a unique pool of merbecoviruses in the Neotropics,
+a recently-discovered lineage of divergent nobecoviruses in Madagascar,
+and---most importantly---hotspots of diversification in southeast Asia,
+sub-Saharan Africa, and the Middle East that correspond to the site of
+previous zoonotic emergence events. Our framework may help identify
+hotspots of future risk that have also been previously overlooked, like
+west Africa and the Indian subcontinent, and may more broadly help
+researchers understand how reservoir ecology shapes the evolution of
+pandemic threats.
 

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@c75934d4c3d789d4d02b04bd3adedf4ed3efeaa9 🚀

@@
 
-Bats are important reservoir hosts for different classes of
-microorganisms, many of which a threat to human
-health.\textsuperscript{38,39} Chiropterans emerged around 64 million
-years ago and are one of the most diverse mammalian orders, with more
-than 1,400 estimated species.\textsuperscript{40,41} They exhibit a
-broad variety of habitat use, behaviour, and feeding strategies, putting
-them at key positions in the delivery and provisioning of several
-ecosystem services, tied to important ecosystem-derived benefits to
-humans.\textsuperscript{42} Over two-thirds of bats are know to be
-either obligate or facultative insectivores, therefore actively
-contributing for agricultural pest control,\textsuperscript{43,44} and
-vectors of pathogens that put a risk on human
-health;\textsuperscript{45,46} some other species are essential links in
-many seed-dispersal networks.\textsuperscript{47} However, many of these
-species face a high risk of extinction, particularly given persecution
-and killings that sometimes follows from messaging about their role in
-disease emergence. Areas where bats, viruses, and humans co-occur are
-not always hotspots of risk for human heath; as such, developing more
-precise ways to map zoonotic hazards can help bats and humans coexist
-safely, and support the conservation of these important and unique
-animals.
+Bats emerged around 64 million years ago, and are one of the most
+diverse mammalian orders, with more than 1,400 estimated
+species.\textsuperscript{38,39} They exhibit a broad variety of habitat
+use, behaviour, and feeding strategies, putting them at key positions in
+the delivery and provisioning of several ecosystem services, tied to
+important ecosystem-derived benefits to humans.\textsuperscript{40} Over
+two-thirds of bats are know to be either obligate or facultative
+insectivores, therefore actively contributing for agricultural pest
+control,\textsuperscript{41,42} and vectors of pathogens that put a risk
+on human health;\textsuperscript{43,44} some other species are essential
+links in many seed-dispersal networks.\textsuperscript{45} However, many
+of these species face a high risk of extinction, particularly given
+persecution and killings that sometimes follows from messaging about
+their role in disease emergence. Areas where bats, viruses, and humans
+co-occur are not always hotspots of risk for human heath; as such,
+developing more precise ways to map zoonotic hazards can help bats and
+humans coexist safely, and support the conservation of these important
+and unique animals.
 
@@
 from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{48}
+lineage).\textsuperscript{46}
 
@@
 hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{28,49} Bats---and the spillover of their
+region.\textsuperscript{28,47} Bats---and the spillover of their
 viruses---are also sensitive to anthropogenic factors others than
@@
 increased stress, and greater contact with potential bridge hosts like
-domesticated species.\textsuperscript{26,50,51} This represents a
+domesticated species.\textsuperscript{26,48,49} This represents a
 challenge for both conservation strategies and pandemic
-prevention,\textsuperscript{52} but identifying areas at risk, and
+prevention,\textsuperscript{50} but identifying areas at risk, and
 protecting the health of bats and ecosystems within those zones, can be
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{53} The range maps were subsequently rasterized
+data.\textsuperscript{51} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{54} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{52} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{55} based on a recent synthetic tree with
+Diversity\textsuperscript{53} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{56} Faith's PD measures the sum of unique
+species.\textsuperscript{54} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{57} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{55} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{58} transformed
+location, and following best practices\textsuperscript{56} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{59} shows that LCBD
+community. Recent results by\textsuperscript{57} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{60} This network stores pairwise values of
+mammals.\textsuperscript{58} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{61} In this approach, every risk component gets
+model.\textsuperscript{59} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{62} in that this allows multiple
+and low viral sharing,\textsuperscript{60} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{63} we measured the risk level by
+extraction,\textsuperscript{61} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{64} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{62} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{65} v1.6.12
-with ModelFinder\textsuperscript{66} ultrafast bootstrap
-approximation\textsuperscript{67} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{63} v1.6.12
+with ModelFinder\textsuperscript{64} ultrafast bootstrap
+approximation\textsuperscript{65} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{68} for each of the viruses in the tree,
+distinctiveness\textsuperscript{66} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{69,70} who
+this idea, we loosely adapted a method from,\textsuperscript{67,68} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{69} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Letko2020BatVir}{}%
+\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
 \CSLLeftMargin{38. }
-\CSLRightInline{Letko, M., Seifert, S. N., Olival, K. J., Plowright, R.
-K. \& Munster, V. J. Bat-borne virus diversity, spillover and emergence.
-\emph{Nature Reviews Microbiology} \textbf{18}, 461--471 (2020).}
-
-\leavevmode\hypertarget{ref-VanBrussel2022ZooDis}{}%
-\CSLLeftMargin{39. }
-\CSLRightInline{Van Brussel, K. \& Holmes, E. C. Zoonotic disease and
-virome diversity in bats. \emph{Current Opinion in Virology}
-\textbf{52}, 192--202 (2022).}
-
-\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{40. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{41. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{51. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a82c6af29dc307643df442305a53bea052c2fe17 🚀

@@
 
-\hypertarget{coevolutionary-regimes-structure-evolutionary-risk-of-zoonotic-emergence}{%
-\subsection{Coevolutionary regimes structure evolutionary risk of
+\hypertarget{coevolutionary-regimes-structure-evolutionary-potential-for-zoonotic-emergence}{%
+\subsection{Coevolutionary regimes structure evolutionary potential for
 zoonotic
-emergence}\label{coevolutionary-regimes-structure-evolutionary-risk-of-zoonotic-emergence}}
+emergence}\label{coevolutionary-regimes-structure-evolutionary-potential-for-zoonotic-emergence}}
 
@@
 bat-betacoronavirus system is spatially fragmented enough to create
-divergent coevolutionary trajectories; in turn, the coevolutionary
-mosaic underlies the risk of zoonotic emergence. These ideas are,
+divergent coevolutionary trajectories; in turn, this coevolutionary
+mosaic may alter the risk of zoonotic emergence. These ideas are,
 respectively, supported by the existence of hotspots of viral uniqueness
-and the diverse origins of human betacoronaviruses. Together, these
-ideas point to a predictable relationship between host community
+and the diverse origins of human betacoronaviruses. Together, this
+framework points to a predictable relationship between host community
 structure and coevolutionary pressure: phylogeographic structure in bat

### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f14ca88f84fed5da917f3bce816a15f3c9be159f 🚀

@@
 coevolutionary response is, in turn, constrained by their opportunities
-for either coevolutionary specialization or diversification through host
-jumps and recombination.
+for either specialization or diversification through host jumps and
+recombination.
 

### 2022-06-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b77c93793d62675209a0550d7411ad7373fd06f4 🚀

@@
 increased stress, and greater contact with potential bridge hosts like
-domesticated species.\textsuperscript{26,48,49} This represents a
-challenge for both conservation strategies and pandemic
-prevention,\textsuperscript{50} but identifying areas at risk, and
-protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both (add cites for DOIs:
+domesticated species\textsuperscript{26,48,49} (add cite for
+10.1038/nature13139). This represents a challenge for both conservation
+strategies and pandemic prevention,\textsuperscript{50} but identifying
+areas at risk, and protecting the health of bats and ecosystems within
+those zones, can be a win-win intervention for both (add cites for DOIs:
 10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).

### 2022-06-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@990f852bfb5132c073bedfca4370007dad21b749 🚀

@@
 Pathogen evolution is one of the least predictable components of disease
-emergence, particularly in zoonotic reservoirs. Here, building on
-principles established by the geographic mosaic theory of coevolution,
-we develop a quantitative, spatially-explicit framework for mapping the
-evolutionary risk of viral emergence. Driven by interest in diseases
-like SARS, MERS, and COVID-19, we examine the global biogeography of
-bat-origin betacoronaviruses, and apply our coevolutionary framework to
-develop a map of emergence risk. We find that our framework helps
-explain patterns like a unique pool of merbecoviruses in the Neotropics,
-a recently-discovered lineage of divergent nobecoviruses in Madagascar,
-and---most importantly---hotspots of diversification in southeast Asia,
-sub-Saharan Africa, and the Middle East that correspond to the site of
-previous zoonotic emergence events. Our framework may help identify
-hotspots of future risk that have also been previously overlooked, like
-west Africa and the Indian subcontinent, and may more broadly help
-researchers understand how reservoir ecology shapes the evolution of
+emergence, particularly in nature. Here, building on principles
+established by the geographic mosaic theory of coevolution, we develop a
+quantitative, spatially-explicit framework for mapping the evolutionary
+risk of viral emergence. Driven by interest in diseases like SARS, MERS,
+and COVID-19, we examine the global biogeography of bat-origin
+betacoronaviruses, and find that our coevolutionary framework generates
+different inferences about risk than simply mapping hotspots of host
+diversity. Further, our framework helps explain patterns like a unique
+pool of merbecoviruses in the Neotropics, a recently-discovered lineage
+of divergent nobecoviruses in Madagascar, and--most
+importantly--hotspots of diversification in southeast Asia, sub-Saharan
+Africa, and the Middle East that correspond to the site of previous
+zoonotic emergence events. Our framework may help identify hotspots of
+future risk that have also been previously overlooked, like west Africa
+and the Indian subcontinent, and may more broadly help researchers
+understand how host ecology shapes the evolution and diversity of
 pandemic threats.

### 2022-06-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e931113738a48c3ce976b2dd9c83a52541713dd8 🚀

@@
 and COVID-19, we examine the global biogeography of bat-origin
-betacoronaviruses, and find that our coevolutionary framework generates
-different inferences about risk than simply mapping hotspots of host
-diversity. Further, our framework helps explain patterns like a unique
-pool of merbecoviruses in the Neotropics, a recently-discovered lineage
-of divergent nobecoviruses in Madagascar, and--most
+betacoronaviruses, and find that coevolutionary principles suggest
+geographies of risk that are distinct from the hotspots and coldspots of
+host richness. Further, our framework helps explain patterns like a
+unique pool of merbecoviruses in the Neotropics, a recently-discovered
+lineage of divergent nobecoviruses in Madagascar, and--most
 importantly--hotspots of diversification in southeast Asia, sub-Saharan

### 2022-06-29 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@a8f5b9e565b086cbaef1eee76be468c67b3a2d0e 🚀

@@
 increased stress, and greater contact with potential bridge hosts like
-domesticated species\textsuperscript{26,48,49} (add cite for
-10.1038/nature13139). This represents a challenge for both conservation
-strategies and pandemic prevention,\textsuperscript{50} but identifying
-areas at risk, and protecting the health of bats and ecosystems within
-those zones, can be a win-win intervention for both (add cites for DOIs:
-10.1038/s41893-020-00640-z; 10.1016/S2542-5196(21)00031-0).
+domesticated species.\textsuperscript{26,48--50} This represents a
+challenge for both conservation strategies and pandemic
+prevention,\textsuperscript{51} but identifying areas at risk, and
+protecting the health of bats and ecosystems within those zones, can be
+a win-win intervention for both.\textsuperscript{52,53}
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{51} The range maps were subsequently rasterized
+data.\textsuperscript{54} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{52} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{53} based on a recent synthetic tree with
+Diversity\textsuperscript{56} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{54} Faith's PD measures the sum of unique
+species.\textsuperscript{57} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{55} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{56} transformed
+location, and following best practices\textsuperscript{59} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{57} shows that LCBD
+community. Recent results by\textsuperscript{60} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{58} This network stores pairwise values of
+mammals.\textsuperscript{61} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{59} In this approach, every risk component gets
+model.\textsuperscript{62} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{60} in that this allows multiple
+and low viral sharing,\textsuperscript{63} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{61} we measured the risk level by
+extraction,\textsuperscript{64} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{62} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{65} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{63} v1.6.12
-with ModelFinder\textsuperscript{64} ultrafast bootstrap
-approximation\textsuperscript{65} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{66} v1.6.12
+with ModelFinder\textsuperscript{67} ultrafast bootstrap
+approximation\textsuperscript{68} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{66} for each of the viruses in the tree,
+distinctiveness\textsuperscript{69} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{67,68} who
+this idea, we loosely adapted a method from,\textsuperscript{70,71} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{69} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{72} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
+\CSLLeftMargin{50. }
+\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
+E. A. \& Daily, G. C. Predicting biodiversity change and averting
+collapse in agricultural landscapes. \emph{Nature} \textbf{509},
+213--217 (2014).}
+
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 
+\leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
+\CSLLeftMargin{52. }
+\CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
+interventions that benefit human health and conservation. \emph{Nature
+Sustainability} \textbf{4}, 298--304 (2021).}
+
+\leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
+\CSLLeftMargin{53. }
+\CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
+spillover: A call to action to safeguard environmental, animal, and
+human health. \emph{The Lancet Planetary Health} \textbf{5}, e237--e245
+(2021).}
+
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{54. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-06-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e8860f7ac217c0966ddaa6eafb18f4552a06303b 🚀

@@
 evidence has indicated human exposure to SARS-like
-viruses.\textsuperscript{34} MERS-CoV was originally detected in Saudi
-Arabia, accompanied by a nearly identical virus sequenced from an
-Egyptian tomb bat (\emph{Taphozous perforatus}),\textsuperscript{35} but
+viruses.\textsuperscript{34} MERS-CoV was first detected in Jordan, but
 is widespread in camels in East Africa and the Middle East, and may have
 reached its bridge host decades earlier than originally
-supposed;\textsuperscript{36} as a result, the geography of the original
+supposed;\textsuperscript{35} as a result, the geography of the original
 bat-to-camel transmission is still widely regarded as uncertain. All of
@@
 populations (i.e., those with regular wildlife
-contact)\textsuperscript{37} for maximum impact.
+contact)\textsuperscript{36} for maximum impact.
 
@@
 diverse mammalian orders, with more than 1,400 estimated
-species.\textsuperscript{38,39} They exhibit a broad variety of habitat
+species.\textsuperscript{37,38} They exhibit a broad variety of habitat
 use, behaviour, and feeding strategies, putting them at key positions in
 the delivery and provisioning of several ecosystem services, tied to
-important ecosystem-derived benefits to humans.\textsuperscript{40} Over
+important ecosystem-derived benefits to humans.\textsuperscript{39} Over
 two-thirds of bats are know to be either obligate or facultative
 insectivores, therefore actively contributing for agricultural pest
-control,\textsuperscript{41,42} and vectors of pathogens that put a risk
-on human health;\textsuperscript{43,44} some other species are essential
-links in many seed-dispersal networks.\textsuperscript{45} However, many
+control,\textsuperscript{40,41} and vectors of pathogens that put a risk
+on human health;\textsuperscript{42,43} some other species are essential
+links in many seed-dispersal networks.\textsuperscript{44} However, many
 of these species face a high risk of extinction, particularly given
@@
 from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{46}
+lineage).\textsuperscript{45}
 
@@
 hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{28,47} Bats---and the spillover of their
+region.\textsuperscript{28,46} Bats---and the spillover of their
 viruses---are also sensitive to anthropogenic factors others than
@@
 increased stress, and greater contact with potential bridge hosts like
-domesticated species.\textsuperscript{26,48--50} This represents a
+domesticated species.\textsuperscript{26,47--49} This represents a
 challenge for both conservation strategies and pandemic
-prevention,\textsuperscript{51} but identifying areas at risk, and
+prevention,\textsuperscript{50} but identifying areas at risk, and
 protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both.\textsuperscript{52,53}
+a win-win intervention for both.\textsuperscript{51,52}
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{54} The range maps were subsequently rasterized
+data.\textsuperscript{53} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{54} at a resolution of approximately
 100kmx100km. For every pixel in the resulting raster where at least one
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{56} based on a recent synthetic tree with
+Diversity\textsuperscript{55} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{57} Faith's PD measures the sum of unique
+species.\textsuperscript{56} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{57} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{59} transformed
+location, and following best practices\textsuperscript{58} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{60} shows that LCBD
+community. Recent results by\textsuperscript{59} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{61} This network stores pairwise values of
+mammals.\textsuperscript{60} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{62} In this approach, every risk component gets
+model.\textsuperscript{61} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{63} in that this allows multiple
+and low viral sharing,\textsuperscript{62} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{64} we measured the risk level by
+extraction,\textsuperscript{63} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{65} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{64} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{66} v1.6.12
-with ModelFinder\textsuperscript{67} ultrafast bootstrap
-approximation\textsuperscript{68} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{65} v1.6.12
+with ModelFinder\textsuperscript{66} ultrafast bootstrap
+approximation\textsuperscript{67} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{69} for each of the viruses in the tree,
+distinctiveness\textsuperscript{68} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{70,71} who
+this idea, we loosely adapted a method from,\textsuperscript{69,70} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{72} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
-\leavevmode\hypertarget{ref-Memish2013MidEas}{}%
+\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
 \CSLLeftMargin{35. }
-\CSLRightInline{Memish, Z. A. \emph{et al.} Middle East respiratory
-syndrome coronavirus in bats, Saudi Arabia. \emph{Emerging infectious
-diseases} \textbf{19}, 1819 (2013).}
-
-\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
-\CSLLeftMargin{36. }
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
@@
 \leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{40. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
@@
 \leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{53. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-07-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f142045940be48fceab8947d82b4319eb5f943d7 🚀

@@
 mammal diversity, with a striking Neotropical hotspot (especially in the
-Amazon basin) and a secondary hotspot centered in the southeast Asian
-peninsula. These hotspots of bat diversity are generally presumed to be
-hotspots of viral adaptive radiation, and therefore areas of concern for
-human health.\textsuperscript{2,18} However, the hotspots of known bat
+Amazon basin) and a secondary hotspot centered in Indochina. These
+hotspots of bat diversity are generally presumed to be hotspots of viral
+adaptive radiation, and therefore areas of concern for human
+health.\textsuperscript{2,18} However, the hotspots of known bat
 betacoronavirus hosts show a distinct pattern, with primary hotspots
-(both in terms of size and higher values) of host richness situated in
+(both in terms of area and higher values) of host richness situated in
 southeast Asia, parts of southern Europe, and to a lesser extent parts
@@
 Based on the geographic mosaic theory of coevolution, we developed a
-trivariate map of three facets of coevolutionary pressure
-(fig.~\ref{fig:trivariate}): (1) \emph{host phylogenetic diversity}: a
-high diversity of evolutionary histories should expose viruses to more
-variation in host immune traits; (2) \emph{host community uniqueness}:
-exposure to greater host trait heterogeneity can drive viral
-diversification, and coevolving with more unique host communities should
-create more unique branches of viral evolution; and (3) propensity for
-\emph{viral sharing}: frequent cross-species transmission may act as a
-buffer on selective pressure, while lower rates of exchange may enable
-more simultaneous trajectories of viral specialization to coexist within
-a given community. We combine global maps of all three to generate a map
-of coevolutionary regimes, where close colors represent similar risks,
-and paler pixels represent overall higher risk (see Methods). We find
-that these regions do not neatly overlap with those defined in
-fig.~\ref{fig:biogeo} or fig.~\ref{fig:richness}, reinforcing the notion
-that local-scale coevolutionary mosaics can form within cophylogenetic
-regions.
+trivariate map of coevolutionary pressure (fig.~\ref{fig:trivariate}):
+(1) \emph{host phylogenetic diversity}: a high diversity of evolutionary
+histories should expose viruses to more variation in host immune traits;
+(2) \emph{host community uniqueness}: exposure to greater host trait
+heterogeneity can drive viral diversification, and coevolving with more
+unique host communities should create more unique branches of viral
+evolution; and (3) propensity for \emph{viral sharing}: frequent
+cross-species transmission may act as a buffer on selective pressure,
+while lower rates of exchange may enable more simultaneous trajectories
+of viral specialization to coexist within a given community. We combine
+global maps of all three to generate a map of coevolutionary regimes,
+where close colors represent similar risks, and paler pixels represent
+overall higher risk (see Methods). We find that these regions do not
+neatly overlap with those defined in fig.~\ref{fig:biogeo} or
+fig.~\ref{fig:richness}, reinforcing the notion that local-scale
+coevolutionary mosaics can form within cophylogenetic regions.
 
@@
 diverse host communities with low viral sharing, \emph{i.e.} pixels
-around yellow. All components are scaled in brightness so that a pixel
-with no sharing, no phylogenetic diversity, and no compositional
-uniqueness would be black, and a pixel with maximal values for each
-would be white.}\label{fig:trivariate}
+around yellow. All components within bat host ranges are scaled in
+brightness so that a pixel with no sharing, no phylogenetic diversity,
+and no compositional uniqueness would be black, and a pixel with maximal
+values for each would be white.}\label{fig:trivariate}
 }
@@
 is mediated by both human-wildlife interfaces (the probability of
-spillover) and opportunities for onward transmission (the probability
-that spillovers become epidemics)\textsuperscript{1}. As a proxy for
-both, we finally overlaid the risk component from the composite map (see
-above) with the proportion of built land, as a proxy for a mix of
-habitat disturbance, potential for bat synanthropy or contact with
-bridge hosts like livestock,\textsuperscript{26,27} and human population
-density and connectivity\textsuperscript{1,28,29}
+spillover) and opportunities for onward horizontal transmission (the
+probability that spillovers become epidemics)\textsuperscript{1}. As a
+proxy for both, we finally overlaid the risk component from the
+composite map (see above) with the proportion of built land, as a proxy
+for a mix of habitat disturbance, potential for bat synanthropy or
+contact with bridge hosts like livestock,\textsuperscript{26,27} and
+human population density and connectivity\textsuperscript{1,28,29}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 of the virus is almost certainly in a divergent lineage of
-sarbecoviruses from the Indochinese peninsula that was poorly
-characterized prior to the pandemic.\textsuperscript{30--32} Similarly,
-the SARS-CoV outbreak began in Guangdong province in 2002, reaching
-humans through small carnivore bridge hosts, but was eventually traced
-back to a set of likely progenitor viruses found in cave-dwelling
-horseshoe bats in Yunnan province;\textsuperscript{33} nearby, antibody
-evidence has indicated human exposure to SARS-like
-viruses.\textsuperscript{34} MERS-CoV was first detected in Jordan, but
-is widespread in camels in East Africa and the Middle East, and may have
-reached its bridge host decades earlier than originally
-supposed;\textsuperscript{35} as a result, the geography of the original
-bat-to-camel transmission is still widely regarded as uncertain. All of
-these are broadly consistent with the risk factors we identify. Notably,
-India and west Africa are additional hotspots that have yet to
-experience the emergence of a bat coronavirus into human populations,
-but may still be at risk---particularly given known gaps in bat
-surveillance,\textsuperscript{20} and a dense population in both regions
-with global connectivity. In any of these regions, surveillance on viral
-reservoirs can be paired with targeted monitoring of high-risk human
-populations (i.e., those with regular wildlife
+sarbecoviruses from Indochina that was poorly characterized prior to the
+pandemic.\textsuperscript{30--32} Similarly, the SARS-CoV outbreak began
+in Guangdong province in 2002, reaching humans through small carnivore
+bridge hosts, but was eventually traced back to a set of likely
+progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
+province;\textsuperscript{33} nearby, antibody evidence has indicated
+human exposure to SARS-like viruses.\textsuperscript{34} MERS-CoV was
+first detected in Jordan, but is widespread in camels in East Africa and
+the Middle East, and may have reached its bridge host decades earlier
+than originally supposed;\textsuperscript{35} as a result, the geography
+of the original bat-to-camel transmission is still widely regarded as
+uncertain. All of these are broadly consistent with the risk factors we
+identify. Notably, India and west Africa are additional hotspots that
+have yet to experience the emergence of a bat coronavirus into human
+populations, but may still be at risk---particularly given known gaps in
+bat surveillance,\textsuperscript{20} and a dense population in both
+regions with global connectivity. In any of these regions, surveillance
+on viral reservoirs can be paired with targeted monitoring of high-risk
+human populations (i.e., those with regular wildlife
 contact)\textsuperscript{36} for maximum impact.
@@
 protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both.\textsuperscript{51,52}
+a win-win intervention for both
+{[}\textsuperscript{51};;\textsuperscript{52}OHHLEP2022OneHea{]}.
 
@@
 \texttt{GDAL}\textsuperscript{54} at a resolution of approximately
-100kmx100km. For every pixel in the resulting raster where at least one
-bat host of \emph{Betacoronavirus} was present, we extract the species
-pool (list of all known bat hosts), which was used to calculate the
-following risk assessment components: bat phylogenetic diversity, bat
-compositional uniqueness, and predicted viral sharing risk.
+100kmx100km at the equator. For every pixel in the resulting raster
+where at least one bat host of \emph{Betacoronavirus} was present, we
+extract the species pool (list of all known bat hosts), which was used
+to calculate the following risk assessment components: bat phylogenetic
+diversity, bat compositional uniqueness, and predicted viral sharing
+risk.
 

### 2022-07-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ca64bd4dd795d92a9f82a51c6cbb81762e178696 🚀

@@
 protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both
-{[}\textsuperscript{51};;\textsuperscript{52}OHHLEP2022OneHea{]}.
+a win-win intervention for both.\textsuperscript{51--53}
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{53} The range maps were subsequently rasterized
+data.\textsuperscript{54} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{54} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
 100kmx100km at the equator. For every pixel in the resulting raster
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{55} based on a recent synthetic tree with
+Diversity\textsuperscript{56} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{56} Faith's PD measures the sum of unique
+species.\textsuperscript{57} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{57} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{58} transformed
+location, and following best practices\textsuperscript{59} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{59} shows that LCBD
+community. Recent results by\textsuperscript{60} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{60} This network stores pairwise values of
+mammals.\textsuperscript{61} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{61} In this approach, every risk component gets
+model.\textsuperscript{62} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{62} in that this allows multiple
+and low viral sharing,\textsuperscript{63} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{63} we measured the risk level by
+extraction,\textsuperscript{64} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{64} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{65} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{65} v1.6.12
-with ModelFinder\textsuperscript{66} ultrafast bootstrap
-approximation\textsuperscript{67} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{66} v1.6.12
+with ModelFinder\textsuperscript{67} ultrafast bootstrap
+approximation\textsuperscript{68} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{68} for each of the viruses in the tree,
+distinctiveness\textsuperscript{69} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{69,70} who
+this idea, we loosely adapted a method from,\textsuperscript{70,71} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{72} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-OHHLEP2022OneHea}{}%
+\CSLLeftMargin{53. }
+\CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
+sustainable and healthy future. \emph{PLOS Pathogens} \textbf{18},
+e1010537 (2022).}
+
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{57. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2022-07-04 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@146a421bb1ff8a4538f5892e0a570ea851def1bd 🚀

@@
 the places where the fewest bat betacoronavirus sequences have been
-generated (cite2), resulting in a sparser phylogenetic tree, and
-artificially inflating distinctiveness; conversely, disproportionate
-research effort in eastern China\textsuperscript{20} may have led to a
-more complete inventory of the local diversity of coronaviruses, again
-inflating these metrics relative to underlying patterns. Even accounting
-for these potential biases, though, there is obvious heterogeneity in
-betacoronavirus evolutionary distinctiveness that is distinct from
-overall bat diversity.
+generated,\textsuperscript{20--22} resulting in a sparser phylogenetic
+tree, and artificially inflating distinctiveness; conversely,
+disproportionate research effort in eastern China\textsuperscript{23}
+may have led to a more complete inventory of the local diversity of
+coronaviruses, again inflating these metrics relative to underlying
+patterns. Even accounting for these potential biases, though, there is
+obvious heterogeneity in betacoronavirus evolutionary distinctiveness
+that is distinct from overall bat diversity.
 
@@
 been of the greatest interest to researchers\textsuperscript{13} (and so
-have been sampled most intensively).\textsuperscript{20} The hotspots of
+have been sampled most intensively).\textsuperscript{23} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{3,21} While comparatively understudied, New
+ancient.\textsuperscript{3,24} While comparatively understudied, New
 World merbecoviruses have been found in the ghost-faced bats
@@
 the explosive adaptive radiations of the phyllostomids are responsible
-for the hotspot of bat diversity in the Amazon.\textsuperscript{22}
+for the hotspot of bat diversity in the Amazon.\textsuperscript{25}
 Together, these clades of New World bats play host to a distinct regime
@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts (and their diverse immune strategies)\textsuperscript{23} creates
+hosts (and their diverse immune strategies)\textsuperscript{26} creates
 a landscape of selective pressure; the trajectory of viruses'
@@
 species are endemic following evolutionary divergence from sister
-species in both African and Asian continents,\textsuperscript{24} as one
-such hotspot; interestingly, a recent study\textsuperscript{25} reported
+species in both African and Asian continents,\textsuperscript{27} as one
+such hotspot; interestingly, a recent study\textsuperscript{28} reported
 a novel and highly divergent lineage of nobecoviruses from
@@
 for a mix of habitat disturbance, potential for bat synanthropy or
-contact with bridge hosts like livestock,\textsuperscript{26,27} and
-human population density and connectivity\textsuperscript{1,28,29}
+contact with bridge hosts like livestock,\textsuperscript{29,30} and
+human population density and connectivity\textsuperscript{1,31,32}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 sarbecoviruses from Indochina that was poorly characterized prior to the
-pandemic.\textsuperscript{30--32} Similarly, the SARS-CoV outbreak began
+pandemic.\textsuperscript{20--22} Similarly, the SARS-CoV outbreak began
 in Guangdong province in 2002, reaching humans through small carnivore
@@
 populations, but may still be at risk---particularly given known gaps in
-bat surveillance,\textsuperscript{20} and a dense population in both
+bat surveillance,\textsuperscript{23} and a dense population in both
 regions with global connectivity. In any of these regions, surveillance
@@
 that is supported by previous work indicating a higher rate of
-codivergence in Latin America.\textsuperscript{2} In contrast, in
+codivergence in Latin America.\textsuperscript{2,45} In contrast, in
 southeast Asia, host richness and viral distinctiveness are high but
@@
 from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{45}
+lineage).\textsuperscript{46}
 
@@
 predictions that they contain many bat hosts of undiscovered
-betacoronaviruses.\textsuperscript{19,20} However, both the evolutionary
+betacoronaviruses.\textsuperscript{19,23} However, both the evolutionary
 and ecological aspects of emergence risk are higher in southeast
@@
 hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{28,46} Bats---and the spillover of their
+region.\textsuperscript{31,47} Bats---and the spillover of their
 viruses---are also sensitive to anthropogenic factors others than
@@
 increased stress, and greater contact with potential bridge hosts like
-domesticated species.\textsuperscript{26,47--49} This represents a
+domesticated species.\textsuperscript{29,48--50} This represents a
 challenge for both conservation strategies and pandemic
-prevention,\textsuperscript{50} but identifying areas at risk, and
+prevention,\textsuperscript{51} but identifying areas at risk, and
 protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both.\textsuperscript{51--53}
+a win-win intervention for both.\textsuperscript{52--54}
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{54} The range maps were subsequently rasterized
+data.\textsuperscript{55} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{55} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
 100kmx100km at the equator. For every pixel in the resulting raster
@@
 
+\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
+\CSLLeftMargin{20. }
+\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
+epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
+doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}
+
+\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
+\CSLLeftMargin{21. }
+\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
+SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
+330--336 (2022).}
+
+\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
+\CSLLeftMargin{22. }
+\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
+SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
+\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}
+
 \leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Olival2020PosRev}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
@@
 \leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
@@
 \leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 evolution} \textbf{32}, 55--67 (2017).}
-
-\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{30. }
-\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
-epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
-doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}
-
-\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{31. }
-\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
-SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
-330--336 (2022).}
-
-\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{32. }
-\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
-SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
-\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}
 
@@
 
+\leavevmode\hypertarget{ref-Caraballo2022CroTra}{}%
+\CSLLeftMargin{45. }
+\CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
+Coronaviruses in the Americas: Contrasting Patterns between
+Alphacoronavirus and Betacoronavirus. \emph{Microbiology Spectrum}
+\textbf{0}, e01411--22 (2022).}
+
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
@@
 \leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
@@
 \leavevmode\hypertarget{ref-OHHLEP2022OneHea}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a

### 2022-08-30 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d74d2d8fe84f81e9431ea1c135d8cb274a599339 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e95f073e8afe0ab4deb3916811d8ca7d3cf74bd9 🚀

@@
 %
-Norma\,Forero Rocio Munoz%
+Norma R\,Forero-Muñoz%
 %
@@
 reservoirs.\textsuperscript{1} Although host richness is often used as a
-superficial proxy for spillover risk,\textsuperscript{2,3} these
+superficial proxy for spillover risk,\textsuperscript{2--4} these
 approaches oversimplify the relevant interspecific heterogeneity in
@@
 host pools that allow for the rapid evolution of highly divergent
-viruses.\textsuperscript{4} In the case of generalist pathogens like
+viruses.\textsuperscript{5} In the case of generalist pathogens like
 betacoronaviruses, there is conceptual and empirical support to the idea
 that these community-level mechanisms are even more
-important,\textsuperscript{5} particularly given that cross-species
+important,\textsuperscript{6} particularly given that cross-species
 transmission may, as a rule, structure viral evolution more than
-co-divergence with hosts.\textsuperscript{6} This creates a disconnect
+co-divergence with hosts.\textsuperscript{7} This creates a disconnect
 between coevolutionary theory and most existing ecological frameworks
@@
 explicitly connect microevolutionary dynamics to the macroecology and
-biogeography of symbiotic interactions.\textsuperscript{7} The GMTC
-posits that coevolutionary processes among pairs\textsuperscript{8} or
-complexes\textsuperscript{9} of species are structured in space by the
+biogeography of symbiotic interactions.\textsuperscript{8} The GMTC
+posits that coevolutionary processes among pairs\textsuperscript{9} or
+complexes\textsuperscript{10} of species are structured in space by the
 rippling effects of abiotic conditions onto evolutionary mechanisms,
 giving rise to fragmented systems with different ecologies over large
-spatial extents.\textsuperscript{10} The GMTC predicts a spatial
+spatial extents.\textsuperscript{11} The GMTC predicts a spatial
 fragmentation of coevolutionary dynamics under the joint action of three
-processes:\textsuperscript{11} coevolutionary hot- and coldspots, which
+processes:\textsuperscript{12} coevolutionary hot- and coldspots, which
 appear when the intensity of \emph{interaction} (in terms of reciprocal
@@
 are more biotically complex) and the local favorability of the
-environment;\textsuperscript{12} and trait remixing, which occurs when
+environment;\textsuperscript{13} and trait remixing, which occurs when
 coevolutionary dynamics change when community-level \emph{functional
@@
 significant events in coronavirus evolutionary
-history;\textsuperscript{2,13} as a result, the mosaic of
+history;\textsuperscript{2,14} as a result, the mosaic of
 betacoronavirus phylogeography is assembled from a set of overlapping
@@
 bats (Vespertilionidae), are considered the primary hosts of the
-subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,13}
+subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,14}
 but in the Americas, where merbecoviruses are the only lineage present,
 they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
-Phyllostomidae).\textsuperscript{14--17} At the coarsest scale, these
+Phyllostomidae).\textsuperscript{15--18} At the coarsest scale, these
 heterogeneities are lost, and betacoronavirus biogeography tracks the
@@
 gradients in their species richness generally track broader patterns of
-mammal diversity, with a striking Neotropical hotspot (especially in the
-Amazon basin) and a secondary hotspot centered in Indochina. These
-hotspots of bat diversity are generally presumed to be hotspots of viral
-adaptive radiation, and therefore areas of concern for human
-health.\textsuperscript{2,18} However, the hotspots of known bat
-betacoronavirus hosts show a distinct pattern, with primary hotspots
-(both in terms of area and higher values) of host richness situated in
-southeast Asia, parts of southern Europe, and to a lesser extent parts
-of Africa in the -25-0 range of latitudes (fig.~\ref{fig:richness};
-top). Although hundreds of species likely host undiscovered
-betacoronaviruses, machine learning predictions have suggested that
-these undiscovered reservoirs should follow the same diversity
-gradient.\textsuperscript{19} In principle, these hotspots of
+mammal diversity,\textsuperscript{\textbf{Tanalgo2022Mapping?}} with a
+striking Neotropical hotspot (especially in the Amazon basin) and a
+secondary hotspot centered in Indochina. These hotspots of bat diversity
+are generally presumed to be hotspots of viral adaptive radiation, and
+therefore areas of concern for human health.\textsuperscript{2,19}
+However, the hotspots of known bat betacoronavirus hosts show a distinct
+pattern, with primary hotspots (both in terms of area and higher values)
+of host richness situated in southeast Asia, parts of southern Europe,
+and to a lesser extent parts of Africa in the -25-0 range of latitudes
+(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
+undiscovered betacoronaviruses, machine learning predictions have
+suggested that these undiscovered reservoirs should follow the same
+diversity gradient.\textsuperscript{20} In principle, these hotspots of
 locally-diverse, virus-rich bat communities should drive more adaptive
@@
 the places where the fewest bat betacoronavirus sequences have been
-generated,\textsuperscript{20--22} resulting in a sparser phylogenetic
+generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
 tree, and artificially inflating distinctiveness; conversely,
-disproportionate research effort in eastern China\textsuperscript{23}
+disproportionate research effort in eastern China\textsuperscript{24}
 may have led to a more complete inventory of the local diversity of
@@
 (subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
-been of the greatest interest to researchers\textsuperscript{13} (and so
-have been sampled most intensively).\textsuperscript{23} The hotspots of
+been of the greatest interest to researchers\textsuperscript{14} (and so
+have been sampled most intensively).\textsuperscript{24} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 richness---are almost entirely driven by viral adaptive radiation
-through host switching within this clade\textsuperscript{3,19}. In
+through host switching within this clade\textsuperscript{3,20}. In
 contrast, the Neotropical hotspot of viral distinctiveness is driven by
@@
 Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{3,24} While comparatively understudied, New
+ancient.\textsuperscript{3,25} While comparatively understudied, New
 World merbecoviruses have been found in the ghost-faced bats
 (Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
+free-tailed bats (Molossidae).\textsuperscript{15--18} The former two
 groups and a clade of the latter are endemic to the Neotropics, while
 the explosive adaptive radiations of the phyllostomids are responsible
-for the hotspot of bat diversity in the Amazon.\textsuperscript{25}
+for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
 Together, these clades of New World bats play host to a distinct regime
@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts (and their diverse immune strategies)\textsuperscript{26} creates
+hosts (and their diverse immune strategies)\textsuperscript{27} creates
 a landscape of selective pressure; the trajectory of viruses'
@@
 species are endemic following evolutionary divergence from sister
-species in both African and Asian continents,\textsuperscript{27} as one
-such hotspot; interestingly, a recent study\textsuperscript{28} reported
+species in both African and Asian continents,\textsuperscript{28} as one
+such hotspot; interestingly, a recent study\textsuperscript{29} reported
 a novel and highly divergent lineage of nobecoviruses from
@@
 for a mix of habitat disturbance, potential for bat synanthropy or
-contact with bridge hosts like livestock,\textsuperscript{29,30} and
-human population density and connectivity\textsuperscript{1,31,32}
+contact with bridge hosts like livestock,\textsuperscript{30,31} and
+human population density and connectivity\textsuperscript{1,32,33}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 sarbecoviruses from Indochina that was poorly characterized prior to the
-pandemic.\textsuperscript{20--22} Similarly, the SARS-CoV outbreak began
+pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
 in Guangdong province in 2002, reaching humans through small carnivore
@@
 progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province;\textsuperscript{33} nearby, antibody evidence has indicated
-human exposure to SARS-like viruses.\textsuperscript{34} MERS-CoV was
+province;\textsuperscript{34} nearby, antibody evidence has indicated
+human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
 first detected in Jordan, but is widespread in camels in East Africa and
 the Middle East, and may have reached its bridge host decades earlier
-than originally supposed;\textsuperscript{35} as a result, the geography
+than originally supposed;\textsuperscript{36} as a result, the geography
 of the original bat-to-camel transmission is still widely regarded as
@@
 populations, but may still be at risk---particularly given known gaps in
-bat surveillance,\textsuperscript{23} and a dense population in both
+bat surveillance,\textsuperscript{24} and a dense population in both
 regions with global connectivity. In any of these regions, surveillance
@@
 human populations (i.e., those with regular wildlife
-contact)\textsuperscript{36} for maximum impact.
+contact)\textsuperscript{37} for maximum impact.
 
@@
 diverse mammalian orders, with more than 1,400 estimated
-species.\textsuperscript{37,38} They exhibit a broad variety of habitat
+species.\textsuperscript{38,39} They exhibit a broad variety of habitat
 use, behaviour, and feeding strategies, putting them at key positions in
 the delivery and provisioning of several ecosystem services, tied to
-important ecosystem-derived benefits to humans.\textsuperscript{39} Over
+important ecosystem-derived benefits to humans.\textsuperscript{40} Over
 two-thirds of bats are know to be either obligate or facultative
 insectivores, therefore actively contributing for agricultural pest
-control,\textsuperscript{40,41} and vectors of pathogens that put a risk
-on human health;\textsuperscript{42,43} some other species are essential
-links in many seed-dispersal networks.\textsuperscript{44} However, many
+control,\textsuperscript{41,42} and vectors of pathogens that put a risk
+on human health;\textsuperscript{43,44} some other species are essential
+links in many seed-dispersal networks.\textsuperscript{45} However, many
 of these species face a high risk of extinction, particularly given
@@
 have predicted a high diversity of unsampled bat
-viruses,\textsuperscript{18} bat coronaviruses,\textsuperscript{2} and
-even specifically betacoronaviruses\textsuperscript{19} in both the
+viruses,\textsuperscript{19} bat coronaviruses,\textsuperscript{2} and
+even specifically betacoronaviruses\textsuperscript{20} in both the
 Amazon and southeast Asia. While we find that both regions are
@@
 that is supported by previous work indicating a higher rate of
-codivergence in Latin America.\textsuperscript{2,45} In contrast, in
+codivergence in Latin America.\textsuperscript{2,46} In contrast, in
 southeast Asia, host richness and viral distinctiveness are high but
@@
 that could generate high local diversity of viruses through host
-switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
+switching and viral recombination (see \emph{e.g.},\textsuperscript{14}
 as well as the discovery of recombinant viruses with genetic material
 from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{46}
+lineage).\textsuperscript{47}
 
@@
 predictions that they contain many bat hosts of undiscovered
-betacoronaviruses.\textsuperscript{19,23} However, both the evolutionary
+betacoronaviruses.\textsuperscript{20,24} However, both the evolutionary
 and ecological aspects of emergence risk are higher in southeast
@@
 hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{31,47} Bats---and the spillover of their
+region.\textsuperscript{32,48} Bats---and the spillover of their
 viruses---are also sensitive to anthropogenic factors others than
@@
 increased stress, and greater contact with potential bridge hosts like
-domesticated species.\textsuperscript{29,48--50} This represents a
+domesticated species.\textsuperscript{30,49--51} This represents a
 challenge for both conservation strategies and pandemic
-prevention,\textsuperscript{51} but identifying areas at risk, and
+prevention,\textsuperscript{52} but identifying areas at risk, and
 protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both.\textsuperscript{52--54}
+a win-win intervention for both.\textsuperscript{53--55}
 
@@
 \texttt{https://www.viralemergence.org/betacov} on
-Apr.~2022,\textsuperscript{19} and filtered it to ``known'' hosts
+Apr.~2022,\textsuperscript{20} and filtered it to ``known'' hosts
 (established before the emergence of SARS-CoV-2) and ``novel'' hosts
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{55} The range maps were subsequently rasterized
+data.\textsuperscript{56} The range maps were subsequently rasterized
 using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
+\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
 100kmx100km at the equator. For every pixel in the resulting raster
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{62} This network stores pairwise values of
+mammals.\textsuperscript{63} This network stores pairwise values of
 viral community similarity. To project viral sharing values into a
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{64} in that this allows multiple
+and low viral sharing,\textsuperscript{65} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{65} we measured the risk level by
+extraction,\textsuperscript{66} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{67} v1.6.12
-with ModelFinder\textsuperscript{68} ultrafast bootstrap
-approximation\textsuperscript{69} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
+with ModelFinder\textsuperscript{69} ultrafast bootstrap
+approximation\textsuperscript{70} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 
+\leavevmode\hypertarget{ref-Sanchez2022Strategy}{}%
+\CSLLeftMargin{4. }
+\CSLRightInline{Sánchez, C. A. \emph{et al.} A strategy to assess
+spillover risk of bat SARS-related coronaviruses in southeast asia.
+\emph{Nature communications} \textbf{13}, 1--12 (2022).}
+
 \leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
-\CSLLeftMargin{4. }
+\CSLLeftMargin{5. }
 \CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
@@
 \leavevmode\hypertarget{ref-Power2004PatSpi}{}%
-\CSLLeftMargin{5. }
+\CSLLeftMargin{6. }
 \CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
@@
 \leavevmode\hypertarget{ref-Geoghegan2017ComAna}{}%
-\CSLLeftMargin{6. }
+\CSLLeftMargin{7. }
 \CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
@@
 \leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
-\CSLLeftMargin{7. }
+\CSLLeftMargin{8. }
 \CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
@@
 \leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
-\CSLLeftMargin{8. }
+\CSLLeftMargin{9. }
 \CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
@@
 \leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
-\CSLLeftMargin{9. }
+\CSLLeftMargin{10. }
 \CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-Price2002MacThe}{}%
-\CSLLeftMargin{10. }
+\CSLLeftMargin{11. }
 \CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
-\CSLLeftMargin{11. }
+\CSLLeftMargin{12. }
 \CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
@@
 \leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
-\CSLLeftMargin{12. }
+\CSLLeftMargin{13. }
 \CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
@@
 \leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-\CSLLeftMargin{13. }
+\CSLLeftMargin{14. }
 \CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
@@
 \leavevmode\hypertarget{ref-Anthony2013CorBat}{}%
-\CSLLeftMargin{14. }
+\CSLLeftMargin{15. }
 \CSLRightInline{Anthony, S. J. \emph{et al.} Coronaviruses in bats from
@@
 \leavevmode\hypertarget{ref-Goes2013NovBat}{}%
-\CSLLeftMargin{15. }
+\CSLLeftMargin{16. }
 \CSLRightInline{Góes, L. G. B. \emph{et al.} Novel Bat Coronaviruses,
@@
 \leavevmode\hypertarget{ref-Goes2016GenDiv}{}%
-\CSLLeftMargin{16. }
+\CSLLeftMargin{17. }
 \CSLRightInline{Góes, L. G. B. \emph{et al.} Genetic diversity of bats
@@
 \leavevmode\hypertarget{ref-Brandao2008CorDet}{}%
-\CSLLeftMargin{17. }
+\CSLLeftMargin{18. }
 \CSLRightInline{Brandão, P. E. \emph{et al.} A coronavirus detected in
@@
 \leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
@@
 \leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
@@
 \leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
@@
 \leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Olival2020PosRev}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
@@
 \leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
@@
 \leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
@@
 \leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
@@
 \leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{40. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{41. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Caraballo2022CroTra}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
@@
 \leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
@@
 \leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
@@
 \leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
@@
 \leavevmode\hypertarget{ref-OHHLEP2022OneHea}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{55. }
 \CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
@@
 \leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{56. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d9e5089ca1aeb57d00b06ed661d22af1941cf001 🚀

@@
 contact, but also by the underlying evolutionary dynamics in viral
-reservoirs.\textsuperscript{1} Although host richness is often used as a
-superficial proxy for spillover risk,\textsuperscript{2--4} these
-approaches oversimplify the relevant interspecific heterogeneity in
-immunology, behavior, and other traits, and therefore overlook unique
+reservoirs.\textsuperscript{\textbf{Plowright2017PatZoo?}} Although host
+richness is often used as a superficial proxy for spillover
+risk,\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Sanchez2022Strategy?}}
+these approaches oversimplify the relevant interspecific heterogeneity
+in immunology, behavior, and other traits, and therefore overlook unique
 host pools that allow for the rapid evolution of highly divergent
-viruses.\textsuperscript{5} In the case of generalist pathogens like
-betacoronaviruses, there is conceptual and empirical support to the idea
-that these community-level mechanisms are even more
-important,\textsuperscript{6} particularly given that cross-species
-transmission may, as a rule, structure viral evolution more than
-co-divergence with hosts.\textsuperscript{7} This creates a disconnect
-between coevolutionary theory and most existing ecological frameworks
-for mapping spillover risk.
+viruses.\textsuperscript{\textbf{Agosta2010HowSpe?}} In the case of
+generalist pathogens like betacoronaviruses, there is conceptual and
+empirical support to the idea that these community-level mechanisms are
+even more important,\textsuperscript{\textbf{Power2004PatSpi?}}
+particularly given that cross-species transmission may, as a rule,
+structure viral evolution more than co-divergence with
+hosts.\textsuperscript{\textbf{Geoghegan2017ComAna?}} This creates a
+disconnect between coevolutionary theory and most existing ecological
+frameworks for mapping spillover risk.
 
@@
 explicitly connect microevolutionary dynamics to the macroecology and
-biogeography of symbiotic interactions.\textsuperscript{8} The GMTC
-posits that coevolutionary processes among pairs\textsuperscript{9} or
-complexes\textsuperscript{10} of species are structured in space by the
-rippling effects of abiotic conditions onto evolutionary mechanisms,
-giving rise to fragmented systems with different ecologies over large
-spatial extents.\textsuperscript{11} The GMTC predicts a spatial
-fragmentation of coevolutionary dynamics under the joint action of three
-processes:\textsuperscript{12} coevolutionary hot- and coldspots, which
-appear when the intensity of \emph{interaction} (in terms of reciprocal
-fitness consequences) varies spatially; selection mosaics, wherein the
-intensity of \emph{selection} varies across space, driven by both the
-biotic complexity of the community (locally diverse hosts and viruses
-are more biotically complex) and the local favorability of the
-environment;\textsuperscript{13} and trait remixing, which occurs when
-coevolutionary dynamics change when community-level \emph{functional
-traits} change through meta-community dynamics.
+biogeography of symbiotic
+interactions.\textsuperscript{\textbf{Thompson2005GeoMos?}} The GMTC
+posits that coevolutionary processes among
+pairs\textsuperscript{\textbf{Thompson1994CoePro?}} or
+complexes\textsuperscript{\textbf{Janzen1980WheIt?}} of species are
+structured in space by the rippling effects of abiotic conditions onto
+evolutionary mechanisms, giving rise to fragmented systems with
+different ecologies over large spatial
+extents.\textsuperscript{\textbf{Price2002MacThe?}} The GMTC predicts a
+spatial fragmentation of coevolutionary dynamics under the joint action
+of three processes:\textsuperscript{\textbf{Gomulkiewicz2007DosDon?}}
+coevolutionary hot- and coldspots, which appear when the intensity of
+\emph{interaction} (in terms of reciprocal fitness consequences) varies
+spatially; selection mosaics, wherein the intensity of \emph{selection}
+varies across space, driven by both the biotic complexity of the
+community (locally diverse hosts and viruses are more biotically
+complex) and the local favorability of the
+environment;\textsuperscript{\textbf{Thrall2007CoeSym?}} and trait
+remixing, which occurs when coevolutionary dynamics change when
+community-level \emph{functional traits} change through meta-community
+dynamics.
 
@@
 of host jumps, recombination among disparate lineages, and, to a lesser
-degree, co-divergence with their hosts---\textsuperscript{2}a mix of
-mechanisms that creates a complex and nonlinear relationship between
-host diversity and viral emergence. Working from a recently published
-database of bat hosts of betacoronaviruses, we test whether spatial
-structure in bat-betacoronavirus coevolution is identifiable at a global
-scale. Aiming to explain these patterns, we develop a generalized
-framework for applying the GMTC to host-virus interactions, with a
-specific emphasis on the potential to create independent coevolutionary
-dynamics (and therefore spatial fragmentation in risk) through
-heterogeneity. We develop a trivariate risk assessment system that
-connects each GMTC mechanism to a quantifiable aspect of host-virus
-interactions: (i) viral sharing rates in host communities, representing
-the strength of potential interaction between viruses and any one host
-(i.e., places where viruses undergo constant host switching may be
-coevolutionary coldspots); (ii) the phylogenetic diversity of hosts, as
-a proxy for variation in the immunological mechanisms that antagonize
-viruses (i.e., the selection mosaic); and (iii) the local uniqueness of
-the bat community, representing the potential for viruses to be exposed
-to novel host traits (e.g., variation in receptor sequences). Together,
-we argue that these can be used to identify and map the evolutionary
-drivers that---in conjunction with transmission processes (e.g., viral
+degree, co-divergence with their
+hosts---\textsuperscript{\textbf{Anthony2017GloPat?}}a mix of mechanisms
+that creates a complex and nonlinear relationship between host diversity
+and viral emergence. Working from a recently published database of bat
+hosts of betacoronaviruses, we test whether spatial structure in
+bat-betacoronavirus coevolution is identifiable at a global scale.
+Aiming to explain these patterns, we develop a generalized framework for
+applying the GMTC to host-virus interactions, with a specific emphasis
+on the potential to create independent coevolutionary dynamics (and
+therefore spatial fragmentation in risk) through heterogeneity. We
+develop a trivariate risk assessment system that connects each GMTC
+mechanism to a quantifiable aspect of host-virus interactions: (i) viral
+sharing rates in host communities, representing the strength of
+potential interaction between viruses and any one host (i.e., places
+where viruses undergo constant host switching may be coevolutionary
+coldspots); (ii) the phylogenetic diversity of hosts, as a proxy for
+variation in the immunological mechanisms that antagonize viruses (i.e.,
+the selection mosaic); and (iii) the local uniqueness of the bat
+community, representing the potential for viruses to be exposed to novel
+host traits (e.g., variation in receptor sequences). Together, we argue
+that these can be used to identify and map the evolutionary drivers
+that---in conjunction with transmission processes (e.g., viral
 prevalence in reservoirs and animal-human contact rates)--- determine
@@
 groups of bat hosts are predictive of coronavirus
-diversity.\textsuperscript{2,3} Projecting bat and betacoronavirus
-phylogeny over space (fig.~\ref{fig:biogeo}), we find support for the
-idea that bat community assembly is directly responsible for a global
-mosaic of viral evolution. The distinct groupings (represented by
-different colors, symbolizing positions in a subspace formed by the
-first two phylogenetic principal components) are essentially equivalent
-between the two groups, and can be coarsely delineated as (1) south and
-southeast Asia; (2) east Asia (including northern China), west Asia, and
-the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
-Africa and Latin America. In some cases, this diverges from expectations
-about coronavirus biogeography: for example, previous work has rarely
-flagged India as a region of interest, but for both bats and
-betacoronaviruses, the subcontinent falls into the same regions as the
-southeast Asian peninsula (and indeed, the region is home to known bat
-hosts of multiple betacoronavirus subgenera, including nobecoviruses,
-sarbecoviruses, and merbecoviruses).\textsuperscript{3}
+diversity.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?}}
+Projecting bat and betacoronavirus phylogeny over space
+(fig.~\ref{fig:biogeo}), we find support for the idea that bat community
+assembly is directly responsible for a global mosaic of viral evolution.
+The distinct groupings (represented by different colors, symbolizing
+positions in a subspace formed by the first two phylogenetic principal
+components) are essentially equivalent between the two groups, and can
+be coarsely delineated as (1) south and southeast Asia; (2) east Asia
+(including northern China), west Asia, and the Mediterranean coast; (3)
+Eurasia above a northing of 40; and (4) Africa and Latin America. In
+some cases, this diverges from expectations about coronavirus
+biogeography: for example, previous work has rarely flagged India as a
+region of interest, but for both bats and betacoronaviruses, the
+subcontinent falls into the same regions as the southeast Asian
+peninsula (and indeed, the region is home to known bat hosts of multiple
+betacoronavirus subgenera, including nobecoviruses, sarbecoviruses, and
+merbecoviruses).\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?}}
 
@@
 Overall, these results suggest that the boundaries of bat and
-betacoronavirus biogeographic regions are largely consistent. This may
-be surprising, given that cross-species transmission may play a stronger
-role in coronavirus diversification than
-cospeciation---\textsuperscript{2}a property that would theoretically
-allow for substantial broad divergence in their biogeography. However,
-host jumps at the family level or higher are relatively rare and
-significant events in coronavirus evolutionary
-history;\textsuperscript{2,14} as a result, the mosaic of
-betacoronavirus phylogeography is assembled from a set of overlapping
-smaller coevolutionary systems, superimposed in space and filtered by
-the importance of different subgroups in local host communities. For
-example, the most speciose and cosmopolitan family of bats, the vesper
-bats (Vespertilionidae), are considered the primary hosts of the
-subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,14}
+betacoronavirus biogeographic regions are broadly consistent at a global
+scale; perfect matching between the biogeographic regions would have
+indicated that the signal of virus distribution is fully predicted by
+bat hosts ranges. Areas for which the biogeographic regions for bats and
+betacoronaviruses differ are primarily (i) southeast Asia and southern
+China, and (ii) the Arabian peninsula, which are both regions where
+zoonotic transmission has been documented (potentially driving a unique
+level of viral sampling effort that generates these patterns). These
+spatially limited mismatches nonwithstanding, the large level of
+congruence may be surprising, given that cross-species transmission may
+play a stronger role in coronavirus diversification than
+cospeciation---\textsuperscript{\textbf{Anthony2017GloPat?}}a property
+that would theoretically allow for substantial broad divergence in their
+biogeography. However, host jumps at the family level or higher are
+relatively rare and significant events in coronavirus evolutionary
+history;\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Latinne2020OriCro?}}
+as a result, the mosaic of betacoronavirus phylogeography is assembled
+from a set of overlapping smaller coevolutionary systems, superimposed
+in space and filtered by the importance of different subgroups in local
+host communities. For example, the most speciose and cosmopolitan family
+of bats, the vesper bats (Vespertilionidae), are considered the primary
+hosts of the subgenus \emph{Merbecovirus} (MERS-like
+viruses);\textsuperscript{\textbf{Latinne2020OriCro?},\textbf{Ruiz-Aravena2022EcoEvo?}}
 but in the Americas, where merbecoviruses are the only lineage present,
 they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
-Phyllostomidae).\textsuperscript{15--18} At the coarsest scale, these
-heterogeneities are lost, and betacoronavirus biogeography tracks the
-deep rifts in bat evolutionary history---but within broad regions, the
-component coevolutionary systems may have very different dynamics.
+Phyllostomidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
+At the coarsest scale, these heterogeneities are lost, and
+betacoronavirus biogeography tracks the deep rifts in bat evolutionary
+history---but within broad regions, the component coevolutionary systems
+may have very different dynamics.
 
@@
 are generally presumed to be hotspots of viral adaptive radiation, and
-therefore areas of concern for human health.\textsuperscript{2,19}
+therefore areas of concern for human
+health.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Olival2017HosVir?}}
 However, the hotspots of known bat betacoronavirus hosts show a distinct
@@
 suggested that these undiscovered reservoirs should follow the same
-diversity gradient.\textsuperscript{20} In principle, these hotspots of
-locally-diverse, virus-rich bat communities should drive more adaptive
-diversification in their viruses.
+diversity gradient.\textsuperscript{\textbf{Becker2022OptPre?}} In
+principle, these hotspots of locally-diverse, virus-rich bat communities
+should drive more adaptive diversification in their viruses.
 
@@
 the places where the fewest bat betacoronavirus sequences have been
-generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
-tree, and artificially inflating distinctiveness; conversely,
-disproportionate research effort in eastern China\textsuperscript{24}
-may have led to a more complete inventory of the local diversity of
-coronaviruses, again inflating these metrics relative to underlying
-patterns. Even accounting for these potential biases, though, there is
-obvious heterogeneity in betacoronavirus evolutionary distinctiveness
-that is distinct from overall bat diversity.
+generated,\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
+resulting in a sparser phylogenetic tree, and artificially inflating
+distinctiveness; conversely, disproportionate research effort in eastern
+China\textsuperscript{\textbf{Cohen2022SamStr?}} may have led to a more
+complete inventory of the local diversity of coronaviruses, again
+inflating these metrics relative to underlying patterns. Even accounting
+for these potential biases, though, there is obvious heterogeneity in
+betacoronavirus evolutionary distinctiveness that is distinct from
+overall bat diversity.
 
@@
 (subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
-been of the greatest interest to researchers\textsuperscript{14} (and so
-have been sampled most intensively).\textsuperscript{24} The hotspots of
+been of the greatest interest to
+researchers\textsuperscript{\textbf{Latinne2020OriCro?}} (and so have
+been sampled most
+intensively).\textsuperscript{\textbf{Cohen2022SamStr?}} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 richness---are almost entirely driven by viral adaptive radiation
-through host switching within this clade\textsuperscript{3,20}. In
-contrast, the Neotropical hotspot of viral distinctiveness is driven by
-isolation by host vicariance. Out of the four main groups of
+through host switching within this
+clade\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Ruiz-Aravena2022EcoEvo?}}.
+In contrast, the Neotropical hotspot of viral distinctiveness is driven
+by isolation by host vicariance. Out of the four main groups of
 betacoronaviruses, only merbecoviruses have been found in animals in the
 Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{3,25} While comparatively understudied, New
-World merbecoviruses have been found in the ghost-faced bats
-(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae).\textsuperscript{15--18} The former two
-groups and a clade of the latter are endemic to the Neotropics, while
-the explosive adaptive radiations of the phyllostomids are responsible
-for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
-Together, these clades of New World bats play host to a distinct regime
-of betacoronavirus coevolution.
+ancient.\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Olival2020PosRev?}}
+While comparatively understudied, New World merbecoviruses have been
+found in the ghost-faced bats (Mormoopidae), Neotropical leaf-nosed bats
+(Phyllostomidae), and free-tailed bats
+(Molossidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
+The former two groups and a clade of the latter are endemic to the
+Neotropics, while the explosive adaptive radiations of the phyllostomids
+are responsible for the hotspot of bat diversity in the
+Amazon.\textsuperscript{\textbf{Ammerman2012FirMol?}} Together, these
+clades of New World bats play host to a distinct regime of
+betacoronavirus coevolution.
+
+Our approach is potentially limited by sampling bias: key hotspots
+identified by our model have, indeed, been sampled intensely following
+major zoonotic emergence events. In these areas, more betacoronavirus
+hosts will have been discovered, leading to higher overall diversity and
+potentially higher sharing. Similarly, hotspots of evolutionary
+uniqueness - as in the Arabian peninsula - could reflect much broader
+lineages that have only been sampled in focal areas for public health.
+While the discovery of new branches of bat-betacoronavirus coevolution
+is certainly likely, and might change some of the observed patterns, our
+framework is likely to be fairly robust: the 126 hosts in our study
+capture nearly 10\% of global bat diversity, and the underlying
+evolutionary patterns they represent are much less sensitive to new
+information than any inferences about viral evolution.
 
@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts (and their diverse immune strategies)\textsuperscript{27} creates
-a landscape of selective pressure; the trajectory of viruses'
+hosts (and their diverse immune
+strategies)\textsuperscript{\textbf{Banerjee2020NovIns?}} creates a
+landscape of selective pressure; the trajectory of viruses'
 coevolutionary response is, in turn, constrained by their opportunities
@@
 therefore maximized under higher phylogenetic diversity (viruses are
-exposes to different host clades), higher host uniqueness (viruses are
+exposed to different host clades), higher host uniqueness (viruses are
 experiencing novel, heterogeneous host traits combinations), and low to
@@
 striking correspondence to regions where zoonotic betacoronaviruses have
-previously emerged.
+previously emerged. Our findings align with predictions regarding the
+spatial location of cross-species transmission. These locations not only
+pose a potential risk of viral jumps that could endanger human health
+but also provide valuable information for monitoring wildlife health.
+This could guide us to determine where and what measures to implement
+for effectively monitoring wildlife and human betacoronavirus outbreaks
+before they escalate to critical levels. Nevertheless, there are
+actually very few documented cases of emergence events, and similarities
+could be some degree of coincidental.
 
@@
 species are endemic following evolutionary divergence from sister
-species in both African and Asian continents,\textsuperscript{28} as one
-such hotspot; interestingly, a recent study\textsuperscript{29} reported
-a novel and highly divergent lineage of nobecoviruses from
-Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
-\emph{Rousettus madagascariensis}), again supporting the predictive
-power of the coevolutionary framework.
+species in both African and Asian
+continents,\textsuperscript{\textbf{Shi2014DeeDiv?}} as one such
+hotspot; interestingly, a recent
+study\textsuperscript{\textbf{Kettenburg2022FulGen?}} reported a novel
+and highly divergent lineage of nobecoviruses from Madagascar-endemic
+pteropid bat species (\emph{Pteropus rufus} and \emph{Rousettus
+madagascariensis}), again supporting the predictive power of the
+coevolutionary framework.
 
@@
 spillover) and opportunities for onward horizontal transmission (the
-probability that spillovers become epidemics)\textsuperscript{1}. As a
-proxy for both, we finally overlaid the risk component from the
-composite map (see above) with the proportion of built land, as a proxy
-for a mix of habitat disturbance, potential for bat synanthropy or
-contact with bridge hosts like livestock,\textsuperscript{30,31} and
-human population density and connectivity\textsuperscript{1,32,33}
+probability that spillovers become
+epidemics)\textsuperscript{\textbf{Plowright2017PatZoo?}}. It must be
+noted that the assesment of risk based on the GMTC mechanisms does not
+account for human presence; for this reason, it represents ``potential''
+level of risk, which must be re-evaluated in the light of human
+presence. As a proxy for both, we finally overlaid the risk component
+from the composite map (see above) with the proportion of built land, as
+a proxy for a mix of habitat disturbance, potential for bat synanthropy
+or contact with bridge hosts like
+livestock,\textsuperscript{\textbf{Rulli2021LanCha?},\textbf{Cui2019OriEvo?}}
+and human population density and
+connectivity\textsuperscript{\textbf{Plowright2017PatZoo?},\textbf{Muylaert2022PreFut?},\textbf{Hassell2017UrbDis?}}
 (fig.~\ref{fig:compound}). Accounting for these factors, most of South
@@
 sarbecoviruses from Indochina that was poorly characterized prior to the
-pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
-in Guangdong province in 2002, reaching humans through small carnivore
-bridge hosts, but was eventually traced back to a set of likely
-progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province;\textsuperscript{34} nearby, antibody evidence has indicated
-human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
-first detected in Jordan, but is widespread in camels in East Africa and
-the Middle East, and may have reached its bridge host decades earlier
-than originally supposed;\textsuperscript{36} as a result, the geography
-of the original bat-to-camel transmission is still widely regarded as
-uncertain. All of these are broadly consistent with the risk factors we
-identify. Notably, India and west Africa are additional hotspots that
-have yet to experience the emergence of a bat coronavirus into human
-populations, but may still be at risk---particularly given known gaps in
-bat surveillance,\textsuperscript{24} and a dense population in both
-regions with global connectivity. In any of these regions, surveillance
-on viral reservoirs can be paired with targeted monitoring of high-risk
-human populations (i.e., those with regular wildlife
-contact)\textsuperscript{37} for maximum impact.
+pandemic.\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
+Similarly, the SARS-CoV outbreak began in Guangdong province in 2002,
+reaching humans through small carnivore bridge hosts, but was eventually
+traced back to a set of likely progenitor viruses found in cave-dwelling
+horseshoe bats in Yunnan
+province;\textsuperscript{\textbf{Hu2017DisRic?}} nearby, antibody
+evidence has indicated human exposure to SARS-like
+viruses.\textsuperscript{\textbf{Wang2018SerEvi?}} MERS-CoV was first
+detected in Jordan, but is widespread in camels in East Africa and the
+Middle East, and may have reached its bridge host decades earlier than
+originally supposed;\textsuperscript{\textbf{Muller2014MerCor?}} as a
+result, the geography of the original bat-to-camel transmission is still
+widely regarded as uncertain. All of these are broadly consistent with
+the risk factors we identify. Notably, India and west Africa are
+additional hotspots that have yet to experience the emergence of a bat
+coronavirus into human populations, but may still be at
+risk---particularly given known gaps in bat
+surveillance,\textsuperscript{\textbf{Cohen2022SamStr?}} and a dense
+population in both regions with global connectivity. In any of these
+regions, surveillance on viral reservoirs can be paired with targeted
+monitoring of high-risk human populations (i.e., those with regular
+wildlife contact)\textsuperscript{\textbf{Xu2004EpiClu?}} for maximum
+impact.
 
@@
 diverse mammalian orders, with more than 1,400 estimated
-species.\textsuperscript{38,39} They exhibit a broad variety of habitat
-use, behaviour, and feeding strategies, putting them at key positions in
-the delivery and provisioning of several ecosystem services, tied to
-important ecosystem-derived benefits to humans.\textsuperscript{40} Over
-two-thirds of bats are know to be either obligate or facultative
-insectivores, therefore actively contributing for agricultural pest
-control,\textsuperscript{41,42} and vectors of pathogens that put a risk
-on human health;\textsuperscript{43,44} some other species are essential
-links in many seed-dispersal networks.\textsuperscript{45} However, many
-of these species face a high risk of extinction, particularly given
+species.\textsuperscript{\textbf{Peixoto2018SynEco?},\textbf{Simmons2020BatSpe?}}
+They exhibit a broad variety of habitat use, behaviour, and feeding
+strategies, putting them at key positions in the delivery and
+provisioning of several ecosystem services, tied to important
+ecosystem-derived benefits to
+humans.\textsuperscript{\textbf{Kasso2013EcoEco?}} Over two-thirds of
+bats are known to be either obligate or facultative insectivores,
+therefore actively contributing for agricultural pest
+control,\textsuperscript{\textbf{Voigt2016BatAnt?},\textbf{Williams-Guillen2008BatLim?}}
+and vectors of pathogens that put a risk on human
+health;\textsuperscript{\textbf{Gonsalves2013MosCon?},\textbf{Gonsalves2013MosInf?}}
+some other species are essential links in many seed-dispersal
+networks.\textsuperscript{\textbf{Mello2011MisPar?}} However, many of
+these species face a high risk of extinction, particularly given
 persecution and killings that sometimes follows from messaging about
@@
 have predicted a high diversity of unsampled bat
-viruses,\textsuperscript{19} bat coronaviruses,\textsuperscript{2} and
-even specifically betacoronaviruses\textsuperscript{20} in both the
-Amazon and southeast Asia. While we find that both regions are
+viruses,\textsuperscript{\textbf{Olival2017HosVir?}} bat
+coronaviruses,\textsuperscript{\textbf{Anthony2017GloPat?}} and even
+specifically
+betacoronaviruses\textsuperscript{\textbf{Becker2022OptPre?}} in both
+the Amazon and southeast Asia. While we find that both regions are
 characterized by unique and diverse communities of both hosts and
@@
 that is supported by previous work indicating a higher rate of
-codivergence in Latin America.\textsuperscript{2,46} In contrast, in
-southeast Asia, host richness and viral distinctiveness are high but
-sharing is low; this suggests a different type of evolutionary dynamics
-that could generate high local diversity of viruses through host
-switching and viral recombination (see \emph{e.g.},\textsuperscript{14}
-as well as the discovery of recombinant viruses with genetic material
-from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{47}
-
-Both of these regions are priority areas for sampling, especially given
-predictions that they contain many bat hosts of undiscovered
-betacoronaviruses.\textsuperscript{20,24} However, both the evolutionary
-and ecological aspects of emergence risk are higher in southeast
-Asia---a fact that will only become more relevant, as bats track
-shifting climates and exchange viruses with other species, creating a
-hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{32,48} Bats---and the spillover of their
-viruses---are also sensitive to anthropogenic factors others than
-climate change, including deforestation and other kinds of habitat loss,
-increased stress, and greater contact with potential bridge hosts like
-domesticated species.\textsuperscript{30,49--51} This represents a
-challenge for both conservation strategies and pandemic
-prevention,\textsuperscript{52} but identifying areas at risk, and
-protecting the health of bats and ecosystems within those zones, can be
-a win-win intervention for both.\textsuperscript{53--55}
+codivergence in Latin
+America.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Caraballo2022CroTra?}}
+In contrast, in southeast Asia, host richness and viral distinctiveness
+are high but sharing is low; this suggests a different type of
+evolutionary dynamics that could generate high local diversity of
+viruses through host switching and viral recombination (see
+\emph{e.g.},\textsuperscript{\textbf{Latinne2020OriCro?}} as well as the
+discovery of recombinant viruses with genetic material from both the
+SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
+lineage).\textsuperscript{\textbf{Wu2021ComSur?}}Both of these regions
+are priority areas for sampling, especially given predictions that they
+contain many bat hosts of undiscovered
+betacoronaviruses.\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Cohen2022SamStr?}}
+However, both the evolutionary and ecological aspects of emergence risk
+are higher in southeast Asia---a fact that will only become more
+relevant, as bats track shifting climates and exchange viruses with
+other species, creating a hotspot of elevated cross-species transmission
+unique to the
+region.\textsuperscript{\textbf{Carlson2022CliCha?},\textbf{Muylaert2022PreFut?}}
+
+Our trivariate additive mapping of components of risk
+(fig.~\ref{fig:trivariate}) aims to elicit the complexity of spatial
+cross-species transmission risk beyond the mere presence or absence of
+the pathogen host in a specific location. By considering coevolutionary
+factors such as viral sharing and host uniqueness, we suggest insights
+that can aid in identifying potential locations for surveillance of
+betacoronavirus circulation and assessing the risk of cross-species
+transmission to other mammals. In communities characterized by diverse
+but unique host populations, with limited viral sharing between them, we
+could encounter viruses that specialize in targeting the immune system
+of specific hosts. This implies a low likelihood of infecting novel
+hosts but, once locally introduced into a new host (either a new
+species, or an immunologically naïve population), the specialized virus
+could spread relatively easily due to encountering little immune
+resistance (CITE PLOWRIGHT). With the right combination of viral traits,
+such as low disease-induced mortality or high transmission rate, this
+could lead to successfully spread within the new host community.
+However, while high adaptation to a specific host can be advantageous,
+it may also lead to maladaptation when the pathogen encounters a new
+unsuitable host, potentially resulting in its extinction.
+
+\begin{quote}
+Plowright RK, Foley P, Field HE, Dobson AP, Foley JE, Eby P, Daszak P.
+Urban habituation, ecological connectivity and epidemic dampening: the
+emergence of Hendra virus from flying foxes (Pteropus spp.). Proceedings
+of the Royal Society B: Biological Sciences. 2011 Dec
+22;278(1725):3703-12.
+\end{quote}
+
+Bats---and the spillover of their viruses---are also sensitive to
+anthropogenic factors others than climate change, including
+deforestation and other kinds of habitat loss, increased stress, and
+greater contact with potential bridge hosts like domesticated
+species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
+This represents a challenge for both conservation strategies and
+pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
+identifying areas at risk, and protecting the health of bats and
+ecosystems within those zones, can be a win-win intervention for
+both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
+As we scale these predictions down in space to finer spatial resolutions
+to guide public health actions,\textsuperscript{1} the incorporation of
+human activity predictors will become more
+importyant.\textsuperscript{2}
 
@@
 \texttt{https://www.viralemergence.org/betacov} on
-Apr.~2022,\textsuperscript{20} and filtered it to ``known'' hosts
-(established before the emergence of SARS-CoV-2) and ``novel'' hosts
-(confirmed through sampling and competence assays since the initial data
-collection). The original database was assembled by a combination of
-data mining and literature surveys, including automated alerts on the
-``bats'' and ``coronavirus'' keywords to identify novel empirical
-evidence of bats-betacoronaviruses associations; this yielded a total of
-126 known hosts, 47 of which were novel hosts.
+Apr.~2022,\textsuperscript{\textbf{Becker2022OptPre?}} and filtered it
+to ``known'' hosts (established before the emergence of SARS-CoV-2) and
+``novel'' hosts (confirmed through sampling and competence assays since
+the initial data collection). The original database was assembled by a
+combination of data mining and literature surveys, including automated
+alerts on the ``bats'' and ``coronavirus'' keywords to identify novel
+empirical evidence of bats-betacoronaviruses associations; this yielded
+a total of 126 known hosts, 47 of which were novel hosts. This
+host--virus list of interactions was obtained through a comprehensive
+aggregation of GenBank data as well as systematic literature
+searches,\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Cohen2022SamStr?}}
+such that we have high confidence in its fitness for the purpose of
+inference at a large spatial scale.
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{56} The range maps were subsequently rasterized
-using the \texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
-100kmx100km at the equator. For every pixel in the resulting raster
-where at least one bat host of \emph{Betacoronavirus} was present, we
-extract the species pool (list of all known bat hosts), which was used
-to calculate the following risk assessment components: bat phylogenetic
-diversity, bat compositional uniqueness, and predicted viral sharing
-risk.
+data.\textsuperscript{\textbf{IUCN2021IucRed?}} The IUCN data have been
+assembled to support wildlife conservation efforts, and therefore we do
+not expect that they are biased by wildlife disease sampling efforts or
+priority. The range maps were subsequently rasterized using the
+\texttt{rasterize} function from
+\texttt{GDAL}\textsuperscript{\textbf{RouaultEven2022GdaOgr?}} at a
+resolution of approximately 100kmx100km at the equator. For every pixel
+in the resulting raster where at least one bat host of
+\emph{Betacoronavirus} was present, we extract the species pool (list of
+all known bat hosts), which was used to calculate the following risk
+assessment components: bat phylogenetic diversity, bat compositional
+uniqueness, and predicted viral sharing risk.
 
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{58} based on a recent synthetic tree with
-robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{59} Faith's PD measures the sum of unique
-branches from an arbitrary root to a set of tips, and comparatively
-larger values indicate a more phylogenetic diverse species pool. We
-measured phylogenetic diversity starting from the root of the entire
-tree (and not from Chiroptera); this bears no consequences on the
-resulting values, since all branches leading up to Chiroptera are only
-counted one per species pool, and (as we explain when describing the
-assembly of the composite risk map), all individual risk components are
-ranged in {[}0,1{]}. This measure incorporates a richness component,
-which we chose not to correct for; the interpretation of the
-phylogenetic diversity is therefore a weighted species richness, that
-accounts for phylogenetic over/under-dispersal in some places.
+Diversity\textsuperscript{\textbf{Faith1992ConEva?}} based on a recent
+synthetic tree with robust time calibration, covering about 6000
+mammalian species.\textsuperscript{\textbf{Upham2019InfMam?}} Faith's PD
+measures the sum of unique branches from an arbitrary root to a set of
+tips, and comparatively larger values indicate a more phylogenetic
+diverse species pool. We measured phylogenetic diversity starting from
+the root of the entire tree (and not from Chiroptera); this bears no
+consequences on the resulting values, since all branches leading up to
+Chiroptera are only counted once per species pool, and (as we explain
+when describing the assembly of the composite risk map), all individual
+risk components are ranged in {[}0,1{]}. This measure incorporates a
+richness component, which we chose not to correct for; the
+interpretation of the phylogenetic diversity is therefore a weighted
+species richness, that accounts for phylogenetic over/under-dispersal in
+some places.
 
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{60} LCBD works from a species-data
-matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
-and sites are columns, and a value of 1 indicates occurrence. We
-extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{61} transformed
-it using Hellinger's distance to account for unequal bat richness at
+Beta-Diversity;\textsuperscript{\textbf{Legendre2013BetDiv?}} LCBD works
+from a species-data matrix (traditionally noted as \(\mathbf{Y}\)),
+where species are rows and sites are columns, and a value of 1 indicates
+occurrence. We extracted the Y matrix assuming that every pixel
+represents a unique location, and following best
+practices\textsuperscript{\textbf{Legendre2019SpaTem?}} transformed it
+using Hellinger's distance to account for unequal bat richness at
 different pixels. The correction of raw community data is particularly
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{62} shows that LCBD
+community. Recent results
+by\textsuperscript{\textbf{Dansereau2022EvaEco?}} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{63} This network stores pairwise values of
-viral community similarity. To project viral sharing values into a
-single value for every pixel, we averaged the pairwise scores. High
-values of the average sharing propensity means that this specific extant
-bat assemblage is likely to be proficient at exchanging viruses.
+mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
+pairwise values of viral community similarity, measured for all hosts
+(to maintain consistency with teh phylogenetic diversity measure) across
+all viruses; therefore, we consider that it accounts for some overall
+similarity in the way hosts deal with viruses, and not only
+betacoronaviruses. There is empirical evidence that capacity for
+cross-species transmission even between divergent species is generally
+high,\textsuperscript{3} especially for
+beta-coronaviruses.\textsuperscript{4} To project viral sharing values
+into a single value for every pixel, we averaged the pairwise scores.
+High values of the average sharing propensity means that this specific
+extant bat assemblage is likely to be proficient at exchanging viruses.
 
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{64} In this approach, every risk component gets
-assigned a component in the RGB color model (phylogenetic diversity is
-green, compositional uniqueness is red, and viral sharing is blue). In
-order to achieve a valid RGB measure, all components are re-scaled to
-the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
-diversity, and no compositional uniqueness is black, and a pixel with
-maximal values for each is white. This additive model conveys both the
-intensity of the overall risk, but also the nature of the risk as colors
-diverge towards combinations of values for three risk components. Out of
-the possible combinations, the most risky in terms or rapid
-diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{65} in that this allows multiple
-independent host-virus coevolutionary dynamics to take place in the same
-location. In the colorimetric space, this correspond to yellow --
-because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{66} we measured the risk level by
-calculating the angular distance of the hue of each pixel to a reference
-value of 60 (yellow), and weighted this risk level by the value
-component. Specifically, given a pixel with colorimetric coordinates
-\((h,s,v)\), its ranged weighted risk value is
+model.\textsuperscript{\textbf{Seekell2018GeoLak?}} In this approach,
+every risk component gets assigned a component in the RGB color model
+(phylogenetic diversity is green, compositional uniqueness is red, and
+viral sharing is blue). In order to achieve a valid RGB measure, all
+components are re-scaled to the {[}0,1{]} interval, so that a pixel with
+no sharing, no phylogenetic diversity, and no compositional uniqueness
+is black, and a pixel with maximal values for each is white. This
+additive model conveys both the intensity of the overall risk, but also
+the nature of the risk as colors diverge towards combinations of values
+for three risk components. Out of the possible combinations, the most
+risky in terms or rapid diversification and spillover potential is high
+phylogenetic diversity and low viral
+sharing,\textsuperscript{\textbf{Gomulkiewicz2000HotSpo?}} in that this
+allows multiple independent host-virus coevolutionary dynamics to take
+place in the same location. In the colorimetric space, this correspond
+to yellow -- because the HSV space is more amenable to calculations for
+feature extraction,\textsuperscript{\textbf{Keke2010StuSki?}} we
+measured the risk level by calculating the angular distance of the hue
+of each pixel to a reference value of 60 (yellow), and weighted this
+risk level by the value component. Specifically, given a pixel with
+colorimetric coordinates \((h,s,v)\), its ranged weighted risk value is
 
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
-200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
-with ModelFinder\textsuperscript{69} ultrafast bootstrap
-approximation\textsuperscript{70} with a general time reversible model
-with empirical base frequencies and the 5-discrete-rate-category
-FreeRaye model of nucleotide substitution (GTR+F+R5).
+MAFFT\textsuperscript{\textbf{Katoh2013MafMul?}} v1.4.0 (Algorithm
+FFT-NS-2, Scoring matrix 200PAM / k=2, gap open penalty 1.53m offset
+value 0.123) and a maximum likelihood tree reconstructed in
+IQ-TREE\textsuperscript{\textbf{Nguyen2015IqtFas?}} v1.6.12 with
+ModelFinder\textsuperscript{\textbf{Kalyaanamoorthy2017ModFas?}}
+ultrafast bootstrap
+approximation\textsuperscript{\textbf{Hoang2018UfbImp?}} with a general
+time reversible model with empirical base frequencies and the
+5-discrete-rate-category FreeRaye model of nucleotide substitution
+(GTR+F+R5).
 
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{71} for each of the viruses in the tree,
-then averaged these at the bat species level, projected these values
-onto their geographic distributions, and averaged across every bat found
-in a given pixel. As such, this can be thought of as a map of the mean
-evolutionary distinctiveness of the known viral community believed to be
-associated with a particular subset of bats present.
+distinctiveness\textsuperscript{\textbf{Isaac2007MamEdg?}} for each of
+the viruses in the tree, then averaged these at the bat species level,
+projected these values onto their geographic distributions, and averaged
+across every bat found in a given pixel. As such, this can be thought of
+as a map of the mean evolutionary distinctiveness of the known viral
+community believed to be associated with a particular subset of bats
+present.
 
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{72,73} who
-proposed a phylogenetic method for the delineation of animal
+this idea, we loosely adapted a method
+from,\textsuperscript{\textbf{Kreft2007GloPat?},\textbf{Kreft2010FraDel?}}
+who proposed a phylogenetic method for the delineation of animal
 biogeographic regions. In their original method, a distance matrix -
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{74} subsequently,
+cophenetic function in
+\texttt{ape};\textsuperscript{\textbf{Paradis2019ApeEnv?}} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\hypertarget{data-availability-statement}{%
+\subsection{Data availability
+statement}\label{data-availability-statement}}
+
+The code to reproduce these analyses, as well as the data (with the
+exception of the IUCN rangemaps, which must be downloaded from their
+website) are available in the
+\href{https://github.com/viralemergence/betamap}{\texttt{viralemergence/betamap}}
+repository on GitHub.
+
 \newpage
@@
 \begin{CSLReferences}{0}{0}
-\leavevmode\hypertarget{ref-Plowright2017PatZoo}{}%
+\leavevmode\hypertarget{ref-Muylaert2022Present}{}%
 \CSLLeftMargin{1. }
-\CSLRightInline{Plowright, R. K. \emph{et al.} Pathways to zoonotic
-spillover. \emph{Nature Reviews Microbiology} \textbf{15}, 502--510
-(2017).}
-
-\leavevmode\hypertarget{ref-Anthony2017GloPat}{}%
+\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
+distribution of bat hosts of sarbecoviruses: Implications for
+conservation and public health. \emph{Proceedings of the Royal Society
+B: Biological Sciences} \textbf{289}, 20220397 (2022).}
+
+\leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
 \CSLLeftMargin{2. }
-\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
-coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}
-
-\leavevmode\hypertarget{ref-Ruiz-Aravena2022EcoEvo}{}%
+\CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
+re-emerging viral infectious diseases. \emph{Microbes and Infection}
+\textbf{8}, 905--916 (2006).}
+
+\leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
 \CSLLeftMargin{3. }
-\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
-spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
-\textbf{20}, 299--314 (2022).}
-
-\leavevmode\hypertarget{ref-Sanchez2022Strategy}{}%
+\CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
+homogenous among taxonomic orders of mammalian and avian reservoir
+hosts. \emph{Proceedings of the National Academy of Sciences}
+\textbf{117}, 9423--9430 (2020).}
+
+\leavevmode\hypertarget{ref-Latinne2020Origin}{}%
 \CSLLeftMargin{4. }
-\CSLRightInline{Sánchez, C. A. \emph{et al.} A strategy to assess
-spillover risk of bat SARS-related coronaviruses in southeast asia.
-\emph{Nature communications} \textbf{13}, 1--12 (2022).}
-
-\leavevmode\hypertarget{ref-Agosta2010HowSpe}{}%
-\CSLLeftMargin{5. }
-\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
-can be generalists: Resolving the {`parasite paradox'} and implications
-for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
-151--162 (2010).}
-
-\leavevmode\hypertarget{ref-Power2004PatSpi}{}%
-\CSLLeftMargin{6. }
-\CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
-Disease Epidemics. \emph{The American Naturalist} \textbf{164}, S79--S89
-(2004).}
-
-\leavevmode\hypertarget{ref-Geoghegan2017ComAna}{}%
-\CSLLeftMargin{7. }
-\CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
-Comparative analysis estimates the relative frequencies of co-divergence
-and cross-species transmission within viral families. \emph{PLOS
-Pathogens} \textbf{13}, e1006215 (2017).}
-
-\leavevmode\hypertarget{ref-Thompson2005GeoMos}{}%
-\CSLLeftMargin{8. }
-\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
-Coevolution}. (University Of Chicago Press, 2005).}
-
-\leavevmode\hypertarget{ref-Thompson1994CoePro}{}%
-\CSLLeftMargin{9. }
-\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
-(University of Chicago Press, 1994).}
-
-\leavevmode\hypertarget{ref-Janzen1980WheIt}{}%
-\CSLLeftMargin{10. }
-\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
-\textbf{34}, 611--612 (1980).}
-
-\leavevmode\hypertarget{ref-Price2002MacThe}{}%
-\CSLLeftMargin{11. }
-\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
-Macroecological Patterns}. (Cambridge University Press, 2002).
-doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}
-
-\leavevmode\hypertarget{ref-Gomulkiewicz2007DosDon}{}%
-\CSLLeftMargin{12. }
-\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
-the geographic mosaic theory of coevolution. \emph{Heredity}
-\textbf{98}, 249--258 (2007).}
-
-\leavevmode\hypertarget{ref-Thrall2007CoeSym}{}%
-\CSLLeftMargin{13. }
-\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
-J. D. Coevolution of symbiotic mutualists and parasites in a community
-context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
-(2007).}
-
-\leavevmode\hypertarget{ref-Latinne2020OriCro}{}%
-\CSLLeftMargin{14. }
 \CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
@@
 
-\leavevmode\hypertarget{ref-Anthony2013CorBat}{}%
-\CSLLeftMargin{15. }
-\CSLRightInline{Anthony, S. J. \emph{et al.} Coronaviruses in bats from
-Mexico. \emph{The Journal of General Virology} \textbf{94}, 1028--1038
-(2013).}
-
-\leavevmode\hypertarget{ref-Goes2013NovBat}{}%
-\CSLLeftMargin{16. }
-\CSLRightInline{Góes, L. G. B. \emph{et al.} Novel Bat Coronaviruses,
-Brazil and Mexico. \emph{Emerging Infectious Diseases} \textbf{19},
-1711--1713 (2013).}
-
-\leavevmode\hypertarget{ref-Goes2016GenDiv}{}%
-\CSLLeftMargin{17. }
-\CSLRightInline{Góes, L. G. B. \emph{et al.} Genetic diversity of bats
-coronaviruses in the Atlantic Forest hotspot biome, Brazil.
-\emph{Infection, Genetics and Evolution: Journal of Molecular
-Epidemiology and Evolutionary Genetics in Infectious Diseases}
-\textbf{44}, 510--513 (2016).}
-
-\leavevmode\hypertarget{ref-Brandao2008CorDet}{}%
-\CSLLeftMargin{18. }
-\CSLRightInline{Brandão, P. E. \emph{et al.} A coronavirus detected in
-the vampire bat Desmodus rotundus. \emph{Brazilian Journal of Infectious
-Diseases} \textbf{12}, 466--468 (2008).}
-
-\leavevmode\hypertarget{ref-Olival2017HosVir}{}%
-\CSLLeftMargin{19. }
-\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
-predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
-646--650 (2017).}
-
-\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
-\CSLLeftMargin{20. }
-\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
-to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
-Microbe} (2022)
-doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}
-
-\leavevmode\hypertarget{ref-Worobey2022HuaMar}{}%
-\CSLLeftMargin{21. }
-\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
-epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
-doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}
-
-\leavevmode\hypertarget{ref-Temmam2022BatCor}{}%
-\CSLLeftMargin{22. }
-\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
-SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
-330--336 (2022).}
-
-\leavevmode\hypertarget{ref-Boni2020EvoOri}{}%
-\CSLLeftMargin{23. }
-\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
-SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
-\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}
-
-\leavevmode\hypertarget{ref-Cohen2022SamStr}{}%
-\CSLLeftMargin{24. }
-\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
-Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
-bat coronaviruses. 2022.06.15.496296 (2022)
-doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}
-
-\leavevmode\hypertarget{ref-Olival2020PosRev}{}%
-\CSLLeftMargin{25. }
-\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
-zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
-study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}
-
-\leavevmode\hypertarget{ref-Ammerman2012FirMol}{}%
-\CSLLeftMargin{26. }
-\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
-molecular phylogenetic insights into the evolution of free-tailed bats
-in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
-Mammalogy} \textbf{93}, 12--28 (2012).}
-
-\leavevmode\hypertarget{ref-Banerjee2020NovIns}{}%
-\CSLLeftMargin{27. }
-\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
-Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}
-
-\leavevmode\hypertarget{ref-Shi2014DeeDiv}{}%
-\CSLLeftMargin{28. }
-\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
-Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
-Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}
-
-\leavevmode\hypertarget{ref-Kettenburg2022FulGen}{}%
-\CSLLeftMargin{29. }
-\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
-Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
-for This Coronavirus Clade. \emph{Frontiers in Public Health}
-\textbf{10}, (2022).}
-
-\leavevmode\hypertarget{ref-Rulli2021LanCha}{}%
-\CSLLeftMargin{30. }
-\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
-Land-use change and the livestock revolution increase the risk of
-zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
-Food} \textbf{2}, 409--416 (2021).}
-
-\leavevmode\hypertarget{ref-Cui2019OriEvo}{}%
-\CSLLeftMargin{31. }
-\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
-pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
-\textbf{17}, 181--192 (2019).}
-
-\leavevmode\hypertarget{ref-Muylaert2022PreFut}{}%
-\CSLLeftMargin{32. }
-\CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
-distribution of bat hosts of sarbecoviruses: Implications for
-conservation and public health. \emph{Proceedings of the Royal Society
-B: Biological Sciences} \textbf{289}, 20220397 (2022).}
-
-\leavevmode\hypertarget{ref-Hassell2017UrbDis}{}%
-\CSLLeftMargin{33. }
-\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
-Urbanization and disease emergence: Dynamics at the
-wildlife--livestock--human interface. \emph{Trends in ecology \&
-evolution} \textbf{32}, 55--67 (2017).}
-
-\leavevmode\hypertarget{ref-Hu2017DisRic}{}%
-\CSLLeftMargin{34. }
-\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
-bat SARS-related coronaviruses provides new insights into the origin of
-SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}
-
-\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
-\CSLLeftMargin{35. }
-\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
-SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
-Sinica} \textbf{33}, 104--107 (2018).}
-
-\leavevmode\hypertarget{ref-Muller2014MerCor}{}%
-\CSLLeftMargin{36. }
-\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
-neutralizing antibodies in camels, Eastern Africa, 1983--1997.
-\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}
-
-\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
-\CSLLeftMargin{37. }
-\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
-Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
-1030--1037 (2004).}
-
-\leavevmode\hypertarget{ref-Peixoto2018SynEco}{}%
-\CSLLeftMargin{38. }
-\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
-synthesis of ecological and evolutionary determinants of bat diversity
-across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}
-
-\leavevmode\hypertarget{ref-Simmons2020BatSpe}{}%
-\CSLLeftMargin{39. }
-\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
-World: A taxonomic and geographic database. \url{https://batnames.org/}
-(2020).}
-
-\leavevmode\hypertarget{ref-Kasso2013EcoEco}{}%
-\CSLLeftMargin{40. }
-\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
-Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
-\textbf{2013}, e187415 (2013).}
-
-\leavevmode\hypertarget{ref-Voigt2016BatAnt}{}%
-\CSLLeftMargin{41. }
-\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
-a Changing World}. (Springer International Publishing, 2016).
-doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}
-
-\leavevmode\hypertarget{ref-Williams-Guillen2008BatLim}{}%
-\CSLLeftMargin{42. }
-\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
-Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
-\textbf{320}, 70--70 (2008).}
-
-\leavevmode\hypertarget{ref-Gonsalves2013MosCon}{}%
-\CSLLeftMargin{43. }
-\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
-Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
-\emph{PLOS ONE} \textbf{8}, e77183 (2013).}
-
-\leavevmode\hypertarget{ref-Gonsalves2013MosInf}{}%
-\CSLLeftMargin{44. }
-\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
-Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
-Research} \textbf{40}, 10--24 (2013).}
-
-\leavevmode\hypertarget{ref-Mello2011MisPar}{}%
-\CSLLeftMargin{45. }
-\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
-Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
-\emph{PLOS ONE} \textbf{6}, e17395 (2011).}
-
-\leavevmode\hypertarget{ref-Caraballo2022CroTra}{}%
-\CSLLeftMargin{46. }
-\CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
-Coronaviruses in the Americas: Contrasting Patterns between
-Alphacoronavirus and Betacoronavirus. \emph{Microbiology Spectrum}
-\textbf{0}, e01411--22 (2022).}
-
-\leavevmode\hypertarget{ref-Wu2021ComSur}{}%
-\CSLLeftMargin{47. }
-\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
-sarbecoviruses across China for the origin tracing of SARS-CoV and
-SARS-CoV-2}. (2021)
-doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}
-
-\leavevmode\hypertarget{ref-Carlson2022CliCha}{}%
-\CSLLeftMargin{48. }
-\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
-cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
-doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
-
-\leavevmode\hypertarget{ref-Alves2018GeoVar}{}%
-\CSLLeftMargin{49. }
-\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
-Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
-relationship between large-scale environmental determinants and bat
-species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
-(2018).}
-
-\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
-\CSLLeftMargin{50. }
-\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
-effect of local land use and loss of forests on bats and nocturnal
-insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}
-
-\leavevmode\hypertarget{ref-Mendenhall2014PreBio}{}%
-\CSLLeftMargin{51. }
-\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
-E. A. \& Daily, G. C. Predicting biodiversity change and averting
-collapse in agricultural landscapes. \emph{Nature} \textbf{509},
-213--217 (2014).}
-
-\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
-\CSLLeftMargin{52. }
-\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
-of bats in emerging zoonoses: Balancing ecology, conservation and public
-health interest}. (FAO, 2011).}
-
-\leavevmode\hypertarget{ref-Hopkins2021HowIde}{}%
-\CSLLeftMargin{53. }
-\CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
-interventions that benefit human health and conservation. \emph{Nature
-Sustainability} \textbf{4}, 298--304 (2021).}
-
-\leavevmode\hypertarget{ref-Plowright2021LanUse}{}%
-\CSLLeftMargin{54. }
-\CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
-spillover: A call to action to safeguard environmental, animal, and
-human health. \emph{The Lancet Planetary Health} \textbf{5}, e237--e245
-(2021).}
-
-\leavevmode\hypertarget{ref-OHHLEP2022OneHea}{}%
-\CSLLeftMargin{55. }
-\CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
-sustainable and healthy future. \emph{PLOS Pathogens} \textbf{18},
-e1010537 (2022).}
-
-\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
-\CSLLeftMargin{56. }
-\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
-\url{https://www.iucnredlist.org} (2021).}
-
-\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
-\CSLLeftMargin{57. }
-\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
-Abstraction software Library}. (Zenodo, 2022).
-doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}
-
-\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
-\CSLLeftMargin{58. }
-\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
-diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}
-
-\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
-\CSLLeftMargin{59. }
-\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
-mammal tree: Species-level sets of phylogenies for questions in ecology,
-evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
-(2019).}
-
-\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
-\CSLLeftMargin{60. }
-\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
-variance of community data: Dissimilarity coefficients and partitioning.
-\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}
-
-\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
-\CSLLeftMargin{61. }
-\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
-of beta diversity in the Barro Colorado Island forest dynamics plot,
-Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}
-
-\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
-\CSLLeftMargin{62. }
-\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
-ecological uniqueness over broad spatial extents using species
-distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}
-
-\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
-\CSLLeftMargin{63. }
-\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
-Predicting the global mammalian viral sharing network using
-phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}
-
-\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
-\CSLLeftMargin{64. }
-\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
-geography of lake carbon cycling. \emph{Limnology and Oceanography
-Letters} \textbf{3}, 49--56 (2018).}
-
-\leavevmode\hypertarget{ref-Gomulkiewicz2000HotSpo}{}%
-\CSLLeftMargin{65. }
-\CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
-S. L. \& Hochberg, M. E. Hot Spots, Cold Spots, and the Geographic
-Mosaic Theory of Coevolution. \emph{The American Naturalist}
-\textbf{156}, 156--174 (2000).}
-
-\leavevmode\hypertarget{ref-Keke2010StuSki}{}%
-\CSLLeftMargin{66. }
-\CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
-image segmentation used by Fuzzy-c-means arithmetic. in \emph{2010
-Seventh International Conference on Fuzzy Systems and Knowledge
-Discovery} vol. 2 612--615 (2010).}
-
-\leavevmode\hypertarget{ref-Katoh2013MafMul}{}%
-\CSLLeftMargin{67. }
-\CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
-Alignment Software Version 7: Improvements in Performance and Usability.
-\emph{Molecular Biology and Evolution} \textbf{30}, 772--780 (2013).}
-
-\leavevmode\hypertarget{ref-Nguyen2015IqtFas}{}%
-\CSLLeftMargin{68. }
-\CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
-B. Q. IQ-TREE: A Fast and Effective Stochastic Algorithm for Estimating
-Maximum-Likelihood Phylogenies. \emph{Molecular Biology and Evolution}
-\textbf{32}, 268--274 (2015).}
-
-\leavevmode\hypertarget{ref-Kalyaanamoorthy2017ModFas}{}%
-\CSLLeftMargin{69. }
-\CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
-Haeseler, A. \& Jermiin, L. S. ModelFinder: Fast model selection for
-accurate phylogenetic estimates. \emph{Nature Methods} \textbf{14},
-587--589 (2017).}
-
-\leavevmode\hypertarget{ref-Hoang2018UfbImp}{}%
-\CSLLeftMargin{70. }
-\CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
-Q. \& Vinh, L. S. UFBoot2: Improving the Ultrafast Bootstrap
-Approximation. \emph{Molecular Biology and Evolution} \textbf{35},
-518--522 (2018).}
-
-\leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
-\CSLLeftMargin{71. }
-\CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
-\& Baillie, J. E. M. Mammals on the EDGE: Conservation Priorities Based
-on Threat and Phylogeny. \emph{PLOS ONE} \textbf{2}, e296 (2007).}
-
-\leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
-\CSLLeftMargin{72. }
-\CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
-of vascular plant diversity. \emph{Proceedings of the National Academy
-of Sciences} \textbf{104}, 5925--5930 (2007).}
-
-\leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
-\CSLLeftMargin{73. }
-\CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
-biogeographical regions based on species distributions. \emph{Journal of
-Biogeography} \textbf{37}, 2029--2053 (2010).}
-
-\leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
-\CSLLeftMargin{74. }
-\CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for
-modern phylogenetics and evolutionary analyses in R.
-\emph{Bioinformatics} \textbf{35}, 526--528 (2019).}
-
 \end{CSLReferences}

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@514dee0640609a4566841d9b3736fc708e991cad 🚀

@@
 contact, but also by the underlying evolutionary dynamics in viral
-reservoirs.\textsuperscript{\textbf{Plowright2017PatZoo?}} Although host
-richness is often used as a superficial proxy for spillover
+reservoirs.\textsuperscript{1} Although host richness is often used as a
+superficial proxy for spillover
 risk,\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Sanchez2022Strategy?}}
@@
 host pools that allow for the rapid evolution of highly divergent
-viruses.\textsuperscript{\textbf{Agosta2010HowSpe?}} In the case of
-generalist pathogens like betacoronaviruses, there is conceptual and
-empirical support to the idea that these community-level mechanisms are
-even more important,\textsuperscript{\textbf{Power2004PatSpi?}}
-particularly given that cross-species transmission may, as a rule,
-structure viral evolution more than co-divergence with
-hosts.\textsuperscript{\textbf{Geoghegan2017ComAna?}} This creates a
-disconnect between coevolutionary theory and most existing ecological
-frameworks for mapping spillover risk.
+viruses.\textsuperscript{2} In the case of generalist pathogens like
+betacoronaviruses, there is conceptual and empirical support to the idea
+that these community-level mechanisms are even more
+important,\textsuperscript{3} particularly given that cross-species
+transmission may, as a rule, structure viral evolution more than
+co-divergence with hosts.\textsuperscript{4} This creates a disconnect
+between coevolutionary theory and most existing ecological frameworks
+for mapping spillover risk.
 
@@
 Phylogeography of bats (top) and viruses (bottom) is categorized based
-on analysis of bat distributions paired with bat or virus phylogeny. The
-different colors show tendencies to separate alongside the first two
+on an analysis of bat distributions, paired with bat or virus phylogeny.
+The different colors show tendencies to separate alongside the first two
 components of a PCoA. Note that the PCoA for the bats and viruses are
 independent, and so cannot be compared directly -- that being said, the
-regions can be compared across maps.}\label{fig:biogeo}
+fact that different regions cluster in the same way across maps be
+directly compared.}\label{fig:biogeo}
 }
@@
 \includegraphics{figures/combined_richness.png}
-\caption{\textbf{Bat and betacoronavirus diversity.} Top panel: relative
-diversity of known bat hosts of betacoronaviruses. This map shows that
-the region with the largest number of possible hosts is South-Eastern
-Asia. Bottom panel: congruence between the evolutionary distinctiveness
-of the hosts (grey to blue) and the viruses (grey to
-red).}\label{fig:richness}
+\caption{\textbf{Bat and betacoronavirus diversity.} Top panel:
+diversity of known bat hosts of betacoronaviruses in our dataset. This
+map shows that the region with the largest number of possible hosts is
+South-Eastern Asia. Bottom panel: congruence between the
+\emph{evolutionary} distinctiveness of the hosts (grey to blue) and the
+viruses (grey to red). Darker areas have higher combined evolutionary
+distinctiveness for the entire bat-virus system.}\label{fig:richness}
 }
@@
 and no compositional uniqueness would be black, and a pixel with maximal
-values for each would be white.}\label{fig:trivariate}
+values for each would be white. The individual layers that compose this
+figure are given in supplementary material.}\label{fig:trivariate}
 }
@@
 value and angular distance to the yellow hue in
-fig.~\ref{fig:trivariate} (see Methods).}\label{fig:risk}
+fig.~\ref{fig:trivariate} (see Methods). Darker pixels represent areas
+where the co-evolutionary mechanisms are likely to introduce a strong
+risk of emergence.}\label{fig:risk}
 }
@@
 As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{1} the incorporation of
+to guide public health actions,\textsuperscript{5} the incorporation of
 human activity predictors will become more
-importyant.\textsuperscript{2}
+importyant.\textsuperscript{6}
 
@@
 cross-species transmission even between divergent species is generally
-high,\textsuperscript{3} especially for
-beta-coronaviruses.\textsuperscript{4} To project viral sharing values
+high,\textsuperscript{7} especially for
+beta-coronaviruses.\textsuperscript{8} To project viral sharing values
 into a single value for every pixel, we averaged the pairwise scores.
@@
 \begin{CSLReferences}{0}{0}
+\leavevmode\hypertarget{ref-Plowright2017Pathways}{}%
+\CSLLeftMargin{1. }
+\CSLRightInline{Plowright, R. K. \emph{et al.} Pathways to zoonotic
+spillover. \emph{Nature Reviews Microbiology} \textbf{15}, 502--510
+(2017).}
+
+\leavevmode\hypertarget{ref-Agosta2010How}{}%
+\CSLLeftMargin{2. }
+\CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
+can be generalists: Resolving the {`parasite paradox'} and implications
+for emerging infectious disease. \emph{Zoologia (Curitiba)} \textbf{27},
+151--162 (2010).}
+
+\leavevmode\hypertarget{ref-Power2004Pathogen}{}%
+\CSLLeftMargin{3. }
+\CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
+Disease Epidemics. \emph{The American Naturalist} \textbf{164}, S79--S89
+(2004).}
+
+\leavevmode\hypertarget{ref-Geoghegan2017Comparative}{}%
+\CSLLeftMargin{4. }
+\CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
+Comparative analysis estimates the relative frequencies of co-divergence
+and cross-species transmission within viral families. \emph{PLOS
+Pathogens} \textbf{13}, e1006215 (2017).}
+
 \leavevmode\hypertarget{ref-Muylaert2022Present}{}%
-\CSLLeftMargin{1. }
+\CSLLeftMargin{5. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{2. }
+\CSLLeftMargin{6. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{3. }
+\CSLLeftMargin{7. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
@@
 \leavevmode\hypertarget{ref-Latinne2020Origin}{}%
-\CSLLeftMargin{4. }
+\CSLLeftMargin{8. }
 \CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@7deac7d13ef5a3958af7107cc311210c3f33b74b 🚀

@@
 explicitly connect microevolutionary dynamics to the macroecology and
-biogeography of symbiotic
-interactions.\textsuperscript{\textbf{Thompson2005GeoMos?}} The GMTC
-posits that coevolutionary processes among
-pairs\textsuperscript{\textbf{Thompson1994CoePro?}} or
-complexes\textsuperscript{\textbf{Janzen1980WheIt?}} of species are
-structured in space by the rippling effects of abiotic conditions onto
-evolutionary mechanisms, giving rise to fragmented systems with
-different ecologies over large spatial
-extents.\textsuperscript{\textbf{Price2002MacThe?}} The GMTC predicts a
-spatial fragmentation of coevolutionary dynamics under the joint action
-of three processes:\textsuperscript{\textbf{Gomulkiewicz2007DosDon?}}
-coevolutionary hot- and coldspots, which appear when the intensity of
-\emph{interaction} (in terms of reciprocal fitness consequences) varies
-spatially; selection mosaics, wherein the intensity of \emph{selection}
-varies across space, driven by both the biotic complexity of the
-community (locally diverse hosts and viruses are more biotically
-complex) and the local favorability of the
-environment;\textsuperscript{\textbf{Thrall2007CoeSym?}} and trait
-remixing, which occurs when coevolutionary dynamics change when
-community-level \emph{functional traits} change through meta-community
-dynamics.
+biogeography of symbiotic interactions.\textsuperscript{5} The GMTC
+posits that coevolutionary processes among pairs\textsuperscript{6} or
+complexes\textsuperscript{7} of species are structured in space by the
+rippling effects of abiotic conditions onto evolutionary mechanisms,
+giving rise to fragmented systems with different ecologies over large
+spatial extents.\textsuperscript{8} The GMTC predicts a spatial
+fragmentation of coevolutionary dynamics under the joint action of three
+processes:\textsuperscript{9} coevolutionary hot- and coldspots, which
+appear when the intensity of \emph{interaction} (in terms of reciprocal
+fitness consequences) varies spatially; selection mosaics, wherein the
+intensity of \emph{selection} varies across space, driven by both the
+biotic complexity of the community (locally diverse hosts and viruses
+are more biotically complex) and the local favorability of the
+environment;\textsuperscript{10} and trait remixing, which occurs when
+coevolutionary dynamics change when community-level \emph{functional
+traits} change through meta-community dynamics.
 
@@
 of host jumps, recombination among disparate lineages, and, to a lesser
-degree, co-divergence with their
-hosts---\textsuperscript{\textbf{Anthony2017GloPat?}}a mix of mechanisms
-that creates a complex and nonlinear relationship between host diversity
-and viral emergence. Working from a recently published database of bat
-hosts of betacoronaviruses, we test whether spatial structure in
-bat-betacoronavirus coevolution is identifiable at a global scale.
-Aiming to explain these patterns, we develop a generalized framework for
-applying the GMTC to host-virus interactions, with a specific emphasis
-on the potential to create independent coevolutionary dynamics (and
-therefore spatial fragmentation in risk) through heterogeneity. We
-develop a trivariate risk assessment system that connects each GMTC
-mechanism to a quantifiable aspect of host-virus interactions: (i) viral
-sharing rates in host communities, representing the strength of
-potential interaction between viruses and any one host (i.e., places
-where viruses undergo constant host switching may be coevolutionary
-coldspots); (ii) the phylogenetic diversity of hosts, as a proxy for
-variation in the immunological mechanisms that antagonize viruses (i.e.,
-the selection mosaic); and (iii) the local uniqueness of the bat
-community, representing the potential for viruses to be exposed to novel
-host traits (e.g., variation in receptor sequences). Together, we argue
-that these can be used to identify and map the evolutionary drivers
-that---in conjunction with transmission processes (e.g., viral
+degree, co-divergence with their hosts---\textsuperscript{11}a mix of
+mechanisms that creates a complex and nonlinear relationship between
+host diversity and viral emergence. Working from a recently published
+database of bat hosts of betacoronaviruses, we test whether spatial
+structure in bat-betacoronavirus coevolution is identifiable at a global
+scale. Aiming to explain these patterns, we develop a generalized
+framework for applying the GMTC to host-virus interactions, with a
+specific emphasis on the potential to create independent coevolutionary
+dynamics (and therefore spatial fragmentation in risk) through
+heterogeneity. We develop a trivariate risk assessment system that
+connects each GMTC mechanism to a quantifiable aspect of host-virus
+interactions: (i) viral sharing rates in host communities, representing
+the strength of potential interaction between viruses and any one host
+(i.e., places where viruses undergo constant host switching may be
+coevolutionary coldspots); (ii) the phylogenetic diversity of hosts, as
+a proxy for variation in the immunological mechanisms that antagonize
+viruses (i.e., the selection mosaic); and (iii) the local uniqueness of
+the bat community, representing the potential for viruses to be exposed
+to novel host traits (e.g., variation in receptor sequences). Together,
+we argue that these can be used to identify and map the evolutionary
+drivers that---in conjunction with transmission processes (e.g., viral
 prevalence in reservoirs and animal-human contact rates)--- determine
@@
 betacoronavirus subgenera, including nobecoviruses, sarbecoviruses, and
-merbecoviruses).\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?}}
+merbecoviruses).\textsuperscript{12}
 
@@
 play a stronger role in coronavirus diversification than
-cospeciation---\textsuperscript{\textbf{Anthony2017GloPat?}}a property
-that would theoretically allow for substantial broad divergence in their
-biogeography. However, host jumps at the family level or higher are
-relatively rare and significant events in coronavirus evolutionary
-history;\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Latinne2020OriCro?}}
-as a result, the mosaic of betacoronavirus phylogeography is assembled
-from a set of overlapping smaller coevolutionary systems, superimposed
-in space and filtered by the importance of different subgroups in local
-host communities. For example, the most speciose and cosmopolitan family
-of bats, the vesper bats (Vespertilionidae), are considered the primary
-hosts of the subgenus \emph{Merbecovirus} (MERS-like
-viruses);\textsuperscript{\textbf{Latinne2020OriCro?},\textbf{Ruiz-Aravena2022EcoEvo?}}
+cospeciation---\textsuperscript{11}a property that would theoretically
+allow for substantial broad divergence in their biogeography. However,
+host jumps at the family level or higher are relatively rare and
+significant events in coronavirus evolutionary
+history;\textsuperscript{11,13} as a result, the mosaic of
+betacoronavirus phylogeography is assembled from a set of overlapping
+smaller coevolutionary systems, superimposed in space and filtered by
+the importance of different subgroups in local host communities. For
+example, the most speciose and cosmopolitan family of bats, the vesper
+bats (Vespertilionidae), are considered the primary hosts of the
+subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{12,13}
 but in the Americas, where merbecoviruses are the only lineage present,
 they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
-Phyllostomidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
-At the coarsest scale, these heterogeneities are lost, and
-betacoronavirus biogeography tracks the deep rifts in bat evolutionary
-history---but within broad regions, the component coevolutionary systems
-may have very different dynamics.
+Phyllostomidae).\textsuperscript{14--17} At the coarsest scale, these
+heterogeneities are lost, and betacoronavirus biogeography tracks the
+deep rifts in bat evolutionary history---but within broad regions, the
+component coevolutionary systems may have very different dynamics.
 
@@
 As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{5} the incorporation of
+to guide public health actions,\textsuperscript{18} the incorporation of
 human activity predictors will become more
-importyant.\textsuperscript{6}
+importyant.\textsuperscript{19}
 
@@
 cross-species transmission even between divergent species is generally
-high,\textsuperscript{7} especially for
-beta-coronaviruses.\textsuperscript{8} To project viral sharing values
+high,\textsuperscript{20} especially for
+beta-coronaviruses.\textsuperscript{13} To project viral sharing values
 into a single value for every pixel, we averaged the pairwise scores.
@@
 
+\leavevmode\hypertarget{ref-Thompson2005Geographic}{}%
+\CSLLeftMargin{5. }
+\CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
+Coevolution}. (University Of Chicago Press, 2005).}
+
+\leavevmode\hypertarget{ref-Thompson1994Coevolutionary}{}%
+\CSLLeftMargin{6. }
+\CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
+(University of Chicago Press, 1994).}
+
+\leavevmode\hypertarget{ref-Janzen1980When}{}%
+\CSLLeftMargin{7. }
+\CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
+\textbf{34}, 611--612 (1980).}
+
+\leavevmode\hypertarget{ref-Price2002Macroevolutionary}{}%
+\CSLLeftMargin{8. }
+\CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
+Macroecological Patterns}. (Cambridge University Press, 2002).
+doi:\href{https://doi.org/10.1017/CBO9780511615030}{10.1017/CBO9780511615030}.}
+
+\leavevmode\hypertarget{ref-Gomulkiewicz2007Dos}{}%
+\CSLLeftMargin{9. }
+\CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
+the geographic mosaic theory of coevolution. \emph{Heredity}
+\textbf{98}, 249--258 (2007).}
+
+\leavevmode\hypertarget{ref-Thrall2007Coevolution}{}%
+\CSLLeftMargin{10. }
+\CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
+J. D. Coevolution of symbiotic mutualists and parasites in a community
+context. \emph{Trends in Ecology \& Evolution} \textbf{22}, 120--126
+(2007).}
+
+\leavevmode\hypertarget{ref-Anthony2017Global}{}%
+\CSLLeftMargin{11. }
+\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
+coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}
+
+\leavevmode\hypertarget{ref-Ruiz-Aravena2022Ecology}{}%
+\CSLLeftMargin{12. }
+\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
+spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
+\textbf{20}, 299--314 (2022).}
+
+\leavevmode\hypertarget{ref-Latinne2020Origin}{}%
+\CSLLeftMargin{13. }
+\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
+transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
+Server for Biology} 2020.05.31.116061 (2020)
+doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
+
+\leavevmode\hypertarget{ref-Anthony2013Coronaviruses}{}%
+\CSLLeftMargin{14. }
+\CSLRightInline{Anthony, S. J. \emph{et al.} Coronaviruses in bats from
+Mexico. \emph{The Journal of General Virology} \textbf{94}, 1028--1038
+(2013).}
+
+\leavevmode\hypertarget{ref-Goes2013Novel}{}%
+\CSLLeftMargin{15. }
+\CSLRightInline{Góes, L. G. B. \emph{et al.} Novel Bat Coronaviruses,
+Brazil and Mexico. \emph{Emerging Infectious Diseases} \textbf{19},
+1711--1713 (2013).}
+
+\leavevmode\hypertarget{ref-Goes2016Genetic}{}%
+\CSLLeftMargin{16. }
+\CSLRightInline{Góes, L. G. B. \emph{et al.} Genetic diversity of bats
+coronaviruses in the Atlantic Forest hotspot biome, Brazil.
+\emph{Infection, Genetics and Evolution: Journal of Molecular
+Epidemiology and Evolutionary Genetics in Infectious Diseases}
+\textbf{44}, 510--513 (2016).}
+
+\leavevmode\hypertarget{ref-Brandao2008Coronavirus}{}%
+\CSLLeftMargin{17. }
+\CSLRightInline{Brandão, P. E. \emph{et al.} A coronavirus detected in
+the vampire bat Desmodus rotundus. \emph{Brazilian Journal of Infectious
+Diseases} \textbf{12}, 466--468 (2008).}
+
 \leavevmode\hypertarget{ref-Muylaert2022Present}{}%
-\CSLLeftMargin{5. }
+\CSLLeftMargin{18. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{6. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{7. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
@@
 
-\leavevmode\hypertarget{ref-Latinne2020Origin}{}%
-\CSLLeftMargin{8. }
-\CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
-transmission of bat coronaviruses in China. \emph{bioRxiv: The Preprint
-Server for Biology} 2020.05.31.116061 (2020)
-doi:\href{https://doi.org/10.1101/2020.05.31.116061}{10.1101/2020.05.31.116061}.}
-
 \end{CSLReferences}

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@ac86af753d6ea60ef333d44f62aa0dd3fb7860c4 🚀

@@
 gradients in their species richness generally track broader patterns of
-mammal diversity,\textsuperscript{\textbf{Tanalgo2022Mapping?}} with a
-striking Neotropical hotspot (especially in the Amazon basin) and a
-secondary hotspot centered in Indochina. These hotspots of bat diversity
-are generally presumed to be hotspots of viral adaptive radiation, and
-therefore areas of concern for human
-health.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Olival2017HosVir?}}
-However, the hotspots of known bat betacoronavirus hosts show a distinct
-pattern, with primary hotspots (both in terms of area and higher values)
-of host richness situated in southeast Asia, parts of southern Europe,
-and to a lesser extent parts of Africa in the -25-0 range of latitudes
+mammal diversity,\textsuperscript{18} with a striking Neotropical
+hotspot (especially in the Amazon basin) and a secondary hotspot
+centered in Indochina. These hotspots of bat diversity are generally
+presumed to be hotspots of viral adaptive radiation, and therefore areas
+of concern for human health.\textsuperscript{11,19} However, the
+hotspots of known bat betacoronavirus hosts show a distinct pattern,
+with primary hotspots (both in terms of area and higher values) of host
+richness situated in southeast Asia, parts of southern Europe, and to a
+lesser extent parts of Africa in the -25-0 range of latitudes
 (fig.~\ref{fig:richness}; top). Although hundreds of species likely host
@@
 suggested that these undiscovered reservoirs should follow the same
-diversity gradient.\textsuperscript{\textbf{Becker2022OptPre?}} In
-principle, these hotspots of locally-diverse, virus-rich bat communities
-should drive more adaptive diversification in their viruses.
+diversity gradient.\textsuperscript{20} In principle, these hotspots of
+locally-diverse, virus-rich bat communities should drive more adaptive
+diversification in their viruses.
 
@@
 As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{18} the incorporation of
+to guide public health actions,\textsuperscript{21} the incorporation of
 human activity predictors will become more
-importyant.\textsuperscript{19}
+importyant.\textsuperscript{22}
 
@@
 cross-species transmission even between divergent species is generally
-high,\textsuperscript{20} especially for
+high,\textsuperscript{23} especially for
 beta-coronaviruses.\textsuperscript{13} To project viral sharing values
@@
 
+\leavevmode\hypertarget{ref-Tanalgo2022Mapping}{}%
+\CSLLeftMargin{18. }
+\CSLRightInline{Tanalgo, K. C., Oliveira, H. F. M. \& Hughes, A. C.
+Mapping global conservation priorities and habitat vulnerabilities for
+cave-dwelling bats in a changing world. \emph{Science of The Total
+Environment} \textbf{843}, 156909 (2022).}
+
+\leavevmode\hypertarget{ref-Olival2017Host}{}%
+\CSLLeftMargin{19. }
+\CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
+predict zoonotic spillover from mammals. \emph{Nature} \textbf{546},
+646--650 (2017).}
+
+\leavevmode\hypertarget{ref-Becker2022Optimising}{}%
+\CSLLeftMargin{20. }
+\CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
+to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
+Microbe} (2022)
+doi:\href{https://doi.org/10.1016/S2666-5247(21)00245-7}{10.1016/S2666-5247(21)00245-7}.}
+
 \leavevmode\hypertarget{ref-Muylaert2022Present}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@3ef21deae4d7de31ea2b5a3daee2d1c1559c4e45 🚀

@@
 the places where the fewest bat betacoronavirus sequences have been
-generated,\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
-resulting in a sparser phylogenetic tree, and artificially inflating
-distinctiveness; conversely, disproportionate research effort in eastern
-China\textsuperscript{\textbf{Cohen2022SamStr?}} may have led to a more
-complete inventory of the local diversity of coronaviruses, again
-inflating these metrics relative to underlying patterns. Even accounting
-for these potential biases, though, there is obvious heterogeneity in
-betacoronavirus evolutionary distinctiveness that is distinct from
-overall bat diversity.
+generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
+tree, and artificially inflating distinctiveness; conversely,
+disproportionate research effort in eastern China\textsuperscript{24}
+may have led to a more complete inventory of the local diversity of
+coronaviruses, again inflating these metrics relative to underlying
+patterns. Even accounting for these potential biases, though, there is
+obvious heterogeneity in betacoronavirus evolutionary distinctiveness
+that is distinct from overall bat diversity.
 
@@
 been of the greatest interest to
-researchers\textsuperscript{\textbf{Latinne2020OriCro?}} (and so have
-been sampled most
-intensively).\textsuperscript{\textbf{Cohen2022SamStr?}} The hotspots of
+researchers\textsuperscript{\textbf{Latinne2020Origins?}} (and so have
+been sampled most intensively;).\textsuperscript{24} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 richness---are almost entirely driven by viral adaptive radiation
-through host switching within this
-clade\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Ruiz-Aravena2022EcoEvo?}}.
-In contrast, the Neotropical hotspot of viral distinctiveness is driven
-by isolation by host vicariance. Out of the four main groups of
+through host switching within this clade\textsuperscript{12,20}. In
+contrast, the Neotropical hotspot of viral distinctiveness is driven by
+isolation by host vicariance. Out of the four main groups of
 betacoronaviruses, only merbecoviruses have been found in animals in the
 Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Olival2020PosRev?}}
-While comparatively understudied, New World merbecoviruses have been
-found in the ghost-faced bats (Mormoopidae), Neotropical leaf-nosed bats
-(Phyllostomidae), and free-tailed bats
-(Molossidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
-The former two groups and a clade of the latter are endemic to the
-Neotropics, while the explosive adaptive radiations of the phyllostomids
-are responsible for the hotspot of bat diversity in the
-Amazon.\textsuperscript{\textbf{Ammerman2012FirMol?}} Together, these
-clades of New World bats play host to a distinct regime of
-betacoronavirus coevolution.
+ancient.\textsuperscript{12,25} While comparatively understudied, New
+World merbecoviruses have been found in the ghost-faced bats
+(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
+free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
+groups and a clade of the latter are endemic to the Neotropics, while
+the explosive adaptive radiations of the phyllostomids are responsible
+for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
+Together, these clades of New World bats play host to a distinct regime
+of betacoronavirus coevolution.
 
@@
 As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{21} the incorporation of
+to guide public health actions,\textsuperscript{27} the incorporation of
 human activity predictors will become more
-importyant.\textsuperscript{22}
+importyant.\textsuperscript{28}
 
@@
 cross-species transmission even between divergent species is generally
-high,\textsuperscript{23} especially for
+high,\textsuperscript{29} especially for
 beta-coronaviruses.\textsuperscript{13} To project viral sharing values
@@
 
+\leavevmode\hypertarget{ref-Worobey2022Huanan}{}%
+\CSLLeftMargin{21. }
+\CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
+epicenter of SARS-CoV-2 emergence. \emph{Zenodo} (2022)
+doi:\href{https://doi.org/10.5281/zenodo.6299600}{10.5281/zenodo.6299600}.}
+
+\leavevmode\hypertarget{ref-Temmam2022Bat}{}%
+\CSLLeftMargin{22. }
+\CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
+SARS-CoV-2 and infectious for human cells. \emph{Nature} \textbf{604},
+330--336 (2022).}
+
+\leavevmode\hypertarget{ref-Boni2020Evolutionary}{}%
+\CSLLeftMargin{23. }
+\CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
+SARS-CoV-2 sarbecovirus lineage responsible for the COVID-19 pandemic.
+\emph{Nature microbiology} \textbf{5}, 1408--1417 (2020).}
+
+\leavevmode\hypertarget{ref-Cohen2022Sampling}{}%
+\CSLLeftMargin{24. }
+\CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
+Becker, D. J. Sampling strategies and pre-pandemic surveillance gaps for
+bat coronaviruses. 2022.06.15.496296 (2022)
+doi:\href{https://doi.org/10.1101/2022.06.15.496296}{10.1101/2022.06.15.496296}.}
+
+\leavevmode\hypertarget{ref-Olival2020Possibility}{}%
+\CSLLeftMargin{25. }
+\CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
+zoonotic transmission of SARS-CoV-2 to free-ranging wildlife: A case
+study of bats. \emph{PLoS pathogens} \textbf{16}, e1008758 (2020).}
+
+\leavevmode\hypertarget{ref-Ammerman2012First}{}%
+\CSLLeftMargin{26. }
+\CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
+molecular phylogenetic insights into the evolution of free-tailed bats
+in the subfamily Molossinae (Molossidae, Chiroptera). \emph{Journal of
+Mammalogy} \textbf{93}, 12--28 (2012).}
+
 \leavevmode\hypertarget{ref-Muylaert2022Present}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@7c7d7768c1e549485c1655de1f1c4b42888ecbdb 🚀

@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts (and their diverse immune
-strategies)\textsuperscript{\textbf{Banerjee2020NovIns?}} creates a
-landscape of selective pressure; the trajectory of viruses'
+hosts (and their diverse immune strategies)\textsuperscript{27} creates
+a landscape of selective pressure; the trajectory of viruses'
 coevolutionary response is, in turn, constrained by their opportunities
@@
 species are endemic following evolutionary divergence from sister
-species in both African and Asian
-continents,\textsuperscript{\textbf{Shi2014DeeDiv?}} as one such
-hotspot; interestingly, a recent
-study\textsuperscript{\textbf{Kettenburg2022FulGen?}} reported a novel
-and highly divergent lineage of nobecoviruses from Madagascar-endemic
-pteropid bat species (\emph{Pteropus rufus} and \emph{Rousettus
-madagascariensis}), again supporting the predictive power of the
-coevolutionary framework.
+species in both African and Asian continents,\textsuperscript{28} as one
+such hotspot; interestingly, a recent study\textsuperscript{29} reported
+a novel and highly divergent lineage of nobecoviruses from
+Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
+\emph{Rousettus madagascariensis}), again supporting the predictive
+power of the coevolutionary framework.
 
@@
 spillover) and opportunities for onward horizontal transmission (the
-probability that spillovers become
-epidemics)\textsuperscript{\textbf{Plowright2017PatZoo?}}. It must be
-noted that the assesment of risk based on the GMTC mechanisms does not
-account for human presence; for this reason, it represents ``potential''
-level of risk, which must be re-evaluated in the light of human
-presence. As a proxy for both, we finally overlaid the risk component
-from the composite map (see above) with the proportion of built land, as
-a proxy for a mix of habitat disturbance, potential for bat synanthropy
-or contact with bridge hosts like
-livestock,\textsuperscript{\textbf{Rulli2021LanCha?},\textbf{Cui2019OriEvo?}}
-and human population density and
-connectivity\textsuperscript{\textbf{Plowright2017PatZoo?},\textbf{Muylaert2022PreFut?},\textbf{Hassell2017UrbDis?}}
-(fig.~\ref{fig:compound}). Accounting for these factors, most of South
-America and Europe are at comparatively lower risk, as--although densely
-populated--settlements tend to be in areas with lower potential risk.
-Conversely, regions like Malaysia and the northern coast of Australia
-have a high evolutionary risk component, but should represent a
-relatively lower effective risk due to low human density. However,
-southeast Asia, the Indian subcontinent, and scattered hotspots in
-sub-Saharan Africa are at high risk due to the overlap between human
-populations and natural opportunities for cross-species transmission of
-betacoronaviruses.
+probability that spillovers become epidemics)\textsuperscript{1}. It
+must be noted that the assesment of risk based on the GMTC mechanisms
+does not account for human presence; for this reason, it represents
+``potential'' level of risk, which must be re-evaluated in the light of
+human presence. As a proxy for both, we finally overlaid the risk
+component from the composite map (see above) with the proportion of
+built land, as a proxy for a mix of habitat disturbance, potential for
+bat synanthropy or contact with bridge hosts like
+livestock,\textsuperscript{30,31} and human population density and
+connectivity\textsuperscript{1,32,33} (fig.~\ref{fig:compound}).
+Accounting for these factors, most of South America and Europe are at
+comparatively lower risk, as--although densely populated--settlements
+tend to be in areas with lower potential risk. Conversely, regions like
+Malaysia and the northern coast of Australia have a high evolutionary
+risk component, but should represent a relatively lower effective risk
+due to low human density. However, southeast Asia, the Indian
+subcontinent, and scattered hotspots in sub-Saharan Africa are at high
+risk due to the overlap between human populations and natural
+opportunities for cross-species transmission of betacoronaviruses.
 
@@
 sarbecoviruses from Indochina that was poorly characterized prior to the
-pandemic.\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
-Similarly, the SARS-CoV outbreak began in Guangdong province in 2002,
-reaching humans through small carnivore bridge hosts, but was eventually
-traced back to a set of likely progenitor viruses found in cave-dwelling
-horseshoe bats in Yunnan
-province;\textsuperscript{\textbf{Hu2017DisRic?}} nearby, antibody
-evidence has indicated human exposure to SARS-like
-viruses.\textsuperscript{\textbf{Wang2018SerEvi?}} MERS-CoV was first
-detected in Jordan, but is widespread in camels in East Africa and the
-Middle East, and may have reached its bridge host decades earlier than
-originally supposed;\textsuperscript{\textbf{Muller2014MerCor?}} as a
-result, the geography of the original bat-to-camel transmission is still
-widely regarded as uncertain. All of these are broadly consistent with
-the risk factors we identify. Notably, India and west Africa are
-additional hotspots that have yet to experience the emergence of a bat
-coronavirus into human populations, but may still be at
-risk---particularly given known gaps in bat
-surveillance,\textsuperscript{\textbf{Cohen2022SamStr?}} and a dense
-population in both regions with global connectivity. In any of these
-regions, surveillance on viral reservoirs can be paired with targeted
-monitoring of high-risk human populations (i.e., those with regular
-wildlife contact)\textsuperscript{\textbf{Xu2004EpiClu?}} for maximum
-impact.
+pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
+in Guangdong province in 2002, reaching humans through small carnivore
+bridge hosts, but was eventually traced back to a set of likely
+progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
+province;\textsuperscript{34} nearby, antibody evidence has indicated
+human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
+first detected in Jordan, but is widespread in camels in East Africa and
+the Middle East, and may have reached its bridge host decades earlier
+than originally supposed;\textsuperscript{36} as a result, the geography
+of the original bat-to-camel transmission is still widely regarded as
+uncertain. All of these are broadly consistent with the risk factors we
+identify. Notably, India and west Africa are additional hotspots that
+have yet to experience the emergence of a bat coronavirus into human
+populations, but may still be at risk---particularly given known gaps in
+bat surveillance,\textsuperscript{24} and a dense population in both
+regions with global connectivity. In any of these regions, surveillance
+on viral reservoirs can be paired with targeted monitoring of high-risk
+human populations (i.e., those with regular wildlife
+contact)\textsuperscript{37} for maximum impact.
 
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to
-humans.\textsuperscript{\textbf{Kasso2013EcoEco?}} Over two-thirds of
-bats are known to be either obligate or facultative insectivores,
-therefore actively contributing for agricultural pest
-control,\textsuperscript{\textbf{Voigt2016BatAnt?},\textbf{Williams-Guillen2008BatLim?}}
-and vectors of pathogens that put a risk on human
-health;\textsuperscript{\textbf{Gonsalves2013MosCon?},\textbf{Gonsalves2013MosInf?}}
-some other species are essential links in many seed-dispersal
-networks.\textsuperscript{\textbf{Mello2011MisPar?}} However, many of
-these species face a high risk of extinction, particularly given
+ecosystem-derived benefits to humans.\textsuperscript{38} Over
+two-thirds of bats are known to be either obligate or facultative
+insectivores, therefore actively contributing for agricultural pest
+control,\textsuperscript{39,40} and vectors of pathogens that put a risk
+on human health;\textsuperscript{41,42} some other species are essential
+links in many seed-dispersal networks.\textsuperscript{43} However, many
+of these species face a high risk of extinction, particularly given
 persecution and killings that sometimes follows from messaging about
@@
 have predicted a high diversity of unsampled bat
-viruses,\textsuperscript{\textbf{Olival2017HosVir?}} bat
-coronaviruses,\textsuperscript{\textbf{Anthony2017GloPat?}} and even
-specifically
-betacoronaviruses\textsuperscript{\textbf{Becker2022OptPre?}} in both
-the Amazon and southeast Asia. While we find that both regions are
+viruses,\textsuperscript{19} bat coronaviruses,\textsuperscript{11} and
+even specifically betacoronaviruses\textsuperscript{20} in both the
+Amazon and southeast Asia. While we find that both regions are
 characterized by unique and diverse communities of both hosts and
@@
 that is supported by previous work indicating a higher rate of
-codivergence in Latin
-America.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Caraballo2022CroTra?}}
-In contrast, in southeast Asia, host richness and viral distinctiveness
-are high but sharing is low; this suggests a different type of
-evolutionary dynamics that could generate high local diversity of
-viruses through host switching and viral recombination (see
-\emph{e.g.},\textsuperscript{\textbf{Latinne2020OriCro?}} as well as the
-discovery of recombinant viruses with genetic material from both the
-SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{\textbf{Wu2021ComSur?}}Both of these regions
-are priority areas for sampling, especially given predictions that they
-contain many bat hosts of undiscovered
-betacoronaviruses.\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Cohen2022SamStr?}}
-However, both the evolutionary and ecological aspects of emergence risk
-are higher in southeast Asia---a fact that will only become more
-relevant, as bats track shifting climates and exchange viruses with
-other species, creating a hotspot of elevated cross-species transmission
-unique to the
-region.\textsuperscript{\textbf{Carlson2022CliCha?},\textbf{Muylaert2022PreFut?}}
+codivergence in Latin America.\textsuperscript{11,44} In contrast, in
+southeast Asia, host richness and viral distinctiveness are high but
+sharing is low; this suggests a different type of evolutionary dynamics
+that could generate high local diversity of viruses through host
+switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
+as well as the discovery of recombinant viruses with genetic material
+from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
+lineage).\textsuperscript{45}Both of these regions are priority areas
+for sampling, especially given predictions that they contain many bat
+hosts of undiscovered betacoronaviruses.\textsuperscript{20,24} However,
+both the evolutionary and ecological aspects of emergence risk are
+higher in southeast Asia---a fact that will only become more relevant,
+as bats track shifting climates and exchange viruses with other species,
+creating a hotspot of elevated cross-species transmission unique to the
+region.\textsuperscript{32,46}
 
@@
 could spread relatively easily due to encountering little immune
-resistance (CITE PLOWRIGHT). With the right combination of viral traits,
-such as low disease-induced mortality or high transmission rate, this
-could lead to successfully spread within the new host community.
-However, while high adaptation to a specific host can be advantageous,
-it may also lead to maladaptation when the pathogen encounters a new
-unsuitable host, potentially resulting in its extinction.
-
-\begin{quote}
-Plowright RK, Foley P, Field HE, Dobson AP, Foley JE, Eby P, Daszak P.
-Urban habituation, ecological connectivity and epidemic dampening: the
-emergence of Hendra virus from flying foxes (Pteropus spp.). Proceedings
-of the Royal Society B: Biological Sciences. 2011 Dec
-22;278(1725):3703-12.
-\end{quote}
+resistance.\textsuperscript{\textbf{Plowright2011Urban?}} With the right
+combination of viral traits, such as low disease-induced mortality or
+high transmission rate, this could lead to successfully spread within
+the new host community. However, while high adaptation to a specific
+host can be advantageous, it may also lead to maladaptation when the
+pathogen encounters a new unsuitable host, potentially resulting in its
+extinction.
 
@@
 greater contact with potential bridge hosts like domesticated
-species.\textsuperscript{\textbf{Alves2018GeoVar?},\textbf{Treitler2016EffLoc?},\textbf{Rulli2021LanCha?},\textbf{Mendenhall2014PreBio?}}
-This represents a challenge for both conservation strategies and
-pandemic prevention,\textsuperscript{\textbf{Amman2011InvRol?}} but
+species.\textsuperscript{30,47--49} This represents a challenge for both
+conservation strategies and pandemic prevention,\textsuperscript{50} but
 identifying areas at risk, and protecting the health of bats and
 ecosystems within those zones, can be a win-win intervention for
-both.\textsuperscript{\textbf{Hopkins2021HowIde?},\textbf{Plowright2021LanUse?},\textbf{OHHLEP2022OneHea?}}
+both.\textsuperscript{51,\textbf{Hopkins2021HowIde?},\textbf{Plowright2021Lan?}}
 As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{27} the incorporation of
+to guide public health actions,\textsuperscript{32} the incorporation of
 human activity predictors will become more
-importyant.\textsuperscript{28}
+importyant.\textsuperscript{52}
 
@@
 cross-species transmission even between divergent species is generally
-high,\textsuperscript{29} especially for
+high,\textsuperscript{53} especially for
 beta-coronaviruses.\textsuperscript{13} To project viral sharing values
@@
 
+\leavevmode\hypertarget{ref-Banerjee2020Novel}{}%
+\CSLLeftMargin{27. }
+\CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
+Systems of Bats. \emph{Frontiers in Immunology} \textbf{11}, 26 (2020).}
+
+\leavevmode\hypertarget{ref-Shi2014Deep}{}%
+\CSLLeftMargin{28. }
+\CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
+Sister Species of Eidolon (Pteropodidae) with Evidence for Widespread
+Panmixia. \emph{Acta Chiropterologica} \textbf{16}, 279--292 (2014).}
+
+\leavevmode\hypertarget{ref-Kettenburg2022Full}{}%
+\CSLLeftMargin{29. }
+\CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
+Sequences From Malagasy Fruit Bats Define a Unique Evolutionary History
+for This Coronavirus Clade. \emph{Frontiers in Public Health}
+\textbf{10}, (2022).}
+
+\leavevmode\hypertarget{ref-Rulli2021Landuse}{}%
+\CSLLeftMargin{30. }
+\CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
+Land-use change and the livestock revolution increase the risk of
+zoonotic coronavirus transmission from rhinolophid bats. \emph{Nature
+Food} \textbf{2}, 409--416 (2021).}
+
+\leavevmode\hypertarget{ref-Cui2019Origin}{}%
+\CSLLeftMargin{31. }
+\CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
+pathogenic coronaviruses. \emph{Nature Reviews Microbiology}
+\textbf{17}, 181--192 (2019).}
+
 \leavevmode\hypertarget{ref-Muylaert2022Present}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 
+\leavevmode\hypertarget{ref-Hassell2017Urbanization}{}%
+\CSLLeftMargin{33. }
+\CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
+Urbanization and disease emergence: Dynamics at the
+wildlife--livestock--human interface. \emph{Trends in ecology \&
+evolution} \textbf{32}, 55--67 (2017).}
+
+\leavevmode\hypertarget{ref-Hu2017Discovery}{}%
+\CSLLeftMargin{34. }
+\CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
+bat SARS-related coronaviruses provides new insights into the origin of
+SARS coronavirus. \emph{PLoS pathogens} \textbf{13}, e1006698 (2017).}
+
+\leavevmode\hypertarget{ref-Wang2018Serological}{}%
+\CSLLeftMargin{35. }
+\CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
+SARS-Related Coronavirus Infection in Humans, China. \emph{Virologica
+Sinica} \textbf{33}, 104--107 (2018).}
+
+\leavevmode\hypertarget{ref-Muller2014Mers}{}%
+\CSLLeftMargin{36. }
+\CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
+neutralizing antibodies in camels, Eastern Africa, 1983--1997.
+\emph{Emerging infectious diseases} \textbf{20}, 2093 (2014).}
+
+\leavevmode\hypertarget{ref-Xu2004Epidemiologic}{}%
+\CSLLeftMargin{37. }
+\CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
+Origin in China. \emph{Emerging Infectious Diseases} \textbf{10},
+1030--1037 (2004).}
+
+\leavevmode\hypertarget{ref-Kasso2013Ecological}{}%
+\CSLLeftMargin{38. }
+\CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
+Importance of Bats (Order Chiroptera). \emph{ISRN Biodiversity}
+\textbf{2013}, e187415 (2013).}
+
+\leavevmode\hypertarget{ref-Voigt2016Bats}{}%
+\CSLLeftMargin{39. }
+\CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
+a Changing World}. (Springer International Publishing, 2016).
+doi:\href{https://doi.org/10.1007/978-3-319-25220-9}{10.1007/978-3-319-25220-9}.}
+
+\leavevmode\hypertarget{ref-Williams-Guillen2008Bats}{}%
+\CSLLeftMargin{40. }
+\CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
+Bats Limit Insects in a Neotropical Agroforestry System. \emph{Science}
+\textbf{320}, 70--70 (2008).}
+
+\leavevmode\hypertarget{ref-Gonsalves2013Mosquito}{}%
+\CSLLeftMargin{41. }
+\CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
+Monamy, V. Mosquito Consumption by Insectivorous Bats: Does Size Matter?
+\emph{PLOS ONE} \textbf{8}, e77183 (2013).}
+
+\leavevmode\hypertarget{ref-Gonsalves2013Mosquitoes}{}%
+\CSLLeftMargin{42. }
+\CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
+Do mosquitoes influence bat activity in coastal habitats? \emph{Wildlife
+Research} \textbf{40}, 10--24 (2013).}
+
+\leavevmode\hypertarget{ref-Mello2011Missing}{}%
+\CSLLeftMargin{43. }
+\CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
+Dispersal Networks: Structure and Robustness of Bat-Fruit Interactions.
+\emph{PLOS ONE} \textbf{6}, e17395 (2011).}
+
+\leavevmode\hypertarget{ref-Caraballo2022Crossspecies}{}%
+\CSLLeftMargin{44. }
+\CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
+Coronaviruses in the Americas: Contrasting Patterns between
+Alphacoronavirus and Betacoronavirus. \emph{Microbiology Spectrum}
+\textbf{0}, e01411--22 (2022).}
+
+\leavevmode\hypertarget{ref-Wu2021Comprehensive}{}%
+\CSLLeftMargin{45. }
+\CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
+sarbecoviruses across China for the origin tracing of SARS-CoV and
+SARS-CoV-2}. (2021)
+doi:\href{https://doi.org/10.21203/rs.3.rs-885194/v1}{10.21203/rs.3.rs-885194/v1}.}
+
+\leavevmode\hypertarget{ref-Carlson2022Climate}{}%
+\CSLLeftMargin{46. }
+\CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
+cross-species viral transmission risk. \emph{Nature} 1--1 (2022)
+doi:\href{https://doi.org/10.1038/s41586-022-04788-w}{10.1038/s41586-022-04788-w}.}
+
+\leavevmode\hypertarget{ref-Alves2018Geographic}{}%
+\CSLLeftMargin{47. }
+\CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
+Souza, K., Gouveia, S. F. \& Villalobos, F. Geographic variation in the
+relationship between large-scale environmental determinants and bat
+species richness. \emph{Basic and Applied Ecology} \textbf{27}, 1--8
+(2018).}
+
+\leavevmode\hypertarget{ref-Treitler2016Effect}{}%
+\CSLLeftMargin{48. }
+\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
+effect of local land use and loss of forests on bats and nocturnal
+insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}
+
+\leavevmode\hypertarget{ref-Mendenhall2014Predictinga}{}%
+\CSLLeftMargin{49. }
+\CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
+E. A. \& Daily, G. C. Predicting biodiversity change and averting
+collapse in agricultural landscapes. \emph{Nature} \textbf{509},
+213--217 (2014).}
+
+\leavevmode\hypertarget{ref-Amman2011Investigating}{}%
+\CSLLeftMargin{50. }
+\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
+of bats in emerging zoonoses: Balancing ecology, conservation and public
+health interest}. (FAO, 2011).}
+
+\leavevmode\hypertarget{ref-OHHLEP2022One}{}%
+\CSLLeftMargin{51. }
+\CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
+sustainable and healthy future. \emph{PLOS Pathogens} \textbf{18},
+e1010537 (2022).}
+
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@5012201741ab7f7cf3b29d541c45d898fad9f315 🚀

@@
 superficial proxy for spillover
-risk,\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Sanchez2022Strategy?}}
-these approaches oversimplify the relevant interspecific heterogeneity
-in immunology, behavior, and other traits, and therefore overlook unique
+risk,\textsuperscript{2,3,\textbf{Ruiz-Aravena2022Ecological?}} these
+approaches oversimplify the relevant interspecific heterogeneity in
+immunology, behavior, and other traits, and therefore overlook unique
 host pools that allow for the rapid evolution of highly divergent
-viruses.\textsuperscript{2} In the case of generalist pathogens like
+viruses.\textsuperscript{4} In the case of generalist pathogens like
 betacoronaviruses, there is conceptual and empirical support to the idea
 that these community-level mechanisms are even more
-important,\textsuperscript{3} particularly given that cross-species
+important,\textsuperscript{5} particularly given that cross-species
 transmission may, as a rule, structure viral evolution more than
-co-divergence with hosts.\textsuperscript{4} This creates a disconnect
+co-divergence with hosts.\textsuperscript{6} This creates a disconnect
 between coevolutionary theory and most existing ecological frameworks
@@
 explicitly connect microevolutionary dynamics to the macroecology and
-biogeography of symbiotic interactions.\textsuperscript{5} The GMTC
-posits that coevolutionary processes among pairs\textsuperscript{6} or
-complexes\textsuperscript{7} of species are structured in space by the
+biogeography of symbiotic interactions.\textsuperscript{7} The GMTC
+posits that coevolutionary processes among pairs\textsuperscript{8} or
+complexes\textsuperscript{9} of species are structured in space by the
 rippling effects of abiotic conditions onto evolutionary mechanisms,
 giving rise to fragmented systems with different ecologies over large
-spatial extents.\textsuperscript{8} The GMTC predicts a spatial
+spatial extents.\textsuperscript{10} The GMTC predicts a spatial
 fragmentation of coevolutionary dynamics under the joint action of three
-processes:\textsuperscript{9} coevolutionary hot- and coldspots, which
+processes:\textsuperscript{11} coevolutionary hot- and coldspots, which
 appear when the intensity of \emph{interaction} (in terms of reciprocal
@@
 are more biotically complex) and the local favorability of the
-environment;\textsuperscript{10} and trait remixing, which occurs when
+environment;\textsuperscript{12} and trait remixing, which occurs when
 coevolutionary dynamics change when community-level \emph{functional
@@
 of host jumps, recombination among disparate lineages, and, to a lesser
-degree, co-divergence with their hosts---\textsuperscript{11}a mix of
+degree, co-divergence with their hosts---\textsuperscript{2}a mix of
 mechanisms that creates a complex and nonlinear relationship between
@@
 betacoronavirus subgenera, including nobecoviruses, sarbecoviruses, and
-merbecoviruses).\textsuperscript{12}
+merbecoviruses).\textsuperscript{13}
 
@@
 play a stronger role in coronavirus diversification than
-cospeciation---\textsuperscript{11}a property that would theoretically
+cospeciation---\textsuperscript{2}a property that would theoretically
 allow for substantial broad divergence in their biogeography. However,
@@
 significant events in coronavirus evolutionary
-history;\textsuperscript{11,13} as a result, the mosaic of
+history;\textsuperscript{2,14} as a result, the mosaic of
 betacoronavirus phylogeography is assembled from a set of overlapping
@@
 bats (Vespertilionidae), are considered the primary hosts of the
-subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{12,13}
+subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{13,14}
 but in the Americas, where merbecoviruses are the only lineage present,
 they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
-Phyllostomidae).\textsuperscript{14--17} At the coarsest scale, these
+Phyllostomidae).\textsuperscript{15--18} At the coarsest scale, these
 heterogeneities are lost, and betacoronavirus biogeography tracks the
@@
 gradients in their species richness generally track broader patterns of
-mammal diversity,\textsuperscript{18} with a striking Neotropical
+mammal diversity,\textsuperscript{19} with a striking Neotropical
 hotspot (especially in the Amazon basin) and a secondary hotspot
@@
 presumed to be hotspots of viral adaptive radiation, and therefore areas
-of concern for human health.\textsuperscript{11,19} However, the
-hotspots of known bat betacoronavirus hosts show a distinct pattern,
-with primary hotspots (both in terms of area and higher values) of host
-richness situated in southeast Asia, parts of southern Europe, and to a
-lesser extent parts of Africa in the -25-0 range of latitudes
+of concern for human health.\textsuperscript{2,20} However, the hotspots
+of known bat betacoronavirus hosts show a distinct pattern, with primary
+hotspots (both in terms of area and higher values) of host richness
+situated in southeast Asia, parts of southern Europe, and to a lesser
+extent parts of Africa in the -25-0 range of latitudes
 (fig.~\ref{fig:richness}; top). Although hundreds of species likely host
@@
 suggested that these undiscovered reservoirs should follow the same
-diversity gradient.\textsuperscript{20} In principle, these hotspots of
+diversity gradient.\textsuperscript{21} In principle, these hotspots of
 locally-diverse, virus-rich bat communities should drive more adaptive
@@
 the places where the fewest bat betacoronavirus sequences have been
-generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
+generated,\textsuperscript{22--24} resulting in a sparser phylogenetic
 tree, and artificially inflating distinctiveness; conversely,
-disproportionate research effort in eastern China\textsuperscript{24}
+disproportionate research effort in eastern China\textsuperscript{25}
 may have led to a more complete inventory of the local diversity of
@@
 researchers\textsuperscript{\textbf{Latinne2020Origins?}} (and so have
-been sampled most intensively;).\textsuperscript{24} The hotspots of
+been sampled most intensively;).\textsuperscript{25} The hotspots of
 host richness and viral diversity in southeast Asia---both of which are
@@
 richness---are almost entirely driven by viral adaptive radiation
-through host switching within this clade\textsuperscript{12,20}. In
+through host switching within this clade\textsuperscript{13,21}. In
 contrast, the Neotropical hotspot of viral distinctiveness is driven by
@@
 Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{12,25} While comparatively understudied, New
+ancient.\textsuperscript{13,26} While comparatively understudied, New
 World merbecoviruses have been found in the ghost-faced bats
 (Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae).\textsuperscript{14--17} The former two
+free-tailed bats (Molossidae).\textsuperscript{15--18} The former two
 groups and a clade of the latter are endemic to the Neotropics, while
 the explosive adaptive radiations of the phyllostomids are responsible
-for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
+for the hotspot of bat diversity in the Amazon.\textsuperscript{27}
 Together, these clades of New World bats play host to a distinct regime
@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts (and their diverse immune strategies)\textsuperscript{27} creates
+hosts (and their diverse immune strategies)\textsuperscript{28} creates
 a landscape of selective pressure; the trajectory of viruses'
@@
 species are endemic following evolutionary divergence from sister
-species in both African and Asian continents,\textsuperscript{28} as one
-such hotspot; interestingly, a recent study\textsuperscript{29} reported
+species in both African and Asian continents,\textsuperscript{29} as one
+such hotspot; interestingly, a recent study\textsuperscript{30} reported
 a novel and highly divergent lineage of nobecoviruses from
@@
 bat synanthropy or contact with bridge hosts like
-livestock,\textsuperscript{30,31} and human population density and
-connectivity\textsuperscript{1,32,33} (fig.~\ref{fig:compound}).
+livestock,\textsuperscript{31,32} and human population density and
+connectivity\textsuperscript{1,33,34} (fig.~\ref{fig:compound}).
 Accounting for these factors, most of South America and Europe are at
@@
 sarbecoviruses from Indochina that was poorly characterized prior to the
-pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
+pandemic.\textsuperscript{22--24} Similarly, the SARS-CoV outbreak began
 in Guangdong province in 2002, reaching humans through small carnivore
@@
 progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
-province;\textsuperscript{34} nearby, antibody evidence has indicated
-human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
+province;\textsuperscript{35} nearby, antibody evidence has indicated
+human exposure to SARS-like viruses.\textsuperscript{36} MERS-CoV was
 first detected in Jordan, but is widespread in camels in East Africa and
 the Middle East, and may have reached its bridge host decades earlier
-than originally supposed;\textsuperscript{36} as a result, the geography
+than originally supposed;\textsuperscript{37} as a result, the geography
 of the original bat-to-camel transmission is still widely regarded as
@@
 populations, but may still be at risk---particularly given known gaps in
-bat surveillance,\textsuperscript{24} and a dense population in both
+bat surveillance,\textsuperscript{25} and a dense population in both
 regions with global connectivity. In any of these regions, surveillance
@@
 human populations (i.e., those with regular wildlife
-contact)\textsuperscript{37} for maximum impact.
+contact)\textsuperscript{38} for maximum impact.
 
@@
 provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to humans.\textsuperscript{38} Over
+ecosystem-derived benefits to humans.\textsuperscript{39} Over
 two-thirds of bats are known to be either obligate or facultative
 insectivores, therefore actively contributing for agricultural pest
-control,\textsuperscript{39,40} and vectors of pathogens that put a risk
-on human health;\textsuperscript{41,42} some other species are essential
-links in many seed-dispersal networks.\textsuperscript{43} However, many
+control,\textsuperscript{40,41} and vectors of pathogens that put a risk
+on human health;\textsuperscript{42,43} some other species are essential
+links in many seed-dispersal networks.\textsuperscript{44} However, many
 of these species face a high risk of extinction, particularly given
@@
 have predicted a high diversity of unsampled bat
-viruses,\textsuperscript{19} bat coronaviruses,\textsuperscript{11} and
-even specifically betacoronaviruses\textsuperscript{20} in both the
+viruses,\textsuperscript{20} bat coronaviruses,\textsuperscript{2} and
+even specifically betacoronaviruses\textsuperscript{21} in both the
 Amazon and southeast Asia. While we find that both regions are
@@
 that is supported by previous work indicating a higher rate of
-codivergence in Latin America.\textsuperscript{11,44} In contrast, in
+codivergence in Latin America.\textsuperscript{2,45} In contrast, in
 southeast Asia, host richness and viral distinctiveness are high but
@@
 that could generate high local diversity of viruses through host
-switching and viral recombination (see \emph{e.g.},\textsuperscript{13}
+switching and viral recombination (see \emph{e.g.},\textsuperscript{14}
 as well as the discovery of recombinant viruses with genetic material
 from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{45}Both of these regions are priority areas
+lineage).\textsuperscript{46}Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
-hosts of undiscovered betacoronaviruses.\textsuperscript{20,24} However,
+hosts of undiscovered betacoronaviruses.\textsuperscript{21,25} However,
 both the evolutionary and ecological aspects of emergence risk are
@@
 creating a hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{32,46}
+region.\textsuperscript{33,47}
 
@@
 could spread relatively easily due to encountering little immune
-resistance.\textsuperscript{\textbf{Plowright2011Urban?}} With the right
-combination of viral traits, such as low disease-induced mortality or
-high transmission rate, this could lead to successfully spread within
-the new host community. However, while high adaptation to a specific
-host can be advantageous, it may also lead to maladaptation when the
-pathogen encounters a new unsuitable host, potentially resulting in its
-extinction.
+resistance.\textsuperscript{48} With the right combination of viral
+traits, such as low disease-induced mortality or high transmission rate,
+this could lead to successfully spread within the new host community.
+However, while high adaptation to a specific host can be advantageous,
+it may also lead to maladaptation when the pathogen encounters a new
+unsuitable host, potentially resulting in its extinction.
 
@@
 greater contact with potential bridge hosts like domesticated
-species.\textsuperscript{30,47--49} This represents a challenge for both
-conservation strategies and pandemic prevention,\textsuperscript{50} but
+species.\textsuperscript{31,49--51} This represents a challenge for both
+conservation strategies and pandemic prevention,\textsuperscript{52} but
 identifying areas at risk, and protecting the health of bats and
 ecosystems within those zones, can be a win-win intervention for
-both.\textsuperscript{51,\textbf{Hopkins2021HowIde?},\textbf{Plowright2021Lan?}}
+both.\textsuperscript{53,\textbf{Hopkins2021HowIde?},\textbf{Plowright2021Lan?}}
 As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{32} the incorporation of
+to guide public health actions,\textsuperscript{33} the incorporation of
 human activity predictors will become more
-importyant.\textsuperscript{52}
+importyant.\textsuperscript{54}
 
@@
 cross-species transmission even between divergent species is generally
-high,\textsuperscript{53} especially for
-beta-coronaviruses.\textsuperscript{13} To project viral sharing values
+high,\textsuperscript{55} especially for
+beta-coronaviruses.\textsuperscript{14} To project viral sharing values
 into a single value for every pixel, we averaged the pairwise scores.
@@
 
+\leavevmode\hypertarget{ref-Anthony2017Global}{}%
+\CSLLeftMargin{2. }
+\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
+coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}
+
+\leavevmode\hypertarget{ref-Sanchez2022Strategy}{}%
+\CSLLeftMargin{3. }
+\CSLRightInline{Sánchez, C. A. \emph{et al.} A strategy to assess
+spillover risk of bat SARS-related coronaviruses in southeast asia.
+\emph{Nature Communications} \textbf{13}, (2022).}
+
 \leavevmode\hypertarget{ref-Agosta2010How}{}%
-\CSLLeftMargin{2. }
+\CSLLeftMargin{4. }
 \CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
@@
 \leavevmode\hypertarget{ref-Power2004Pathogen}{}%
-\CSLLeftMargin{3. }
+\CSLLeftMargin{5. }
 \CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
@@
 \leavevmode\hypertarget{ref-Geoghegan2017Comparative}{}%
-\CSLLeftMargin{4. }
+\CSLLeftMargin{6. }
 \CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
@@
 \leavevmode\hypertarget{ref-Thompson2005Geographic}{}%
-\CSLLeftMargin{5. }
+\CSLLeftMargin{7. }
 \CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
@@
 \leavevmode\hypertarget{ref-Thompson1994Coevolutionary}{}%
-\CSLLeftMargin{6. }
+\CSLLeftMargin{8. }
 \CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
@@
 \leavevmode\hypertarget{ref-Janzen1980When}{}%
-\CSLLeftMargin{7. }
+\CSLLeftMargin{9. }
 \CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-Price2002Macroevolutionary}{}%
-\CSLLeftMargin{8. }
+\CSLLeftMargin{10. }
 \CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2007Dos}{}%
-\CSLLeftMargin{9. }
+\CSLLeftMargin{11. }
 \CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
@@
 \leavevmode\hypertarget{ref-Thrall2007Coevolution}{}%
-\CSLLeftMargin{10. }
+\CSLLeftMargin{12. }
 \CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
@@
 
-\leavevmode\hypertarget{ref-Anthony2017Global}{}%
-\CSLLeftMargin{11. }
-\CSLRightInline{Anthony, S. J. \emph{et al.} Global patterns in
-coronavirus diversity. \emph{Virus Evolution} \textbf{3}, (2017).}
-
 \leavevmode\hypertarget{ref-Ruiz-Aravena2022Ecology}{}%
-\CSLLeftMargin{12. }
+\CSLLeftMargin{13. }
 \CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
@@
 \leavevmode\hypertarget{ref-Latinne2020Origin}{}%
-\CSLLeftMargin{13. }
+\CSLLeftMargin{14. }
 \CSLRightInline{Latinne, A. \emph{et al.} Origin and cross-species
@@
 \leavevmode\hypertarget{ref-Anthony2013Coronaviruses}{}%
-\CSLLeftMargin{14. }
+\CSLLeftMargin{15. }
 \CSLRightInline{Anthony, S. J. \emph{et al.} Coronaviruses in bats from
@@
 \leavevmode\hypertarget{ref-Goes2013Novel}{}%
-\CSLLeftMargin{15. }
+\CSLLeftMargin{16. }
 \CSLRightInline{Góes, L. G. B. \emph{et al.} Novel Bat Coronaviruses,
@@
 \leavevmode\hypertarget{ref-Goes2016Genetic}{}%
-\CSLLeftMargin{16. }
+\CSLLeftMargin{17. }
 \CSLRightInline{Góes, L. G. B. \emph{et al.} Genetic diversity of bats
@@
 \leavevmode\hypertarget{ref-Brandao2008Coronavirus}{}%
-\CSLLeftMargin{17. }
+\CSLLeftMargin{18. }
 \CSLRightInline{Brandão, P. E. \emph{et al.} A coronavirus detected in
@@
 \leavevmode\hypertarget{ref-Tanalgo2022Mapping}{}%
-\CSLLeftMargin{18. }
+\CSLLeftMargin{19. }
 \CSLRightInline{Tanalgo, K. C., Oliveira, H. F. M. \& Hughes, A. C.
@@
 \leavevmode\hypertarget{ref-Olival2017Host}{}%
-\CSLLeftMargin{19. }
+\CSLLeftMargin{20. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Host and viral traits
@@
 \leavevmode\hypertarget{ref-Becker2022Optimising}{}%
-\CSLLeftMargin{20. }
+\CSLLeftMargin{21. }
 \CSLRightInline{Becker, D. J. \emph{et al.} Optimising predictive models
@@
 \leavevmode\hypertarget{ref-Worobey2022Huanan}{}%
-\CSLLeftMargin{21. }
+\CSLLeftMargin{22. }
 \CSLRightInline{Worobey, M. \emph{et al.} The Huanan market was the
@@
 \leavevmode\hypertarget{ref-Temmam2022Bat}{}%
-\CSLLeftMargin{22. }
+\CSLLeftMargin{23. }
 \CSLRightInline{Temmam, S. \emph{et al.} Bat coronaviruses related to
@@
 \leavevmode\hypertarget{ref-Boni2020Evolutionary}{}%
-\CSLLeftMargin{23. }
+\CSLLeftMargin{24. }
 \CSLRightInline{Boni, M. F. \emph{et al.} Evolutionary origins of the
@@
 \leavevmode\hypertarget{ref-Cohen2022Sampling}{}%
-\CSLLeftMargin{24. }
+\CSLLeftMargin{25. }
 \CSLRightInline{Cohen, L. E., Fagre, A. C., Chen, B., Carlson, C. J. \&
@@
 \leavevmode\hypertarget{ref-Olival2020Possibility}{}%
-\CSLLeftMargin{25. }
+\CSLLeftMargin{26. }
 \CSLRightInline{Olival, K. J. \emph{et al.} Possibility for reverse
@@
 \leavevmode\hypertarget{ref-Ammerman2012First}{}%
-\CSLLeftMargin{26. }
+\CSLLeftMargin{27. }
 \CSLRightInline{Ammerman, L. K., Lee, D. N. \& Tipps, T. M. First
@@
 \leavevmode\hypertarget{ref-Banerjee2020Novel}{}%
-\CSLLeftMargin{27. }
+\CSLLeftMargin{28. }
 \CSLRightInline{Banerjee, A. \emph{et al.} Novel Insights Into Immune
@@
 \leavevmode\hypertarget{ref-Shi2014Deep}{}%
-\CSLLeftMargin{28. }
+\CSLLeftMargin{29. }
 \CSLRightInline{Shi, J. J. \emph{et al.} A Deep Divergence Time between
@@
 \leavevmode\hypertarget{ref-Kettenburg2022Full}{}%
-\CSLLeftMargin{29. }
+\CSLLeftMargin{30. }
 \CSLRightInline{Kettenburg, G. \emph{et al.} Full Genome Nobecovirus
@@
 \leavevmode\hypertarget{ref-Rulli2021Landuse}{}%
-\CSLLeftMargin{30. }
+\CSLLeftMargin{31. }
 \CSLRightInline{Rulli, M. C., D'Odorico, P., Galli, N. \& Hayman, D. T.
@@
 \leavevmode\hypertarget{ref-Cui2019Origin}{}%
-\CSLLeftMargin{31. }
+\CSLLeftMargin{32. }
 \CSLRightInline{Cui, J., Li, F. \& Shi, Z.-L. Origin and evolution of
@@
 \leavevmode\hypertarget{ref-Muylaert2022Present}{}%
-\CSLLeftMargin{32. }
+\CSLLeftMargin{33. }
 \CSLRightInline{Muylaert, R. L. \emph{et al.} Present and future
@@
 \leavevmode\hypertarget{ref-Hassell2017Urbanization}{}%
-\CSLLeftMargin{33. }
+\CSLLeftMargin{34. }
 \CSLRightInline{Hassell, J. M., Begon, M., Ward, M. J. \& Fèvre, E. M.
@@
 \leavevmode\hypertarget{ref-Hu2017Discovery}{}%
-\CSLLeftMargin{34. }
+\CSLLeftMargin{35. }
 \CSLRightInline{Hu, B. \emph{et al.} Discovery of a rich gene pool of
@@
 \leavevmode\hypertarget{ref-Wang2018Serological}{}%
-\CSLLeftMargin{35. }
+\CSLLeftMargin{36. }
 \CSLRightInline{Wang, N. \emph{et al.} Serological Evidence of Bat
@@
 \leavevmode\hypertarget{ref-Muller2014Mers}{}%
-\CSLLeftMargin{36. }
+\CSLLeftMargin{37. }
 \CSLRightInline{Müller, M. A. \emph{et al.} MERS coronavirus
@@
 \leavevmode\hypertarget{ref-Xu2004Epidemiologic}{}%
-\CSLLeftMargin{37. }
+\CSLLeftMargin{38. }
 \CSLRightInline{Xu, R.-H. \emph{et al.} Epidemiologic Clues to SARS
@@
 \leavevmode\hypertarget{ref-Kasso2013Ecological}{}%
-\CSLLeftMargin{38. }
+\CSLLeftMargin{39. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Voigt2016Bats}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{40. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008Bats}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013Mosquito}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{42. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013Mosquitoes}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Mello2011Missing}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Caraballo2022Crossspecies}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
@@
 \leavevmode\hypertarget{ref-Wu2021Comprehensive}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022Climate}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 
+\leavevmode\hypertarget{ref-Plowright2011Urban}{}%
+\CSLLeftMargin{48. }
+\CSLRightInline{Plowright, R. K. \emph{et al.} Urban habituation,
+ecological connectivity and epidemic dampening: The emergence of hendra
+virus from flying foxes (pteropus spp.). \emph{Proceedings of the Royal
+Society B: Biological Sciences} \textbf{278}, 3703--3712 (2011).}
+
 \leavevmode\hypertarget{ref-Alves2018Geographic}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016Effect}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Mendenhall2014Predictinga}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
@@
 \leavevmode\hypertarget{ref-Amman2011Investigating}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-OHHLEP2022One}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{53. }
 \CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@19136a1ac60e36dcbf1bf724a135b57c30f310f2 🚀

@@
 reservoirs.\textsuperscript{1} Although host richness is often used as a
-superficial proxy for spillover
-risk,\textsuperscript{2,3,\textbf{Ruiz-Aravena2022Ecological?}} these
+superficial proxy for spillover risk,\textsuperscript{2--4} these
 approaches oversimplify the relevant interspecific heterogeneity in
@@
 groups of bat hosts are predictive of coronavirus
-diversity.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?}}
-Projecting bat and betacoronavirus phylogeny over space
-(fig.~\ref{fig:biogeo}), we find support for the idea that bat community
-assembly is directly responsible for a global mosaic of viral evolution.
-The distinct groupings (represented by different colors, symbolizing
-positions in a subspace formed by the first two phylogenetic principal
-components) are essentially equivalent between the two groups, and can
-be coarsely delineated as (1) south and southeast Asia; (2) east Asia
-(including northern China), west Asia, and the Mediterranean coast; (3)
-Eurasia above a northing of 40; and (4) Africa and Latin America. In
-some cases, this diverges from expectations about coronavirus
-biogeography: for example, previous work has rarely flagged India as a
-region of interest, but for both bats and betacoronaviruses, the
-subcontinent falls into the same regions as the southeast Asian
-peninsula (and indeed, the region is home to known bat hosts of multiple
-betacoronavirus subgenera, including nobecoviruses, sarbecoviruses, and
-merbecoviruses).\textsuperscript{13}
+diversity.\textsuperscript{2,3} Projecting bat and betacoronavirus
+phylogeny over space (fig.~\ref{fig:biogeo}), we find support for the
+idea that bat community assembly is directly responsible for a global
+mosaic of viral evolution. The distinct groupings (represented by
+different colors, symbolizing positions in a subspace formed by the
+first two phylogenetic principal components) are essentially equivalent
+between the two groups, and can be coarsely delineated as (1) south and
+southeast Asia; (2) east Asia (including northern China), west Asia, and
+the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
+Africa and Latin America. In some cases, this diverges from expectations
+about coronavirus biogeography: for example, previous work has rarely
+flagged India as a region of interest, but for both bats and
+betacoronaviruses, the subcontinent falls into the same regions as the
+southeast Asian peninsula (and indeed, the region is home to known bat
+hosts of multiple betacoronavirus subgenera, including nobecoviruses,
+sarbecoviruses, and merbecoviruses).\textsuperscript{3}
 
@@
 bats (Vespertilionidae), are considered the primary hosts of the
-subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{13,14}
+subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,14}
 but in the Americas, where merbecoviruses are the only lineage present,
@@
 (subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
-been of the greatest interest to
-researchers\textsuperscript{\textbf{Latinne2020Origins?}} (and so have
-been sampled most intensively;).\textsuperscript{25} The hotspots of
-host richness and viral diversity in southeast Asia---both of which are
-disproportionately high, considering the global landscape of bat species
-richness---are almost entirely driven by viral adaptive radiation
-through host switching within this clade\textsuperscript{13,21}. In
-contrast, the Neotropical hotspot of viral distinctiveness is driven by
-isolation by host vicariance. Out of the four main groups of
-betacoronaviruses, only merbecoviruses have been found in animals in the
-Americas--- an introduction that is generally presumed to be
-ancient.\textsuperscript{13,26} While comparatively understudied, New
-World merbecoviruses have been found in the ghost-faced bats
-(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
-free-tailed bats (Molossidae).\textsuperscript{15--18} The former two
-groups and a clade of the latter are endemic to the Neotropics, while
-the explosive adaptive radiations of the phyllostomids are responsible
-for the hotspot of bat diversity in the Amazon.\textsuperscript{27}
-Together, these clades of New World bats play host to a distinct regime
-of betacoronavirus coevolution.
+been of the greatest interest to researchers\textsuperscript{14} (and so
+have been sampled most intensively;).\textsuperscript{25} The hotspots
+of host richness and viral diversity in southeast Asia---both of which
+are disproportionately high, considering the global landscape of bat
+species richness---are almost entirely driven by viral adaptive
+radiation through host switching within this
+clade\textsuperscript{3,21}. In contrast, the Neotropical hotspot of
+viral distinctiveness is driven by isolation by host vicariance. Out of
+the four main groups of betacoronaviruses, only merbecoviruses have been
+found in animals in the Americas--- an introduction that is generally
+presumed to be ancient.\textsuperscript{3,26} While comparatively
+understudied, New World merbecoviruses have been found in the
+ghost-faced bats (Mormoopidae), Neotropical leaf-nosed bats
+(Phyllostomidae), and free-tailed bats
+(Molossidae).\textsuperscript{15--18} The former two groups and a clade
+of the latter are endemic to the Neotropics, while the explosive
+adaptive radiations of the phyllostomids are responsible for the hotspot
+of bat diversity in the Amazon.\textsuperscript{27} Together, these
+clades of New World bats play host to a distinct regime of
+betacoronavirus coevolution.
 
@@
 structure and coevolutionary pressure: phylogeographic structure in bat
-hosts (and their diverse immune strategies)\textsuperscript{28} creates
+hosts (and their diverse immune strategies;)\textsuperscript{28} creates
 a landscape of selective pressure; the trajectory of viruses'
@@
 
+\leavevmode\hypertarget{ref-Ruiz-Aravena2022Ecology}{}%
+\CSLLeftMargin{3. }
+\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
+spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
+\textbf{20}, 299--314 (2022).}
+
 \leavevmode\hypertarget{ref-Sanchez2022Strategy}{}%
-\CSLLeftMargin{3. }
+\CSLLeftMargin{4. }
 \CSLRightInline{Sánchez, C. A. \emph{et al.} A strategy to assess
@@
 \leavevmode\hypertarget{ref-Agosta2010How}{}%
-\CSLLeftMargin{4. }
+\CSLLeftMargin{5. }
 \CSLRightInline{Agosta, S. J., Janz, N. \& Brooks, D. R. How specialists
@@
 \leavevmode\hypertarget{ref-Power2004Pathogen}{}%
-\CSLLeftMargin{5. }
+\CSLLeftMargin{6. }
 \CSLRightInline{Power, A. G. \& Mitchell, C. E. Pathogen Spillover in
@@
 \leavevmode\hypertarget{ref-Geoghegan2017Comparative}{}%
-\CSLLeftMargin{6. }
+\CSLLeftMargin{7. }
 \CSLRightInline{Geoghegan, J. L., Duchêne, S. \& Holmes, E. C.
@@
 \leavevmode\hypertarget{ref-Thompson2005Geographic}{}%
-\CSLLeftMargin{7. }
+\CSLLeftMargin{8. }
 \CSLRightInline{Thompson, J. N. \emph{The Geographic Mosaic of
@@
 \leavevmode\hypertarget{ref-Thompson1994Coevolutionary}{}%
-\CSLLeftMargin{8. }
+\CSLLeftMargin{9. }
 \CSLRightInline{Thompson, J. N. \emph{The Coevolutionary Process}.
@@
 \leavevmode\hypertarget{ref-Janzen1980When}{}%
-\CSLLeftMargin{9. }
+\CSLLeftMargin{10. }
 \CSLRightInline{Janzen, D. H. When is it Coevolution? \emph{Evolution}
@@
 \leavevmode\hypertarget{ref-Price2002Macroevolutionary}{}%
-\CSLLeftMargin{10. }
+\CSLLeftMargin{11. }
 \CSLRightInline{Price, P. W. \emph{Macroevolutionary Theory on
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2007Dos}{}%
-\CSLLeftMargin{11. }
+\CSLLeftMargin{12. }
 \CSLRightInline{Gomulkiewicz, R. \emph{et al.} Dos and don'ts of testing
@@
 \leavevmode\hypertarget{ref-Thrall2007Coevolution}{}%
-\CSLLeftMargin{12. }
+\CSLLeftMargin{13. }
 \CSLRightInline{Thrall, P. H., Hochberg, M. E., Burdon, J. J. \& Bever,
@@
 (2007).}
-
-\leavevmode\hypertarget{ref-Ruiz-Aravena2022Ecology}{}%
-\CSLLeftMargin{13. }
-\CSLRightInline{Ruiz-Aravena, M. \emph{et al.} Ecology, evolution and
-spillover of coronaviruses from bats. \emph{Nature Reviews Microbiology}
-\textbf{20}, 299--314 (2022).}
 

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6e36402c1b5194bfa5ebc5948ea574144eaefacd 🚀

@@
 \texttt{https://www.viralemergence.org/betacov} on
-Apr.~2022,\textsuperscript{\textbf{Becker2022OptPre?}} and filtered it
-to ``known'' hosts (established before the emergence of SARS-CoV-2) and
-``novel'' hosts (confirmed through sampling and competence assays since
-the initial data collection). The original database was assembled by a
-combination of data mining and literature surveys, including automated
-alerts on the ``bats'' and ``coronavirus'' keywords to identify novel
-empirical evidence of bats-betacoronaviruses associations; this yielded
-a total of 126 known hosts, 47 of which were novel hosts. This
-host--virus list of interactions was obtained through a comprehensive
-aggregation of GenBank data as well as systematic literature
-searches,\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Cohen2022SamStr?}}
+Apr.~2022,\textsuperscript{21} and filtered it to ``known'' hosts
+(established before the emergence of SARS-CoV-2) and ``novel'' hosts
+(confirmed through sampling and competence assays since the initial data
+collection). The original database was assembled by a combination of
+data mining and literature surveys, including automated alerts on the
+``bats'' and ``coronavirus'' keywords to identify novel empirical
+evidence of bats-betacoronaviruses associations; this yielded a total of
+126 known hosts, 47 of which were novel hosts. This host--virus list of
+interactions was obtained through a comprehensive aggregation of GenBank
+data as well as systematic literature searches,\textsuperscript{21,25}
 such that we have high confidence in its fitness for the purpose of
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{\textbf{IUCN2021IucRed?}} The IUCN data have been
+data.\textsuperscript{\textbf{IUCN2021Iuc?}} The IUCN data have been
 assembled to support wildlife conservation efforts, and therefore we do
@@
 priority. The range maps were subsequently rasterized using the
-\texttt{rasterize} function from
-\texttt{GDAL}\textsuperscript{\textbf{RouaultEven2022GdaOgr?}} at a
+\texttt{rasterize} function from \texttt{GDAL}\textsuperscript{55} at a
 resolution of approximately 100kmx100km at the equator. For every pixel
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{\textbf{Faith1992ConEva?}} based on a recent
-synthetic tree with robust time calibration, covering about 6000
-mammalian species.\textsuperscript{\textbf{Upham2019InfMam?}} Faith's PD
-measures the sum of unique branches from an arbitrary root to a set of
-tips, and comparatively larger values indicate a more phylogenetic
-diverse species pool. We measured phylogenetic diversity starting from
-the root of the entire tree (and not from Chiroptera); this bears no
-consequences on the resulting values, since all branches leading up to
-Chiroptera are only counted once per species pool, and (as we explain
-when describing the assembly of the composite risk map), all individual
-risk components are ranged in {[}0,1{]}. This measure incorporates a
-richness component, which we chose not to correct for; the
-interpretation of the phylogenetic diversity is therefore a weighted
-species richness, that accounts for phylogenetic over/under-dispersal in
-some places.
+Diversity\textsuperscript{56} based on a recent synthetic tree with
+robust time calibration, covering about 6000 mammalian
+species.\textsuperscript{57} Faith's PD measures the sum of unique
+branches from an arbitrary root to a set of tips, and comparatively
+larger values indicate a more phylogenetic diverse species pool. We
+measured phylogenetic diversity starting from the root of the entire
+tree (and not from Chiroptera); this bears no consequences on the
+resulting values, since all branches leading up to Chiroptera are only
+counted once per species pool, and (as we explain when describing the
+assembly of the composite risk map), all individual risk components are
+ranged in {[}0,1{]}. This measure incorporates a richness component,
+which we chose not to correct for; the interpretation of the
+phylogenetic diversity is therefore a weighted species richness, that
+accounts for phylogenetic over/under-dispersal in some places.
 
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{\textbf{Legendre2013BetDiv?}} LCBD works
-from a species-data matrix (traditionally noted as \(\mathbf{Y}\)),
-where species are rows and sites are columns, and a value of 1 indicates
-occurrence. We extracted the Y matrix assuming that every pixel
-represents a unique location, and following best
-practices\textsuperscript{\textbf{Legendre2019SpaTem?}} transformed it
-using Hellinger's distance to account for unequal bat richness at
+Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
+matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
+and sites are columns, and a value of 1 indicates occurrence. We
+extracted the Y matrix assuming that every pixel represents a unique
+location, and following best practices\textsuperscript{59} transformed
+it using Hellinger's distance to account for unequal bat richness at
 different pixels. The correction of raw community data is particularly
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results
-by\textsuperscript{\textbf{Dansereau2022EvaEco?}} shows that LCBD
+community. Recent results by\textsuperscript{60} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{\textbf{Albery2020PreGlo?}} This network stores
-pairwise values of viral community similarity, measured for all hosts
-(to maintain consistency with teh phylogenetic diversity measure) across
-all viruses; therefore, we consider that it accounts for some overall
-similarity in the way hosts deal with viruses, and not only
-betacoronaviruses. There is empirical evidence that capacity for
-cross-species transmission even between divergent species is generally
-high,\textsuperscript{55} especially for
-beta-coronaviruses.\textsuperscript{14} To project viral sharing values
-into a single value for every pixel, we averaged the pairwise scores.
-High values of the average sharing propensity means that this specific
-extant bat assemblage is likely to be proficient at exchanging viruses.
+mammals.\textsuperscript{61} This network stores pairwise values of
+viral community similarity, measured for all hosts (to maintain
+consistency with teh phylogenetic diversity measure) across all viruses;
+therefore, we consider that it accounts for some overall similarity in
+the way hosts deal with viruses, and not only betacoronaviruses. There
+is empirical evidence that capacity for cross-species transmission even
+between divergent species is generally high,\textsuperscript{62}
+especially for beta-coronaviruses.\textsuperscript{14} To project viral
+sharing values into a single value for every pixel, we averaged the
+pairwise scores. High values of the average sharing propensity means
+that this specific extant bat assemblage is likely to be proficient at
+exchanging viruses.
 
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{\textbf{Seekell2018GeoLak?}} In this approach,
-every risk component gets assigned a component in the RGB color model
-(phylogenetic diversity is green, compositional uniqueness is red, and
-viral sharing is blue). In order to achieve a valid RGB measure, all
-components are re-scaled to the {[}0,1{]} interval, so that a pixel with
-no sharing, no phylogenetic diversity, and no compositional uniqueness
-is black, and a pixel with maximal values for each is white. This
-additive model conveys both the intensity of the overall risk, but also
-the nature of the risk as colors diverge towards combinations of values
-for three risk components. Out of the possible combinations, the most
-risky in terms or rapid diversification and spillover potential is high
-phylogenetic diversity and low viral
-sharing,\textsuperscript{\textbf{Gomulkiewicz2000HotSpo?}} in that this
-allows multiple independent host-virus coevolutionary dynamics to take
-place in the same location. In the colorimetric space, this correspond
-to yellow -- because the HSV space is more amenable to calculations for
-feature extraction,\textsuperscript{\textbf{Keke2010StuSki?}} we
-measured the risk level by calculating the angular distance of the hue
-of each pixel to a reference value of 60 (yellow), and weighted this
-risk level by the value component. Specifically, given a pixel with
-colorimetric coordinates \((h,s,v)\), its ranged weighted risk value is
+model.\textsuperscript{63} In this approach, every risk component gets
+assigned a component in the RGB color model (phylogenetic diversity is
+green, compositional uniqueness is red, and viral sharing is blue). In
+order to achieve a valid RGB measure, all components are re-scaled to
+the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
+diversity, and no compositional uniqueness is black, and a pixel with
+maximal values for each is white. This additive model conveys both the
+intensity of the overall risk, but also the nature of the risk as colors
+diverge towards combinations of values for three risk components. Out of
+the possible combinations, the most risky in terms or rapid
+diversification and spillover potential is high phylogenetic diversity
+and low viral sharing,\textsuperscript{64} in that this allows multiple
+independent host-virus coevolutionary dynamics to take place in the same
+location. In the colorimetric space, this correspond to yellow --
+because the HSV space is more amenable to calculations for feature
+extraction,\textsuperscript{65} we measured the risk level by
+calculating the angular distance of the hue of each pixel to a reference
+value of 60 (yellow), and weighted this risk level by the value
+component. Specifically, given a pixel with colorimetric coordinates
+\((h,s,v)\), its ranged weighted risk value is
 
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{\textbf{Katoh2013MafMul?}} v1.4.0 (Algorithm
-FFT-NS-2, Scoring matrix 200PAM / k=2, gap open penalty 1.53m offset
-value 0.123) and a maximum likelihood tree reconstructed in
-IQ-TREE\textsuperscript{\textbf{Nguyen2015IqtFas?}} v1.6.12 with
-ModelFinder\textsuperscript{\textbf{Kalyaanamoorthy2017ModFas?}}
-ultrafast bootstrap
-approximation\textsuperscript{\textbf{Hoang2018UfbImp?}} with a general
+MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
+likelihood tree reconstructed in
+IQ-TREE\textsuperscript{\textbf{Nguyen2015IqTree?}} v1.6.12 with
+ModelFinder\textsuperscript{\textbf{Kalyaanamoorthy2017ModelFinder?}}
+ultrafast bootstrap approximation\textsuperscript{67} with a general
 time reversible model with empirical base frequencies and the
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{\textbf{Isaac2007MamEdg?}} for each of
-the viruses in the tree, then averaged these at the bat species level,
-projected these values onto their geographic distributions, and averaged
-across every bat found in a given pixel. As such, this can be thought of
-as a map of the mean evolutionary distinctiveness of the known viral
-community believed to be associated with a particular subset of bats
-present.
+distinctiveness\textsuperscript{68} for each of the viruses in the tree,
+then averaged these at the bat species level, projected these values
+onto their geographic distributions, and averaged across every bat found
+in a given pixel. As such, this can be thought of as a map of the mean
+evolutionary distinctiveness of the known viral community believed to be
+associated with a particular subset of bats present.
 
@@
 
+\leavevmode\hypertarget{ref-RouaultEven2022Gdal}{}%
+\CSLLeftMargin{55. }
+\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
+Abstraction software Library}. (Zenodo, 2022).
+doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}
+
+\leavevmode\hypertarget{ref-Faith1992Conservation}{}%
+\CSLLeftMargin{56. }
+\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
+diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}
+
+\leavevmode\hypertarget{ref-Upham2019Inferring}{}%
+\CSLLeftMargin{57. }
+\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
+mammal tree: Species-level sets of phylogenies for questions in ecology,
+evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
+(2019).}
+
+\leavevmode\hypertarget{ref-Legendre2013Beta}{}%
+\CSLLeftMargin{58. }
+\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
+variance of community data: Dissimilarity coefficients and partitioning.
+\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}
+
+\leavevmode\hypertarget{ref-Legendre2019Spatial}{}%
+\CSLLeftMargin{59. }
+\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
+of beta diversity in the Barro Colorado Island forest dynamics plot,
+Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}
+
+\leavevmode\hypertarget{ref-Dansereau2022Evaluating}{}%
+\CSLLeftMargin{60. }
+\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
+ecological uniqueness over broad spatial extents using species
+distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}
+
+\leavevmode\hypertarget{ref-Albery2020Predicting}{}%
+\CSLLeftMargin{61. }
+\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
+Predicting the global mammalian viral sharing network using
+phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}
+
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
@@
 
+\leavevmode\hypertarget{ref-Seekell2018Geography}{}%
+\CSLLeftMargin{63. }
+\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
+geography of lake carbon cycling. \emph{Limnology and Oceanography
+Letters} \textbf{3}, 49--56 (2018).}
+
+\leavevmode\hypertarget{ref-Gomulkiewicz2000Hot}{}%
+\CSLLeftMargin{64. }
+\CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
+S. L. \& Hochberg, M. E. Hot Spots, Cold Spots, and the Geographic
+Mosaic Theory of Coevolution. \emph{The American Naturalist}
+\textbf{156}, 156--174 (2000).}
+
+\leavevmode\hypertarget{ref-Keke2010Study}{}%
+\CSLLeftMargin{65. }
+\CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
+image segmentation used by Fuzzy-c-means arithmetic. in \emph{2010
+Seventh International Conference on Fuzzy Systems and Knowledge
+Discovery} vol. 2 612--615 (2010).}
+
+\leavevmode\hypertarget{ref-Katoh2013Mafft}{}%
+\CSLLeftMargin{66. }
+\CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
+Alignment Software Version 7: Improvements in Performance and Usability.
+\emph{Molecular Biology and Evolution} \textbf{30}, 772--780 (2013).}
+
+\leavevmode\hypertarget{ref-Hoang2018Ufboot2}{}%
+\CSLLeftMargin{67. }
+\CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
+Q. \& Vinh, L. S. UFBoot2: Improving the Ultrafast Bootstrap
+Approximation. \emph{Molecular Biology and Evolution} \textbf{35},
+518--522 (2018).}
+
+\leavevmode\hypertarget{ref-Isaac2007Mammals}{}%
+\CSLLeftMargin{68. }
+\CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
+\& Baillie, J. E. M. Mammals on the EDGE: Conservation Priorities Based
+on Threat and Phylogeny. \emph{PLOS ONE} \textbf{2}, e296 (2007).}
+
 \end{CSLReferences}

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@4a3bbbe3ed7da6b936352d3d80f46c926d4b515c 🚀

@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method
-from,\textsuperscript{\textbf{Kreft2007GloPat?},\textbf{Kreft2010FraDel?}}
-who proposed a phylogenetic method for the delineation of animal
+this idea, we loosely adapted a method from,\textsuperscript{69,70} who
+proposed a phylogenetic method for the delineation of animal
 biogeographic regions. In their original method, a distance matrix -
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in
-\texttt{ape};\textsuperscript{\textbf{Paradis2019ApeEnv?}} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Kreft2007Global}{}%
+\CSLLeftMargin{69. }
+\CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
+of vascular plant diversity. \emph{Proceedings of the National Academy
+of Sciences} \textbf{104}, 5925--5930 (2007).}
+
+\leavevmode\hypertarget{ref-Kreft2010Framework}{}%
+\CSLLeftMargin{70. }
+\CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
+biogeographical regions based on species distributions. \emph{Journal of
+Biogeography} \textbf{37}, 2029--2053 (2010).}
+
+\leavevmode\hypertarget{ref-Paradis2019Ape}{}%
+\CSLLeftMargin{71. }
+\CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for
+modern phylogenetics and evolutionary analyses in R.
+\emph{Bioinformatics} \textbf{35}, 526--528 (2019).}
+
 \end{CSLReferences}

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@8cd06e661973e53a534530b583905c6fc05c4bd5 🚀

@@
 ecosystems within those zones, can be a win-win intervention for
-both.\textsuperscript{53,\textbf{Hopkins2021HowIde?},\textbf{Plowright2021Lan?}}
-As we scale these predictions down in space to finer spatial resolutions
-to guide public health actions,\textsuperscript{33} the incorporation of
-human activity predictors will become more
-importyant.\textsuperscript{54}
+both.\textsuperscript{53--55} As we scale these predictions down in
+space to finer spatial resolutions to guide public health
+actions,\textsuperscript{33} the incorporation of human activity
+predictors will become more importyant.\textsuperscript{56}
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{\textbf{IUCN2021Iuc?}} The IUCN data have been
-assembled to support wildlife conservation efforts, and therefore we do
-not expect that they are biased by wildlife disease sampling efforts or
-priority. The range maps were subsequently rasterized using the
-\texttt{rasterize} function from \texttt{GDAL}\textsuperscript{55} at a
-resolution of approximately 100kmx100km at the equator. For every pixel
-in the resulting raster where at least one bat host of
-\emph{Betacoronavirus} was present, we extract the species pool (list of
-all known bat hosts), which was used to calculate the following risk
-assessment components: bat phylogenetic diversity, bat compositional
-uniqueness, and predicted viral sharing risk.
+data.\textsuperscript{57} The IUCN data have been assembled to support
+wildlife conservation efforts, and therefore we do not expect that they
+are biased by wildlife disease sampling efforts or priority. The range
+maps were subsequently rasterized using the \texttt{rasterize} function
+from \texttt{GDAL}\textsuperscript{58} at a resolution of approximately
+100kmx100km at the equator. For every pixel in the resulting raster
+where at least one bat host of \emph{Betacoronavirus} was present, we
+extract the species pool (list of all known bat hosts), which was used
+to calculate the following risk assessment components: bat phylogenetic
+diversity, bat compositional uniqueness, and predicted viral sharing
+risk.
 
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{56} based on a recent synthetic tree with
+Diversity\textsuperscript{59} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{57} Faith's PD measures the sum of unique
+species.\textsuperscript{60} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{58} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{61} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{59} transformed
+location, and following best practices\textsuperscript{62} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{60} shows that LCBD
+community. Recent results by\textsuperscript{63} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{61} This network stores pairwise values of
+mammals.\textsuperscript{64} This network stores pairwise values of
 viral community similarity, measured for all hosts (to maintain
@@
 is empirical evidence that capacity for cross-species transmission even
-between divergent species is generally high,\textsuperscript{62}
+between divergent species is generally high,\textsuperscript{65}
 especially for beta-coronaviruses.\textsuperscript{14} To project viral
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{63} In this approach, every risk component gets
+model.\textsuperscript{66} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{64} in that this allows multiple
+and low viral sharing,\textsuperscript{67} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{65} we measured the risk level by
+extraction,\textsuperscript{68} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{69} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in
-IQ-TREE\textsuperscript{\textbf{Nguyen2015IqTree?}} v1.6.12 with
-ModelFinder\textsuperscript{\textbf{Kalyaanamoorthy2017ModelFinder?}}
-ultrafast bootstrap approximation\textsuperscript{67} with a general
-time reversible model with empirical base frequencies and the
-5-discrete-rate-category FreeRaye model of nucleotide substitution
-(GTR+F+R5).
+likelihood tree reconstructed in IQ-TREE\textsuperscript{70} v1.6.12
+with ModelFinder\textsuperscript{71} ultrafast bootstrap
+approximation\textsuperscript{72} with a general time reversible model
+with empirical base frequencies and the 5-discrete-rate-category
+FreeRaye model of nucleotide substitution (GTR+F+R5).
 
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{68} for each of the viruses in the tree,
+distinctiveness\textsuperscript{73} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{69,70} who
+this idea, we loosely adapted a method from,\textsuperscript{74,75} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{71} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{76} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Hopkins2021How}{}%
+\CSLLeftMargin{53. }
+\CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
+interventions that benefit human health and conservation. \emph{Nature
+Sustainability} \textbf{4}, 298--304 (2021).}
+
+\leavevmode\hypertarget{ref-Plowright2021Land}{}%
+\CSLLeftMargin{54. }
+\CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
+spillover: A call to action to safeguard environmental, animal, and
+human health. \emph{The Lancet Planetary Health} \textbf{5}, e237--e245
+(2021).}
+
 \leavevmode\hypertarget{ref-OHHLEP2022One}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{55. }
 \CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 
+\leavevmode\hypertarget{ref-IUCN2021Iucn}{}%
+\CSLLeftMargin{57. }
+\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
+\url{https://www.iucnredlist.org} (2021).}
+
 \leavevmode\hypertarget{ref-RouaultEven2022Gdal}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992Conservation}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{59. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019Inferring}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013Beta}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019Spatial}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022Evaluating}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020Predicting}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
@@
 \leavevmode\hypertarget{ref-Seekell2018Geography}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000Hot}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010Study}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013Mafft}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 
+\leavevmode\hypertarget{ref-Nguyen2015Iqtree}{}%
+\CSLLeftMargin{70. }
+\CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
+B. Q. IQ-TREE: A Fast and Effective Stochastic Algorithm for Estimating
+Maximum-Likelihood Phylogenies. \emph{Molecular Biology and Evolution}
+\textbf{32}, 268--274 (2015).}
+
+\leavevmode\hypertarget{ref-Kalyaanamoorthy2017Modelfinder}{}%
+\CSLLeftMargin{71. }
+\CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
+Haeseler, A. \& Jermiin, L. S. ModelFinder: Fast model selection for
+accurate phylogenetic estimates. \emph{Nature Methods} \textbf{14},
+587--589 (2017).}
+
 \leavevmode\hypertarget{ref-Hoang2018Ufboot2}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007Mammals}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{73. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007Global}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{74. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010Framework}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{75. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019Ape}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{76. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for

### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f2a5f0e2f65ea0f1f7e79efb46a46d9f20839e4e 🚀

@@
 diverse mammalian orders, with more than 1,400 estimated
-species.\textsuperscript{\textbf{Peixoto2018SynEco?},\textbf{Simmons2020BatSpe?}}
-They exhibit a broad variety of habitat use, behaviour, and feeding
-strategies, putting them at key positions in the delivery and
-provisioning of several ecosystem services, tied to important
-ecosystem-derived benefits to humans.\textsuperscript{39} Over
+species.\textsuperscript{39,40} They exhibit a broad variety of habitat
+use, behaviour, and feeding strategies, putting them at key positions in
+the delivery and provisioning of several ecosystem services, tied to
+important ecosystem-derived benefits to humans.\textsuperscript{41} Over
 two-thirds of bats are known to be either obligate or facultative
 insectivores, therefore actively contributing for agricultural pest
-control,\textsuperscript{40,41} and vectors of pathogens that put a risk
-on human health;\textsuperscript{42,43} some other species are essential
-links in many seed-dispersal networks.\textsuperscript{44} However, many
+control,\textsuperscript{42,43} and vectors of pathogens that put a risk
+on human health;\textsuperscript{44,45} some other species are essential
+links in many seed-dispersal networks.\textsuperscript{46} However, many
 of these species face a high risk of extinction, particularly given
@@
 that is supported by previous work indicating a higher rate of
-codivergence in Latin America.\textsuperscript{2,45} In contrast, in
+codivergence in Latin America.\textsuperscript{2,47} In contrast, in
 southeast Asia, host richness and viral distinctiveness are high but
@@
 from both the SARS-CoV and SARS-CoV-2 branches of the Sarbecovirus
-lineage).\textsuperscript{46}Both of these regions are priority areas
+lineage).\textsuperscript{48}Both of these regions are priority areas
 for sampling, especially given predictions that they contain many bat
@@
 creating a hotspot of elevated cross-species transmission unique to the
-region.\textsuperscript{33,47}
+region.\textsuperscript{33,49}
 
@@
 could spread relatively easily due to encountering little immune
-resistance.\textsuperscript{48} With the right combination of viral
+resistance.\textsuperscript{50} With the right combination of viral
 traits, such as low disease-induced mortality or high transmission rate,
@@
 greater contact with potential bridge hosts like domesticated
-species.\textsuperscript{31,49--51} This represents a challenge for both
-conservation strategies and pandemic prevention,\textsuperscript{52} but
+species.\textsuperscript{31,51--53} This represents a challenge for both
+conservation strategies and pandemic prevention,\textsuperscript{54} but
 identifying areas at risk, and protecting the health of bats and
 ecosystems within those zones, can be a win-win intervention for
-both.\textsuperscript{53--55} As we scale these predictions down in
+both.\textsuperscript{55--57} As we scale these predictions down in
 space to finer spatial resolutions to guide public health
 actions,\textsuperscript{33} the incorporation of human activity
-predictors will become more importyant.\textsuperscript{56}
+predictors will become more importyant.\textsuperscript{58}
 
@@
 from the previous step, according to recent IUCN
-data.\textsuperscript{57} The IUCN data have been assembled to support
+data.\textsuperscript{59} The IUCN data have been assembled to support
 wildlife conservation efforts, and therefore we do not expect that they
@@
 maps were subsequently rasterized using the \texttt{rasterize} function
-from \texttt{GDAL}\textsuperscript{58} at a resolution of approximately
+from \texttt{GDAL}\textsuperscript{60} at a resolution of approximately
 100kmx100km at the equator. For every pixel in the resulting raster
@@
 For every pixel, we measured Faith's Phylogenetic
-Diversity\textsuperscript{59} based on a recent synthetic tree with
+Diversity\textsuperscript{61} based on a recent synthetic tree with
 robust time calibration, covering about 6000 mammalian
-species.\textsuperscript{60} Faith's PD measures the sum of unique
+species.\textsuperscript{62} Faith's PD measures the sum of unique
 branches from an arbitrary root to a set of tips, and comparatively
@@
 For every species pool, we measured its Local Contribution to
-Beta-Diversity;\textsuperscript{61} LCBD works from a species-data
+Beta-Diversity;\textsuperscript{63} LCBD works from a species-data
 matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
@@
 extracted the Y matrix assuming that every pixel represents a unique
-location, and following best practices\textsuperscript{62} transformed
+location, and following best practices\textsuperscript{64} transformed
 it using Hellinger's distance to account for unequal bat richness at
@@
 what is expected knowing the entire matrix, i.e.~a more unique
-community. Recent results by\textsuperscript{63} shows that LCBD
+community. Recent results by\textsuperscript{65} shows that LCBD
 measures are robust with regards to spatial scale, and are therefore
@@
 of phylogenetic distance and geographic range overlap across
-mammals.\textsuperscript{64} This network stores pairwise values of
+mammals.\textsuperscript{66} This network stores pairwise values of
 viral community similarity, measured for all hosts (to maintain
@@
 is empirical evidence that capacity for cross-species transmission even
-between divergent species is generally high,\textsuperscript{65}
+between divergent species is generally high,\textsuperscript{67}
 especially for beta-coronaviruses.\textsuperscript{14} To project viral
@@
 uniqueness, and viral sharing) using an additive color
-model.\textsuperscript{66} In this approach, every risk component gets
+model.\textsuperscript{68} In this approach, every risk component gets
 assigned a component in the RGB color model (phylogenetic diversity is
@@
 diversification and spillover potential is high phylogenetic diversity
-and low viral sharing,\textsuperscript{67} in that this allows multiple
+and low viral sharing,\textsuperscript{69} in that this allows multiple
 independent host-virus coevolutionary dynamics to take place in the same
@@
 because the HSV space is more amenable to calculations for feature
-extraction,\textsuperscript{68} we measured the risk level by
+extraction,\textsuperscript{70} we measured the risk level by
 calculating the angular distance of the hue of each pixel to a reference
@@
 betacoronavirus RdRp sequences were then aligned using
-MAFFT\textsuperscript{69} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
+MAFFT\textsuperscript{71} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
 200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
-likelihood tree reconstructed in IQ-TREE\textsuperscript{70} v1.6.12
-with ModelFinder\textsuperscript{71} ultrafast bootstrap
-approximation\textsuperscript{72} with a general time reversible model
+likelihood tree reconstructed in IQ-TREE\textsuperscript{72} v1.6.12
+with ModelFinder\textsuperscript{73} ultrafast bootstrap
+approximation\textsuperscript{74} with a general time reversible model
 with empirical base frequencies and the 5-discrete-rate-category
@@
 viruses. To calculate this, we derived the fair proportions evolutionary
-distinctiveness\textsuperscript{73} for each of the viruses in the tree,
+distinctiveness\textsuperscript{75} for each of the viruses in the tree,
 then averaged these at the bat species level, projected these values
@@
 betacoronaviruses should track the biogeography of their hosts. To test
-this idea, we loosely adapted a method from,\textsuperscript{74,75} who
+this idea, we loosely adapted a method from,\textsuperscript{76,77} who
 proposed a phylogenetic method for the delineation of animal
@@
 between the different viruses in the betacoronaviruses tree by using the
-cophenetic function in \texttt{ape};\textsuperscript{76} subsequently,
+cophenetic function in \texttt{ape};\textsuperscript{78} subsequently,
 we take a principal components analysis of that distance matrix (readily
@@
 
+\leavevmode\hypertarget{ref-Peixoto2018Synthesis}{}%
+\CSLLeftMargin{39. }
+\CSLRightInline{Peixoto, F. P., Braga, P. H. P. \& Mendes, P. A
+synthesis of ecological and evolutionary determinants of bat diversity
+across spatial scales. \emph{BMC ecology} \textbf{18}, 18 (2018).}
+
+\leavevmode\hypertarget{ref-Simmons2020Bat}{}%
+\CSLLeftMargin{40. }
+\CSLRightInline{Simmons, N. B. \& Cirranello, A. L. Bat Species of the
+World: A taxonomic and geographic database. \url{https://batnames.org/}
+(2020).}
+
 \leavevmode\hypertarget{ref-Kasso2013Ecological}{}%
-\CSLLeftMargin{39. }
+\CSLLeftMargin{41. }
 \CSLRightInline{Kasso, M. \& Balakrishnan, M. Ecological and Economic
@@
 \leavevmode\hypertarget{ref-Voigt2016Bats}{}%
-\CSLLeftMargin{40. }
+\CSLLeftMargin{42. }
 \CSLRightInline{\emph{Bats in the Anthropocene: Conservation of Bats in
@@
 \leavevmode\hypertarget{ref-Williams-Guillen2008Bats}{}%
-\CSLLeftMargin{41. }
+\CSLLeftMargin{43. }
 \CSLRightInline{Williams-Guillén, K., Perfecto, I. \& Vandermeer, J.
@@
 \leavevmode\hypertarget{ref-Gonsalves2013Mosquito}{}%
-\CSLLeftMargin{42. }
+\CSLLeftMargin{44. }
 \CSLRightInline{Gonsalves, L., Bicknell, B., Law, B., Webb, C. \&
@@
 \leavevmode\hypertarget{ref-Gonsalves2013Mosquitoes}{}%
-\CSLLeftMargin{43. }
+\CSLLeftMargin{45. }
 \CSLRightInline{Gonsalves, L., Lamb, S., Webb, C., Law, B. \& Monamy, V.
@@
 \leavevmode\hypertarget{ref-Mello2011Missing}{}%
-\CSLLeftMargin{44. }
+\CSLLeftMargin{46. }
 \CSLRightInline{Mello, M. A. R. \emph{et al.} The Missing Part of Seed
@@
 \leavevmode\hypertarget{ref-Caraballo2022Crossspecies}{}%
-\CSLLeftMargin{45. }
+\CSLLeftMargin{47. }
 \CSLRightInline{Caraballo, D. A. Cross-Species Transmission of Bat
@@
 \leavevmode\hypertarget{ref-Wu2021Comprehensive}{}%
-\CSLLeftMargin{46. }
+\CSLLeftMargin{48. }
 \CSLRightInline{Wu, Z. \emph{et al.} \emph{A comprehensive survey of bat
@@
 \leavevmode\hypertarget{ref-Carlson2022Climate}{}%
-\CSLLeftMargin{47. }
+\CSLLeftMargin{49. }
 \CSLRightInline{Carlson, C. J. \emph{et al.} Climate change increases
@@
 \leavevmode\hypertarget{ref-Plowright2011Urban}{}%
-\CSLLeftMargin{48. }
+\CSLLeftMargin{50. }
 \CSLRightInline{Plowright, R. K. \emph{et al.} Urban habituation,
@@
 \leavevmode\hypertarget{ref-Alves2018Geographic}{}%
-\CSLLeftMargin{49. }
+\CSLLeftMargin{51. }
 \CSLRightInline{Alves, D. M. C. C., Diniz-Filho, J. A. F., da Silva e
@@
 \leavevmode\hypertarget{ref-Treitler2016Effect}{}%
-\CSLLeftMargin{50. }
+\CSLLeftMargin{52. }
 \CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
@@
 \leavevmode\hypertarget{ref-Mendenhall2014Predictinga}{}%
-\CSLLeftMargin{51. }
+\CSLLeftMargin{53. }
 \CSLRightInline{Mendenhall, C. D., Karp, D. S., Meyer, C. F. J., Hadly,
@@
 \leavevmode\hypertarget{ref-Amman2011Investigating}{}%
-\CSLLeftMargin{52. }
+\CSLLeftMargin{54. }
 \CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
@@
 \leavevmode\hypertarget{ref-Hopkins2021How}{}%
-\CSLLeftMargin{53. }
+\CSLLeftMargin{55. }
 \CSLRightInline{Hopkins, S. R. \emph{et al.} How to identify win--win
@@
 \leavevmode\hypertarget{ref-Plowright2021Land}{}%
-\CSLLeftMargin{54. }
+\CSLLeftMargin{56. }
 \CSLRightInline{Plowright, R. K. \emph{et al.} Land use-induced
@@
 \leavevmode\hypertarget{ref-OHHLEP2022One}{}%
-\CSLLeftMargin{55. }
+\CSLLeftMargin{57. }
 \CSLRightInline{OHHLEP \emph{et al.} One Health: A new definition for a
@@
 \leavevmode\hypertarget{ref-Ka-WaiHui2006Reasons}{}%
-\CSLLeftMargin{56. }
+\CSLLeftMargin{58. }
 \CSLRightInline{Ka-Wai Hui, E. Reasons for the increase in emerging and
@@
 \leavevmode\hypertarget{ref-IUCN2021Iucn}{}%
-\CSLLeftMargin{57. }
+\CSLLeftMargin{59. }
 \CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
@@
 \leavevmode\hypertarget{ref-RouaultEven2022Gdal}{}%
-\CSLLeftMargin{58. }
+\CSLLeftMargin{60. }
 \CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
@@
 \leavevmode\hypertarget{ref-Faith1992Conservation}{}%
-\CSLLeftMargin{59. }
+\CSLLeftMargin{61. }
 \CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
@@
 \leavevmode\hypertarget{ref-Upham2019Inferring}{}%
-\CSLLeftMargin{60. }
+\CSLLeftMargin{62. }
 \CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
@@
 \leavevmode\hypertarget{ref-Legendre2013Beta}{}%
-\CSLLeftMargin{61. }
+\CSLLeftMargin{63. }
 \CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
@@
 \leavevmode\hypertarget{ref-Legendre2019Spatial}{}%
-\CSLLeftMargin{62. }
+\CSLLeftMargin{64. }
 \CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
@@
 \leavevmode\hypertarget{ref-Dansereau2022Evaluating}{}%
-\CSLLeftMargin{63. }
+\CSLLeftMargin{65. }
 \CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
@@
 \leavevmode\hypertarget{ref-Albery2020Predicting}{}%
-\CSLLeftMargin{64. }
+\CSLLeftMargin{66. }
 \CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
@@
 \leavevmode\hypertarget{ref-Mollentze2020Viral}{}%
-\CSLLeftMargin{65. }
+\CSLLeftMargin{67. }
 \CSLRightInline{Mollentze, N. \& Streicker, D. G. Viral zoonotic risk is
@@
 \leavevmode\hypertarget{ref-Seekell2018Geography}{}%
-\CSLLeftMargin{66. }
+\CSLLeftMargin{68. }
 \CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
@@
 \leavevmode\hypertarget{ref-Gomulkiewicz2000Hot}{}%
-\CSLLeftMargin{67. }
+\CSLLeftMargin{69. }
 \CSLRightInline{Gomulkiewicz, R., Thompson, J. N., Holt, R. D., Nuismer,
@@
 \leavevmode\hypertarget{ref-Keke2010Study}{}%
-\CSLLeftMargin{68. }
+\CSLLeftMargin{70. }
 \CSLRightInline{Keke, S., Peng, Z. \& Guohui, L. Study on skin color
@@
 \leavevmode\hypertarget{ref-Katoh2013Mafft}{}%
-\CSLLeftMargin{69. }
+\CSLLeftMargin{71. }
 \CSLRightInline{Katoh, K. \& Standley, D. M. MAFFT Multiple Sequence
@@
 \leavevmode\hypertarget{ref-Nguyen2015Iqtree}{}%
-\CSLLeftMargin{70. }
+\CSLLeftMargin{72. }
 \CSLRightInline{Nguyen, L.-T., Schmidt, H. A., von Haeseler, A. \& Minh,
@@
 \leavevmode\hypertarget{ref-Kalyaanamoorthy2017Modelfinder}{}%
-\CSLLeftMargin{71. }
+\CSLLeftMargin{73. }
 \CSLRightInline{Kalyaanamoorthy, S., Minh, B. Q., Wong, T. K. F., von
@@
 \leavevmode\hypertarget{ref-Hoang2018Ufboot2}{}%
-\CSLLeftMargin{72. }
+\CSLLeftMargin{74. }
 \CSLRightInline{Hoang, D. T., Chernomor, O., von Haeseler, A., Minh, B.
@@
 \leavevmode\hypertarget{ref-Isaac2007Mammals}{}%
-\CSLLeftMargin{73. }
+\CSLLeftMargin{75. }
 \CSLRightInline{Isaac, N. J. B., Turvey, S. T., Collen, B., Waterman, C.
@@
 \leavevmode\hypertarget{ref-Kreft2007Global}{}%
-\CSLLeftMargin{74. }
+\CSLLeftMargin{76. }
 \CSLRightInline{Kreft, H. \& Jetz, W. Global patterns and determinants
@@
 \leavevmode\hypertarget{ref-Kreft2010Framework}{}%
-\CSLLeftMargin{75. }
+\CSLLeftMargin{77. }
 \CSLRightInline{Kreft, H. \& Jetz, W. A framework for delineating
@@
 \leavevmode\hypertarget{ref-Paradis2019Ape}{}%
-\CSLLeftMargin{76. }
+\CSLLeftMargin{78. }
 \CSLRightInline{Paradis, E. \& Schliep, K. Ape 5.0: An environment for
